### 2019-09-07 paper/virtual_cytoplasm.tex
paper file

\documentclass[journal=jacsat,manuscript=article]{achemso}
% \linespread{1.5}
% \usepackage{amssymb,amsmath,graphics,epsfig}
\usepackage{amsmath}
\usepackage[font=small,labelfont=bf]{caption}
\setkeys{acs}{super = true}
\setcitestyle{super,open={},close={}}
\def\citenumfont{}

% \usepackage[version=3]{mhchem} 
\usepackage{color}
\newcommand{\eqnref}[1]{Eqn.~\plainref{#1}}
\newcommand{\figref}[1]{Fig.~\plainref{#1}}
\newcommand{\tabref}[1]{Table~\plainref{#1}}

\title{Cytoplasm Simulation}

\author{David van der Spoel$^1$}

\begin{document}

\maketitle

% \begin{affiliations}
%  \item Science for Life Laboratory, Department of Cell and Molecular Biology. Uppsala University, SE-751 05 Uppsala, Sweden
%  \item Department of Biological Science and Engineering, School of Chemistry and Biological Engineering, University of Science and Technology Beijing, 100083 Beijing, China
% \end{affiliations}

\begin{abstract}
%For Nature, the abstract is really an introductory paragraph set
%in bold type.  This paragraph must be ``fully referenced'' and
%less than 180 words for Letters.  This is the thing that is
%supposed to be aimed at people from other disciplines and is
%arguably the most important part to getting your paper past the
%editors.  End this paragraph with a sentence like ``Here we
%show...'' or something similar.

\end{abstract}
\section*{Introduction}

Biomolecules move and function in an environment densely packed with high concentrations of macromolecules. The presence of macromolecules leads to steric effect due to excluded volume effect and intermolecular attrative/repulsive forces due to distributed charges on the surface of macromolecules.

Structure and dynamics of biomolecules are well characterized {\em in vitro}, however how these features differ {\em in vivo} remains unclear. 
Investigating biomolecular properties {\em in vivo} is possible through development in the fields of nuclear magnetic resonance~\cite{reckel2007,pielak2008}, or fluoroscence spectroscopies~\cite{ignatova2004,xie2008,English2011}.
An alternative is to use computational models and simulations techniques. Biomolecular simulations are often carried out under dilute conditions or simple models of macromolecular crowdings~\cite{Spiga2014a,Harada2012a,Nawrocki2017a}. However, more attemps in modeling bacterial cytoplasm have been made recently~\cite{Mcguffee2010,Yu2016a}.

Here we report on a model of {\em Escherichia coli} cytoplasm at atomistic level. The challenges in modeling and simulation with Molecular Dynamics are pointed and discussed and solutions are provided.

This work is the first model of {\em E. coli} at atomistic resolution that spans cellular dynamics on a microsecond scale.

\section*{Material and Method}
\subsection*{Cytoplasm model}
In order to investigate the effects of crowding on cellular components in a more realistic way, we built a  cytoplasm model based on {\em Escheria coli} as a model organism~\cite{Dong1996,Bennett2009,Link1997,Mcguffee2010}

Because of restriction of computational resources, incorporating a full list of proteins, nucleic acids and metabolites in full atomistic model is prohibitive.

In order to build the RNA fraction of our cytoplasm model, we consider data showing that 2.9\% of the dry weight of \textit{E. coli} is composed by tRNAs, which correspond to 74\% of the dry weight of non-ribosomal RNAs, while 55.0\% of it is composed of proteins~\cite{schaechter2006}. Thus, the tRNA weight should correspond to 5\% of the total weight of all protein chains used in our model. Specifically, tRNA(Phe) was selected as a representative due to the availability of a recent crystallographic structure~\cite{Byrne2015}. Copies were added to the simulation box to account for the correct protein/RNA weight ratio (Table X).

The metabolite fraction of our cytoplasm model was build based on data showing that the number of metabolite molecules in the cytoplasm of \textit{E. coli} is about 42.86 times higher than the number of proteins. We considered the most abundant molecule as representative of each metabolite class, i.e. Glutamate for amino acids, ATP for nucleotides, FBP for central carbon intermediates and Glutathione redox cofactors. The copy number for each molecule was calculated as the ratio of their experimentally observed concentration in {\em E. coli}.

\subsection*{Components Preparation}
The proteins and tRNA were downloaded from Protein Data Bank (PDB). We looked for the proteins structures that were either from or expressed in {\em E-coli}. In case of 1U22 (MetE) and 2EIP (Ppa) we used a loop-closure modelng tool based on Random Coordinate Descent (RCD) method~\cite{Chys2013} to correct the information for missing residues. The four metabolites were parametrized using GAFF and Antechamber.

\subsection*{All-Atom Molecular Dynamics Simulation}

% General Simulation setup:
% \begin{itemize}
% \item Force Field
% \item Water model
% \item Gromacs
% \item Temperature and pressure
% \item (details of soup box)
% \item Minimization 
% \item Equilibration
% \item Production
% \end{itemize}

% \begin{itemize}
% \item Box building
% \item Droplets
% \item Shrinking
% \end{itemize}

{\bf General Simulation Setup: }The proteins were simulated at 30\% biomolecular mass fraction in  a physiological salt concentration (0.15M NaCl). For all simulations, Amber99SB-ws force field was used~\cite{Best2014a} in combination with the TIP4P/2005 water model~\cite{Abascal2005b}. Electrostatic interactions were treated using the particle mesh Ewald algorithm~\cite{Essmann1995a}. All chemical bonds were constrained at their equilibrium length using the LINCS algorithm~\cite{Hess2008b} allowing an integration time step of 2 fs. Temperature was controlled at 310 K using the v-rescale algorithm~\cite{Bussi2007a} and a coupling time of 0.5 ps. The pressure was controlled at 1 bar using the Parrinello-Rahman algorithm~\cite{Parrinello1981a} with a time constant of 10 ps.

\begin{table}[h!]
\label{tbl:component}
\centering
\begin{tabular}{lcc}
\hline
Class & Name (PDB ID) & Number\\
\hline
8*Protein & TufA (1DG1~\cite{Abel1996}) & 6\\
  & MetE (1U22~\cite{Ferrer2004}) & 7\\
  & IcdA (1P8F~\cite{Mesecar2000}) & 2\\
  & AhpC (1YEP~\cite{Parsonage2005}) & 1\\
  & CspC (1MJC~\cite{Schindelin1994}) & 3\\
  & Ppa (2EIP~\cite{Kankare1996}) & 1\\
  & GapA (1S7C~\cite{ShinXXX}) & 1\\
  & Eno (1E9I~\cite{Kuhnel2001}) & 1\\
\hline
RNA & tRNA$^{\text{Phe}}$ (4YCO~\cite{Byrne2015}) & 5\\
\hline
4*Metabolite & GLU & 1436\\
  & ATP & 144\\
  & FBP & 225\\
  & GSH & 255\\
\hline
3*Inorganic Ion & K$^{+}$ & 4602\\
  & Mg$^{2+}$ & 400\\
  & Cl$^{-}$ & 1320\\
\hline
Solvent & Water & 306221\\
\hline
\end{tabular}
\caption{Cytoplasm Components}
\end{table}

Initially, each component was prepared with the same simulation setup as cytoplasm in a box of water. After minimization and equilibration steps, each component together with a layer of water molecule with specific size were extracted. The size of the water shell was probed to make up the total number of water molecules the model needed to make a total biomolecular concentration of 30\%. Possible ions in the water shell were not discarded and at the end counted for the total ionic strength of the cytoplasm model. The components were placed in a preliminary box of size $L=30 nm$ according to their abundancy. The resulting box were equilibrated with temperature and pressure coupling for total time of $200 ns$, this step is called shrinking in this text. The size of the box shrank to $L=22.90 nm$ after $150 ns$.

For error analysis, each simulations were repeated three times with independent starting velocities.

All simulations were performed with Gromacs 2018. Single simulations were started from crystal conformations. The cytoplasm simulations starting conformation were taken from the equilibrated conformation of each single simulation.

{\bf Single Simulation Setup: } Each component was simulated with the same parameter as the cytoplasm for $200 ns$.

\subsection*{Analysis}
Before any analysis the periodic boundary condition (pbc) artifacts have been removed. We used GROMACS tools to do the analysis. For single component simulations, first the components were made whole and jump removed and then all the atom were put inside the compact box. The same treatment were applied to the cytoplasm simulations. Additionally, each component's trajectory were extracted and fitted by rotation and translation for later rotational correlation time analysis.

A Mean Square Displacement (MSD) analysis was used to calculate the translational diffusion coefficient~\cite{Allen1987a}. The diffusion coefficients were extracted by a linear fit to MSD analysis by averaging blocks with a length of $10 \,ns$. In principle diffusion coefficient needs to be corrected for finite size effects~\cite{Yeh2004} but due to relatively large simulation boxes this correction is negligible.

% We followed the Lipari-Szabo approach to calculate the rotational correlation time, using a second order Legendre polynomial $P_2({\mathbf r}_{NH})$ to be able to compare the results to experimental relaxation data~\cite{Lipari1982b}. First the average order parameter $<\!\!S^2\!\!>$ over all ${\mathbf r}_{NH}$ vectors for each protein was determined from trajectories after the biomolecular rotation- and translation-movements were removed. Then the rotational correlation function $C(t)$ of the original trajectories was fitted to 
% \begin{equation}
% C(t) ~=~ <\!\!S^2\!\!> \exp(-t/\tau_M) + (1-<\!\!S^2\!\!>)\exp(-t/\tau_T)
% \end{equation}
% where $\tau_M$ is the rotational correlation time (tumbling) and $\tau_T = (\tau_M^{-1} + \tau_e^{-1})^{-1}$, where $\tau_e$ is the internal bond movement.

Protein protein interaction: Who is interacting with who?! (Residue)

\section*{Results}

\begin{itemize}
	\item Translational Diffusion
		\begin{itemize}
			\item Translation diffusion as a function of size for both soup components and ingredients in the same plot
			\item D/D0 for all components
			% \item D-x, D-y, D-z for metabolites, then a combination of 2-d Diffusion
		\end{itemize}
	% \item Rotational Diffusion
	% 	\begin{itemize}
	% 		\item Rotational correlation time for non-metabolite components and comparison with ingredients
	% 		\item C-R/C-R0
	% 	\end{itemize}
	\item RMSD analysis
		\begin{itemize}
			\item Comparison of RMSD in  the soup and ingredients
			\item RMSD for metabolites around macromolecules
		\end{itemize}
	\item Protein-Protein interaction
	\item Metabolite-Protein interaction
\end{itemize}

\section*{Discussion}\label{sec:dissc}

\bibliography{../bibtex/library}

% \begin{addendum}
%  \item  
%  \item[Competing Interests] 
%  \item[Correspondence]  
% \end{addendum}

\end{document}

% Cytoplasm setup:
% \begin{itemize}
% \item Constitutents of the cytoplasm in ratio, (non-ribomosal) protein, (non-ribosomal) RNA and metabolites {\bf review paper}
% \item Proteins fraction 
% \item Nucleic Acid fraction
% \item Metabolites fraction
% \item Physiological condition of the cell
% \item Biomolecular fraction 
% \end{itemize}

% General Simulation setup:
% \begin{itemize}
% \item Force Field
% \item Water model
% \item Gromacs
% \item Temperature and pressure
% \item (details of soup box)
% \item Minimization 
% \item Equilibration
% \item Production
% \end{itemize}

% \begin{itemize}
% \item Box building
% \item Droplets
% \item Shrinking
% \end{itemize}

% {\it Cytoplasm Model: }The most abundant non-ribosomal proteins that account for 50\% of the {\em Escherichia coli} xxx cytoplasm's protein composition were chosen for our cytoplasm model~\cite{}. Specifically TufA (pdb 1dg1), MetE (pdb 1u22), IcdA (pdb 1p8f), AhpC(pdb 1yep), CspC (pdb 1mjc), Ppa (pdb 2eip), GapA (pdb 1s7c) and .

% The proteins were simulated at 30\% biomolecular mass fraction in  a physiological salt concentration (0.15M NaCl).

% For all simulations the Amber99sb-ildn force field was used~\cite{Hornak2006, Lindorff2010} in combination with the TIP4P/2005 water model~\cite{Abascal2005b}. Electrostatic interactions were treated using the particle mesh Ewald algorithm~\cite{Essmann1995a} and XXXX. All chemical bonds were constrained at their equilibrium length using the LINCS algorithm~\cite{Hess2008b} allowing an integration time step of 2 fs. Temperature was controlled at 310 K using the v-rescale algorithm~\cite{Bussi2007a} and a coupling time of 0.5 ps. The pressure was controlled at 1 bar using the Parrinello-Rahman algorithm~\cite{Parrinello1981a} with a time constant of 10 ps.

% Among proteins, (the corrections)
% The metabolites structures have been prepared by ... .

% For tRNA and metabolites the vector ... was used.

% \begin{table}
% \centering
% \caption{Proteins PDB ID}
% \label{tbl:pdb}
% \begin{tabular}{lcc}
% \hline
% Protein (pdb)  \\
% \hline
% crowder\_n1 & 1dg1~\cite{Abel1996}   \\ 
% crowder\_n2 & 1u22~\cite{Ferrer2004} \\
% crowder\_n3 & 1p8f~\cite{Mesecar2000} \\
% crowder\_n4 & 1yep~\cite{Parsonage2005}\\
% crowder\_n5 & 1mjc~\cite{Schindelin1994}\\
% crowder\_n6 & 2eip~\cite{Kankare1996}\\
% crowder\_n7 & 1s7c~\cite{ShinXXX}\\
% crowder\_n8 & 1sl4~\cite{Guo2004d}\\
% crowder\_n9 & 1wz2~\cite{FukunagaXXX}\\
% \hline
% \end{tabular}
% \end{table}

% Cytoplasm setup:
% \begin{itemize}
% \item Constitutents of the cytoplasm in ratio, (non-ribomosal) protein, (non-ribosomal) RNA and metabolites {\bf review paper}
% \item Proteins fraction 
% \item Nucleic Acid fraction
% \item Metabolites fraction
% \item Physiological condition of the cell
% \item Biomolecular fraction 
% \end{itemize}

### 2019-09-09 ingredients-maker.py
Add files via upload

@@ -1,0 +1,286 @@
+#!/usr/bin/env python3
+
+import argparse, os, glob, math
+import subprocess
+import numpy as np
+
+path="/home/zahedeh/shared_soup/our_soup/"
+path_mdp= "../mdp/" 
+# path_mdp= "%s/mdp/" % (path)
+
+def parseArguments():
+	parser = argparse.ArgumentParser()
+	parser.add_argument("-gmx", "--gromacs_command",  type=str,  default="gmx")
+	parser.add_argument("-cr_n", "--n", help="crowder index",  type=int,  default=0)
+	
+	args = parser.parse_args()
+	return args
+
+
+
+# export GMX_MAXBACKUP=-1
+
+if __name__ == '__main__':
+    args  = parseArguments()
+    gromacs_command = args.gromacs_command
+    n=args.n 				#crowder index
+    crowder_total=n
+    rep_first=1
+    rep_total=3
+    flag_build=1
+    flag_eq=0
+    flag_eq_nvt=0
+    flag_eq_npt=0
+    flag_prod=1
+    mdp_min_none='min_none.mdp'
+    mdp_min='min.mdp'
+    mdp_nvt='nvt.mdp'
+    mdp_npt='npt.mdp'
+    mdp_prod='prod.mdp'
+    topology_template='%s/topology_template.top' % (path)
+    while n <= crowder_total:
+        print('Crowder No. %s:\n' % (str(n)))
+        print('\tChecking the ingredient ... ')
+        # Creates a temporary index file to check what type of molecule is the crowder
+        with open('q.txt', 'w') as outfile:
+            outfile.write("q\n")
+        os.system('%s make_ndx -f crowder_n%s/%s_*pdb -o temporary.ndx < q.txt  >/dev/null 2>/dev/null ' % (gromacs_command, str(n), str(n) ))
+        os.system('rm q.txt')
+        with open('temporary.ndx', 'r') as infile:
+            l=[]
+            for line in infile:
+                if line.find('[') > -1:
+                    l.append(line.strip())
+            if '[ Protein ]' in l:
+                crowder_type = "Protein"
+            elif '[ RNA ]' in l:
+                crowder_type = "RNA"
+            elif '[ Other ]' in l and '[ ATP ]' in l:
+                crowder_type = "Other"
+                is_atp = 1
+            elif '[ Other ]' in l and not '[ ATP ]' in l:
+                crowder_type = "Other"
+                is_atp = 0
+            else:
[…]

### 2019-09-09 soup-maker.py
soup-maker.py added

@@ -1,0 +1,454 @@
+#!/usr/bin/env python3
+
+import argparse, os, glob, math
+import subprocess
+import numpy as np
+
+export GMX_MAXBACKUP=-1
+
+def parseArguments():
+    parser = argparse.ArgumentParser()
+    parser.add_argument("-gmx", "--gromacs_command",  type=str,  default="gmx")
+    # parser.add_argument("-cr_n", "--n", help="crowder index",  type=int,  default=0)
+    
+    args = parser.parse_args()
+    return args
+
+box_size_initial=35		# in nm
+box_size_step=2	# nm
+
+droplet_thickness_initial=0.6
+
+droplet_shell_precision=2	# percentage as integer
+
+crowder_total=13
+fraction=0.30
+ionic_strength=0.150	# M
+
+make_droplets=0
+make_box=1
+make_top=1
+make_neutral=1
+make_is=1
+check_fraction=1
+
+path=os.getcwd()
+path_crowders=path
+path_mdp='mdp'
+
+mdp_min='min.mdp'
+
+path_output='soup_4'
+
+# number of gro files for each crowder
+num_crowder[1]=6
+num_crowder[2]=7
+num_crowder[3]=2
+num_crowder[4]=1
+num_crowder[5]=3
+num_crowder[6]=1
+num_crowder[7]=1
+num_crowder[8]=1
+num_crowder[9]=5
+num_crowder[10]=1436
+num_crowder[11]=144
+num_crowder[12]=225
+num_crowder[13]=255
+crowder_type_arr = []
+crowder_mass = []
+# number of gro files for each crowder
+
+
+print('Preliminary steps\n')
+print('\tGathering data about each crowder\n')
+
+crowder=1
+for crowder in range(1, crowder_total+1):
+	# Creates a temporary index file to check what type of molecule is the crowder
+	print( 'q\n | gmx make_ndx -f crowder_n%s/npt.1.gro -o temporary.ndx >/dev/null 2>/dev/null ' %(crowder))
+	with open('q.txt', 'w') as outfile:
+            outfile.write("q\n")
+    os.system('%s make_ndx -f crowder_n%s/npt.1.gro -o temporary.ndx < q.txt  >/dev/null 2>/dev/null ' % (gromacs_command, str(n), str(n) ))
+    os.remove('q.txt')
+    with open('temporary.ndx', 'r') as infile:
+        l=[]
+	# Is this crowder a protein, RNA or metabolite?
+        for line in infile:
+            if line.find('[') > -1:
[…]

### 2019-09-09 soup-maker.py
soup-maker.py

@@ -105,29 +105,28 @@ for crowder in range(1, crowder_total+1):
 
 
 print('\t\tTotal biomolecular mass =%0.3f' % (total_nonwater_mass))
+print('\tCalculating the number of waters necessary to achieve the biomolecular fraction of %.0f %' % str(fraction*100))
 
-# echo $fraction | awk '{printf "\tCalculating the number of waters necessary to achieve the biomolecular fraction of %.0f %%\n", $1*100}'
+water_necessary = (float(total_biomol_mass)/fraction - float(total_biomol_mass))/18.0
+total_water_mass = water_necessary*18.
 
-# water_necessary=$(echo $total_biomol_mass $fraction | awk '{printf "%.0f\n", (($1/$2)-$1)/18 }')
-
-# total_water_mass=$(echo $water_necessary | awk '{printf "%.2f\n",$1*18}')
-
-# echo -e "\t\t"'It is necessary to add '"$water_necessary"' water molecules to the final box'
-# echo -e "\t\t"'That corresponds to '"$total_water_mass"' Da'
-# echo
-# echo
-# echo
+print('\t\tIt is necessary to add %s water molecules to the final box' % str(water_necessary))
+print('\t\tThat corresponds to %s Da'% str(total_water_mass))
+print('\n')
+print('\n')
+print('\n')
 
 
 
 if make_droplets == 1:
     print('Making droplets for each crowder')
     print('\tFinding the optimum droplet shell thickness which comprises a total of %s water molecules for all crowders\n' % str(water_necessary))
-    print('\tThe maximum difference that is allowed between the number of water molecules inside the droplets and the number of molecules we need is %s %%' % str(droplet_shell_precision))
+    print('\tThe maximum difference that is allowed between the number of water molecules inside the droplets and the number of molecules we need is %s %' % str(droplet_shell_precision))
 # if [ $make_droplets -eq 1 ]; then
 	
     ####Zaza: convert the line below
 # 	echo 'thickness[nm] biomol-fract[%] added-perc[%] error-perc[%]' | awk '{printf "\t\t%16s\t%16s\t%16s\t%16s\n",$1,$2,$3,$4}'
+    # print('\t\tthickness[nm] biomol-fract[%]')
 	
 	droplet_thickness_step=0.1	# nm
 	droplet_thickness=droplet_thickness_initial		# nm
@@ -337,7 +336,6 @@ if make_box == 1:
 
[…]

### 2019-09-10 paper/virtual_cytoplasm_v2.tex
New version of the manuscript

--- 2019-09-07 paper/virtual_cytoplasm.tex
@@
 
-\author{David van der Spoel$^1$}
+\author{Zahedeh}
+\affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology. Uppsala University, SE-751 05 Uppsala, Sweden}
+
+\author{Leandro Oliveira Bortot}
+\affiliation{Laboratory of Biological Physics, School of Pharmaceutical Sciences of Ribeirão Preto, University of São Paulo, Ribeirão Preto, Brazil}
+
+\author{David van der Spoel}
+\affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology. Uppsala University, SE-751 05 Uppsala, Sweden}
+\email{aaa@aaa.com}
 
@@
 
-% \begin{affiliations}
-%  \item Science for Life Laboratory, Department of Cell and Molecular Biology. Uppsala University, SE-751 05 Uppsala, Sweden
+%\begin{affiliations}
+%\item 
 %  \item Department of Biological Science and Engineering, School of Chemistry and Biological Engineering, University of Science and Technology Beijing, 100083 Beijing, China
-% \end{affiliations}
+%\end{affiliations}
 
@@
 
-\section*{Material and Method}
-\subsection*{Cytoplasm model}
-In order to investigate the effects of crowding on cellular components in a more realistic way, we built a  cytoplasm model based on {\em Escheria coli} as a model organism~\cite{Dong1996,Bennett2009,Link1997,Mcguffee2010}
+\section*{Materials and Methods}
 
-Because of restriction of computational resources, incorporating a full list of proteins, nucleic acids and metabolites in full atomistic model is prohibitive.
-
-In order to build the RNA fraction of our cytoplasm model, we consider data showing that 2.9\% of the dry weight of \textit{E.
-coli} is composed by tRNAs, which correspond to 74\% of the dry weight of non-ribosomal RNAs, while 55.0\% of it is composed of proteins~\cite{schaechter2006}.
-Thus, the tRNA weight should correspond to 5\% of the total weight of all protein chains used in our model.
-Specifically, tRNA(Phe) was selected as a representative due to the availability of a recent crystallographic structure~\cite{Byrne2015}.
-Copies were added to the simulation box to account for the correct protein/RNA weight ratio (Table X).
-
-The metabolite fraction of our cytoplasm model was build based on data showing that the number of metabolite molecules in the cytoplasm of \textit{E.
-coli} is about 42.86 times higher than the number of proteins.
-We considered the most abundant molecule as representative of each metabolite class, i.e.
-Glutamate for amino acids, ATP for nucleotides, FBP for central carbon intermediates and Glutathione redox cofactors.
-The copy number for each molecule was calculated as the ratio of their experimentally observed concentration in {\em E.
-coli}.
-
-\subsection*{Components Preparation}
+\subsection*{Initial structures}
 The proteins and tRNA were downloaded from Protein Data Bank (PDB).
@@
 
-\subsection*{All-Atom Molecular Dynamics Simulation}
+\subsection*{Molecular Dynamics Simulation}
 
@@
 The pressure was controlled at 1 bar using the Parrinello-Rahman algorithm~\cite{Parrinello1981a} with a time constant of 10 ps.
-
-\begin{table}[h!]
-\label{tbl:component}
-\centering
-\begin{tabular}{lcc}
-\hline
-Class & Name (PDB ID) & Number\\
-\hline
-8*Protein & TufA (1DG1~\cite{Abel1996}) & 6\\
-  & MetE (1U22~\cite{Ferrer2004}) & 7\\
-  & IcdA (1P8F~\cite{Mesecar2000}) & 2\\
-  & AhpC (1YEP~\cite{Parsonage2005}) & 1\\
-  & CspC (1MJC~\cite{Schindelin1994}) & 3\\
-  & Ppa (2EIP~\cite{Kankare1996}) & 1\\
-  & GapA (1S7C~\cite{ShinXXX}) & 1\\
-  & Eno (1E9I~\cite{Kuhnel2001}) & 1\\
-\hline
-RNA & tRNA$^{\text{Phe}}$ (4YCO~\cite{Byrne2015}) & 5\\
-\hline
-4*Metabolite & GLU & 1436\\
-  & ATP & 144\\
-  & FBP & 225\\
-  & GSH & 255\\
-\hline
-3*Inorganic Ion & K$^{+}$ & 4602\\
-  & Mg$^{2+}$ & 400\\
-  & Cl$^{-}$ & 1320\\
-\hline
-Solvent & Water & 306221\\
-\hline
-\end{tabular}
-\caption{Cytoplasm Components}
-\end{table}
-
-Initially, each component was prepared with the same simulation setup as cytoplasm in a box of water.
-After minimization and equilibration steps, each component together with a layer of water molecule with specific size were extracted.
-The size of the water shell was probed to make up the total number of water molecules the model needed to make a total biomolecular concentration of 30\%.
-Possible ions in the water shell were not discarded and at the end counted for the total ionic strength of the cytoplasm model.
-The components were placed in a preliminary box of size $L=30 nm$ according to their abundancy.
-The resulting box were equilibrated with temperature and pressure coupling for total time of $200 ns$, this step is called shrinking in this text.
-The size of the box shrank to $L=22.90 nm$ after $150 ns$.
 
@@
 
-Protein protein interaction: Who is interacting with who?! (Residue)
-
 \section*{Results}
 
-\begin{itemize}
-	\item Translational Diffusion
-		\begin{itemize}
-			\item Translation diffusion as a function of size for both soup components and ingredients in the same plot
-			\item D/D0 for all components
-			% \item D-x, D-y, D-z for metabolites, then a combination of 2-d Diffusion
-		\end{itemize}
-	% \item Rotational Diffusion
-	% 	\begin{itemize}
-	% 		\item Rotational correlation time for non-metabolite components and comparison with ingredients
-	% 		\item C-R/C-R0
-	% 	\end{itemize}
-	\item RMSD analysis
-		\begin{itemize}
-			\item Comparison of RMSD in  the soup and ingredients
-			\item RMSD for metabolites around macromolecules
-		\end{itemize}
-	\item Protein-Protein interaction
-	\item Metabolite-Protein interaction
-\end{itemize}
+\subsection{Cytoplasm model}
+In this section we describe the rationale behind the composition of our model, which has five fractions: protein, RNA, metabolites, water and ions.
+We highlight that we didn't add lipids and DNAs to our model because we are considering only elements that are free to diffuse through the cytosol.
+We gathered data from several sources in order to build a computational model that is representative of the cytoplasm of {\em Escheria coli}~\cite{Dong1996,Bennett2009,Link1997,Mcguffee2010}.
+
+\subsubsection{Protein fraction}
+We selected a group of eight proteins that account for 50\% of the abundance of non-ribosomal proteins in the cytoplasm of Escherichia coli K-12 \cite{ref_prot_abundance}.
+The two most abundant proteins, TufA and MetE, account for 20\% and 12\% of the total protein count in E.coli K-12, respectively, while the other six proteins contribute with less than 5\% each (Table ~\ref{tbl:protein_fraction}).
+
+\begin{table}
+\centering
+\begin{tabular}{lcc}
+\hline
+Protein & Abundance in E.Coli K12 \\ 
+\hline
+TufA	 & 20 \\
+MetE	 & 12 \\
+IcdA	 & 5 \\
+AhpC	 & 4 \\
+CspC	 & 4 \\
+Ppa & 3 \\
+GapA	 & 2 \\
+Eno & 2 \\
+\hline
+\end{tabular}
+\caption{Protein fraction}
+\label{tbl:protein_fraction}
+\end{table}
+
+\subsubsection{RNA fraction}
+In order to build the RNA fraction of our cytoplasm model, we consider the following data:
+Since 74\% of the dry weight of non-ribosomal RNAs is composed by tRNA, we chose to model the RNA presence with tRNA molecules.
+The protein and RNA content of the total dry weight of \textit{E.coli} is 55\% and 2.9\%, respectively.
+That is, the total RNA weight corresponds to 5\% of the total protein weight.
+We incorporated this data in our model by considering the tRNA(Phe) molecule as a representative of RNAs due to the availability of a recent crystallographic structure~\cite{Byrne2015}.
+The correct number of copies were added to the simulation box to account for the correct protein/RNA weight ratio (Table ).
+
+\subsubsection{Metabolites fraction}
+The total number of metabolite molecules was calculated considering data showing that the number of metabolite molecules in the cytoplasm of \textit{E.
+coli} is about 42.86 times higher than the number of proteins.
+We considered the most abundant molecule as representative of each metabolite class, i.e.
+Glutamate for amino acids, ATP for nucleotides, FBP for central carbon intermediates and Glutathione redox cofactors.
+The copy number for each molecule was calculated from the ratio of their experimentally observed concentration in {\em E.
+coli}.
+
+\subsubsection{Water fraction}
+The number of water molecules was calculated according to the desired biomolecular concentration, which ranges from XX to YY in biological systems such as E.
+coli cytoplasm.
+In our case, we choose the biomolecular concentration of 30\%, that is, the number of molecules necessary to reach a ratio of total biomolecular mass to water mass of 30\%.
+
+\subsubsection{Inorganic ions fraction}
+Finally, Mg2+ was used as counter-ions for tRNA and K+ cations were added to neutralize the charges of the simulation box.
+The number of water molecules was used to calculate the number of K+ and Cl- that were necessary to reach the ionic strength of 0.150 M.
+
+\begin{table}
+\centering
+\begin{tabular}{lcc}
+\hline
+Class & Name (PDB ID) & Number\\
+\hline
+Protein & TufA (1DG1~\cite{Abel1996}) & 6\\
+  & MetE (1U22~\cite{Ferrer2004}) & 7\\
+  & IcdA (1P8F~\cite{Mesecar2000}) & 2\\
+  & AhpC (1YEP~\cite{Parsonage2005}) & 1\\
+  & CspC (1MJC~\cite{Schindelin1994}) & 3\\
+  & Ppa (2EIP~\cite{Kankare1996}) & 1\\
+  & GapA (1S7C~\cite{ShinXXX}) & 1\\
+  & Eno (1E9I~\cite{Kuhnel2001}) & 1\\
+\hline
+RNA & tRNA$^{\text{Phe}}$ (4YCO~\cite{Byrne2015}) & 5\\
+\hline
+Metabolite & GLU & 1436\\
+  & ATP & 144\\
+  & FBP & 225\\
+  & GSH & 255\\
+\hline
+Solvent & Water & 306221\\
+\hline
+Inorganic Ion & K$^{+}$ & 4602\\
+  & Mg$^{2+}$ & 400\\
+  & Cl$^{-}$ & 1320\\
+\hline
+\end{tabular}
+\caption{Cytoplasm Components}
+\label{tbl:soup_components}
+\end{table}
+
+\subsection{Building the simulation box}
+All components can be put in the same simulation box by inserting each of them in random orientations in a cubic box of side L that is initially empty.
+However, that is not a trivial process.
+We need to use a box size that is big enough to allow the random insertions to succeed without structural overlapping and without creating artificial interactions between the elements, that is, we need L to be big enough so that different elements are not placed too close from each other.
+On the other hand, as the box gets bigger, there will be more space between the components and the system will be harder to equilibrate due to larger variations in box size.
+
+We devised an iterative process that solves both problems simultaneously.
+We start with a box size L that is too small to allow all systems to fit in the box by random insertion.
+In our case, we started with L=30nm.
+Then, we allow 100 insertion trials for each element.
+If all trials fail for any of them, we increase the box by a dL step of 1nm and start again until all insertions succeed.
+In our case, all insertions succeeded after increasing L to 35nm.
+Additionally, instead of adding only the protein, RNA or metabolite molecules in the empty box in each trial, we actually add a droplet of water and counter ions in which the molecule of interest is embedded.
+Such droplets are taken from molecular dynamics simulations in which each component was previously simulated.
+The benefits of using such droplets are threefold: i.
+it prevents artificial contacts between the components as consequence of packing in the simulation box because the water droplet acts as natural protecting layer.
+ii.
+it is a natural way to place water molecules and counter-ions in the simulation box around each component.
+iii.
+the components are already pre-equilibrated, which will help us to perform the equilibration of the whole simulation box.
+
+In order to do this, the droplets around each component are also constructed iteractively.
+The number of water molecules that we must place in the simulation box is known.
+However, we don't know the thickness of the water layer around each component, l, that accounts for such amount of water molecules.
+Since l depends on a series of factors such as the shape, size and abundancy of each element, we define it iteratively.
+We start by taking a droplet of thickness l=3A around all elements and counting the number of water that we would add to the box.
+If it is smaller than the number we need, l is increased by dl=1A.
+If it is larger, we take a step back and reduce dl by 10 times.
+We can carry this process until an arbitrary precision cutoff is satisfied.
+
+< IMAGE SHOWING DROPLETS, l, dl, THE EMPTY BOX AND INSERTIONS, L, dL >
+
+After all droplets are successfully inserted in a simulation box by the iterative process described above, we proceed to add ions to neutralize the net charge of the system and to reach the desired ionic strength.
+Then, we perform energy minimization and a short simulation step of 500ps in which all molecules of the box are free to move.
+In this step the box shrinks to its optimum size.
+In our case, the simulation box shrank from the initial L=35nm to L=22.9nm.
+From this point, the system is ready to be submitted to the default simulation steps such as thermalization and production run (please check the materials section for details about the parameters we used).
+
+< SHRINKING PLOTS SHOWING BOX VOLUME/DENSITY ? >
+
+Python scripts and all topology and structure files necessary to perform all these steps are available at github.com/dspoel...
+The cytoplasm model that we built here can be used as a tool to answer many research questions about the effects of crowding on specific systems of interest.
+With the files we are providing here it is possible, for example, to add a probe protein to investigate the effect of crowding on it, to add new macromolecular or small crowders, add new solvents and change the biomolecular concentration to increase or decrease the intensity of the crowding effect.
+
+\subsection{The effect of crowded systems}
+
+In order to investigate the effects of crowding in the E.
+coli cytoplasm model on the behavior of its elements, we constructed three boxes that have different orientations for each of its elements and submitted each to molecular dynamics simulations of 1 us.
+We also performed 200 ns molecular dynamics simulations for each isolated element (please check the materials section for details about the parameters).
+
+\subsubsection{Aggregation}
+We observed aggregation of the metabolites that contain phosphate, ATP and FBP, with Mg2+ that was used as initially used as counter-ion for ATP and tRNA.
+After running simulations of small boxes containing just these metabolites and Mg2+, we found that completely protonating their phosphate groups can prevent aggregation.
+Thus, we advise caution when dealing with small molecules that are highly charged, such as ATP and FBP.
+Trial simulations with small boxes in which these metabolites are in high concentration to verify if they'll aggregate with Mg2+ or other charged species that are present in the box.
+In the specific case of ATP and FBP modeled with GAFF in boxes containing Mg2+, we advise that they should be completely protonated even though that is not their physiological protonation state.
+
+< FIGURE SHOWING THE AGGREGATION >
+< FIGURE SHOWING THAT PROTONATED PHOSPHATE DOESN'T AGGREGATE >
+
+\subsubsection{Translational diffusion}
+D/D0 for all crowders
+
+\subsubsection{Structural integrity of the crowders}
+RMSD and SASA plots, show that the crowders didn't unfold
+
+\subsubsection{Structural dynamics}
+RMSF. Are there differences between the soup and single sim.?
 

### 2019-09-10 paper/barplots.py
barplot script added to make hb sasa rmsd plots. some plots added as pdf. Makefile

@@ -1,0 +1,90 @@
+#!/usr/bin/env python3
+import argparse
+import matplotlib
+import matplotlib.pyplot as plt
+import numpy as np
+
+
+
+
+def parseArguments():
+    parser = argparse.ArgumentParser()
+    parser.add_argument("-prop", "--prop",  type=str,  default="")
+    parser.add_argument("-title", "--title",  type=str,  default="")
+    parser.add_argument("-ylabel", "--ylabel",  type=str,  default="")
+    parser.add_argument("-xlabel", "--xlabel",  type=str,  default="crowder")
+    
+    args = parser.parse_args()
+    return args
+
+
+men_means, men_std = (20, 35, 30, 35, 27), (2, 3, 4, 1, 2)
+women_means, women_std = (25, 32, 34, 20, 25), (3, 5, 2, 3, 3)
+
+
+
+
+def autolabel(rects, xpos='center'):
+    """
+    Attach a text label above each bar in *rects*, displaying its height.
+
+    *xpos* indicates which side to place the text w.r.t. the center of
+    the bar. It can be one of the following {'center', 'right', 'left'}.
+    """
+
+    ha = {'center': 'center', 'right': 'left', 'left': 'right'}
+    offset = {'center': 0, 'right': 1, 'left': -1}
+
+    # for rect in rects:
+    #     height = rect.get_height()
+    #     ax.annotate('{}'.format(height),
+    #                 xy=(rect.get_x() + rect.get_width() / 2, height),
+    #                 xytext=(offset[xpos]*3, 3),  # use 3 points offset
+    #                 textcoords="offset points",  # in both directions
+    #                 ha=ha[xpos], va='bottom')
+
+if __name__ == '__main__':
+    olygomers = ('1', '3', '4', '6', '7', '8')
+    crowders =('1', '2', '3', '4', '5', '6', '7', '8')
+    args  = parseArguments()
+    singles_mean, singles_std, cyto_mean, cyto_std = [], [], [], []
+    for env in ["singles", "cyto"]:
+        with open("%s_%s.dat" % (args.prop, env), 'r') as infile:
+            for line in infile:
+                if line.find("crowder_n") > -1 and line.find("-") <= 0:
+                    line = line.split()
+                    n = line[0].split("crowder_n")[1]
+                    if env == "singles":
+                        singles_mean.append(float(line[1]))
+                        singles_std.append(float(line[2]))
[…]

### 2019-09-10 paper/diff_plots.py
diff data added

@@ -1,0 +1,58 @@
+#!/usr/bin/env python3
+
+import argparse, os, glob, math
+import matplotlib.pyplot as plt
+import numpy as np
+
+markers = ['o', 'v','*', 'd', '^', '<', '>', '1', '2', '3', '4', '8', 's', 'p' , 'h', 'H', '+', 'x', 'D',  '|', '_', 'P', 'X', ',', '.']
+
+atom_crowder = [11890, 11825, 12906, 25592, 1009, 16284, 19976, 12814, 2377]
+atom_s_b = sorted(atom_crowder)
+
+order_crowder=[4, 7, 6, 3, 8, 1, 2, 9, 5] #biggest to smallest
+order_s_b=[5, 9, 2, 1, 8, 3, 6, 7, 4] #smallest to biggest
+
+avg_cyto, ste_cyto =[],[]
+for i in range(0, len(order_s_b)):
+	with open('msd_cyto.dat', 'r') as infile:
+		for line in infile:
+			if "crowder_n"+str(order_s_b[i]) in line:
+				avg_cyto.append(float(line.split()[1]))
+				ste_cyto.append(float(line.split()[2]))
+plt.errorbar(atom_s_b, avg_cyto, ste_cyto, marker='o', markersize=10, linestyle=' ')
+plt.xlabel('# atoms', fontsize=12)
+plt.ylabel('$D_{cyto} (10^{-5} cm^2/s)$', fontsize=12)
+plt.savefig('diff_cyto.pdf', bbox_inches="tight")
+plt.close()
+
+
+avg_sing, ste_sing = [],[]
+for i in range(0, len(order_s_b)):
+	with open('msd_singles.dat', 'r') as infile:
+		for line in infile:
+			if "crowder_n"+str(order_s_b[i]) in line:
+				avg_sing.append(float(line.split()[1]))
+				ste_sing.append(float(line.split()[2]))
+plt.errorbar(atom_s_b, avg_sing, ste_sing, marker='o', markersize=10, linestyle=' ')
+plt.xlabel('# atoms', fontsize=12)
+plt.ylabel('$D_{0} (10^{-5} cm^2/s)$', fontsize=12)
+plt.savefig('diff_singles.pdf', bbox_inches="tight")
+plt.close()
+
+AVG, STD = [],[]
+
+for i in range(0, len(avg_sing)):
+	AVG.append(avg_cyto[i]/avg_sing[i])
[…]

### 2019-09-10 paper/virtual_cytoplasm_v2.tex
small correction
barplot script added to make hb sasa rmsd plots. some plots added as pdf. Makefile
diff data added
diff plots added and manuscript updated

@@
 
+\subsubsection{Translational diffusion}
+D/D0 for all crowders
+
+\begin{figure}
+\includegraphics[scale=0.6]{diff_singles.pdf} &
+\includegraphics[scale=0.6]{diff_cyto.pdf} &
+\includegraphics[scale=0.6]{diff_cyto_over_singles.pdf}
+\caption{A) Diffusion coefficient of crowders from single molecule simulations.
+B) Diffusion coefficient of crowders from cytoplasm simulations.
+C) The ratio between diffusion coefficient obtained from cytoplasm to single molecule simulations.
+The crowders are sorted according to their sizes on the x-axis.}
+\end{figure}
+
+\subsubsection{Structural integrity of the crowders}
+RMSD and SASA plots, show that the crowders didn't unfold
+
+\begin{figure}
+\includegraphics[scale=0.6]{rmsd.pdf}
+\caption{Root mean squared displacement of 8 crowders from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.6]{sasa.pdf}
+\caption{Solvent accessible surface area of 8 crowders from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).}
+
+\end{figure}
+
+\subsubsection{Structural dynamics}
+RMSF. Are there differences between the soup and single sim.?
+
+\begin{figure}
+\includegraphics[scale=0.6]{rmsf.pdf}
+\caption{Root mean squared fluctuations of 8 crowders from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). x-axis shows the residues of each chain of crowders.}
+\end{figure}
+
+\subsubsection{Olygomers under crowded condition}
+\begin{figure}
+\includegraphics[scale=0.6]{hb.pdf}
+\caption{Intramolecular hydrogen bonds for crowders that appear as olygomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). The cytoplasm data are not correct!}
+\end{figure}
+
 \subsubsection{Aggregation}
@@
 In the specific case of ATP and FBP modeled with GAFF in boxes containing Mg2+, we advise that they should be completely protonated even though that is not their physiological protonation state.
+The topology and structure files we are providing for ATP and FBP on github are already protonated.
 
@@
 
-\subsubsection{Translational diffusion}
-D/D0 for all crowders
+\begin{figure}
+\includegraphics[scale=0.6]{rdf_RNA_metabolites.pdf}
+\caption{Radial distribution function showing the probability of finding a metabolite around RNA molecules.}
+\end{figure}
 
-\subsubsection{Structural integrity of the crowders}
-RMSD and SASA plots, show that the crowders didn't unfold
+\begin{figure}
+\includegraphics[scale=0.6]{avclust.pdf}
+\caption{Number of metabolite molecules that form a cluster in cytoplasm simulations.}
+\end{figure}
 
-\subsubsection{Structural dynamics}
-RMSF. Are there differences between the soup and single sim.?
+\begin{figure}
+\includegraphics[scale=0.6]{nclust.pdf}
+\caption{Number of clusters that form during the cytoplasm simulations.}
+\end{figure}
 

### 2019-09-10 paper/diff_plots.py

@@ -22,6 +22,7 @@ for i in range(0, len(order_s_b)):
 plt.errorbar(atom_s_b, avg_cyto, ste_cyto, marker='o', markersize=10, linestyle=' ')
 plt.xlabel('# atoms', fontsize=12)
 plt.ylabel('$D_{cyto} (10^{-5} cm^2/s)$', fontsize=12)
+plt.text(-5000,0.018,'B',fontsize=14)
 plt.savefig('diff_cyto.pdf', bbox_inches="tight")
 plt.close()
 
@@ -36,6 +37,7 @@ for i in range(0, len(order_s_b)):
 plt.errorbar(atom_s_b, avg_sing, ste_sing, marker='o', markersize=10, linestyle=' ')
 plt.xlabel('# atoms', fontsize=12)
 plt.ylabel('$D_{0} (10^{-5} cm^2/s)$', fontsize=12)
+plt.text(-5000,0.14, 'A',fontsize=14)
 plt.savefig('diff_singles.pdf', bbox_inches="tight")
 plt.close()
 
@@ -49,6 +51,7 @@ for i in range(0, len(avg_sing)):
 plt.errorbar(atom_s_b, AVG, STD, marker='o', markersize=10, linestyle=' ')
 plt.xlabel('# atoms', fontsize=12)
 plt.ylabel('$D_{cyto}/D_{0}$', fontsize=12)
+plt.text(-5000,0.2,'C',fontsize=14)
 plt.savefig('diff_cyto_over_singles.pdf', bbox_inches="tight")
 plt.close()
 

### 2019-09-11 paper/barplots.py
update on Makefile hb_cyto.dat barplots.py

@@ -12,18 +12,13 @@ def parseArguments():
     parser.add_argument("-prop", "--prop",  type=str,  default="")
     parser.add_argument("-title", "--title",  type=str,  default="")
     parser.add_argument("-ylabel", "--ylabel",  type=str,  default="")
-    parser.add_argument("-xlabel", "--xlabel",  type=str,  default="crowder")
+    parser.add_argument("-xlabel", "--xlabel",  type=str,  default="")
     
     args = parser.parse_args()
     return args
 
 
-men_means, men_std = (20, 35, 30, 35, 27), (2, 3, 4, 1, 2)
-women_means, women_std = (25, 32, 34, 20, 25), (3, 5, 2, 3, 3)
-
-
-
-
+crowder_name=["TufA", "MetE", "IcdA", "AhpC", "CspC", "Ppa ", "GapA", "Eno", "RNA$^{Phe}$"]
 def autolabel(rects, xpos='center'):
     """
     Attach a text label above each bar in *rects*, displaying its height.
@@ -44,8 +39,9 @@ def autolabel(rects, xpos='center'):
     #                 ha=ha[xpos], va='bottom')
 
 if __name__ == '__main__':
-    olygomers = ('1', '3', '4', '6', '7', '8')
-    crowders =('1', '2', '3', '4', '5', '6', '7', '8')
+    oligomers = (crowder_name[0],crowder_name[2],crowder_name[3],crowder_name[5],crowder_name[6],crowder_name[7])
+    protein_name =crowder_name[0:8]
+
     args  = parseArguments()
     singles_mean, singles_std, cyto_mean, cyto_std = [], [], [], []
     for env in ["singles", "cyto"]:
@@ -76,10 +72,12 @@ if __name__ == '__main__':
     ax.set_xlabel(args.xlabel)
     ax.set_title(args.title)
     ax.set_xticks(ind)
-    if args.prop != "hb":
-        ax.set_xticklabels(crowders)
+    if args.prop == "hb":
+        ax.set_xticklabels(oligomers)
+    elif args.prop == "msd":
+        ax.set_xticklabels(crowder_name)
     else:
-        ax.set_xticklabels(olygomers)
+        ax.set_xticklabels(protein_name)
 
     ax.legend()
 

### 2019-09-11 paper/virtual_cytoplasm_v2.tex
Small technical cleanups

@@
 
-\title{Cytoplasm Simulation}
+\title{Making Soup: Preparing and Validating Molecular Simulations of the Bacterial Cytoplasm}
 
-\author{Zahedeh}
+\author{Zahedeh Bashardanesh}
 \affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology. Uppsala University, SE-751 05 Uppsala, Sweden}
@@
 \author{Leandro Oliveira Bortot}
-\affiliation{Laboratory of Biological Physics, School of Pharmaceutical Sciences of Ribeirão Preto, University of São Paulo, Ribeirão Preto, Brazil}
+\affiliation{Laboratory of Biological Physics, School of Pharmaceutical Sciences of Ribeir{\~a}o Preto, University of S{\~a}o Paulo, Ribeir{\~a}o Preto, Brazil}
 
@@
 \affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology. Uppsala University, SE-751 05 Uppsala, Sweden}
-\email{aaa@aaa.com}
+\email{david.vanderspoel@icm.uu.se}
 
@@
 \begin{figure}
-\includegraphics[scale=0.6]{diff_singles.pdf} &
-\includegraphics[scale=0.6]{diff_cyto.pdf} &
-\includegraphics[scale=0.6]{diff_cyto_over_singles.pdf}
+\includegraphics[scale=0.5]{diff_singles.pdf} \\
+\includegraphics[scale=0.5]{diff_cyto.pdf} \\
+\includegraphics[scale=0.5]{diff_cyto_over_singles.pdf}
 \caption{A) Diffusion coefficient of crowders from single molecule simulations.
@@
 \begin{figure}
-\includegraphics[scale=0.6]{rmsf.pdf}
+\includegraphics[width=0.5\textwidth]{rmsf.pdf}
 \caption{Root mean squared fluctuations of 8 crowders from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). x-axis shows the residues of each chain of crowders.}
@@
 
-\bibliography{../bibtex/library}
+\bibliography{library}
 

### 2019-09-12 paper/virtual_cytoplasm_v2.tex
Makefile, manuscript update

--- 2019-09-10 paper/virtual_cytoplasm_v2.tex
@@
 
+\subsubsection{Translational diffusion}
+D/D0 for all crowders
+
+\begin{figure}
+\includegraphics[scale=0.6]{msd.pdf}
+\caption{Diffusion coefficient of crowders from single molecule simulations (blue bars) and cytoplasm simulations (orange bars).}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.6]{diff_cyto_over_singles.pdf}
+\caption{The ratio between diffusion coefficient obtained from cytoplasm to single molecule simulations. The crowders are sorted according to their sizes on the x-axis.}
+\end{figure}
+
+\subsubsection{Structural integrity of the crowders}
+RMSD and SASA plots, show that the crowders didn't unfold
+
+\begin{figure}
+\includegraphics[scale=0.6]{rmsd.pdf}
+\caption{Root mean squared displacement (rmsd) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
+The average rmsd for each protein in single molecule simulation is over number of chains of proteins and three replica.
+The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.
+The error bars show the standard errors.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.6]{sasa.pdf}
+\caption{Solvent accessible surface area (sasa) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
+The average sasa for each protein in single molecule simulation is over number of chains of proteins and three replica.
+The average sasa for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.
+The error bars show the standard errors.}
+
+\end{figure}
+
+\subsubsection{Structural dynamics}
+RMSF. Are there differences between the soup and single sim.?
+
+\begin{figure}
+\includegraphics[scale=0.6]{rmsf.pdf}
+\caption{Root mean squared fluctuations (rmsf) for 8 crowders from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
+x-axis shows the residues for each chain of proteins.
+The average rmsf for each protein in single molecule simulation is over number of chains of proteins and three replica.
+The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.
+The error bars show the standard errors.}
+\end{figure}
+
+\subsubsection{Oligomers under crowded condition}
+\begin{figure}
+\includegraphics[scale=0.6]{hb.pdf}
+\caption{Intramolecular hydrogen bonds (hb) for oligomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
+The average hydrogen bonds for each oligomer in single molecule simulation is over three replica.
+The average hydrogen bonds for each oligomer in cytoplasm simulation is over number of appearance of the oligomer in three replica.
+The error bars show the standard errors.}
+\end{figure}
+
 \subsubsection{Aggregation}

### 2019-09-12 paper/diff_plots.py

@@ -51,7 +51,7 @@ for i in range(0, len(avg_sing)):
 plt.errorbar(atom_s_b, AVG, STD, marker='o', markersize=10, linestyle=' ')
 plt.xlabel('# atoms', fontsize=12)
 plt.ylabel('$D_{cyto}/D_{0}$', fontsize=12)
-plt.text(-5000,0.2,'C',fontsize=14)
+# plt.text(-5000,0.2,'C',fontsize=14)
 plt.savefig('diff_cyto_over_singles.pdf', bbox_inches="tight")
 plt.close()
 

### 2019-09-12 paper/virtual_cytoplasm_v2.tex

@@
 \begin{figure}
+
 \includegraphics[scale=0.6]{msd.pdf}

### 2019-09-12 paper/virtual_cytoplasm_v2.tex
Makefile
radial distribution function for metabolites and rna with reference to mg is added
adjustment to figure lables, rmsf plot is updated

@@
 \maketitle
-
-%\begin{affiliations}
-%\item 
-%  \item Department of Biological Science and Engineering, School of Chemistry and Biological Engineering, University of Science and Technology Beijing, 100083 Beijing, China
-%\end{affiliations}
 
@@
 
-% General Simulation setup:
-% \begin{itemize}
-% \item Force Field
-% \item Water model
-% \item Gromacs
-% \item Temperature and pressure
-% \item (details of soup box)
-% \item Minimization 
-% \item Equilibration
-% \item Production
-% \end{itemize}
-
-% \begin{itemize}
-% \item Box building
-% \item Droplets
-% \item Shrinking
-% \end{itemize}
-
 {\bf General Simulation Setup: }The proteins were simulated at 30\% biomolecular mass fraction in  a physiological salt concentration (0.15M NaCl).
@@
 In principle diffusion coefficient needs to be corrected for finite size effects~\cite{Yeh2004} but due to relatively large simulation boxes this correction is negligible.
-
-% We followed the Lipari-Szabo approach to calculate the rotational correlation time, using a second order Legendre polynomial $P_2({\mathbf r}_{NH})$ to be able to compare the results to experimental relaxation data~\cite{Lipari1982b}.
-First the average order parameter $<\!\!S^2\!\!>$ over all ${\mathbf r}_{NH}$ vectors for each protein was determined from trajectories after the biomolecular rotation- and translation-movements were removed.
-Then the rotational correlation function $C(t)$ of the original trajectories was fitted to
-% \begin{equation}
-% C(t) ~=~ <\!\!S^2\!\!> \exp(-t/\tau_M) + (1-<\!\!S^2\!\!>)\exp(-t/\tau_T)
-% \end{equation}
-% where $\tau_M$ is the rotational correlation time (tumbling) and $\tau_T = (\tau_M^{-1} + \tau_e^{-1})^{-1}$, where $\tau_e$ is the internal bond movement.
 
@@
 \begin{figure}
-\includegraphics[scale=0.6]{avclust.pdf}
-\caption{Number of metabolite molecules that form a cluster in cytoplasm simulations.}
+\includegraphics[scale=0.6]{rdf_mg.pdf}
+\caption{Radial distribution function showing the probability of finding a metabolite or an RNA around MG2+.}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[scale=0.6]{nclust.pdf}
-\caption{Number of clusters that form during the cytoplasm simulations.}
+\includegraphics[scale=0.6]{avclust.pdf}
+\caption{Number of metabolite molecules that form a cluster in cytoplasm simulations. The solid colored lines are the average of three replica (shown in grey).}
 \end{figure}
@@
 \bibliography{library}
-
 % \begin{addendum}
@@
 \end{document}
-
-% Cytoplasm setup:
-% \begin{itemize}
-% \item Constitutents of the cytoplasm in ratio, (non-ribomosal) protein, (non-ribosomal) RNA and metabolites {\bf review paper}
-% \item Proteins fraction 
-% \item Nucleic Acid fraction
-% \item Metabolites fraction
-% \item Physiological condition of the cell
-% \item Biomolecular fraction 
-% \end{itemize}
-
-% General Simulation setup:
-% \begin{itemize}
-% \item Force Field
-% \item Water model
-% \item Gromacs
-% \item Temperature and pressure
-% \item (details of soup box)
-% \item Minimization 
-% \item Equilibration
-% \item Production
-% \end{itemize}
-
-% \begin{itemize}
-% \item Box building
-% \item Droplets
-% \item Shrinking
-% \end{itemize}
-
-% {\it Cytoplasm Model: }The most abundant non-ribosomal proteins that account for 50\% of the {\em Escherichia coli} xxx cytoplasm's protein composition were chosen for our cytoplasm model~\cite{}.
-Specifically TufA (pdb 1dg1), MetE (pdb 1u22), IcdA (pdb 1p8f), AhpC(pdb 1yep), CspC (pdb 1mjc), Ppa (pdb 2eip), GapA (pdb 1s7c) and .
-
-% The proteins were simulated at 30\% biomolecular mass fraction in  a physiological salt concentration (0.15M NaCl).
-
-% For all simulations the Amber99sb-ildn force field was used~\cite{Hornak2006, Lindorff2010} in combination with the TIP4P/2005 water model~\cite{Abascal2005b}.
-Electrostatic interactions were treated using the particle mesh Ewald algorithm~\cite{Essmann1995a} and XXXX.
-All chemical bonds were constrained at their equilibrium length using the LINCS algorithm~\cite{Hess2008b} allowing an integration time step of 2 fs.
-Temperature was controlled at 310 K using the v-rescale algorithm~\cite{Bussi2007a} and a coupling time of 0.5 ps.
-The pressure was controlled at 1 bar using the Parrinello-Rahman algorithm~\cite{Parrinello1981a} with a time constant of 10 ps.
-
-% Among proteins, (the corrections)
-% The metabolites structures have been prepared by ... .
-
-% For tRNA and metabolites the vector ... was used.
-
-% \begin{table}
-% \centering
-% \caption{Proteins PDB ID}
-% \label{tbl:pdb}
-% \begin{tabular}{lcc}
-% \hline
-% Protein (pdb)  \\
-% \hline
-% crowder\_n1 & 1dg1~\cite{Abel1996}   \\ 
-% crowder\_n2 & 1u22~\cite{Ferrer2004} \\
-% crowder\_n3 & 1p8f~\cite{Mesecar2000} \\
-% crowder\_n4 & 1yep~\cite{Parsonage2005}\\
-% crowder\_n5 & 1mjc~\cite{Schindelin1994}\\
-% crowder\_n6 & 2eip~\cite{Kankare1996}\\
-% crowder\_n7 & 1s7c~\cite{ShinXXX}\\
-% crowder\_n8 & 1sl4~\cite{Guo2004d}\\
-% crowder\_n9 & 1wz2~\cite{FukunagaXXX}\\
-% \hline
-% \end{tabular}
-% \end{table}
-
-% Cytoplasm setup:
-% \begin{itemize}
-% \item Constitutents of the cytoplasm in ratio, (non-ribomosal) protein, (non-ribosomal) RNA and metabolites {\bf review paper}
-% \item Proteins fraction 
-% \item Nucleic Acid fraction
-% \item Metabolites fraction
-% \item Physiological condition of the cell
-% \item Biomolecular fraction 
-% \end{itemize}

### 2019-09-12 paper/barplots.py

@@ -38,6 +38,7 @@ def autolabel(rects, xpos='center'):
     #                 textcoords="offset points",  # in both directions
     #                 ha=ha[xpos], va='bottom')
 
+
 if __name__ == '__main__':
     oligomers = (crowder_name[0],crowder_name[2],crowder_name[3],crowder_name[5],crowder_name[6],crowder_name[7])
     protein_name =crowder_name[0:8]
@@ -46,11 +47,11 @@ if __name__ == '__main__':
     ylabel = args.ylabel
     if len(ylabel) == 0:
         if args.prop == "sasa":
-            ylabel = "\$nm^{2}\$"
+            ylabel = "$nm^{2}$"
         elif args.prop == "rmsd" or args.prop == "rg":
-            ylabel = "\$nm\$"
+            ylabel = "$nm$"
         elif args.prop == "msd":
-            ylabel = '\$D (10^{-5} cm^2/s)\$'
+            ylabel = '$D (10^{-5} cm^2/s)$'
         elif args.prop == "hb":
             ylabel = "#HB"
         else:
@@ -82,7 +83,7 @@ if __name__ == '__main__':
                     label='Cyto.')
     
     # # Add some text for labels, title and custom x-axis tick labels, etc.
-    ax.set_ylabel(args.ylabel)
+    ax.set_ylabel(ylabel)
     ax.set_xlabel(args.xlabel)
     ax.set_title(args.title)
     ax.set_xticks(ind)

### 2019-09-12 paper/diff_plots.py

@@ -52,6 +52,7 @@ plt.errorbar(atom_s_b, AVG, STD, marker='o', markersize=10, linestyle=' ')
 plt.xlabel('# atoms', fontsize=12)
 plt.ylabel('$D_{cyto}/D_{0}$', fontsize=12)
 # plt.text(-5000,0.2,'C',fontsize=14)
+plt.ylim([0, .5])
 plt.savefig('diff_cyto_over_singles.pdf', bbox_inches="tight")
 plt.close()
 

### 2019-09-12 paper/virtual_cytoplasm_v2.tex
shrinking volume and density plots added

@@
 
-< SHRINKING PLOTS SHOWING BOX VOLUME/DENSITY ? >
+\begin{figure}
+\includegraphics[scale=0.6]{shrinking_volume.pdf} 
+\includegraphics[scale=0.6]{shrinking_density.pdf} 
+\caption{A) The volume of the box with length size of $L=35 \,\, nm$ reduces during the energy minimization in the first $300 \,\, ps$ B) The density of the same box increases within the first $300\,\, ps$.
+The repeated experiments are shown in different colors.
+}
+\end{figure}
 
@@
 \begin{figure}
-
 \includegraphics[scale=0.6]{msd.pdf}
@@
 \includegraphics[scale=0.6]{rdf_mg.pdf}
-\caption{Radial distribution function showing the probability of finding a metabolite or an RNA around MG2+.}
+\caption{Radial distribution function (rdf) showing the probability of finding a metabolite or an RNA around MG2+. The inset plot shows the rdf for short distance ($ r < 0.6 \,\, nm$)}
 \end{figure}

### 2019-09-12 paper/virtual_cytoplasm_v2.tex
manuscript updated
Discussion added

@@
 
+\author{Leandro Oliveira Bortot}
+\affiliation{Laboratory of Biological Physics, School of Pharmaceutical Sciences of Ribeir{\~a}o Preto, University of S{\~a}o Paulo, Ribeir{\~a}o Preto, Brazil}
+
 \author{Zahedeh Bashardanesh}
 \affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology. Uppsala University, SE-751 05 Uppsala, Sweden}
-
-\author{Leandro Oliveira Bortot}
-\affiliation{Laboratory of Biological Physics, School of Pharmaceutical Sciences of Ribeir{\~a}o Preto, University of S{\~a}o Paulo, Ribeir{\~a}o Preto, Brazil}
 
@@
 
-Structure and dynamics of biomolecules are well characterized {\em in vitro}, however how these features differ {\em in vivo} remains unclear. 
-Investigating biomolecular properties {\em in vivo} is possible through development in the fields of nuclear magnetic resonance~\cite{reckel2007,pielak2008}, or fluoroscence spectroscopies~\cite{ignatova2004,xie2008,English2011}.
-An alternative is to use computational models and simulations techniques.
+Structure and dynamics of biomolecules are well characterized {\em in vitro}, however these studies {\em in vivo} are still evolving. 
+Investigating biomolecular properties {\em in vivo} is possible through developments in the fields of nuclear magnetic resonance~\cite{reckel2007,pielak2008}, or fluoroscence spectroscopies~\cite{ignatova2004,xie2008,English2011}.
+An alternative method is to use computational models and simulation techniques.
 Biomolecular simulations are often carried out under dilute conditions or simple models of macromolecular crowdings~\cite{Spiga2014a,Harada2012a,Nawrocki2017a}.
 However, more attemps in modeling bacterial cytoplasm have been made recently~\cite{Mcguffee2010,Yu2016a}.
+
+\begin{itemize}
+   \item How a computational model can be useful
+   \item From a simple extrapolation in time, it has been estimated that the simulation of an {\em E.
+coli} bacterium would be possible in 25 years from now for one nanosecond with $10^{11}$ atoms~\cite{vanGunsteren2006a}
+\end{itemize}
 
@@
 \begin{figure}
-\includegraphics[scale=0.6]{shrinking_volume.pdf} 
-\includegraphics[scale=0.6]{shrinking_density.pdf} 
+\includegraphics[scale=0.7]{shrinking_volume.pdf} 
+\includegraphics[scale=0.7]{shrinking_density.pdf} 
 \caption{A) The volume of the box with length size of $L=35 \,\, nm$ reduces during the energy minimization in the first $300 \,\, ps$ B) The density of the same box increases within the first $300\,\, ps$.
@@
 
+ \begin{itemize}
+  \item Use of force field should be taken with care due to aggregation that was reported repeatedly.
+We used AMBER scaled force field for protein water interaction.
+In other atomistic model of Mycoplasma genitalium, the CHARRM scaled protein water interaction was used for the same reason.
+  \item No significant structural changes (such as unfoldin or partial unfolding) can be seen from properties such as sasa (fig 5), and gyration radius  (WE HAVEN'T ADD THIS FIG, SHALL WE?).
+  \item intramoleclar protein protein interactions are not affected by the presence of crowders. This shows for oligomers (1,3,4,6,7,8 NAME THEM), the crowding doesn't cause dissociation of chains. 
+  \item  intermolecular protein protein interactions can be seen in forms of  rmsd (fig 4) and rmsf (fig 6).
+As proteins interact with their environment, their structure flutuate more than in dilute cases.
+However, the effect is not large enough to cause unfolding (fig 5, and gyration)
+  \item rmsf in all cases are higher for crowders in cytoplasm than in dilute codition
+  \item AREN'T RMSD AND RMSF CORRELATED? 
+  \item Diffusion coefficient is significantly different, and it's independent of the size of the proteins (D_cyto/D_single plot in Supp.
+Info.).
+Based on the observations above (sasa, gyration, rmsd, rmsf) the drop is, to a greater amount, due to the excluded volume effect, and to a lesser amount, due electristatic interactions that has been compromised because of choice of force field.
+  \item the greater drop in diffusion coefficient of RNA is due to confinement of RNA by metabolites (fig 9 and 10).
+  \item NOT SURE ABOUT THIS: THE FACT THAT METABOLITES ARE POLYMERIZED DOES NOT SABOTAGE THE WHOLE SIMULATION AND WE STIL COULD ANALYSE THE SIMULATION AND GET MEANINGFUL OUTCOME, AND THIS WOULD NOT (MOST PROBABLY) BE AFFECTED BY NON-CLUSTERED METABOLITE.
+INDEED WE DID NOT ANALYZE CHEMICAL OR DYNAMIC PROPERTIES OF METABOLITES
+ \end{itemize}
+
 \bibliography{library}

### 2019-09-18 paper/virtual_cytoplasm_v2.tex
Discussion added / essential files for running the ingredinet-make and soup-maker
wrote more about results

@@
 \usepackage[font=small,labelfont=bf]{caption}
+\usepackage{multirow}
+\usepackage{xcolor}
+\usepackage{siunitx}
+\usepackage{float}
+\usepackage{upgreek}
 \setkeys{acs}{super = true}
@@
 In this section we describe the rationale behind the composition of our model, which has five fractions: protein, RNA, metabolites, water and ions.
-We highlight that we didn't add lipids and DNAs to our model because we are considering only elements that are free to diffuse through the cytosol.
+We highlight that we didn't add lipids and DNA to our model because we are considering only elements that are free to diffuse through the cytosol.
 We gathered data from several sources in order to build a computational model that is representative of the cytoplasm of {\em Escheria coli}~\cite{Dong1996,Bennett2009,Link1997,Mcguffee2010}.
@@
 \subsubsection{Protein fraction}
-We selected a group of eight proteins that account for 50\% of the abundance of non-ribosomal proteins in the cytoplasm of Escherichia coli K-12 \cite{ref_prot_abundance}.
-The two most abundant proteins, TufA and MetE, account for 20\% and 12\% of the total protein count in E.coli K-12, respectively, while the other six proteins contribute with less than 5\% each (Table ~\ref{tbl:protein_fraction}).
+We selected a group of eight proteins that account for 50\% of the abundance of non-ribosomal proteins in the cytoplasm of {\em Escherichia coli} \cite{ref_prot_abundance}.
+The two most abundant proteins, TufA and MetE, account for about 20\% and 12\% of the total protein abundance in \em{E.
+coli}, respectively, while the other six proteins contribute with less than 5\% each (Table ~\ref{tbl:protein_fraction}).
+Their crystallogaphic structures were used to insert them in the cytoplasm model.
+The number of copies of each protein in the model was calculated as the rounded up fold increase in their abundance when compared to the least abundant protein, which was set to have the copy number of 1.
+Then, this number was divided by the oligomeric state of each protein.
+This was done because the original oligomeric state for each protein was kept as reported in their crystallogaphic structure.
 
@@
 \centering
-\begin{tabular}{lcc}
+\begin{tabular}{ l c c }
 \hline
-Protein & Abundance in E.Coli K12 \\ 
+\multirow{2}{*}{Protein} & 	\multicolumn{2}{c}{Fraction [\%]} \\
+& Absolute & Cumulative \\ 
 \hline
-TufA	 & 20 \\
-MetE	 & 12 \\
-IcdA	 & 5 \\
-AhpC	 & 4 \\
-CspC	 & 4 \\
-Ppa & 3 \\
-GapA	 & 2 \\
-Eno & 2 \\
+TufA & 19.7 & 19.7 \\
+MetE & 11.6 & 31.4 \\
+IcdA & 4.7  & 36.1 \\
+AhpC & 4.1  & 40.2 \\
+CspC & 4.0  & 44.2 \\
+Ppa  & 2.9  & 47.0 \\
+GapA & 2.1  & 49.1 \\
+Eno  & 1.9  & 51.1 \\
 \hline
 \end{tabular}
-\caption{Protein fraction}
+\caption{Absolute and cumulative fraction of the total abundance of non-ribosomal proteins in the cytosol of {\em E.
+coli} K12 to which each protein that was selected to compose our cytoplasm model corresponds to.}
 \label{tbl:protein_fraction}
@@
 \subsubsection{RNA fraction}
-In order to build the RNA fraction of our cytoplasm model, we consider the following data:
-Since 74\% of the dry weight of non-ribosomal RNAs is composed by tRNA, we chose to model the RNA presence with tRNA molecules.
+Transporter RNAs (tRNAs) account for 74\% of the dry weight of non-ribosomal RNAs.
+Thus, we chose to model the RNA presence in the cytoplasm with tRNA molecules.
+Specifically, we considered the tRNA(Phe) molecule as a representative of tRNAs due to the availability of a recent crystallographic structure~\cite{Byrne2015}.
 The protein and RNA content of the total dry weight of \textit{E.coli} is 55\% and 2.9\%, respectively.
 That is, the total RNA weight corresponds to 5\% of the total protein weight.
-We incorporated this data in our model by considering the tRNA(Phe) molecule as a representative of RNAs due to the availability of a recent crystallographic structure~\cite{Byrne2015}.
-The correct number of copies were added to the simulation box to account for the correct protein/RNA weight ratio (Table ).
+This protein/RNA weight ratio was used to calculate the correct number of tRNA(Phe) molecules that were added to the cytoplasm model \colorbox{red!50}{(Table XXX)}.
 
 \subsubsection{Metabolites fraction}
-The total number of metabolite molecules was calculated considering data showing that the number of metabolite molecules in the cytoplasm of \textit{E.
-coli} is about 42.86 times higher than the number of proteins.
-We considered the most abundant molecule as representative of each metabolite class, i.e.
-Glutamate for amino acids, ATP for nucleotides, FBP for central carbon intermediates and Glutathione redox cofactors.
-The copy number for each molecule was calculated from the ratio of their experimentally observed concentration in {\em E.
-coli}.
+We considered the most abundant molecules of each metabolite class as representatives, i.e.
+Glutamate for amino acids, ATP for nucleotides, FBP for central carbon intermediates and Glutathione redox cofactors ~\cite{metabolite_proportion}.
+The total number of molecules was calculated considering data showing that the number of metabolite molecules in the cytoplasm of \textit{E.
+coli} is about 42.86 times higher than the number of proteins ~\cite{metabolite_proportion}.
+The copy number for each molecule was calculated from the ratios between their experimentally observed concentration in \em{E.
+coli} ~\cite{metabolite_proportion}.
 
 \subsubsection{Water fraction}
-The number of water molecules was calculated according to the desired biomolecular concentration, which ranges from XX to YY in biological systems such as E.
-coli cytoplasm.
-In our case, we choose the biomolecular concentration of 30\%, that is, the number of molecules necessary to reach a ratio of total biomolecular mass to water mass of 30\%.
+The number of water molecules was calculated according to the desired biomolecular concentration, which is a parameter of the cytoplasm model.
+This number ranges from \colorbox{red!50}{XX to YY} in biological systems such as \textit{E.
+coli} cytoplasm.
+In our case, we choose the biomolecular concentration of 30\%, that is, the number of molecules necessary to reach a ratio of total biomolecular mass to water mass of 30\% was inserted into the cytoplasm model.
 
 \subsubsection{Inorganic ions fraction}
-Finally, Mg2+ was used as counter-ions for tRNA and K+ cations were added to neutralize the charges of the simulation box.
-The number of water molecules was used to calculate the number of K+ and Cl- that were necessary to reach the ionic strength of 0.150 M.
+Finally, Mg\textsuperscript{2+} was used as counter-ions for tRNA and ATP.
+KCl was added to neutralize the charges of the simulation box and to reach the ionic strength of 0.150 mol/L by substution of randomly selected water molecules.
 
@@
 \end{tabular}
-\caption{Cytoplasm Components}
+\caption{Number of copies for each component of the cytoplasm model built at the biomolecular fraction of 30\%}
 \label{tbl:soup_components}
@@
 \subsection{Building the simulation box}
-All components can be put in the same simulation box by inserting each of them in random orientations in a cubic box of side L that is initially empty.
+All components can be put in the same simulation box by inserting each of them in random orientations in a cubic box of side $L$ that is initially empty.
 However, that is not a trivial process.
-We need to use a box size that is big enough to allow the random insertions to succeed without structural overlapping and without creating artificial interactions between the elements, that is, we need L to be big enough so that different elements are not placed too close from each other.
-On the other hand, as the box gets bigger, there will be more space between the components and the system will be harder to equilibrate due to larger variations in box size.
+We need to use a box size that is big enough to allow the random insertions to succeed without structural overlapping and without creating artificial interactions between the elements by placing them too close from each other.
+On the other hand, as the box gets bigger, it gets harder to equilibrate the system because the empty space between the components will induce the barostat to reduce the box volume quickly during the simulation.
 
 We devised an iterative process that solves both problems simultaneously.
-We start with a box size L that is too small to allow all systems to fit in the box by random insertion.
-In our case, we started with L=30nm.
+We start with a box size $L$ that is too small to allow all components to fit in the box by random insertion.
+In our case, we started with $L = 30$ nm.
 Then, we allow 100 insertion trials for each element.
-If all trials fail for any of them, we increase the box by a dL step of 1nm and start again until all insertions succeed.
-In our case, all insertions succeeded after increasing L to 35nm.
+If all trials fail for any of them, we start the whole process again with an empty cubic box that is larger by a step size $dL$ of 1 nm.
+We repeat this process until all insertions succeed.
+In our case, all insertions succeeded after increasing $L$ to 35 nm.
 Additionally, instead of adding only the protein, RNA or metabolite molecules in the empty box in each trial, we actually add a droplet of water and counter ions in which the molecule of interest is embedded.
-Such droplets are taken from molecular dynamics simulations in which each component was previously simulated.
-The benefits of using such droplets are threefold: i.
-it prevents artificial contacts between the components as consequence of packing in the simulation box because the water droplet acts as natural protecting layer.
-ii.
-it is a natural way to place water molecules and counter-ions in the simulation box around each component.
-iii.
-the components are already pre-equilibrated, which will help us to perform the equilibration of the whole simulation box.
+Such droplets are taken from molecular dynamics simulations in which each component was previously equilibrated.
+The benefits of using such droplets are threefold: \textit{i.} it acts as natural protecting layer that prevents artificial contacts between the components that could be created due to the random insertions.
+\textit{ii.} it is a natural way to place water molecules and counter-ions in the simulation box around each component.
+\textit{iii.} the components are already pre-equilibrated, which will help us to perform the equilibration of the whole cytoplasm model.
 
 In order to do this, the droplets around each component are also constructed iteractively.
-The number of water molecules that we must place in the simulation box is known.
-However, we don't know the thickness of the water layer around each component, l, that accounts for such amount of water molecules.
-Since l depends on a series of factors such as the shape, size and abundancy of each element, we define it iteratively.
-We start by taking a droplet of thickness l=3A around all elements and counting the number of water that we would add to the box.
-If it is smaller than the number we need, l is increased by dl=1A.
-If it is larger, we take a step back and reduce dl by 10 times.
-We can carry this process until an arbitrary precision cutoff is satisfied.
-
-< IMAGE SHOWING DROPLETS, l, dl, THE EMPTY BOX AND INSERTIONS, L, dL >
+The number of water molecules that we must place in the simulation box is known from the desired biomolecular fraction, which is a parameter of the model, and the total biomolecular mass of the system.
+However, we don't know the thickness of the water layer around each component, $l$, that accounts for such amount of water molecules.
+Since $l$ depends on a series of factors such as the shape, size and abundancy of each element, we define it iteratively.
+We start by taking a droplet of thickness $l$ = 3 \r{A} around all elements and counting the number of water molecules that we would add to the box with such droplets.
+If it is smaller than the number we need, $l$ is increased by $dl$ = 1 \r{A}.
+If it is larger, we reduce the droplet thickness by $dl$, reduce $dl$ by 10 times and then increate $l$ with the new smaller $dl$.
+We can carry this process until an arbitrary precision cutoff, such as 5\%, is satisfied.
 
@@
 In this step the box shrinks to its optimum size.
-In our case, the simulation box shrank from the initial L=35nm to L=22.9nm.
+In our case, the simulation box shrank from the initial box size of 35 nm to 22.9 nm, which correspond to a volume change from 42875 nm\textsuperscript{3} to 12009 nm\textsuperscript{3}.
 From this point, the system is ready to be submitted to the default simulation steps such as thermalization and production run (please check the materials section for details about the parameters we used).
 
-\begin{figure}
-\includegraphics[scale=0.7]{shrinking_volume.pdf} 
-\includegraphics[scale=0.7]{shrinking_density.pdf} 
+\begin{figure}[H]
+\includegraphics[scale=0.5]{shrinking_volume.pdf} 
+\includegraphics[scale=0.5]{shrinking_density.pdf} 
 \caption{A) The volume of the box with length size of $L=35 \,\, nm$ reduces during the energy minimization in the first $300 \,\, ps$ B) The density of the same box increases within the first $300\,\, ps$.
@@
 
-Python scripts and all topology and structure files necessary to perform all these steps are available at github.com/dspoel...
-The cytoplasm model that we built here can be used as a tool to answer many research questions about the effects of crowding on specific systems of interest.
-With the files we are providing here it is possible, for example, to add a probe protein to investigate the effect of crowding on it, to add new macromolecular or small crowders, add new solvents and change the biomolecular concentration to increase or decrease the intensity of the crowding effect.
+Python scripts and all topology and structure files necessary to enable any researcher to build its own cytoplasm models are publicly available at github.com/dspoel/soup.
+Such models can be used as tools that answer many research questions about the effects of crowding on specific systems of interest.
+With the files we are providing it is possible, for example, to add a probe protein to investigate the effect of crowding on it, to add new macromolecular or small crowders, verify the effects of different temperatures and change the biomolecular concentration to increase or decrease the intensity of the crowding effect.
 
@@
 
-In order to investigate the effects of crowding in the E.
-coli cytoplasm model on the behavior of its elements, we constructed three boxes that have different orientations for each of its elements and submitted each to molecular dynamics simulations of 1 us.
-We also performed 200 ns molecular dynamics simulations for each isolated element (please check the materials section for details about the parameters).
+In order to investigate the effects of crowding in the \textit{E.
+coli} cytoplasm model on the structural integrity and dynamics of its elements, we constructed three completely independent cytoplasm models that have different orientations for each of its elements.
+We submitted these systems, each composed by more than 1.5 million atoms, to molecular dynamics simulations of \SI{1}{\micro\second}.
+We also performed \SI{200}{\nano\second} molecular dynamics simulations for each isolated element in order to represent the non-crowded, i.e.
+dilute, environment (please check the materials section for details about the parameters).
+
+% -------------------- TRANSLATIONAL DIFFUSION --------------------
 
 \subsubsection{Translational diffusion}
-D/D0 for all crowders
 
-\begin{figure}
-\includegraphics[scale=0.6]{msd.pdf}
-\caption{Diffusion coefficient of crowders from single molecule simulations (blue bars) and cytoplasm simulations (orange bars).}
+Under crowded conditions, molecules are constrained to a smaller volume due to collisions with the other components of the crowded environment.
+This extent of this effect is evident when we compare the translational diffusion constant of each component in the cytoplasm model, $D_{cytoplasm}$, and in a dilute system, $D_{dilute}$ (Figure ~\ref{fig:translational_diffusion}A).
+The ration of these quantities is independent of the protein size and is always close to 0.13 (Figure ~\ref{fig:translational_diffusion}B).
+However, the effect of crowding over the translation diffusion of tRNA is two times higher, indicating that its movement is specially constrained in the crowded environment.
+
+\begin{figure}[H]
+\includegraphics[scale=0.5]{msd.pdf}
+\includegraphics[scale=0.5]{diff_cyto_over_singles.pdf}
+\caption{A)Diffusion coefficient of crowders from single molecule simulations (blue bars) and cytoplasm simulations (orange bars).
+B) The ratio between diffusion coefficient obtained from cytoplasm to single molecule simulations.
+The crowders are sorted according to their sizes on the x-axis.
+\colorbox{blue!50}{ $<<$ Change the x-axis of D/D0 plot to kDa $>>$ }}
+\label{fig:translational_diffusion}
 \end{figure}
 
-\begin{figure}
-\includegraphics[scale=0.6]{diff_cyto_over_singles.pdf}
-\caption{The ratio between diffusion coefficient obtained from cytoplasm to single molecule simulations. The crowders are sorted according to their sizes on the x-axis.}
+\colorbox{green!50}{ Are we going to calculate and discuss rotational diffusion? }
+
+% -------------------- tRNA IS AN OUTLIER DUE TO AGGREGATION --------------------
+
+\subsubsection{tRNA is aggregating}
+After inspecting the trajectories of the cytoplasm models, we found that the reason why the translational diffusion constant for tRNA was reduced to a greater extent than for the other components of the cytoplasm model is that it is forming aggregates with Mg$^{2+}$, ATP and FBP (Figure ~\ref{fig:tRNA_aggregation}).
+Thus, in the next sections we will ignore the RNA molecules in the analyses about the structural integrity of the crowders.
+Additionally, in the last section we will further investigate such aggregation and we will show a way to prevent it.
+
+\begin{figure}[H]
+\includegraphics[scale=0.5]{rdf_RNA_metabolites.pdf}
+\caption{Radial distribution function showing the probability of finding metabolites or cations around RNA molecules in the cytoplasm model.}
+\label{fig:tRNA_aggregation}
 \end{figure}
 
-\subsubsection{Structural integrity of the crowders}
-RMSD and SASA plots, show that the crowders didn't unfold
+% -------------------- STRUCTURAL INTEGRITY OF SINGLE CHAINS --------------------
 
-\begin{figure}
-\includegraphics[scale=0.6]{rmsd.pdf}
-\caption{Root mean squared displacement (rmsd) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
+\subsubsection{Structural integrity of individual chains}
+The structure of the individual chains of the crowders that were used in our cytoplasm model were not affected to a great extent by the crowded environment.
+The solvent accessible surface area (SASA) value for all crowders in the cytoplasm model is within 5\% of its value under the dilute condition.
+The root mean square deviation (RMSD) considering only their C$\upalpha$ atoms show a similar behavior, except for TufA, MetE and IcdA, which have RMSD values 105, 25 and 33\% higher in the crowded environment than in the dilute condition.
+Visual inspection of their trajectories showed that there is partial unfolding of their N- or C-terminal residues \colorbox{red!50}{DESCRIBE CONF.
+CHANGES}.
+\begin{figure}[H]
+\includegraphics[scale=0.5]{sasa.pdf}
+\includegraphics[scale=0.5]{rmsd.pdf}
+
+\caption{A) Solvent accessible surface area (sasa) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
+The average sasa for each protein in single molecule simulation is over number of chains of proteins and three replica.
+The average sasa for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.
+The error bars show the standard errors.
+B) Root mean squared displacement (rmsd) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
 The average rmsd for each protein in single molecule simulation is over number of chains of proteins and three replica.
@@
 
-\begin{figure}
-\includegraphics[scale=0.6]{sasa.pdf}
-\caption{Solvent accessible surface area (sasa) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
-The average sasa for each protein in single molecule simulation is over number of chains of proteins and three replica.
-The average sasa for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.
+% -------------------- STRUCTURAL INTEGRITY OF OLIGOMERS --------------------
+
+\subsubsection{Structural integrity of oligomers}
+
+\colorbox{red!50}{WE STILL NEED TO MAKE THIS PLOT}
+
+%\begin{figure}[H]
+%\includegraphics[scale=0.5]{hb.pdf}
+%\caption{Intramolecular hydrogen bonds (hb) for oligomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
+The average hydrogen bonds for each oligomer in single molecule simulation is over three replica.
+The average hydrogen bonds for each oligomer in cytoplasm simulation is over number of appearance of the oligomer in three replica.
 The error bars show the standard errors.}
+%\end{figure}
 
-\end{figure}
+% -------------------- STRUCTURAL DYNAMICS --------------------
 
 \subsubsection{Structural dynamics}
-RMSF. Are there differences between the soup and single sim.?
+Proteins are slighly more flexible in the crowded environment.
+Root mean square fluction (RMSF) calculations show that the flexibility profile of the proteins is not affected by crowding, except in the cases in which there are conformational changes, TufA ...
+as shown before (Figure ~\ref{fig:rmsf}).
+This is in agreement with our results showing that the structural integrity of the proteins is not significantly affected.
+In addition to that, our results show that the proteins are slighly more flexible in the crowded environment than in the dilute condition (Figure ~\ref{fig:rmsf}).
+This is likely due to the nonspecific interations and collisions that often happen under crowding conditions \cite{aasd}.
 
-\begin{figure}
-\includegraphics[scale=0.6]{rmsf.pdf}
+\begin{figure}[H]
+\includegraphics[scale=0.5]{rmsf.pdf}
 \caption{Root mean squared fluctuations (rmsf) for 8 crowders from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
@@
 The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.
-The error bars show the standard errors.}
+The error bars show the standard errors.
+\colorbox{blue!50}{$<<$ Make this plot dividing the plots in 2 columns and 4 rows to save space$>>$}}
+\label{fig:rmsf}
 \end{figure}
 
-\subsubsection{Oligomers under crowded condition}
-\begin{figure}
-\includegraphics[scale=0.6]{hb.pdf}
-\caption{Intramolecular hydrogen bonds (hb) for oligomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
-The average hydrogen bonds for each oligomer in single molecule simulation is over three replica.
-The average hydrogen bonds for each oligomer in cytoplasm simulation is over number of appearance of the oligomer in three replica.
-The error bars show the standard errors.}
+% -------------------- AVOIDING AGGREGATION --------------------
+
+\subsubsection{Aggregation can be avoided by protonating the metabolites}
+
+The initial detection of aggregation in our cytoplasm model was due to the outlier behavior of the translational diffusion constant of tRNA when compared to the other crowders.
+However, tRNA is not causing such phenomenon.
+Instead, it is triggered by the exaggerated interaction between the phosphate-containing metabolites, ATP and FBP, with Mg$^{2+}$.
+tRNA gets involved in the aggregates because Mg$^{2+}$ is present in the box only as counter-ion for ATP and tRNA, and so tRNA is one of the few "sources" of Mg$^{2+}$ in the whole system.
+We looked for ways of avoind aggregation with simulations of small systems composed by these metabolites in high concentration, Mg$^{2+}$ and water (please check the methods section for details about the parameters we used).
+
+We found that completely protonating the phosphate groups of ATP and FBP is enough prevent their aggregation with Mg$^{2+}$ (Figure ~\ref{fig:avoiding_aggregation}).
+While RDFs show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$, this is not observed for ATPH\textsubscript{3} and FBPH\textsubscript{4}.
+\colorbox{red!50}{We further tested this hypothesis by simulating a system composed by tRNA, FBP, ATP and Mg$^{2+}$.
+In this setup, protonating ATP and FBP was also enough to avoid aggregation with Mg$^{2+}$ and the metabolites didn't aggregated around tRNA, which acts as a source of Mg$^{2+}$.}
+
+\begin{figure}[H]
+\includegraphics[scale=0.5]{rdf_mg.pdf}
+\caption{Radial distribution function (rdf) showing the probability of finding a metabolite or an RNA around MG2+. The inset plot shows the rdf for short distance ($ r < 0.6 \,\, nm$)}
+\label{fig:avoiding_aggregation}
 \end{figure}
 
-\subsubsection{Aggregation}
-We observed aggregation of the metabolites that contain phosphate, ATP and FBP, with Mg2+ that was used as initially used as counter-ion for ATP and tRNA.
-After running simulations of small boxes containing just these metabolites and Mg2+, we found that completely protonating their phosphate groups can prevent aggregation.
-Thus, we advise caution when dealing with small molecules that are highly charged, such as ATP and FBP.
-Trial simulations with small boxes in which these metabolites are in high concentration to verify if they'll aggregate with Mg2+ or other charged species that are present in the box.
-In the specific case of ATP and FBP modeled with GAFF in boxes containing Mg2+, we advise that they should be completely protonated even though that is not their physiological protonation state.
+ % -------------------- DISCUSSION --------------------
+
+\section*{Discussion}\label{sec:dissc}
+
+ In the specific case of ATP and FBP modeled with GAFF in boxes containing Mg$^{2+}$, we advise that they should be completely protonated even though that is not their physiological protonation state.
 The topology and structure files we are providing for ATP and FBP on github are already protonated.
 
-< FIGURE SHOWING THE AGGREGATION >
-< FIGURE SHOWING THAT PROTONATED PHOSPHATE DOESN'T AGGREGATE >
+\colorbox{blue!50}{ $<<$ Discuss the translational diffusion results: }
 
-\begin{figure}
-\includegraphics[scale=0.6]{rdf_RNA_metabolites.pdf}
-\caption{Radial distribution function showing the probability of finding a metabolite around RNA molecules.}
-\end{figure}
+\colorbox{blue!50}{what is the meaning of the straight line in D/D0? Is this expected? }
 
-\begin{figure}
-\includegraphics[scale=0.6]{rdf_mg.pdf}
-\caption{Radial distribution function (rdf) showing the probability of finding a metabolite or an RNA around MG2+. The inset plot shows the rdf for short distance ($ r < 0.6 \,\, nm$)}
-\end{figure}
-
-\begin{figure}
-\includegraphics[scale=0.6]{avclust.pdf}
-\caption{Number of metabolite molecules that form a cluster in cytoplasm simulations. The solid colored lines are the average of three replica (shown in grey).}
-\end{figure}
-
-\section*{Discussion}\label{sec:dissc}
+\colorbox{blue!50}{Is this in agreement/disagreement with which models/theories? $>>$ }
 
@@
   \item rmsf in all cases are higher for crowders in cytoplasm than in dilute codition
+  \item RMSFs for cyto have higher statistical error ! (even if we have more copies in cyto.) e.g. crowder-n5
   \item AREN'T RMSD AND RMSF CORRELATED? 
-  \item Diffusion coefficient is significantly different, and it's independent of the size of the proteins (D_cyto/D_single plot in Supp.
+  \item Diffusion coefficient is significantly different, and it's independent of the size of the proteins (D cyto/D single plot in Supp.
 Info.).
@@
  \end{itemize}
+
+% -------------------- CONCLUSIONS --------------------
+
+\section*{Conclusions}\label{sec:dissc}
 

### 2019-09-21 paper/virtual_cytoplasm_v2.tex
Updates
italic text fixed

@@
 We selected a group of eight proteins that account for 50\% of the abundance of non-ribosomal proteins in the cytoplasm of {\em Escherichia coli} \cite{ref_prot_abundance}.
-The two most abundant proteins, TufA and MetE, account for about 20\% and 12\% of the total protein abundance in \em{E.
+The two most abundant proteins, TufA and MetE, account for about 20\% and 12\% of the total protein abundance in {\em E.
 coli}, respectively, while the other six proteins contribute with less than 5\% each (Table ~\ref{tbl:protein_fraction}).
@@
 coli} is about 42.86 times higher than the number of proteins ~\cite{metabolite_proportion}.
-The copy number for each molecule was calculated from the ratios between their experimentally observed concentration in \em{E.
+The copy number for each molecule was calculated from the ratios between their experimentally observed concentration in {\em E.
 coli} ~\cite{metabolite_proportion}.
@@
 \includegraphics[scale=0.5]{shrinking_volume.pdf} 
-\includegraphics[scale=0.5]{shrinking_density.pdf} 
+\includegraphics[scale=0.11]{box_big.png}
+\includegraphics[scale=0.11]{box_small_.png}
+%\includegraphics[scale=0.5]{shrinking_density.pdf}
+
 \caption{A) The volume of the box with length size of $L=35 \,\, nm$ reduces during the energy minimization in the first $300 \,\, ps$ B) The density of the same box increases within the first $300\,\, ps$.

### 2019-09-21 paper/virtual_cytoplasm_v2.tex
rmsf plot, Dcyto/D0 plot, added discussion, cleaned a bit the text

@@
 B) The ratio between diffusion coefficient obtained from cytoplasm to single molecule simulations.
-The crowders are sorted according to their sizes on the x-axis.
-\colorbox{blue!50}{ $<<$ Change the x-axis of D/D0 plot to kDa $>>$ }}
+The crowders are sorted according to their sizes on the x-axis.}
 \label{fig:translational_diffusion}
@@
 
-\colorbox{green!50}{ Are we going to calculate and discuss rotational diffusion? }
+% \colorbox{green!50}{ Are we going to calculate and discuss rotational diffusion? No, not now at least.}
 
@@
 
-\colorbox{red!50}{WE STILL NEED TO MAKE THIS PLOT}
+% \colorbox{red!50}{WE STILL NEED TO MAKE THIS PLOT}
 
-%\begin{figure}[H]
-%\includegraphics[scale=0.5]{hb.pdf}
-%\caption{Intramolecular hydrogen bonds (hb) for oligomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
-The average hydrogen bonds for each oligomer in single molecule simulation is over three replica.
-The average hydrogen bonds for each oligomer in cytoplasm simulation is over number of appearance of the oligomer in three replica.
+\begin{figure}[H]
+\includegraphics[scale=0.5]{oligomer.pdf}
+\caption{Inter-chain hydrogen bonds (hb) for oligomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
+The average hydrogen bonds for each oligomer in single molecule simulation is over number of chains and three replica.
+The average hydrogen bonds for each oligomer in cytoplasm simulation is over number of appearance of the chain in the oligomer in three replica.
 The error bars show the standard errors.}
-%\end{figure}
+\end{figure}
 
@@
 \begin{figure}[H]
-\includegraphics[scale=0.5]{rmsf.pdf}
+\includegraphics[scale=0.5]{rmsf_2cols.pdf}
 \caption{Root mean squared fluctuations (rmsf) for 8 crowders from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
@@
 The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.
-The error bars show the standard errors.
-\colorbox{blue!50}{$<<$ Make this plot dividing the plots in 2 columns and 4 rows to save space$>>$}}
+The error bars show the standard errors.}
 \label{fig:rmsf}
@@
 \colorbox{blue!50}{ $<<$ Discuss the translational diffusion results: } 
+\colorbox{blue!50}{what is the meaning of the straight line in D/D0? Is this expected? }
+\colorbox{blue!50}{Is this in agreement/disagreement with which models/theories? $>>$ }
 
-\colorbox{blue!50}{what is the meaning of the straight line in D/D0? Is this expected? }
-
-\colorbox{blue!50}{Is this in agreement/disagreement with which models/theories? $>>$ }
+Diffusion coefficient of crowders from single molecules to cytoplasm are decreased in a range of 84\% (Ppa) to 93\% (tRNA) (Fig.~\ref{fig:translational_diffusion} A).
+The straight (fitted) line with the slope very close to zero  ($m = 0.00028$) shows that the drop in diffusion coefficient is independent of the size of proteins.
+The largest drop of diffusion coefficient corresponds to tRNA and we found that the aggregation of ATP$^{-3}$ and FBP$^{-4}$ around Mg$^{2+}$ which was used as cation for tRNA leads to its confinement.
+In fact, by removing diffusion coefficient of tRNA from the data points the slope of the line decreases by half ($m=0.00014$).
 

### 2019-09-21 paper/barplots.py

@@ -52,7 +52,7 @@ if __name__ == '__main__':
             ylabel = "$nm$"
         elif args.prop == "msd":
             ylabel = '$D (10^{-5} cm^2/s)$'
-        elif args.prop == "hb":
+        elif args.prop == "oligomer" or args.prop == "hb":
             ylabel = "#HB"
         else:
             print("Unknown property " + args.prop)
@@ -87,7 +87,7 @@ if __name__ == '__main__':
     ax.set_xlabel(args.xlabel)
     ax.set_title(args.title)
     ax.set_xticks(ind)
-    if args.prop == "hb":
+    if args.prop == "oligomer" or args.prop == "hb":
         ax.set_xticklabels(oligomers)
     elif args.prop == "msd":
         ax.set_xticklabels(crowder_name)

### 2019-09-21 paper/diff_plots.py

@@ -11,6 +11,10 @@ atom_s_b = sorted(atom_crowder)
 
 order_crowder=[4, 7, 6, 3, 8, 1, 2, 9, 5] #biggest to smallest
 order_s_b=[5, 9, 2, 1, 8, 3, 6, 7, 4] #smallest to biggest
+ 
+mass_crowder = [84320.000,  83764.400,  91489.800,  184409.000,  7271.120,  115920.600,  141602.400,  90884.000,  23761.100] #  146.124,  504.164,  336.080,  306.318]
+mass_sorted = sorted(mass_crowder)
+
 
 avg_cyto, ste_cyto =[],[]
 for i in range(0, len(order_s_b)):
@@ -19,12 +23,12 @@ for i in range(0, len(order_s_b)):
 			if "crowder_n"+str(order_s_b[i]) in line:
 				avg_cyto.append(float(line.split()[1]))
 				ste_cyto.append(float(line.split()[2]))
-plt.errorbar(atom_s_b, avg_cyto, ste_cyto, marker='o', markersize=10, linestyle=' ')
-plt.xlabel('# atoms', fontsize=12)
-plt.ylabel('$D_{cyto} (10^{-5} cm^2/s)$', fontsize=12)
-plt.text(-5000,0.018,'B',fontsize=14)
-plt.savefig('diff_cyto.pdf', bbox_inches="tight")
-plt.close()
+# plt.errorbar(atom_s_b, avg_cyto, ste_cyto, marker='o', markersize=10, linestyle=' ')
+# plt.xlabel('# atoms', fontsize=12)
+# plt.ylabel('$D_{cyto} (10^{-5} cm^2/s)$', fontsize=12)
+# plt.text(-5000,0.018,'B',fontsize=14)
+# plt.savefig('diff_cyto.pdf', bbox_inches="tight")
+# plt.close()
 
 
 avg_sing, ste_sing = [],[]
@@ -34,29 +38,31 @@ for i in range(0, len(order_s_b)):
 			if "crowder_n"+str(order_s_b[i]) in line:
 				avg_sing.append(float(line.split()[1]))
 				ste_sing.append(float(line.split()[2]))
-plt.errorbar(atom_s_b, avg_sing, ste_sing, marker='o', markersize=10, linestyle=' ')
-plt.xlabel('# atoms', fontsize=12)
[…]

### 2019-09-22 paper/virtual_cytoplasm_v2.tex

@@
 Under crowded conditions, molecules are constrained to a smaller volume due to collisions with the other components of the crowded environment.
-This extent of this effect is evident when we compare the translational diffusion constant of each component in the cytoplasm model, $D_{cytoplasm}$, and in a dilute system, $D_{dilute}$ (Figure ~\ref{fig:translational_diffusion}A).
+This extent of this effect is evident when we compare the translational diffusion constant of each component in the cytoplasm model, $D_{cyto}$, and in a dilute system, $D_{dil}$ (Figure ~\ref{fig:translational_diffusion}A).
 The ration of these quantities is independent of the protein size and is always close to 0.13 (Figure ~\ref{fig:translational_diffusion}B).

### 2019-09-22 paper/virtual_cytoplasm_v2.tex
manuscript update

@@
 \subsubsection{Protein fraction}
-We selected a group of eight proteins that account for 50\% of the abundance of non-ribosomal proteins in the cytoplasm of {\em Escherichia coli} \cite{ref_prot_abundance}.
+We selected a group of eight proteins that account for 50\% of the abundance of non-ribosomal proteins in the cytoplasm of {\em Escherichia coli} \cite{Link1997}.
 The two most abundant proteins, TufA and MetE, account for about 20\% and 12\% of the total protein abundance in {\em E.
-coli}, respectively, while the other six proteins contribute with less than 5\% each (Table ~\ref{tbl:protein_fraction}).
+coli}, respectively, while the other six proteins contribute with less than 5\% each (Table~\ref{tbl:protein_fraction}).
 Their crystallogaphic structures were used to insert them in the cytoplasm model.
@@
 We considered the most abundant molecules of each metabolite class as representatives, i.e.
-Glutamate for amino acids, ATP for nucleotides, FBP for central carbon intermediates and Glutathione redox cofactors ~\cite{metabolite_proportion}.
+Glutamate for amino acids, ATP for nucleotides, FBP for central carbon intermediates and Glutathione redox cofactors~\cite{Bennett2009}.
 The total number of molecules was calculated considering data showing that the number of metabolite molecules in the cytoplasm of \textit{E.
-coli} is about 42.86 times higher than the number of proteins ~\cite{metabolite_proportion}.
+coli} is about 42.86 times higher than the number of proteins~\cite{Bennett2009}.
 The copy number for each molecule was calculated from the ratios between their experimentally observed concentration in {\em E.
-coli} ~\cite{metabolite_proportion}.
+coli}~\cite{Bennett2009}.
 
@@
 Under crowded conditions, molecules are constrained to a smaller volume due to collisions with the other components of the crowded environment.
-This extent of this effect is evident when we compare the translational diffusion constant of each component in the cytoplasm model, $D_{cyto}$, and in a dilute system, $D_{dil}$ (Figure ~\ref{fig:translational_diffusion}A).
-The ration of these quantities is independent of the protein size and is always close to 0.13 (Figure ~\ref{fig:translational_diffusion}B).
+This extent of this effect is evident when we compare the translational diffusion constant of each component in the cytoplasm model, $D_{cyto}$, and in a dilute system, $D_{dil}$ (Figure~\ref{fig:translational_diffusion}A).
+The ration of these quantities is independent of the protein size and is always close to 0.13 (Figure~\ref{fig:translational_diffusion}B).
 However, the effect of crowding over the translation diffusion of tRNA is two times higher, indicating that its movement is specially constrained in the crowded environment.
@@
 \subsubsection{tRNA is aggregating}
-After inspecting the trajectories of the cytoplasm models, we found that the reason why the translational diffusion constant for tRNA was reduced to a greater extent than for the other components of the cytoplasm model is that it is forming aggregates with Mg$^{2+}$, ATP and FBP (Figure ~\ref{fig:tRNA_aggregation}).
+After inspecting the trajectories of the cytoplasm models, we found that the reason why the translational diffusion constant for tRNA was reduced to a greater extent than for the other components of the cytoplasm model is that it is forming aggregates with Mg$^{2+}$, ATP and FBP (Figure~\ref{fig:tRNA_aggregation}).
 Thus, in the next sections we will ignore the RNA molecules in the analyses about the structural integrity of the crowders.
@@
 Root mean square fluction (RMSF) calculations show that the flexibility profile of the proteins is not affected by crowding, except in the cases in which there are conformational changes, TufA ...
-as shown before (Figure ~\ref{fig:rmsf}).
+as shown before (Figure~\ref{fig:rmsf}).
 This is in agreement with our results showing that the structural integrity of the proteins is not significantly affected.
-In addition to that, our results show that the proteins are slighly more flexible in the crowded environment than in the dilute condition (Figure ~\ref{fig:rmsf}).
+In addition to that, our results show that the proteins are slighly more flexible in the crowded environment than in the dilute condition (Figure~\ref{fig:rmsf}).
 This is likely due to the nonspecific interations and collisions that often happen under crowding conditions \cite{aasd}.
@@
 
-We found that completely protonating the phosphate groups of ATP and FBP is enough prevent their aggregation with Mg$^{2+}$ (Figure ~\ref{fig:avoiding_aggregation}).
+We found that completely protonating the phosphate groups of ATP and FBP is enough prevent their aggregation with Mg$^{2+}$ (Figure~\ref{fig:avoiding_aggregation}).
 While RDFs show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$, this is not observed for ATPH\textsubscript{3} and FBPH\textsubscript{4}.
@@
 
+\begin{itemize}
+  \item Use of force field should be taken with care due to aggregation that was reported repeatedly.
+We used AMBER scaled force field for protein water interaction.
+In other atomistic model of Mycoplasma genitalium, the CHARRM scaled protein water interaction was used for the same reason.
+\end{itemize}
+
  In the specific case of ATP and FBP modeled with GAFF in boxes containing Mg$^{2+}$, we advise that they should be completely protonated even though that is not their physiological protonation state.
@@
 \colorbox{blue!50}{ $<<$ Discuss the translational diffusion results: }
+
+Diffusion coefficient of crowders from single molecules to cytoplasm are decreased in a range of 84\% (Ppa) to 93\% (tRNA) (Fig.~\ref{fig:translational_diffusion} A).
+The fitted line with the slope close to zero  ($m = 0.00028$) shows that the drop in diffusion coefficient is independent of the size of the proteins.
+The largest drop of diffusion coefficient corresponds to tRNA and we found that the aggregation of ATP$^{-3}$ and FBP$^{-4}$ around Mg$^{2+}$ which was used as cation for tRNA leads to its confinement.
+In fact, by removing diffusion coefficient of tRNA from the data points the slope of the line decreases by half ($m=0.00014$).
+Our finding on the drop of diffusion coefficient is closely in line with {\em in vivo}.
+The translational diffusion coefficient of Green Fluorescent Protein (GFP) in {\em E.
+coli} that was measured by fluorescence revovery after photobleaching (FRAP) was decreased 90\% comparing to {\em in vitro} measurement~\cite{Elowitz1999,Konopka2006}.
+
 \colorbox{blue!50}{what is the meaning of the straight line in D/D0? Is this expected? }
+
 \colorbox{blue!50}{Is this in agreement/disagreement with which models/theories? $>>$ }
 
-Diffusion coefficient of crowders from single molecules to cytoplasm are decreased in a range of 84\% (Ppa) to 93\% (tRNA) (Fig.~\ref{fig:translational_diffusion} A).
-The straight (fitted) line with the slope very close to zero  ($m = 0.00028$) shows that the drop in diffusion coefficient is independent of the size of proteins.
-The largest drop of diffusion coefficient corresponds to tRNA and we found that the aggregation of ATP$^{-3}$ and FBP$^{-4}$ around Mg$^{2+}$ which was used as cation for tRNA leads to its confinement.
-In fact, by removing diffusion coefficient of tRNA from the data points the slope of the line decreases by half ($m=0.00014$).
+HARD SPHERE THEORETICAL MODEL PREDICTs THE DROP IN DIFFUSION COEFFICIENT AS A FUNCTION OF VOLUME AVAILABLE TO PARTICLES.
+THEREFORE, THE VOLUME AVAILABLE TO OUR PROTEINS IN OUR MODEL WOULD BE THE SAME IF THE SIMULATION IS LONG ENOUGH FOR ALL PROTEINS TO SPAN THE WHOLE SPACE.
 
  \begin{itemize}
-  \item Use of force field should be taken with care due to aggregation that was reported repeatedly.
-We used AMBER scaled force field for protein water interaction.
-In other atomistic model of Mycoplasma genitalium, the CHARRM scaled protein water interaction was used for the same reason.
   \item No significant structural changes (such as unfoldin or partial unfolding) can be seen from properties such as sasa (fig 5), and gyration radius  (WE HAVEN'T ADD THIS FIG, SHALL WE?).
-  \item intramoleclar protein protein interactions are not affected by the presence of crowders. This shows for oligomers (1,3,4,6,7,8 NAME THEM), the crowding doesn't cause dissociation of chains. 
+  \item inter-chain hydrogen interactions. 
   \item  intermolecular protein protein interactions can be seen in forms of  rmsd (fig 4) and rmsf (fig 6).
@@
   \item RMSFs for cyto have higher statistical error ! (even if we have more copies in cyto.) e.g. crowder-n5
-  \item AREN'T RMSD AND RMSF CORRELATED? 
-  \item Diffusion coefficient is significantly different, and it's independent of the size of the proteins (D cyto/D single plot in Supp.
-Info.).
-Based on the observations above (sasa, gyration, rmsd, rmsf) the drop is, to a greater amount, due to the excluded volume effect, and to a lesser amount, due electristatic interactions that has been compromised because of choice of force field.
-  \item the greater drop in diffusion coefficient of RNA is due to confinement of RNA by metabolites (fig 9 and 10).
   \item NOT SURE ABOUT THIS: THE FACT THAT METABOLITES ARE POLYMERIZED DOES NOT SABOTAGE THE WHOLE SIMULATION AND WE STIL COULD ANALYSE THE SIMULATION AND GET MEANINGFUL OUTCOME, AND THIS WOULD NOT (MOST PROBABLY) BE AFFECTED BY NON-CLUSTERED METABOLITE.

### 2019-09-23 paper/virtual_cytoplasm_v2.tex
Wrote more on discussion and started conclusion

@@
 Investigating biomolecular properties {\em in vivo} is possible through developments in the fields of nuclear magnetic resonance~\cite{reckel2007,pielak2008}, or fluoroscence spectroscopies~\cite{ignatova2004,xie2008,English2011}.
+
 An alternative method is to use computational models and simulation techniques.
@@
 
-\begin{table}
+\begin{table}[H]
 \centering
@@
 \subsubsection{RNA fraction}
-Transporter RNAs (tRNAs) account for 74\% of the dry weight of non-ribosomal RNAs.
+Transporter RNAs (tRNAs) account for 74\% of the dry weight of non-ribosomal RNAs \colorbox{red!50}{REF}.
 Thus, we chose to model the RNA presence in the cytoplasm with tRNA molecules.
 Specifically, we considered the tRNA(Phe) molecule as a representative of tRNAs due to the availability of a recent crystallographic structure~\cite{Byrne2015}.
-The protein and RNA content of the total dry weight of \textit{E.coli} is 55\% and 2.9\%, respectively.
+The protein and RNA content of the total dry weight of \textit{E.coli} is 55\% and 2.9\%, respectively \colorbox{red!50}{REF}.
 That is, the total RNA weight corresponds to 5\% of the total protein weight.
@@
 Finally, Mg\textsuperscript{2+} was used as counter-ions for tRNA and ATP.
-KCl was added to neutralize the charges of the simulation box and to reach the ionic strength of 0.150 mol/L by substution of randomly selected water molecules.
+K$^{+}$ and Cl$^{-}$ were added to neutralize the charges of the simulation box and to reach the ionic strength of 0.150 mol/L by substution of randomly selected water molecules.
 
@@
 
-\subsection{The effect of crowded systems}
+\subsection{Effects of macromolecular crowding}
 
@@
 Under crowded conditions, molecules are constrained to a smaller volume due to collisions with the other components of the crowded environment.
-This extent of this effect is evident when we compare the translational diffusion constant of each component in the cytoplasm model, $D_{cyto}$, and in a dilute system, $D_{dil}$ (Figure~\ref{fig:translational_diffusion}A).
-The ration of these quantities is independent of the protein size and is always close to 0.13 (Figure~\ref{fig:translational_diffusion}B).
-However, the effect of crowding over the translation diffusion of tRNA is two times higher, indicating that its movement is specially constrained in the crowded environment.
+The extent of this effect is evident when we compare the translational diffusion constant of each component in the cytoplasm model, $D_{cyto}$, and in a dilute system, $D_{dil}$ (Figure~\ref{fig:translational_diffusion}A).
+The ratio of these quantities is independent of the protein size and is always close to 0.13 (Figure~\ref{fig:translational_diffusion}B).
+However, the effect of crowding over the translation diffusion of tRNA is two times higher.
+Its $D_{cyto}/D_{dil}$ is 0.07, indicating that its movement is specially constrained in the crowded environment (Figure~\ref{fig:translational_diffusion}B, crowder with 25 kDa).
 
@@
 B) The ratio between diffusion coefficient obtained from cytoplasm to single molecule simulations.
-The crowders are sorted according to their sizes on the x-axis.}
+The crowders are sorted according to their sizes on the x-axis.
+\textbf{\color{red}{THE Dcyto/Ddil PLOT IS "SQUEEZED" BECAUSE OF ITS STRANGE SCALE.
+IT WOULD BE BETTER TO CENTER THE DATAPOITS, E.G.
+USE THE Y-AXIS RANGE OF 0.0-0.2 OR SOMETHING SIMILAR.
+THIS SCALE WILL MAKE THE DIFFERENCE OF THE BEHAVIOR OF tRNA MORE VISIBLE AS WELL}}}
 \label{fig:translational_diffusion}
@@
 \includegraphics[scale=0.5]{rdf_RNA_metabolites.pdf}
-\caption{Radial distribution function showing the probability of finding metabolites or cations around RNA molecules in the cytoplasm model.}
+\caption{Radial distribution function showing the probability of finding metabolites or cations around RNA molecules in the cytoplasm model.\color{red}\textbf{THIS PLOT NEEDS REVISION.
+I DON'T SEE ERROR BARS (HOW WAS STATISTICS DONE FOR THE RDFs?).I SUBMITTED NEW RDF CALCULATIONS CENTERED AROUND tRNA BECAUSE I DIDN'T FIND THE XVG FILES THAT WERE USED TO GENERATE THIS PLOT}}
 \label{fig:tRNA_aggregation}
@@
 The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.
-The error bars show the standard errors.}
+The error bars show the standard errors.\textbf{\color{red}{THE MEASUREMENT NAME, SASA AND RMSD, IS MISSING IN THE Y-AXIS OF BOTH PLOTS}}}
+\label{fig:structural_integrity_chain}
 \end{figure}
@@
 
-\begin{figure}[H]
-\includegraphics[scale=0.5]{oligomer.pdf}
-\caption{Inter-chain hydrogen bonds (hb) for oligomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
+%\begin{figure}[H]
+%\includegraphics[scale=0.5]{oligomer.pdf}
+%\caption{Inter-chain hydrogen bonds (hb) for oligomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
 The average hydrogen bonds for each oligomer in single molecule simulation is over number of chains and three replica.
@@
 The error bars show the standard errors.}
-\end{figure}
+%\end{figure}
 
@@
 In addition to that, our results show that the proteins are slighly more flexible in the crowded environment than in the dilute condition (Figure~\ref{fig:rmsf}).
-This is likely due to the nonspecific interations and collisions that often happen under crowding conditions \cite{aasd}.
+This is likely due to the nonspecific interations and collisions that often happen under crowding conditions \textbf{\color{red}{REF}}.
 
@@
 The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.
-The error bars show the standard errors.}
+The error bars show the standard errors.\textbf{\color{red}{CAN YOU ADD THE PROTEIN NAME AS A LABEL INSIDE EACH PLOT?
+I THINK THE VARIABLE NAME SHOULD BE IN THE Y-AXIS (RMSF NM2)}}}
 \label{fig:rmsf}
@@
 While RDFs show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$, this is not observed for ATPH\textsubscript{3} and FBPH\textsubscript{4}.
-\colorbox{red!50}{We further tested this hypothesis by simulating a system composed by tRNA, FBP, ATP and Mg$^{2+}$.
-In this setup, protonating ATP and FBP was also enough to avoid aggregation with Mg$^{2+}$ and the metabolites didn't aggregated around tRNA, which acts as a source of Mg$^{2+}$.}
 
@@
 \includegraphics[scale=0.5]{rdf_mg.pdf}
-\caption{Radial distribution function (rdf) showing the probability of finding a metabolite or an RNA around MG2+. The inset plot shows the rdf for short distance ($ r < 0.6 \,\, nm$)}
+\caption{Radial distribution function (rdf) showing the probability of finding a metabolite or an RNA around MG2+.
+The inset plot shows the rdf for short distance ($ r < 0.6 \,\, nm$) \color{red}{\textbf{HOW WAS STATISTICS DONE FOR THE RDFs?
+I DON'T SEE ERROR BARS}}}
 \label{fig:avoiding_aggregation}
@@
 
-\begin{itemize}
-  \item Use of force field should be taken with care due to aggregation that was reported repeatedly.
-We used AMBER scaled force field for protein water interaction.
-In other atomistic model of Mycoplasma genitalium, the CHARRM scaled protein water interaction was used for the same reason.
-\end{itemize}
+The main difference between the behavior of the macromolecules that compose our model in the crowded system and in the dilute condition is their translational diffusion coefficient, which drops by more around 85\% for the proteins and 93\% for the tRNA (Fig.~\ref{fig:translational_diffusion} A).
+The slope of a linear regression over values of $D_{cyto}/D_{dil}$ is close to zero ($m = 0.00028$), showing that the drop in diffusion coefficient is independent of the size of the macromolecules.
+We found that the aggregation of ATP$^{-3}$ and FBP$^{-4}$ around Mg$^{2+}$, which was used as cation for tRNA, leads to its exaggerated confinement.
+Thus, tRNA is considered as an outlier in our cytoplasm model.
+In fact, by removing diffusion coefficient of tRNA from the data points of $D_{cyto}/D_{dil}$, the slope of the line changes by 50\% and gets even closer to zero ($m=0.00014$).
+These results are in agreement with experimental results showing that the translational diffusion coefficient of Green Fluorescent Protein (GFP) in {\em E.
+coli}, as measured by fluorescence recovery after photobleaching (FRAP), decreased 90\% when compared to {\em in vitro} measurements~\cite{Elowitz1999,Konopka2006}.
+
+It has been demonstrated that each of the force fields that are most widely used to perform molecular dynamics simulations of macromolecules have different results for highly concentrated amino acid solutions ~ \cite{andrews2013molecular}.
+Previous all-atom simulations of high scale cytoplasm models used parameters from the CHARMM family of force fields ~\cite{Yu2016,Mcguffee2010}.
+In this work we used, for the first time, the AMBER family of force fields to perform all-atom explicit solvent of a cytoplasm model for the total of \SI{3}{\micro\second}.
+
+Except for some local unfolding for TufA, all crowders were stable during the simulations (Figure \ref{fig:structural_integrity_chain} and their flexibility profile doesn't change significantly (Figure \ref{fig:rmsf}.
+This shows that our cytoplasm model is stable even in the microsecond range.
+
+\textbf{\color{red}{MISSING RESULTS: WHOLE CROWDER STRUCTURAL INTEGRITY USING BURIED OLIGOMER AREA}}
+
+However, we observed aggregation of Mg, ATP and FBP...
+
+% -------------------- CONCLUSIONS --------------------
+
+\section*{Conclusions}\label{sec:concl}
+
+In this work we built a cytoplasm model...
+
+the model was validated by reproducing experimental data about the reduction of the translational diffusion coefficient of proteins in E.coli.
+
+the model is stable.
+
+however, aggregation!
 
@@
 
-\colorbox{blue!50}{ $<<$ Discuss the translational diffusion results: }
+\bibliography{library}
 
-Diffusion coefficient of crowders from single molecules to cytoplasm are decreased in a range of 84\% (Ppa) to 93\% (tRNA) (Fig.~\ref{fig:translational_diffusion} A).
-The fitted line with the slope close to zero  ($m = 0.00028$) shows that the drop in diffusion coefficient is independent of the size of the proteins.
-The largest drop of diffusion coefficient corresponds to tRNA and we found that the aggregation of ATP$^{-3}$ and FBP$^{-4}$ around Mg$^{2+}$ which was used as cation for tRNA leads to its confinement.
-In fact, by removing diffusion coefficient of tRNA from the data points the slope of the line decreases by half ($m=0.00014$).
-Our finding on the drop of diffusion coefficient is closely in line with {\em in vivo}.
-The translational diffusion coefficient of Green Fluorescent Protein (GFP) in {\em E.
-coli} that was measured by fluorescence revovery after photobleaching (FRAP) was decreased 90\% comparing to {\em in vitro} measurement~\cite{Elowitz1999,Konopka2006}.
-
-\colorbox{blue!50}{what is the meaning of the straight line in D/D0? Is this expected? }
-
-\colorbox{blue!50}{Is this in agreement/disagreement with which models/theories? $>>$ }
-
-HARD SPHERE THEORETICAL MODEL PREDICTs THE DROP IN DIFFUSION COEFFICIENT AS A FUNCTION OF VOLUME AVAILABLE TO PARTICLES.
-THEREFORE, THE VOLUME AVAILABLE TO OUR PROTEINS IN OUR MODEL WOULD BE THE SAME IF THE SIMULATION IS LONG ENOUGH FOR ALL PROTEINS TO SPAN THE WHOLE SPACE.
-
- \begin{itemize}
-  \item No significant structural changes (such as unfoldin or partial unfolding) can be seen from properties such as sasa (fig 5), and gyration radius  (WE HAVEN'T ADD THIS FIG, SHALL WE?).
-  \item inter-chain hydrogen interactions. 
-  \item  intermolecular protein protein interactions can be seen in forms of  rmsd (fig 4) and rmsf (fig 6).
-As proteins interact with their environment, their structure flutuate more than in dilute cases.
-However, the effect is not large enough to cause unfolding (fig 5, and gyration)
-  \item rmsf in all cases are higher for crowders in cytoplasm than in dilute codition
-  \item RMSFs for cyto have higher statistical error ! (even if we have more copies in cyto.) e.g. crowder-n5
-  \item NOT SURE ABOUT THIS: THE FACT THAT METABOLITES ARE POLYMERIZED DOES NOT SABOTAGE THE WHOLE SIMULATION AND WE STIL COULD ANALYSE THE SIMULATION AND GET MEANINGFUL OUTCOME, AND THIS WOULD NOT (MOST PROBABLY) BE AFFECTED BY NON-CLUSTERED METABOLITE.
-INDEED WE DID NOT ANALYZE CHEMICAL OR DYNAMIC PROPERTIES OF METABOLITES
- \end{itemize}
-
-% -------------------- CONCLUSIONS --------------------
-
-\section*{Conclusions}\label{sec:dissc}
-
-\bibliography{library}
 % \begin{addendum}

### 2019-09-24 paper/virtual_cytoplasm_v2.tex
discussion

@@
 The main difference between the behavior of the macromolecules that compose our model in the crowded system and in the dilute condition is their translational diffusion coefficient, which drops by more around 85\% for the proteins and 93\% for the tRNA (Fig.~\ref{fig:translational_diffusion} A).
-The slope of a linear regression over values of $D_{cyto}/D_{dil}$ is close to zero ($m = 0.00028$), showing that the drop in diffusion coefficient is independent of the size of the macromolecules.
-We found that the aggregation of ATP$^{-3}$ and FBP$^{-4}$ around Mg$^{2+}$, which was used as cation for tRNA, leads to its exaggerated confinement.
+The slope of the line generated by fitting a linear regression model to values of $D_{cyto}/D_{dil}$ is close to zero (slope $= 0.00028$), showing that the drop in diffusion coefficient is independent of the size of the macromolecules.
+We found that the aggregation of ATP$^{-3}$ and FBP$^{-4}$ around Mg$^{2+}$, which was used as counter-ion for tRNA, leads to its exaggerated confinement.
 Thus, tRNA is considered as an outlier in our cytoplasm model.
-In fact, by removing diffusion coefficient of tRNA from the data points of $D_{cyto}/D_{dil}$, the slope of the line changes by 50\% and gets even closer to zero ($m=0.00014$).
+In fact, by removing diffusion coefficient of tRNA from the data, the slope of the linear regression line fitted to $D_{cyto}/D_{dil}$, is reduced by by 50\% and gets even closer to zero (slope $= 0.00014$).
 These results are in agreement with experimental results showing that the translational diffusion coefficient of Green Fluorescent Protein (GFP) in {\em E.
@@
 
-Except for some local unfolding for TufA, all crowders were stable during the simulations (Figure \ref{fig:structural_integrity_chain} and their flexibility profile doesn't change significantly (Figure \ref{fig:rmsf}.
+Except for some local unfolding for TufA, all crowders were stable during the simulations (Figure \ref{fig:structural_integrity_chain}) and their flexibility profile doesn't change significantly (Figure \ref{fig:rmsf}) when compared to the dilute condition.
 This shows that our cytoplasm model is stable even in the microsecond range.
-
-\textbf{\color{red}{MISSING RESULTS: WHOLE CROWDER STRUCTURAL INTEGRITY USING BURIED OLIGOMER AREA}}
+The oligomeric state of the crowders... \textbf{\color{red}{MISSING RESULTS: whole crowder structural integrity using buried oligomer area}}
 

### 2019-09-24 paper/virtual_cytoplasm_v2.tex
Discussion is done.

@@
 \affiliation{Laboratory of Biological Physics, School of Pharmaceutical Sciences of Ribeir{\~a}o Preto, University of S{\~a}o Paulo, Ribeir{\~a}o Preto, Brazil}
+\altaffiliation{These authors contributed equally to this work}
 
@@
 \affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology. Uppsala University, SE-751 05 Uppsala, Sweden}
+\altaffiliation{These authors contributed equally to this work}
 
@@
 
-The main difference between the behavior of the macromolecules that compose our model in the crowded system and in the dilute condition is their translational diffusion coefficient, which drops by more around 85\% for the proteins and 93\% for the tRNA (Fig.~\ref{fig:translational_diffusion} A).
+The main difference between the behavior of the macromolecules that compose our model in the crowded system and in the dilute condition is their translational diffusion coefficient, which drops by more around 85\% for the proteins and 93\% for the tRNA (Fig.~\ref{fig:translational_diffusion}A).
 The slope of the line generated by fitting a linear regression model to values of $D_{cyto}/D_{dil}$ is close to zero (slope $= 0.00028$), showing that the drop in diffusion coefficient is independent of the size of the macromolecules.
@@
 Thus, tRNA is considered as an outlier in our cytoplasm model.
-In fact, by removing diffusion coefficient of tRNA from the data, the slope of the linear regression line fitted to $D_{cyto}/D_{dil}$, is reduced by by 50\% and gets even closer to zero (slope $= 0.00014$).
-These results are in agreement with experimental results showing that the translational diffusion coefficient of Green Fluorescent Protein (GFP) in {\em E.
+In fact, the slope of the linear regression line fitted to $D_{cyto}/D_{dil}$ is reduced by by 50\% and gets even closer to zero (slope $= 0.00014$) after removing the diffusion coefficient of tRNA from the data.
+Our data is in agreement with experimental results showing that the translational diffusion coefficient of Green Fluorescent Protein (GFP) in {\em E.
 coli}, as measured by fluorescence recovery after photobleaching (FRAP), decreased 90\% when compared to {\em in vitro} measurements~\cite{Elowitz1999,Konopka2006}.
@@
 It has been demonstrated that each of the force fields that are most widely used to perform molecular dynamics simulations of macromolecules have different results for highly concentrated amino acid solutions ~ \cite{andrews2013molecular}.
-Previous all-atom simulations of high scale cytoplasm models used parameters from the CHARMM family of force fields ~\cite{Yu2016,Mcguffee2010}.
-In this work we used, for the first time, the AMBER family of force fields to perform all-atom explicit solvent of a cytoplasm model for the total of \SI{3}{\micro\second}.
+Previous all-atom simulations of high scale cytoplasm models used parameters from the CHARMM family of force fields ~\cite{yu2016biomolecular,mcguffee2010diffusion}.
+In this work we use, for the first time, the AMBER force fields to study crowding effects on a cytoplasm model.
+All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of \textit{Mycoplasma genitalium} composed by 103 and 12 million atoms that were simulated for 20 and 60 \SI{}{\nano\second}, respectively ~\cite{yu2016biomolecular}.
+The cytoplasm model described in this work has 1.5 million atoms and was simulated for the total of \SI{3}{\micro\second}, an unprecedented time scale to study crowding effects on a heterogeneous cytoplasm model in atomic detail.
 
-Except for some local unfolding for TufA, all crowders were stable during the simulations (Figure \ref{fig:structural_integrity_chain}) and their flexibility profile doesn't change significantly (Figure \ref{fig:rmsf}) when compared to the dilute condition.
-This shows that our cytoplasm model is stable even in the microsecond range.
+All crowders are as compact in the cytoplasm model as they are in dilute conditions (Figure \ref{fig:structural_integrity_chain}A).
+Despite this overall stability, three crowders (TufA, MetE and IcdA) displayed greater deviation from their crystallographic structure in the cytoplasm model when compared to the dilute condition, indicating that there was some degree of local unfolding (Figure \ref{fig:structural_integrity_chain}B).
+Interestingly, the flexibility profiles calculated considering the C$\upalpha$ for each chain of the elements of our cytoplasm model show that, in the crowded condition, most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are slightly more flexible than in the dilute condition (Figure \ref{fig:rmsf}).
+This is in agreement with simulation ~\cite{feig2011variable} and experimental data ~\cite{miklos2011protein,wang2012macromolecular} showing that crowders can destabilize the native structure of proteins due to the numerous unspecific short-lived interactions that are formed between them in a crowded environment.
+
 The oligomeric state of the crowders... \textbf{\color{red}{MISSING RESULTS: whole crowder structural integrity using buried oligomer area}}
 
-However, we observed aggregation of Mg, ATP and FBP...
+Magnesium cations were added to our cytoplasm model as counter-ions for tRNA and ATP.
+However, as the simulation progressed, FBP and ATP aggregated around Mg$^{2+}$ and, by consequence, around tRNA.
+After identifying this problem, we used simulations of small systems containing only the metabolites, Mg$^{2+}$ and water to test different strategies to prevent aggregation in future studies using long simulations of cytoplasm models.
+We found that protonating the phosphate-containing metabolites is enough to reach this goal (Figure \ref{fig:avoiding_aggregation}).
+Even though ATPH$_3$ and FBPH$_4$ are not the biologically relevant forms of these molecules in physiological pH, we consider that using their protonated forms is a good compromise between parametrization complexity and practical results.
+The biggest impact of this approach is on long range interactions of these metabolites with other molecules due to the neutralization of their net charge.
+However, molecules in a crowded environment are restrained to smaller volumes and shorter inter molecular distances, which mitigates the effect of charge neutralization.
+
+To the best of our knowledge, there is only one work in the literature which studied large scale cytoplasm models that included metabolites with all-atom explicit solvent molecular dynamics ~\cite{yu2016biomolecular}.
+Although they didn't report aggregation, they mentioned that the translational diffusion of highly charged phosphate-containing metabolites was "much slower" than expected.
+They attributed this observation to non-specific interactions between such metabolites and proteins.
+Maybe aggregation also contributed to their slower diffusion but it was not evident during \SI{140}{\nano\second}, which was the length of their longest simulation.
 
@@
 
+In principle it is possible to develop new parameters for phosphate and Mg that doesn't lead to aggregation. However this is outside of the scope of this study and
+
 \bibliography{library}

### 2019-09-24 paper/diff_plots.py
diff_plots updated

@@ -59,7 +59,7 @@ plt.plot([m / 1000. for m in mass_sorted], z[1] + [z[0]*m / 1000. for m in mass_
 plt.xlabel('$kDa$', fontsize=12)
 plt.ylabel('$D_{cyto}/D_{dil}$', fontsize=12)
 # plt.text(-5000,0.2,'C',fontsize=14)
-plt.ylim([0, .5])
+plt.ylim([0, .25])
 plt.savefig('diff_cyto_over_singles.pdf', bbox_inches="tight")
 plt.close()
 

### 2019-09-24 paper/virtual_cytoplasm_v2.tex
plots improved

@@
 B) The ratio between diffusion coefficient obtained from cytoplasm to single molecule simulations.
-The crowders are sorted according to their sizes on the x-axis.
-\textbf{\color{red}{THE Dcyto/Ddil PLOT IS "SQUEEZED" BECAUSE OF ITS STRANGE SCALE.
-IT WOULD BE BETTER TO CENTER THE DATAPOITS, E.G.
-USE THE Y-AXIS RANGE OF 0.0-0.2 OR SOMETHING SIMILAR.
-THIS SCALE WILL MAKE THE DIFFERENCE OF THE BEHAVIOR OF tRNA MORE VISIBLE AS WELL}}}
+The crowders are sorted according to their sizes on the x-axis.}
 \label{fig:translational_diffusion}
@@
 \includegraphics[scale=0.5]{rdf_RNA_metabolites.pdf}
-\caption{Radial distribution function showing the probability of finding metabolites or cations around RNA molecules in the cytoplasm model.\color{red}\textbf{THIS PLOT NEEDS REVISION.
-I DON'T SEE ERROR BARS (HOW WAS STATISTICS DONE FOR THE RDFs?).I SUBMITTED NEW RDF CALCULATIONS CENTERED AROUND tRNA BECAUSE I DIDN'T FIND THE XVG FILES THAT WERE USED TO GENERATE THIS PLOT}}
+\caption{Radial distribution function showing the probability of finding metabolites or cations around RNA molecules in the cytoplasm model.\color{red}\textbf{I SUBMITTED NEW RDF CALCULATIONS CENTERED AROUND tRNA BECAUSE I DIDN'T FIND THE XVG FILES THAT WERE USED TO GENERATE THIS PLOT}}
 \label{fig:tRNA_aggregation}
@@
 The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.
-The error bars show the standard errors.\textbf{\color{red}{THE MEASUREMENT NAME, SASA AND RMSD, IS MISSING IN THE Y-AXIS OF BOTH PLOTS}}}
+The error bars show the standard errors.}
 \label{fig:structural_integrity_chain}
@@
 \begin{figure}[H]
-\includegraphics[scale=0.5]{rmsf_2cols.pdf}
+\includegraphics[scale=0.5]{rmsf.pdf}
 \caption{Root mean squared fluctuations (rmsf) for 8 crowders from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).

### 2019-09-24 paper/barplots.py

@@ -4,7 +4,7 @@ import matplotlib
 import matplotlib.pyplot as plt
 import numpy as np
 
-
+fs = 14
 
 
 def parseArguments():
@@ -18,7 +18,7 @@ def parseArguments():
     return args
 
 
-crowder_name=["TufA", "MetE", "IcdA", "AhpC", "CspC", "Ppa ", "GapA", "Eno", "RNA$^{Phe}$"]
+crowder_name=["TufA", "MetE", "IcdA", "AhpC", "CspC", "Ppa ", "GapA", "Eno", "tRNA$^{Phe}$"]
 def autolabel(rects, xpos='center'):
     """
     Attach a text label above each bar in *rects*, displaying its height.
@@ -47,9 +47,11 @@ if __name__ == '__main__':
     ylabel = args.ylabel
     if len(ylabel) == 0:
         if args.prop == "sasa":
-            ylabel = "$nm^{2}$"
-        elif args.prop == "rmsd" or args.prop == "rg":
-            ylabel = "$nm$"
+            ylabel = "SASA ($nm^{2}$)"
+        elif args.prop == "rmsd": 
+            ylabel = "RMSD ($nm$)"
+        elif args.prop == "rg":
+            ylabel = "Rg ($nm$)"
         elif args.prop == "msd":
             ylabel = '$D (10^{-5} cm^2/s)$'
         elif args.prop == "oligomer" or args.prop == "hb":
@@ -83,19 +85,25 @@ if __name__ == '__main__':
                     label='Cyto.')
     
     # # Add some text for labels, title and custom x-axis tick labels, etc.
-    ax.set_ylabel(ylabel)
-    ax.set_xlabel(args.xlabel)
-    ax.set_title(args.title)
+    deg=30
+    ax.set_ylabel(ylabel, fontsize = fs)
+    ax.set_xlabel(args.xlabel, fontsize = fs)
+    # ax.set_title(args.title)
     ax.set_xticks(ind)
     if args.prop == "oligomer" or args.prop == "hb":
-        ax.set_xticklabels(oligomers)
+        ax.set_xticklabels(oligomers, fontsize = fs, rotation=deg,  ha='right')
     elif args.prop == "msd":
-        ax.set_xticklabels(crowder_name)
+        ax.set_xticklabels(crowder_name, fontsize = fs, rotation=deg)
+        plt.text(-2.25,0.135,'A', fontsize=fs+2)
     else:
[…]

### 2019-09-24 paper/diff_plots.py

@@ -14,7 +14,7 @@ order_s_b=[5, 9, 2, 1, 8, 3, 6, 7, 4] #smallest to biggest
  
 mass_crowder = [84320.000,  83764.400,  91489.800,  184409.000,  7271.120,  115920.600,  141602.400,  90884.000,  23761.100] #  146.124,  504.164,  336.080,  306.318]
 mass_sorted = sorted(mass_crowder)
-
+fs = 14
 
 avg_cyto, ste_cyto =[],[]
 for i in range(0, len(order_s_b)):
@@ -24,8 +24,8 @@ for i in range(0, len(order_s_b)):
 				avg_cyto.append(float(line.split()[1]))
 				ste_cyto.append(float(line.split()[2]))
 # plt.errorbar(atom_s_b, avg_cyto, ste_cyto, marker='o', markersize=10, linestyle=' ')
-# plt.xlabel('# atoms', fontsize=12)
-# plt.ylabel('$D_{cyto} (10^{-5} cm^2/s)$', fontsize=12)
+# plt.xlabel('# atoms', fontsize=fs)
+# plt.ylabel('$D_{cyto} (10^{-5} cm^2/s)$', fontsize=fs)
 # plt.text(-5000,0.018,'B',fontsize=14)
 # plt.savefig('diff_cyto.pdf', bbox_inches="tight")
 # plt.close()
@@ -39,8 +39,8 @@ for i in range(0, len(order_s_b)):
 				avg_sing.append(float(line.split()[1]))
 				ste_sing.append(float(line.split()[2]))
 # plt.errorbar(atom_s_b, avg_sing, ste_sing, marker='o', markersize=10, linestyle=' ')
-# plt.xlabel('# atoms', fontsize=12)
-# plt.ylabel('$D_{0} (10^{-5} cm^2/s)$', fontsize=12)
+# plt.xlabel('# atoms', fontsize=fs)
+# plt.ylabel('$D_{0} (10^{-5} cm^2/s)$', fontsize=fs)
 # plt.text(-5000,0.14, 'A',fontsize=14)
 # plt.savefig('diff_singles.pdf', bbox_inches="tight")
 # plt.close()
@@ -56,9 +56,10 @@ z = np.polyfit([m / 1000. for m in mass_sorted], AVG, 1)
 
[…]

### 2019-09-24 paper/virtual_cytoplasm_v2.tex
shrinking xvg added, changes in plots, captions red texts removed

@@
 The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.
-The error bars show the standard errors.\textbf{\color{red}{CAN YOU ADD THE PROTEIN NAME AS A LABEL INSIDE EACH PLOT?
-I THINK THE VARIABLE NAME SHOULD BE IN THE Y-AXIS (RMSF NM2)}}}
+The error bars show the standard errors.}
 \label{fig:rmsf}
@@
 \includegraphics[scale=0.5]{rdf_mg.pdf}
-\caption{Radial distribution function (rdf) showing the probability of finding a metabolite or an RNA around MG2+.
-The inset plot shows the rdf for short distance ($ r < 0.6 \,\, nm$) \color{red}{\textbf{HOW WAS STATISTICS DONE FOR THE RDFs?
-I DON'T SEE ERROR BARS}}}
+\caption{Radial distribution function (rdf) showing the probability of finding a metabolite or an RNA around MG2+. The inset plot shows the rdf for short distance ($ r < 0.6 \,\, nm$)}
 \label{fig:avoiding_aggregation}

### 2019-09-24 paper/barplots.py

@@ -4,7 +4,7 @@ import matplotlib
 import matplotlib.pyplot as plt
 import numpy as np
 
-fs = 14
+fs=16
 
 
 def parseArguments():

### 2019-09-24 paper/diff_plots.py

@@ -14,7 +14,7 @@ order_s_b=[5, 9, 2, 1, 8, 3, 6, 7, 4] #smallest to biggest
  
 mass_crowder = [84320.000,  83764.400,  91489.800,  184409.000,  7271.120,  115920.600,  141602.400,  90884.000,  23761.100] #  146.124,  504.164,  336.080,  306.318]
 mass_sorted = sorted(mass_crowder)
-fs = 14
+fs=16
 
 avg_cyto, ste_cyto =[],[]
 for i in range(0, len(order_s_b)):

### 2019-09-24 paper/shrinking_plot.py

@@ -1,0 +1,39 @@
+#!/usr/bin/env python3
+
+import argparse, os, glob, math
+import matplotlib.pyplot as plt
+import numpy as np
+
+glrt = 1.618
+width  = 5
+height = width/glrt
+fs=16
+y = [15000, 20000, 25000, 30000, 35000, 40000]
+ylbl =["15", "20", "25", "30", "35", "40"]
+
+for prop in ["volume", "density"]:
+	fig, ax = plt.subplots()
+	fig.set_figheight(height)
+	fig.set_figwidth(width)
+	for rep in range(1,4):
+		v = []
+		with open('shrinking_%s.%s.xvg' % (prop, str(rep)), 'r') as infile: 
+			for line in infile:
+				if line.find('#') < 0 and line.find("@") < 0:
+					data = line.split()
+					v.append(float(data[1]))
+		plt.plot(v, label="replica "+str(rep))
+	plt.xlabel('Time(ps)', fontsize=fs)
+	if prop == "volume":
+		plt.ylabel('%s ($10^3 nm^3$)' % prop, fontsize=fs)
+		plt.yticks(y, ylbl)
+		plt.text(-120,43100,'A', fontsize=fs)
+	else:
+		plt.ylabel('%s ($kg/m^3$)' % prop, fontsize=fs)
+		plt.text(-80,1150,'B', fontsize=fs)
+
+	plt.legend(fontsize=fs)
+	plt.tick_params(labelsize=fs-2);
+	plt.savefig('shrinking_%s.pdf' % prop, bbox_inches="tight")
+	plt.close()
+	# plt.plot(v, label=("replica %s" %(str(rep)) )

### 2019-09-24 paper/virtual_cytoplasm_v2.tex
references in tex

@@
 \subsubsection{RNA fraction}
-Transporter RNAs (tRNAs) account for 74\% of the dry weight of non-ribosomal RNAs \colorbox{red!50}{REF}.
+Transporter RNAs (tRNAs) account for 74\% of the dry weight of non-ribosomal RNAs~\cite{phillips2012}.
 Thus, we chose to model the RNA presence in the cytoplasm with tRNA molecules.
 Specifically, we considered the tRNA(Phe) molecule as a representative of tRNAs due to the availability of a recent crystallographic structure~\cite{Byrne2015}.
-The protein and RNA content of the total dry weight of \textit{E.coli} is 55\% and 2.9\%, respectively \colorbox{red!50}{REF}.
+The protein and RNA content of the total dry weight of \textit{E.coli} is 55\% and 2.9\%, respectively~\cite{phillips2012}.
 That is, the total RNA weight corresponds to 5\% of the total protein weight.
-This protein/RNA weight ratio was used to calculate the correct number of tRNA(Phe) molecules that were added to the cytoplasm model \colorbox{red!50}{(Table XXX)}.
+This protein/RNA weight ratio was used to calculate the correct number of tRNA(Phe) molecules that were added to the cytoplasm model \tabref{tbl:soup_components}.
 
@@
 In this work we use, for the first time, the AMBER force fields to study crowding effects on a cytoplasm model.
-All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of \textit{Mycoplasma genitalium} composed by 103 and 12 million atoms that were simulated for 20 and 60 \SI{}{\nano\second}, respectively ~\cite{yu2016biomolecular}.
+All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of \textit{Mycoplasma genitalium} composed by 103 and 12 million atoms that were simulated for 20 and 60 \SI{}{\nano\second}, respectively~\cite{yu2016biomolecular}.
 The cytoplasm model described in this work has 1.5 million atoms and was simulated for the total of \SI{3}{\micro\second}, an unprecedented time scale to study crowding effects on a heterogeneous cytoplasm model in atomic detail.
@@
 Interestingly, the flexibility profiles calculated considering the C$\upalpha$ for each chain of the elements of our cytoplasm model show that, in the crowded condition, most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are slightly more flexible than in the dilute condition (Figure \ref{fig:rmsf}).
-This is in agreement with simulation ~\cite{feig2011variable} and experimental data ~\cite{miklos2011protein,wang2012macromolecular} showing that crowders can destabilize the native structure of proteins due to the numerous unspecific short-lived interactions that are formed between them in a crowded environment.
+This is in agreement with simulation~\cite{feig2011variable} and experimental data~\cite{miklos2011protein,wang2012macromolecular} showing that crowders can destabilize the native structure of proteins due to the numerous unspecific short-lived interactions that are formed between them in a crowded environment.
 
@@
 
-To the best of our knowledge, there is only one work in the literature which studied large scale cytoplasm models that included metabolites with all-atom explicit solvent molecular dynamics ~\cite{yu2016biomolecular}.
+To the best of our knowledge, there is only one work in the literature which studied large scale cytoplasm models that included metabolites with all-atom explicit solvent molecular dynamics~\cite{Yu2016a}.
 Although they didn't report aggregation, they mentioned that the translational diffusion of highly charged phosphate-containing metabolites was "much slower" than expected.

### 2019-09-24 paper/shrinking_plot.py

@@ -11,7 +11,8 @@ fs=16
 y = [15000, 20000, 25000, 30000, 35000, 40000]
 ylbl =["15", "20", "25", "30", "35", "40"]
 
-for prop in ["volume", "density"]:
+# for prop in ["volume", "density"]:
+for prop in ["volume"]:
 	fig, ax = plt.subplots()
 	fig.set_figheight(height)
 	fig.set_figwidth(width)

### 2019-09-25 paper/virtual_cytoplasm_v2.tex
I'm working on the introduction

--- 2019-09-24 paper/virtual_cytoplasm_v2.tex
@@
 
-Biomolecules move and function in an environment densely packed with high concentrations of macromolecules.
+%Biomolecules move and function in an environment densely packed with high concentrations of macromolecules.
 The presence of macromolecules leads to steric effect due to excluded volume effect and intermolecular attrative/repulsive forces due to distributed charges on the surface of macromolecules.
 
-Structure and dynamics of biomolecules are well characterized {\em in vitro}, however these studies {\em in vivo} are still evolving. 
-Investigating biomolecular properties {\em in vivo} is possible through developments in the fields of nuclear magnetic resonance~\cite{reckel2007,pielak2008}, or fluoroscence spectroscopies~\cite{ignatova2004,xie2008,English2011}.
+Crowding ... ~\cite{feig2017crowding} ~\citep{ostrowska2019modeling}
 
-An alternative method is to use computational models and simulation techniques.
-Biomolecular simulations are often carried out under dilute conditions or simple models of macromolecular crowdings~\cite{Spiga2014a,Harada2012a,Nawrocki2017a}.
-However, more attemps in modeling bacterial cytoplasm have been made recently~\cite{Mcguffee2010,Yu2016a}.
+-> Crowding affects hydration around proteins ~\cite{Harada2012a,king2013crowding}
+-> leads to stabilization of native structure due to the excluded volume effect ~\cite{cheung2005molecular}
+-> can also destabilize proteins due to  ~\cite{feig2011variable,miklos2011protein,wang2012macromolecular}
 
-\begin{itemize}
-   \item How a computational model can be useful
-   \item From a simple extrapolation in time, it has been estimated that the simulation of an {\em E.
-coli} bacterium would be possible in 25 years from now for one nanosecond with $10^{11}$ atoms~\cite{vanGunsteren2006a}
-\end{itemize}
+While the structure and dynamics of biomolecules is well characterized \textit{in vitro}, the understanding of the influences of crowded environments for the \textit{in vivo} are still evolving.
+Techniques such as nuclear magnetic resonance~\cite{reckel2007,pielak2008} and fluoroscence spectroscopy~\cite{ignatova2004,xie2008,English2011} are among the most relevant to the field due to their ability to probe cells in a non destructive way.
+Another essential class of methods come from computational modeling and simulation.
+By extrapolating the advances in computational power in the last decades to the future, it has been estimated that the atomistic simulation of an entire \textit{Escherichia coli} bacterium for \SI{1}{\nano\second} will be possible in 15 years from now ~\cite{vanGunsteren2006a}.
 
-Here we report on a model of {\em Escherichia coli} cytoplasm at atomistic level.
-The challenges in modeling and simulation with Molecular Dynamics are pointed and discussed and solutions are provided.
+As the available computational power increases, models become increasingly complex and are better representations of reality.
+The most simple class of computational models apply brownian dynamics calculations to probe macromolecules imersed in a solution of crowders that are treated as hard spheres with varying radius to represent different types of molecules ~\cite{ando2010crowding}.
+These models usually use an implicit representation of the solvent ~\cite{mcguffee2010diffusion}.More recently, all-atom explicit solvent molecular dynamics simulations have been employed on crowded systems composed by several copies of one or a few macromolecules ~\cite{wang2017influence}.
+However, these models doesn't reflect the true heterogeneity of biologically relevant crowded environments such as the cytoplasm.
+This was addressed by the development of a detailed model for the cytoplasm of the bacterium \textit{Mycoplasma genitalium} containing 103 million atoms ~\cite{feig2015complete}.
+This model, as well as a smaller version with 13 million atoms, was submitted to molecular dynamics simulations ~\cite{yu2016biomolecular}.
+Despite the size of these systems, the simulation times were too short, \SI{20}{\nano\second} for the complete model and \SI{140}{\nano\second} for the smaller version, to accurately characterize the effects of the slow diffusion of its constituents.
 
-This work is the first model of {\em E. coli} at atomistic resolution that spans cellular dynamics on a microsecond scale.
+In this work we report an all-atom model of the cytoplasm of \textit{Escherichia coli} that was built to reflect the real biological system.
+We discuss the difficulties that arise when building such system and provide a set of Python scripts and files that anyone can use to build its own crowded systems with custom parameters and elements.
+Additionally, we performed long molecular dynamics simulations of our cytoplasm model, reaching \SI{3}{\micro\second} of total simulation time.
+This work is the first to explore the structural dynamics of an all-atom explicit solvent cytoplasm model with microsecond scale molecular dynamics simulations.
 
@@
 \includegraphics[scale=0.5]{rdf_RNA_metabolites.pdf}
-\caption{Radial distribution function showing the probability of finding metabolites or cations around RNA molecules in the cytoplasm model.\color{red}\textbf{I SUBMITTED NEW RDF CALCULATIONS CENTERED AROUND tRNA BECAUSE I DIDN'T FIND THE XVG FILES THAT WERE USED TO GENERATE THIS PLOT}}
+\caption{Radial distribution function showing the probability of finding metabolites or cations around RNA molecules in the cytoplasm model.\color{red}\textbf{THIS PLOT NEEDS REVISION.
+I DON'T SEE ERROR BARS (HOW WAS STATISTICS DONE FOR THE RDFs?).I SUBMITTED NEW RDF CALCULATIONS CENTERED AROUND tRNA BECAUSE I DIDN'T FIND THE XVG FILES THAT WERE USED TO GENERATE THIS PLOT}}
 \label{fig:tRNA_aggregation}
@@
 The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.
-The error bars show the standard errors.}
+The error bars show the standard errors.\textbf{\color{red}{THE MEASUREMENT NAME, SASA AND RMSD, IS MISSING IN THE Y-AXIS OF BOTH PLOTS}}}
 \label{fig:structural_integrity_chain}
@@
 In this work we use, for the first time, the AMBER force fields to study crowding effects on a cytoplasm model.
-All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of \textit{Mycoplasma genitalium} composed by 103 and 12 million atoms that were simulated for 20 and 60 \SI{}{\nano\second}, respectively ~\cite{yu2016biomolecular}.
+All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of \textit{Mycoplasma genitalium} composed by 103 and 12 million atoms that were simulated for 20 and 140 \SI{}{\nano\second}, respectively ~\cite{yu2016biomolecular}.
 The cytoplasm model described in this work has 1.5 million atoms and was simulated for the total of \SI{3}{\micro\second}, an unprecedented time scale to study crowding effects on a heterogeneous cytoplasm model in atomic detail.

### 2019-09-25 paper/virtual_cytoplasm_v2.tex

@@
 In this work we use, for the first time, the AMBER force fields to study crowding effects on a cytoplasm model.
-All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of \textit{Mycoplasma genitalium} composed by 103 and 12 million atoms that were simulated for 20 and 140 \SI{}{\nano\second}, respectively ~\cite{yu2016biomolecular}.
+All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of \textit{Mycoplasma genitalium} composed by 103 and 12 million atoms that were simulated for 20 and 60 \SI{}{\nano\second}, respectively~\cite{yu2016biomolecular}.
 The cytoplasm model described in this work has 1.5 million atoms and was simulated for the total of \SI{3}{\micro\second}, an unprecedented time scale to study crowding effects on a heterogeneous cytoplasm model in atomic detail.

### 2019-09-25 paper/virtual_cytoplasm_v2.tex
updates in sasa plot, rdf_RNA plot

@@
 \usepackage{upgreek}
+\usepackage{soul}
 \setkeys{acs}{super = true}
@@
 \includegraphics[scale=0.5]{rdf_RNA_metabolites.pdf}
-\caption{Radial distribution function showing the probability of finding metabolites or cations around RNA molecules in the cytoplasm model.\color{red}\textbf{I SUBMITTED NEW RDF CALCULATIONS CENTERED AROUND tRNA BECAUSE I DIDN'T FIND THE XVG FILES THAT WERE USED TO GENERATE THIS PLOT}}
+\caption{Radial distribution function showing the probability of finding metabolites or cations around RNA molecules in the cytoplasm model.}
 \label{fig:tRNA_aggregation}
@@
 \caption{A) Solvent accessible surface area (sasa) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
-The average sasa for each protein in single molecule simulation is over number of chains of proteins and three replica.
-The average sasa for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.
+\st{The average sasa for each protein in single molecule simulation is over number of chains of proteins and three replica.
+The average sasa for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.} The average bar value for single molecule simulations shows the sum over sasa of each protein chains averaged over the number of replica minus the sasa of the whole proein averaged over the number of replica.
+The average bar value for cytoplasm simulation shows the sum over sasa of each protein chains averaged over the number of replica and the abundancy of the protein minus the sasa of the whole proein averaged over the number of replica and the abundancy of the protein.
 The error bars show the standard errors.
@@
 \includegraphics[scale=0.5]{rdf_mg.pdf}
-\caption{Radial distribution function (rdf) showing the probability of finding a metabolite or an RNA around MG2+. The inset plot shows the rdf for short distance ($ r < 0.6 \,\, nm$)}
+\caption{Radial distribution function (rdf) showing the probability of finding a metabolite or an RNA around MG2+. The inset plot shows the rdf for short distance ($ r < 0.7 \,\, nm$)}
 \label{fig:avoiding_aggregation}

### 2019-09-25 paper/barplots.py

@@ -98,7 +98,7 @@ if __name__ == '__main__':
     else:
         ax.set_xticklabels(protein_name, fontsize = fs, rotation=deg)
         if args.prop == "sasa":
-            plt.text(-2., 360,'A', fontsize=fs+2)
+            plt.text(-2., 160,'A', fontsize=fs+2)
         elif args.prop == "rmsd":
             plt.text(-2., 0.48,'B', fontsize=fs+2)
 

### 2019-09-26 paper/virtual_cytoplasm_v2.tex
tex revised

@@
 
-\caption{A) Solvent accessible surface area (sasa) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
+\caption{A) Interface area for oligomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
 \st{The average sasa for each protein in single molecule simulation is over number of chains of proteins and three replica.
-The average sasa for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.} The average bar value for single molecule simulations shows the sum over sasa of each protein chains averaged over the number of replica minus the sasa of the whole proein averaged over the number of replica.
-The average bar value for cytoplasm simulation shows the sum over sasa of each protein chains averaged over the number of replica and the abundancy of the protein minus the sasa of the whole proein averaged over the number of replica and the abundancy of the protein.
+The average sasa for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.} The average bar value for single molecule simulations shows the sum over sasa of each protein chains averaged over the number of replica minus the sasa of the whole protein averaged over the number of replica.
+The average bar value for cytoplasm simulation shows the sum over sasa of each protein chains averaged over the number of replica and the abundancy of the protein minus the sasa of the whole protein averaged over the number of replica and the abundancy of the protein.
 The error bars show the standard errors.

### 2019-09-26 paper/barplots.py

@@ -47,7 +47,7 @@ if __name__ == '__main__':
     ylabel = args.ylabel
     if len(ylabel) == 0:
         if args.prop == "sasa":
-            ylabel = "SASA ($nm^{2}$)"
+            ylabel = "Interface Area ($nm^{2}$)"
         elif args.prop == "rmsd": 
             ylabel = "RMSD ($nm$)"
         elif args.prop == "rg":
@@ -98,6 +98,7 @@ if __name__ == '__main__':
     else:
         ax.set_xticklabels(protein_name, fontsize = fs, rotation=deg)
         if args.prop == "sasa":
+            ax.set_xticklabels(oligomers, fontsize = fs, rotation=deg)
             plt.text(-2., 160,'A', fontsize=fs+2)
         elif args.prop == "rmsd":
             plt.text(-2., 0.48,'B', fontsize=fs+2)

### 2019-09-27 paper/virtual_cytoplasm_v2.tex
saving tex changes

@@
 While the structure and dynamics of biomolecules is well characterized \textit{in vitro}, the understanding of the influences of crowded environments for the \textit{in vivo} are still evolving.
-Techniques such as nuclear magnetic resonance~\cite{reckel2007,pielak2008} and fluoroscence spectroscopy~\cite{ignatova2004,xie2008,English2011} are among the most relevant to the field due to their ability to probe cells in a non destructive way.
+Techniques such as nuclear magnetic resonance~\cite{reckel2007,pielak2008} and fluorescence spectroscopy~\cite{ignatova2004,xie2008,English2011} are among the most relevant to the field due to their ability to probe cells in a non destructive way.
 Another essential class of methods come from computational modeling and simulation.
@@
 As the available computational power increases, models become increasingly complex and are better representations of reality.
-The most simple class of computational models apply brownian dynamics calculations to probe macromolecules imersed in a solution of crowders that are treated as hard spheres with varying radius to represent different types of molecules ~\cite{ando2010crowding}.
-These models usually use an implicit representation of the solvent ~\cite{mcguffee2010diffusion}.More recently, all-atom explicit solvent molecular dynamics simulations have been employed on crowded systems composed by several copies of one or a few macromolecules ~\cite{wang2017influence}.
+The most simple class of computational models apply Brownian Dynamics calculations to probe macromolecules immersed in a solution of crowders that are treated as hard spheres with varying radius to represent different types of molecules ~\cite{ando2010crowding}.
+These models usually use an implicit representation of the solvent ~\cite{mcguffee2010diffusion}.
+More recently, all-atom explicit solvent molecular dynamics simulations have been employed on crowded systems composed by several copies of one or a few macromolecules ~\cite{wang2017influence}.
 However, these models doesn't reflect the true heterogeneity of biologically relevant crowded environments such as the cytoplasm.
@@
 We looked for the proteins structures that were either from or expressed in {\em E-coli}.
-In case of 1U22 (MetE) and 2EIP (Ppa) we used a loop-closure modelng tool based on Random Coordinate Descent (RCD) method~\cite{Chys2013} to correct the information for missing residues.
+In case of 1U22 (MetE) and 2EIP (Ppa) we used a loop-closure modeling tool based on Random Coordinate Descent (RCD) method~\cite{Chys2013} to correct the information for missing residues.
 The four metabolites were parametrized using GAFF and Antechamber.
@@
 coli}, respectively, while the other six proteins contribute with less than 5\% each (Table~\ref{tbl:protein_fraction}).
-Their crystallogaphic structures were used to insert them in the cytoplasm model.
+Their crystallographic structures were used to insert them in the cytoplasm model.
 The number of copies of each protein in the model was calculated as the rounded up fold increase in their abundance when compared to the least abundant protein, which was set to have the copy number of 1.
 Then, this number was divided by the oligomeric state of each protein.
-This was done because the original oligomeric state for each protein was kept as reported in their crystallogaphic structure.
+This was done because the original oligomeric state for each protein was kept as reported in their crystallographic structure.
 
@@
 Finally, Mg\textsuperscript{2+} was used as counter-ions for tRNA and ATP.
-K$^{+}$ and Cl$^{-}$ were added to neutralize the charges of the simulation box and to reach the ionic strength of 0.150 mol/L by substution of randomly selected water molecules.
+K$^{+}$ and Cl$^{-}$ were added to neutralize the charges of the simulation box and to reach the ionic strength of 0.150 mol/L by substitution of randomly selected water molecules.
 
@@
 
-In order to do this, the droplets around each component are also constructed iteractively.
+In order to do this, the droplets around each component are also constructed iteratively.
 The number of water molecules that we must place in the simulation box is known from the desired biomolecular fraction, which is a parameter of the model, and the total biomolecular mass of the system.
 However, we don't know the thickness of the water layer around each component, $l$, that accounts for such amount of water molecules.
-Since $l$ depends on a series of factors such as the shape, size and abundancy of each element, we define it iteratively.
+Since $l$ depends on a series of factors such as the shape, size and abundance of each element, we define it iteratively.
 We start by taking a droplet of thickness $l$ = 3 \r{A} around all elements and counting the number of water molecules that we would add to the box with such droplets.
 If it is smaller than the number we need, $l$ is increased by $dl$ = 1 \r{A}.
-If it is larger, we reduce the droplet thickness by $dl$, reduce $dl$ by 10 times and then increate $l$ with the new smaller $dl$.
+If it is larger, we reduce the droplet thickness by $dl$, reduce $dl$ by 10 times and then increase $l$ with the new smaller $dl$.
 We can carry this process until an arbitrary precision cutoff, such as 5\%, is satisfied.
@@
 \includegraphics[scale=0.5]{shrinking_volume.pdf} 
-\includegraphics[scale=0.11]{box_big.png}
-\includegraphics[scale=0.11]{box_small_.png}
+\includegraphics[scale=0.5]{box.pdf} 
+% \includegraphics[scale=0.11]{box_big.png}
+% \includegraphics[scale=0.11]{box_small_.png}
 %\includegraphics[scale=0.5]{shrinking_density.pdf}
@@
 
-\caption{A) Interface area for oligomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
-\st{The average sasa for each protein in single molecule simulation is over number of chains of proteins and three replica.
-The average sasa for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.} The average bar value for single molecule simulations shows the sum over sasa of each protein chains averaged over the number of replica minus the sasa of the whole protein averaged over the number of replica.
-The average bar value for cytoplasm simulation shows the sum over sasa of each protein chains averaged over the number of replica and the abundancy of the protein minus the sasa of the whole protein averaged over the number of replica and the abundancy of the protein.
+\caption{A) Chain interface area for oligomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
+The average bar value for single molecule simulations shows the sum over sasa of each protein chains averaged over the number of replica minus the sasa of the whole protein averaged over the number of replica.
+The average bar value for cytoplasm simulation shows the sum over sasa of each protein chains averaged over the number of replica and the abundance of the protein minus the sasa of the whole protein averaged over the number of replica and the abundance of the protein.
 The error bars show the standard errors.
@@
 \subsubsection{Structural dynamics}
-Proteins are slighly more flexible in the crowded environment.
-Root mean square fluction (RMSF) calculations show that the flexibility profile of the proteins is not affected by crowding, except in the cases in which there are conformational changes, TufA ...
+Proteins are slightly more flexible in the crowded environment.
+Root mean square fluctuation (RMSF) calculations show that the flexibility profile of the proteins is not affected by crowding, except in the cases in which there are conformational changes, TufA ...
 as shown before (Figure~\ref{fig:rmsf}).
 This is in agreement with our results showing that the structural integrity of the proteins is not significantly affected.
-In addition to that, our results show that the proteins are slighly more flexible in the crowded environment than in the dilute condition (Figure~\ref{fig:rmsf}).
-This is likely due to the nonspecific interations and collisions that often happen under crowding conditions \textbf{\color{red}{REF}}.
+In addition to that, our results show that the proteins are slightly more flexible in the crowded environment than in the dilute condition (Figure~\ref{fig:rmsf}).
+This is likely due to the nonspecific interactions and collisions that often happen under crowding conditions \textbf{\color{red}{REF}}.
 
@@
 
-The initial detection of aggregation in our cytoplasm model was due to the outlier behavior of the translational diffusion constant of tRNA when compared to the other crowders.
+The initial detection of aggregation in our cytoplasm model was due to the out-lier behavior of the translational diffusion constant of tRNA when compared to the other crowders.
 However, tRNA is not causing such phenomenon.
@@
 tRNA gets involved in the aggregates because Mg$^{2+}$ is present in the box only as counter-ion for ATP and tRNA, and so tRNA is one of the few "sources" of Mg$^{2+}$ in the whole system.
-We looked for ways of avoind aggregation with simulations of small systems composed by these metabolites in high concentration, Mg$^{2+}$ and water (please check the methods section for details about the parameters we used).
+We looked for ways of avoid aggregation with simulations of small systems composed by these metabolites in high concentration, Mg$^{2+}$ and water (please check the methods section for details about the parameters we used).
 
 We found that completely protonating the phosphate groups of ATP and FBP is enough prevent their aggregation with Mg$^{2+}$ (Figure~\ref{fig:avoiding_aggregation}).
-While RDFs show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$, this is not observed for ATPH\textsubscript{3} and FBPH\textsubscript{4}.
+While RDFs show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$, this is not observed for ATPH\textsubscript{3} and FBPH\textsubscript{4}(Figure~\ref{fig:protonation}).
 
@@
 \label{fig:avoiding_aggregation}
+\end{figure}
+
+\begin{figure}[H]
+\includegraphics[scale=0.5]{rdf_mg_metabolites.pdf}
+\caption{A) Radial distribution function (rdf) showing the probability of finding an ATP$^{-3}$ or an ATPH\textsubscript{3} around Mg$^{2+}$.
+B) Radial distribution function (rdf) showing the probability of finding an FBP$^{-4}$ or an FBPH\textsubscript{4} around Mg$^{2+}$.
+}
+\label{fig:protonation}
 \end{figure}
@@
 We found that the aggregation of ATP$^{-3}$ and FBP$^{-4}$ around Mg$^{2+}$, which was used as counter-ion for tRNA, leads to its exaggerated confinement.
-Thus, tRNA is considered as an outlier in our cytoplasm model.
+Thus, tRNA is considered as an out-lier in our cytoplasm model.
 In fact, the slope of the linear regression line fitted to $D_{cyto}/D_{dil}$ is reduced by by 50\% and gets even closer to zero (slope $= 0.00014$) after removing the diffusion coefficient of tRNA from the data.
 Our data is in agreement with experimental results showing that the translational diffusion coefficient of Green Fluorescent Protein (GFP) in {\em E.
-coli}, as measured by fluorescence recovery after photobleaching (FRAP), decreased 90\% when compared to {\em in vitro} measurements~\cite{Elowitz1999,Konopka2006}.
+coli}, as measured by fluorescence recovery after photo-bleaching (FRAP), decreased 90\% when compared to {\em in vitro} measurements~\cite{Elowitz1999,Konopka2006}.
 

### 2019-09-27 paper/barplots.py
figure 1

@@ -47,7 +47,7 @@ if __name__ == '__main__':
     ylabel = args.ylabel
     if len(ylabel) == 0:
         if args.prop == "sasa":
-            ylabel = "Interface Area ($nm^{2}$)"
+            ylabel = "Chain Interface Area ($nm^{2}$)"
         elif args.prop == "rmsd": 
             ylabel = "RMSD ($nm$)"
         elif args.prop == "rg":

### 2019-09-27 paper/shrinking_plot.py

@@ -28,13 +28,50 @@ for prop in ["volume"]:
 	if prop == "volume":
 		plt.ylabel('%s ($10^3 nm^3$)' % prop, fontsize=fs)
 		plt.yticks(y, ylbl)
-		plt.text(-120,43100,'A', fontsize=fs)
+		plt.text(-120,43100,'B', fontsize=fs)
 	else:
 		plt.ylabel('%s ($kg/m^3$)' % prop, fontsize=fs)
-		plt.text(-80,1150,'B', fontsize=fs)
+		# plt.text(-80,1150,'B', fontsize=fs)
 
 	plt.legend(fontsize=fs)
 	plt.tick_params(labelsize=fs-2);
 	plt.savefig('shrinking_%s.pdf' % prop, bbox_inches="tight")
 	plt.close()
 	# plt.plot(v, label=("replica %s" %(str(rep)) )
+
+
+import matplotlib.image as mpimg
+fig = plt.figure()
+# print(fig.get_size_inches())
+pic = ['box_big.png', 'box_small_.png']
+x1=[0, 0]
+x2=[0, 1000]
+x3=[1000, 1000]
+x4=[1000, 0]
+image = mpimg.imread(pic[0])
+ax1 = fig.add_subplot(121)
+plt.imshow(image)
+plt.axis('off')
+plt.text(-100,-5, 'A', fontsize = fs)
+plt.plot(x1,x2, 'k-')
+plt.plot(x2,x3, 'k-')
+plt.plot(x4,x1, 'k-')
+plt.axis('off')
+
+image = mpimg.imread(pic[1])
+ax2 = fig.add_subplot(122)
+plt.imshow(image)
+# plt.title('After')
+plt.axis('off')
+# plt.plot(x1,x2, 'k-')
+plt.plot(x2,x3, 'k-')
+plt.plot(x3,x4, 'k-')
+plt.plot(x4,x1, 'k-')
+plt.axis('off')
+fig.subplots_adjust(wspace=-.1, hspace=0)
+# plt.tight_layout()
+plt.savefig('box.pdf',bbox_inches="tight")
+plt.close()
+
+
+# plt.show()

### 2019-09-28 paper/virtual_cytoplasm_v2.tex
figure 1
Introduction

@@
 
-Crowding ... ~\cite{feig2017crowding} ~\citep{ostrowska2019modeling}
-
--> Crowding affects hydration around proteins ~\cite{Harada2012a,king2013crowding}
--> leads to stabilization of native structure due to the excluded volume effect ~\cite{cheung2005molecular}
--> can also destabilize proteins due to  ~\cite{feig2011variable,miklos2011protein,wang2012macromolecular}
+The intracellular microenvironment is markedly different from the\textit{in vitro} simplification from which most of our current knowledge about the biochemical and biophysical behavior of macromolecules was derived~\cite{feig2017crowding}.
+Instead of simple and diluted, the \textit{in vivo} microenvironment is highly heterogeneous and concentrated, leading to spatial constrains that directly affects biological activities~\citep{ostrowska2019modeling}.
+Macromolecular crowding directly influences the stability of macromolecules.
+There is an equilibrium between stabilization of their native structure due to the entropic component given by the excluded volume effect~\cite{cheung2005molecular} and destabilizationg due to enthalpic component given by unspecific interactions with the crowders~\cite{feig2011variable,miklos2011protein,wang2012macromolecular}.
+Additionally, crowding affects the structure and dynamics of water around proteins, which has an indirect but important effect in their biological activities~\cite{Harada2012a,king2013crowding}.
 
@@
 \begin{figure}[H]
+\includegraphics[scale=0.5]{box.pdf} 
 \includegraphics[scale=0.5]{shrinking_volume.pdf} 
-\includegraphics[scale=0.5]{box.pdf} 
 % \includegraphics[scale=0.11]{box_big.png}

### 2019-09-29 paper/virtual_cytoplasm_v2.tex
tex update

--- 2019-09-27 paper/virtual_cytoplasm_v2.tex
@@
 % \linespread{1.5}
-% \usepackage{amssymb,amsmath,graphics,epsfig}
-\usepackage{amsmath}
+\usepackage{amssymb,amsmath,graphics,epsfig}
+% \usepackage{amsmath}
 \usepackage[font=small,labelfont=bf]{caption}
@@
 
-Crowding ... ~\cite{feig2017crowding} ~\citep{ostrowska2019modeling}
-
--> Crowding affects hydration around proteins ~\cite{Harada2012a,king2013crowding}
--> leads to stabilization of native structure due to the excluded volume effect ~\cite{cheung2005molecular}
--> can also destabilize proteins due to  ~\cite{feig2011variable,miklos2011protein,wang2012macromolecular}
+The intracellular microenvironment is markedly different from the \textit{in vitro} simplification from which most of our current knowledge about the biochemical and biophysical behavior of macromolecules was derived~\cite{Feig2017a}.
+Instead of simple and diluted, the \textit{in vivo} microenvironment is highly heterogeneous and concentrated, leading to spatial constrains that directly affects biological activities~\citep{ostrowska2019}.
+Macromolecular crowding directly influences the stability of macromolecules.
+There is an equilibrium between stabilization of their native structure due to the entropic component given by the excluded volume effect~\cite{cheung2005} and destabilizationg due to enthalpic component given by unspecific interactions with the crowders~\cite{Feig2011,miklos2011,Wang2012b}.
+Additionally, crowding affects the structure and dynamics of water around proteins, which has an indirect but important effect in their biological activities~\cite{Harada2012a,king2013}.
 
@@
 Another essential class of methods come from computational modeling and simulation.
-By extrapolating the advances in computational power in the last decades to the future, it has been estimated that the atomistic simulation of an entire \textit{Escherichia coli} bacterium for \SI{1}{\nano\second} will be possible in 15 years from now ~\cite{vanGunsteren2006a}.
+By extrapolating the advances in computational power in the last decades to the future, it has been estimated that the atomistic simulation of an entire \textit{Escherichia coli} bacterium for \SI{1}{\nano\second} will be possible in 15 years from now~\cite{vanGunsteren2006a}.
 
 As the available computational power increases, models become increasingly complex and are better representations of reality.
-The most simple class of computational models apply Brownian Dynamics calculations to probe macromolecules immersed in a solution of crowders that are treated as hard spheres with varying radius to represent different types of molecules ~\cite{ando2010crowding}.
-These models usually use an implicit representation of the solvent ~\cite{mcguffee2010diffusion}.
-More recently, all-atom explicit solvent molecular dynamics simulations have been employed on crowded systems composed by several copies of one or a few macromolecules ~\cite{wang2017influence}.
+The most simple class of computational models apply Brownian Dynamics calculations to probe macromolecules immersed in a solution of crowders that are treated as hard spheres with varying radius to represent different types of molecules~\cite{Ando2010}.
+These models usually use an implicit representation of the solvent~\cite{Mcguffee2010}.
+More recently, all-atom explicit solvent molecular dynamics simulations have been employed on crowded systems composed of several copies of one or a few macromolecules~\cite{Wang2017c}.
 However, these models doesn't reflect the true heterogeneity of biologically relevant crowded environments such as the cytoplasm.
-This was addressed by the development of a detailed model for the cytoplasm of the bacterium \textit{Mycoplasma genitalium} containing 103 million atoms ~\cite{feig2015complete}.
-This model, as well as a smaller version with 13 million atoms, was submitted to molecular dynamics simulations ~\cite{yu2016biomolecular}.
+This was addressed by the development of a detailed model for the cytoplasm of the bacterium \textit{Mycoplasma genitalium} containing 103 million atoms~\cite{Feig2015}.
+This model, as well as a smaller version with 13 million atoms, was submitted to molecular dynamics simulations~\cite{Yu2016a}.
 Despite the size of these systems, the simulation times were too short, \SI{20}{\nano\second} for the complete model and \SI{140}{\nano\second} for the smaller version, to accurately characterize the effects of the slow diffusion of its constituents.
@@
 \centering
-\begin{tabular}{ l c c }
+\begin{tabular}{ l c c c}
 \hline
-\multirow{2}{*}{Protein} & 	\multicolumn{2}{c}{Fraction [\%]} \\
+\multirow{2}{*}{Protein} & 	\multicolumn{2}{c}{Fraction [\%]} & \multirow{2}{*}{Oligomeric State}\\
 & Absolute & Cumulative \\ 
 \hline
-TufA & 19.7 & 19.7 \\
-MetE & 11.6 & 31.4 \\
-IcdA & 4.7  & 36.1 \\
-AhpC & 4.1  & 40.2 \\
-CspC & 4.0  & 44.2 \\
-Ppa  & 2.9  & 47.0 \\
-GapA & 2.1  & 49.1 \\
-Eno  & 1.9  & 51.1 \\
+TufA & 19.7 & 19.7 & 2-mer \\
+MetE & 11.6 & 31.4 & 1-mer \\
+IcdA & 4.7  & 36.1 & 2-mer \\
+AhpC & 4.1  & 40.2 & 10-mer \\
+CspC & 4.0  & 44.2 & 1-mer \\
+Ppa  & 2.9  & 47.0 & 6-mer \\
+GapA & 2.1  & 49.1 & 4-mer \\
+Eno  & 1.9  & 51.1 & 2-mer \\
 \hline
@@
 The number of water molecules was calculated according to the desired biomolecular concentration, which is a parameter of the cytoplasm model.
-This number ranges from \colorbox{red!50}{XX to YY} in biological systems such as \textit{E.
-coli} cytoplasm.
+This number ranges from 300 to 400 $mg/ml$ in biological systems such as \textit{E.
+coli} cytoplasm~\cite{Zimmerman1991,Ellis2003a}.
 In our case, we choose the biomolecular concentration of 30\%, that is, the number of molecules necessary to reach a ratio of total biomolecular mass to water mass of 30\% was inserted into the cytoplasm model.
@@
 In our case, all insertions succeeded after increasing $L$ to 35 nm.
-Additionally, instead of adding only the protein, RNA or metabolite molecules in the empty box in each trial, we actually add a droplet of water and counter ions in which the molecule of interest is embedded.
+Additionally, instead of adding only the protein, tRNA or metabolite molecules in the empty box in each trial, we actually add a droplet of water and counter ions in which the molecule of interest is embedded.
 Such droplets are taken from molecular dynamics simulations in which each component was previously equilibrated.
@@
 \begin{figure}[H]
-\includegraphics[scale=0.5]{shrinking_volume.pdf} 
-\includegraphics[scale=0.5]{box.pdf} 
-% \includegraphics[scale=0.11]{box_big.png}
-% \includegraphics[scale=0.11]{box_small_.png}
-%\includegraphics[scale=0.5]{shrinking_density.pdf}
-
+\includegraphics[scale=0.5]{shrinking.pdf} 
 \caption{A) The volume of the box with length size of $L=35 \,\, nm$ reduces during the energy minimization in the first $300 \,\, ps$ B) The density of the same box increases within the first $300\,\, ps$.
@@
 coli} cytoplasm model on the structural integrity and dynamics of its elements, we constructed three completely independent cytoplasm models that have different orientations for each of its elements.
-We submitted these systems, each composed by more than 1.5 million atoms, to molecular dynamics simulations of \SI{1}{\micro\second}.
-We also performed \SI{200}{\nano\second} molecular dynamics simulations for each isolated element in order to represent the non-crowded, i.e.
-dilute, environment (please check the materials section for details about the parameters).
+We submitted these systems, each composed of more than 1.5 million atoms, to molecular dynamics simulations of \SI{1}{\micro\second}.
+We also performed \SI{200}{\nano\second} molecular dynamics simulations for each isolated element (proteins and tRNA) in order to represent the non-crowded, i.e.
+dilute environment (please check the materials section for details about the parameters).
 
@@
 Under crowded conditions, molecules are constrained to a smaller volume due to collisions with the other components of the crowded environment.
-The extent of this effect is evident when we compare the translational diffusion constant of each component in the cytoplasm model, $D_{cyto}$, and in a dilute system, $D_{dil}$ (Figure~\ref{fig:translational_diffusion}A).
-The ratio of these quantities is independent of the protein size and is always close to 0.13 (Figure~\ref{fig:translational_diffusion}B).
-However, the effect of crowding over the translation diffusion of tRNA is two times higher.
-Its $D_{cyto}/D_{dil}$ is 0.07, indicating that its movement is specially constrained in the crowded environment (Figure~\ref{fig:translational_diffusion}B, crowder with 25 kDa).
+The extent of this effect is evident when we compare the translational diffusion constant of each component in the cytoplasm model, $D_{cyto}$, and in a dilute system, $D_{dil}$ (Fig.~\ref{fig:translational_diffusion}A).
+The ratio of these quantities is independent of the protein size and is always close to 0.13 (Fig.~\ref{fig:translational_diffusion}B).
+However, the effect of crowding on the translational diffusion of tRNA is two times larger.
+Its $D_{cyto}/D_{dil}$ is 0.07, indicating that its movement is specially constrained in the crowded environment (Fig.~\ref{fig:translational_diffusion}B, crowder with 25 kDa).
 
@@
 \includegraphics[scale=0.5]{diff_cyto_over_singles.pdf}
-\caption{A)Diffusion coefficient of crowders from single molecule simulations (blue bars) and cytoplasm simulations (orange bars).
+\caption{A) Diffusion coefficient of crowders from single molecule simulations (blue bars) and cytoplasm simulations (orange bars).
 B) The ratio between diffusion coefficient obtained from cytoplasm to single molecule simulations.
@@
 \subsubsection{tRNA is aggregating}
-After inspecting the trajectories of the cytoplasm models, we found that the reason why the translational diffusion constant for tRNA was reduced to a greater extent than for the other components of the cytoplasm model is that it is forming aggregates with Mg$^{2+}$, ATP and FBP (Figure~\ref{fig:tRNA_aggregation}).
-Thus, in the next sections we will ignore the RNA molecules in the analyses about the structural integrity of the crowders.
+After inspecting the trajectories of the cytoplasm models, we found that the reason why the translational diffusion constant for tRNA was reduced to a greater extent than for the other components of the cytoplasm model is that it is forming aggregates with Mg$^{2+}$, ATP and FBP (Fig.~\ref{fig:tRNA_aggregation}).
+Thus, in the next sections we will ignore the tRNA molecules in the analyses about the structural integrity of the crowders.
 Additionally, in the last section we will further investigate such aggregation and we will show a way to prevent it.
@@
 \includegraphics[scale=0.5]{rdf_RNA_metabolites.pdf}
-\caption{Radial distribution function showing the probability of finding metabolites or cations around RNA molecules in the cytoplasm model.}
+\caption{Radial distribution function showing the probability of finding metabolites or Magnesium cations around tRNA molecules in the cytoplasm model.}
 \label{fig:tRNA_aggregation}
@@
 
-\subsubsection{Structural integrity of individual chains}
+\subsubsection{Structural integrity}
 The structure of the individual chains of the crowders that were used in our cytoplasm model were not affected to a great extent by the crowded environment.
-The solvent accessible surface area (SASA) value for all crowders in the cytoplasm model is within 5\% of its value under the dilute condition.
-The root mean square deviation (RMSD) considering only their C$\upalpha$ atoms show a similar behavior, except for TufA, MetE and IcdA, which have RMSD values 105, 25 and 33\% higher in the crowded environment than in the dilute condition.
-Visual inspection of their trajectories showed that there is partial unfolding of their N- or C-terminal residues \colorbox{red!50}{DESCRIBE CONF.
-CHANGES}.
+%\st{The solvent accessible surface area (SASA) value for all crowders in the cytoplasm model is within 5\% of its value under the dilute condition.}
+The root mean square deviation (RMSD) considering only their C$\upalpha$ atoms show that crowders remain intacts under crowded condition, except for TufA, MetE and IcdA, which have RMSD values 105, 25 and 33\% higher in the crowded environment than in the dilute condition.
+Visual inspection of their trajectories showed that there is partial unfolding of their N- or C-terminal residues.
 \begin{figure}[H]
+\includegraphics[scale=0.5]{rmsd.pdf}
 \includegraphics[scale=0.5]{sasa.pdf}
-\includegraphics[scale=0.5]{rmsd.pdf}
 
-\caption{A) Chain interface area for oligomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
-The average bar value for single molecule simulations shows the sum over sasa of each protein chains averaged over the number of replica minus the sasa of the whole protein averaged over the number of replica.
-The average bar value for cytoplasm simulation shows the sum over sasa of each protein chains averaged over the number of replica and the abundance of the protein minus the sasa of the whole protein averaged over the number of replica and the abundance of the protein.
-The error bars show the standard errors.
-B) Root mean squared displacement (rmsd) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
+\caption{A) Root mean squared displacement (rmsd) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
 The average rmsd for each protein in single molecule simulation is over number of chains of proteins and three replica.
 The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.
-The error bars show the standard errors.}
+The error bars show the standard errors.
+B) Chain interface area for oligomers from single molecule simulations (blue bars) and cytoplasm simulations (orange bars).
+The average value from single molecule simulation is over three replica.
+The average value from cytoplasm simulation is over the number of proteins and three replica.
+The error bars show the standard errors.
+}
 \label{fig:structural_integrity_chain}
 \end{figure}
-
-% -------------------- STRUCTURAL INTEGRITY OF OLIGOMERS --------------------
-
-\subsubsection{Structural integrity of oligomers}
-
-% \colorbox{red!50}{WE STILL NEED TO MAKE THIS PLOT}
-
-%\begin{figure}[H]
-%\includegraphics[scale=0.5]{oligomer.pdf}
-%\caption{Inter-chain hydrogen bonds (hb) for oligomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
-The average hydrogen bonds for each oligomer in single molecule simulation is over number of chains and three replica.
-The average hydrogen bonds for each oligomer in cytoplasm simulation is over number of appearance of the chain in the oligomer in three replica.
-The error bars show the standard errors.}
-%\end{figure}
 
@@
 Proteins are slightly more flexible in the crowded environment.
-Root mean square fluctuation (RMSF) calculations show that the flexibility profile of the proteins is not affected by crowding, except in the cases in which there are conformational changes, TufA ...
-as shown before (Figure~\ref{fig:rmsf}).
-This is in agreement with our results showing that the structural integrity of the proteins is not significantly affected.
-In addition to that, our results show that the proteins are slightly more flexible in the crowded environment than in the dilute condition (Figure~\ref{fig:rmsf}).
-This is likely due to the nonspecific interactions and collisions that often happen under crowding conditions \textbf{\color{red}{REF}}.
+Root mean square fluctuation (RMSF) calculations show that the flexibility profile of the proteins is not affected by crowding, except in the cases in which there are conformational changes (TufA, MetE and IcdA) (Fig.~\ref{fig:rmsf}).
+This is in agreement with our results showing that the structural integrity of the proteins is not significantly affected (Fig.~\ref{fig:structural_integrity_chain}A).
+%In addition to that, our results show that the proteins are slightly more flexible in the crowded environment than in the dilute condition. 
+This is likely due to the nonspecific interactions that often happen under crowding conditions~\cite{Feig2011, miklos2011, Wang2012b}.
+The effec of crowded environment on inter-chain interaction of type hydrophobic, hydrogen bonding and salt bridges can be assessed by integrity of oligomeric chains.
+This is measured as chain interface area (Fig.~\ref{fig:structural_integrity_chain}B).
+Slightly lower buried area in cytoplasm simulations manifests destabilizing effect of crowded environment on chains of oligomers except except IcdA, GapA and Eno (Fig.~\ref{fig:structural_integrity_chain}B).
+These three oligomers also show the least changes in their RMSF profiles (Fig.~\ref{fig:rmsf}).
 
@@
 
-The initial detection of aggregation in our cytoplasm model was due to the out-lier behavior of the translational diffusion constant of tRNA when compared to the other crowders.
-However, tRNA is not causing such phenomenon.
+The initial detection of aggregation in our cytoplasm model was due to the outlier behavior of the translational diffusion constant of tRNA when compared to the other crowders.
+However, tRNA is not causing such phenomenon (Fig.~\ref{fig:avoiding_aggregation}A).
 Instead, it is triggered by the exaggerated interaction between the phosphate-containing metabolites, ATP and FBP, with Mg$^{2+}$.
-tRNA gets involved in the aggregates because Mg$^{2+}$ is present in the box only as counter-ion for ATP and tRNA, and so tRNA is one of the few "sources" of Mg$^{2+}$ in the whole system.
-We looked for ways of avoid aggregation with simulations of small systems composed by these metabolites in high concentration, Mg$^{2+}$ and water (please check the methods section for details about the parameters we used).
+tRNA gets involved in the aggregates because Mg$^{2+}$ is present in the box only as counter-ion for ATP and tRNA, and so tRNA is one of the few ``sources'' of Mg$^{2+}$ in the whole system.
+We looked for ways of avoiding aggregation by simulations of small systems composed of these metabolites in high concentration, Mg$^{2+}$ and water (please check the methods section for details about the parameters we used).
 
-We found that completely protonating the phosphate groups of ATP and FBP is enough prevent their aggregation with Mg$^{2+}$ (Figure~\ref{fig:avoiding_aggregation}).
-While RDFs show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$, this is not observed for ATPH\textsubscript{3} and FBPH\textsubscript{4}(Figure~\ref{fig:protonation}).
+% We found that completely protonating the phosphate groups of ATP and FBP is enough to prevent their aggregation with Mg$^{2+}$ (Fig.~\ref{fig:avoiding_aggregation}B-C).
+While RDFs show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$ (Fig.~\ref{fig:avoiding_aggregation}A), this is not observed for ATPH\textsubscript{3} and FBPH\textsubscript{4}(Fig.~\ref{fig:avoiding_aggregation}B-C).
+
+While RDFs show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$ (Fig.~\ref{fig:avoiding_aggregation}A), we found that completely protonating the phosphate groups of ATP and FBP is enough to prevent their aggregation with Mg$^{2+}$ (Fig.~\ref{fig:avoiding_aggregation}B-C).
 
 \begin{figure}[H]
-\includegraphics[scale=0.5]{rdf_mg.pdf}
-\caption{Radial distribution function (rdf) showing the probability of finding a metabolite or an RNA around MG2+. The inset plot shows the rdf for short distance ($ r < 0.7 \,\, nm$)}
+\hspace*{-2cm}\includegraphics[scale=0.5]{rdf_mg.pdf}
+\caption{Radial distribution function (rdf) showing the probability of finding A) metabolites, B) ATP$^{-3}$ or an ATPH\textsubscript{3}, C) FBP$^{-4}$ or an FBPH\textsubscript{4} around MG2+.
+The averages are over three replica and the shaded regions are the standard errors.}
 \label{fig:avoiding_aggregation}
-\end{figure}
-
-\begin{figure}[H]
-\includegraphics[scale=0.5]{rdf_mg_metabolites.pdf}
-\caption{A) Radial distribution function (rdf) showing the probability of finding an ATP$^{-3}$ or an ATPH\textsubscript{3} around Mg$^{2+}$.
-B) Radial distribution function (rdf) showing the probability of finding an FBP$^{-4}$ or an FBPH\textsubscript{4} around Mg$^{2+}$.
-}
-\label{fig:protonation}
 \end{figure}
@@
 We found that the aggregation of ATP$^{-3}$ and FBP$^{-4}$ around Mg$^{2+}$, which was used as counter-ion for tRNA, leads to its exaggerated confinement.
-Thus, tRNA is considered as an out-lier in our cytoplasm model.
+Thus, tRNA is considered as an outlier in our cytoplasm model.
 In fact, the slope of the linear regression line fitted to $D_{cyto}/D_{dil}$ is reduced by by 50\% and gets even closer to zero (slope $= 0.00014$) after removing the diffusion coefficient of tRNA from the data.
@@
 
-It has been demonstrated that each of the force fields that are most widely used to perform molecular dynamics simulations of macromolecules have different results for highly concentrated amino acid solutions ~ \cite{andrews2013molecular}.
-Previous all-atom simulations of high scale cytoplasm models used parameters from the CHARMM family of force fields ~\cite{yu2016biomolecular,mcguffee2010diffusion}.
-In this work we use, for the first time, the AMBER force fields to study crowding effects on a cytoplasm model.
-All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of \textit{Mycoplasma genitalium} composed by 103 and 12 million atoms that were simulated for 20 and 60 \SI{}{\nano\second}, respectively~\cite{yu2016biomolecular}.
+It has been demonstrated that each of the force fields that are most widely used to perform molecular dynamics simulations of macromolecules have different results for highly concentrated amino acid solutions~\cite{andrews2013}.
+Moreover, it has been reported that simulations of highly concentrated environment leads to aggregation by using additive force fields~\cite{Petrov2014a,Abriata2015a, Nawrocki2017a}.
+Several solutions has been proposed to eliminate aggregation~\cite{Best2014a,Piana2015a, Bashardanesh2018b}.
+Previous all-atom simulations of high scale cytoplasm models used parameters from the CHARMM family of force fields~\cite{Yu2016a}, whose solvent-solute interactions are strengthened according to the solution proposed by Best {\em et al.}~\cite{Best2014a}.
+In this work we use the AMBER force fields~\cite{Best2014a} to study crowding effects on a cytoplasm model.
+All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of \textit{Mycoplasma genitalium} composed of 103 and 12 million atoms that were simulated for 20 and 60 \SI{}{\nano\second}, respectively~\cite{Yu2016a}.
 The cytoplasm model described in this work has 1.5 million atoms and was simulated for the total of \SI{3}{\micro\second}, an unprecedented time scale to study crowding effects on a heterogeneous cytoplasm model in atomic detail.
 
-All crowders are as compact in the cytoplasm model as they are in dilute conditions (Figure \ref{fig:structural_integrity_chain}A).
-Despite this overall stability, three crowders (TufA, MetE and IcdA) displayed greater deviation from their crystallographic structure in the cytoplasm model when compared to the dilute condition, indicating that there was some degree of local unfolding (Figure \ref{fig:structural_integrity_chain}B).
-Interestingly, the flexibility profiles calculated considering the C$\upalpha$ for each chain of the elements of our cytoplasm model show that, in the crowded condition, most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are slightly more flexible than in the dilute condition (Figure \ref{fig:rmsf}).
-This is in agreement with simulation~\cite{feig2011variable} and experimental data~\cite{miklos2011protein,wang2012macromolecular} showing that crowders can destabilize the native structure of proteins due to the numerous unspecific short-lived interactions that are formed between them in a crowded environment.
+All crowders are as compact in the cytoplasm model as they are in dilute conditions (Fig.~\ref{fig:structural_integrity_chain}).
+Despite this overall stability, three crowders (TufA, MetE and IcdA) displayed greater deviation from their crystallographic structure in the cytoplasm model when compared to the dilute condition, indicating that there was some degree of local unfolding (Fig.~\ref{fig:structural_integrity_chain}).
+Interestingly, the flexibility profiles calculated considering the C$\upalpha$ for each chain of the elements of our cytoplasm model show that, in the crowded condition, most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are slightly more flexible than in the dilute condition (Fig.~\ref{fig:rmsf}).
+This is in agreement with simulation~\cite{Feig2011} and experiment~\cite{miklos2011,Wang2012b} showing that crowders can destabilize the native structure of proteins due to the numerous unspecific short-lived interactions that are formed between them in a crowded environment.
 
-The oligomeric state of the crowders... \textbf{\color{red}{MISSING RESULTS: whole crowder structural integrity using buried oligomer area}}
+Having variation of oligomeric state of the crowders in our model (Table~\ref{tbl:protein_fraction}) enabled us to study the effec of crowding on inter-chain interactions.
+We found out that the oligomers are either slightly destabilized, or no effect is seen (GapA and Eno) (Fig.~\ref{fig:structural_integrity_chain}B).
+This is in agreement with rmsf results (Fig.~\ref{fig:rmsf}).
 
@@
 After identifying this problem, we used simulations of small systems containing only the metabolites, Mg$^{2+}$ and water to test different strategies to prevent aggregation in future studies using long simulations of cytoplasm models.
-We found that protonating the phosphate-containing metabolites is enough to reach this goal (Figure \ref{fig:avoiding_aggregation}).
+We found that protonating the phosphate-containing metabolites is enough to reach this goal (Fig.~\ref{fig:avoiding_aggregation}).
 Even though ATPH$_3$ and FBPH$_4$ are not the biologically relevant forms of these molecules in physiological pH, we consider that using their protonated forms is a good compromise between parametrization complexity and practical results.
@@
 To the best of our knowledge, there is only one work in the literature which studied large scale cytoplasm models that included metabolites with all-atom explicit solvent molecular dynamics~\cite{Yu2016a}.
-Although they didn't report aggregation, they mentioned that the translational diffusion of highly charged phosphate-containing metabolites was "much slower" than expected.
+Although they didn't report aggregation, they mentioned that the translational diffusion of highly charged phosphate-containing metabolites was ``much slower'' than expected.
 They attributed this observation to non-specific interactions between such metabolites and proteins.

### 2019-09-29 paper/barplots.py

@@ -99,9 +99,9 @@ if __name__ == '__main__':
         ax.set_xticklabels(protein_name, fontsize = fs, rotation=deg)
         if args.prop == "sasa":
             ax.set_xticklabels(oligomers, fontsize = fs, rotation=deg)
-            plt.text(-2., 160,'A', fontsize=fs+2)
+            plt.text(-2., 160,'B', fontsize=fs+2)
         elif args.prop == "rmsd":
-            plt.text(-2., 0.48,'B', fontsize=fs+2)
+            plt.text(-2., 0.48,'A', fontsize=fs+2)
 
     ax.legend(fontsize = fs)
     ax.tick_params(labelsize=fs-2);

### 2019-09-29 paper/diff_plots.py

@@ -14,6 +14,7 @@ order_s_b=[5, 9, 2, 1, 8, 3, 6, 7, 4] #smallest to biggest
  
 mass_crowder = [84320.000,  83764.400,  91489.800,  184409.000,  7271.120,  115920.600,  141602.400,  90884.000,  23761.100] #  146.124,  504.164,  336.080,  306.318]
 mass_sorted = sorted(mass_crowder)
+mass_sorted2 = sorted(mass_crowder[0:8])
 fs=16
 
 avg_cyto, ste_cyto =[],[]
@@ -52,15 +53,17 @@ for i in range(0, len(avg_sing)):
 	STD.append((avg_cyto[i]/avg_sing[i])* math.sqrt( (ste_cyto[i]/avg_cyto[i])**2 + (ste_sing[i]/avg_sing[i])**2 ) )
 
 z = np.polyfit([m / 1000. for m in mass_sorted], AVG, 1)
-# z = np.polyfit([m / 1000. for m in mass_sorted], AVG[0:1]+AVG[2:9], 1)
+z2 = np.polyfit([m / 1000. for m in mass_sorted2], AVG[0:1]+AVG[2:9], 1)
 
-plt.errorbar([m / 1000. for m in mass_sorted] , AVG, STD , marker='o', markersize=10, linestyle=' ')
-plt.plot([m / 1000. for m in mass_sorted], z[1] + [z[0]*m / 1000. for m in mass_sorted], color='black')
+plt.errorbar([m / 1000. for m in mass_sorted] , AVG, STD , marker='o', markersize=10, linestyle=' ', label="data points")
+plt.plot([m / 1000. for m in mass_sorted], z[1] + [z[0]*m / 1000. for m in mass_sorted], color='black', label="fitted w outlier")
+plt.plot([m / 1000. for m in mass_sorted2], z2[1] + [z2[0]*m / 1000. for m in mass_sorted2], color='k', linestyle='--', label="fitted w/o outlier")
 plt.xlabel('$kDa$', fontsize=fs)
 plt.ylabel('$D_{cyto}/D_{dil}$', fontsize=fs)
 plt.text(-35,0.25,'B',fontsize=fs+2)
 plt.tick_params(labelsize=fs-2);
 plt.ylim([0, .25])
+plt.legend(loc = 2, fontsize=fs)
 plt.savefig('diff_cyto_over_singles.pdf', bbox_inches="tight")
 plt.close()
 

### 2019-09-29 paper/shrinking_plot.py

@@ -28,7 +28,7 @@ for prop in ["volume"]:
 	if prop == "volume":
 		plt.ylabel('%s ($10^3 nm^3$)' % prop, fontsize=fs)
 		plt.yticks(y, ylbl)
-		plt.text(-120,43100,'B', fontsize=fs)
+		plt.text(-120,47000,'B', fontsize=fs)
 	else:
 		plt.ylabel('%s ($kg/m^3$)' % prop, fontsize=fs)
 		# plt.text(-80,1150,'B', fontsize=fs)
@@ -52,7 +52,7 @@ image = mpimg.imread(pic[0])
 ax1 = fig.add_subplot(121)
 plt.imshow(image)
 plt.axis('off')
-plt.text(-100,-5, 'A', fontsize = fs)
+plt.text(-100,-50, 'A', fontsize = fs)
 plt.plot(x1,x2, 'k-')
 plt.plot(x2,x3, 'k-')
 plt.plot(x4,x1, 'k-')

### 2019-09-30 paper/virtual_cytoplasm_v2.tex
Ready!

@@
 \begin{abstract}
+Biomolecular crowding affects the biophysical and biochemical behavior of macromolecules when compared to the dilute environment present in experiments made with isolated proteins.
+Computational modeling and simulation are essential tools in the investigation of how crowding affects the structural dynamics and biological activities of macromolecules.
+As computational power increased, modeling and simulating large scale all-atom explicit solvent models of the prokaryote cytoplasm become possible.
+In this work, we build an atomistic model of the cytoplasm of \textit{Escherichia coli} composed by 1.5 million atoms and submit it to a total of \SI{3}{\micro\second} of molecular dynamics simulations, which is unprecedented for a system of this scale.
+Our model reproduces experimental data about the diffusion coefficient and stability of macromolecules under crowded conditions.
+We provide a Python script and a set of files that enables any researcher to build their own \textit{E.
+coli} cytoplasm models to address questions related to crowding.
 %For Nature, the abstract is really an introductory paragraph set
@@
 Instead of simple and diluted, the \textit{in vivo} microenvironment is highly heterogeneous and concentrated, leading to spatial constrains that directly affects biological activities~\citep{ostrowska2019}.
-Macromolecular crowding directly influences the stability of macromolecules.
-There is an equilibrium between stabilization of their native structure due to the entropic component given by the excluded volume effect~\cite{cheung2005} and destabilizationg due to enthalpic component given by unspecific interactions with the crowders~\cite{Feig2011,miklos2011,Wang2012b}.
+This effect, called macromolecular crowding, directly influences the stability of macromolecules through an equilibrium between stabilization of their native structure due to the entropic component given by the excluded volume effect~\cite{cheung2005} and destabilizationg due to enthalpic component given by unspecific interactions with the crowders~\cite{Feig2011,miklos2011,Wang2012b}.
 Additionally, crowding affects the structure and dynamics of water around proteins, which has an indirect but important effect in their biological activities~\cite{Harada2012a,king2013}.
 
-While the structure and dynamics of biomolecules is well characterized \textit{in vitro}, the understanding of the influences of crowded environments for the \textit{in vivo} are still evolving.
+While the structure and dynamics of biomolecules is well characterized \textit{in vitro}, the understanding of the influences of crowded environments \textit{in vivo} are still evolving.
 Techniques such as nuclear magnetic resonance~\cite{reckel2007,pielak2008} and fluorescence spectroscopy~\cite{ignatova2004,xie2008,English2011} are among the most relevant to the field due to their ability to probe cells in a non destructive way.
 Another essential class of methods come from computational modeling and simulation.
-By extrapolating the advances in computational power in the last decades to the future, it has been estimated that the atomistic simulation of an entire \textit{Escherichia coli} bacterium for \SI{1}{\nano\second} will be possible in 15 years from now~\cite{vanGunsteren2006a}.
+By extrapolating the advances in computational power in the last decades to the future, it was estimated that the atomistic simulation of an entire \textit{Escherichia coli} bacterium for \SI{1}{\nano\second} will be possible in 15 years from now~\cite{vanGunsteren2006a}.
 
 As the available computational power increases, models become increasingly complex and are better representations of reality.
-The most simple class of computational models apply Brownian Dynamics calculations to probe macromolecules immersed in a solution of crowders that are treated as hard spheres with varying radius to represent different types of molecules~\cite{Ando2010}.
+The most simple class of computational models apply Brownian Dynamics calculations to probe macromolecules immersed in a solution of crowders that are treated as hard spheres with varying radii to represent different types of molecules~\cite{Ando2010}.
 These models usually use an implicit representation of the solvent~\cite{Mcguffee2010}.
@@
 This model, as well as a smaller version with 13 million atoms, was submitted to molecular dynamics simulations~\cite{Yu2016a}.
-Despite the size of these systems, the simulation times were too short, \SI{20}{\nano\second} for the complete model and \SI{140}{\nano\second} for the smaller version, to accurately characterize the effects of the slow diffusion of its constituents.
+Despite the size of these systems, the simulation times were too short, \SI{20}{\nano\second} for the complete model and \SI{140}{\nano\second} for the smaller version, to accurately characterize dynamical effects that can take place in such a large scale.
 
-In this work we report an all-atom model of the cytoplasm of \textit{Escherichia coli} that was built to reflect the real biological system.
+In this work, we report an all-atom model of the cytoplasm of \textit{Escherichia coli} that was built to reflect the real biological system.
 We discuss the difficulties that arise when building such system and provide a set of Python scripts and files that anyone can use to build its own crowded systems with custom parameters and elements.
@@
 
-\subsection*{Initial structures}
+\subsection{Initial structures}
 The proteins and tRNA were downloaded from Protein Data Bank (PDB).
-We looked for the proteins structures that were either from or expressed in {\em E-coli}.
-In case of 1U22 (MetE) and 2EIP (Ppa) we used a loop-closure modeling tool based on Random Coordinate Descent (RCD) method~\cite{Chys2013} to correct the information for missing residues.
-The four metabolites were parametrized using GAFF and Antechamber.
+We looked for the proteins structures that were either from {\em E.
+coli} or its closest homologues (Table \ref{tbl:protein_fraction}).
+In case of PDB IDs 1U22 (MetE protein) and 2EIP (Ppa protein), we used a loop-closure modeling tool based on the Random Coordinate Descent (RCD) method~\cite{Chys2013} to model missing residues.
+The four metabolites were parametrized using the General Amber Force Field (GAFF)~\cite{sprenger2015general} and Antechamber~\cite{Wang2005b} using RESP charges~\cite{Bayly1993a} that were derived from electron density calculated at the HF/6-31G* theory level with Gaussian16.
 
-\subsection*{Molecular Dynamics Simulation}
+\subsection{General simulation setup}
+The cytoplasm model was built at 30\% biomolecular mass fraction at the physiological salt concentration of 0.15 mol/L KCl.
+All simulations were performed with the GROMACS 2018 package~\cite{Abraham2015a} with Amber99SB-ws~\cite{Best2014a}, a modified version of the Amber99SB force field, in combination with the TIP4P/2005 water model~\cite{Abascal2005b}.
+All interactions were calculated explicitly inside a 1 nm radius and long range electrostatic interactions were treated using the particle mesh Ewald algorithm~\cite{Essmann1995a}.
+Corrections due to long range effects of dispersion interactions were made to both energy and pressure.
+All chemical bonds were constrained at their equilibrium length using the LINCS algorithm~\cite{Hess2008b} allowing an integration time step of 2 fs.
+Temperature was kept at 310 K by the v-rescale thermostat~\cite{Bussi2007a} with a coupling constant of 0.5 ps.
+Pressure was kept at 1 bar by the Berendsen barostat~\cite{Berendsen1984a} with the coupling constant of 10.0 ps during thermalization steps and by the Parrinello-Rahman barostat~\cite{Parrinello1981a} with the same coupling constant during production steps.
 
-{\bf General Simulation Setup: }The proteins were simulated at 30\% biomolecular mass fraction in  a physiological salt concentration (0.15M NaCl).
-For all simulations, Amber99SB-ws force field was used~\cite{Best2014a} in combination with the TIP4P/2005 water model~\cite{Abascal2005b}.
-Electrostatic interactions were treated using the particle mesh Ewald algorithm~\cite{Essmann1995a}.
-All chemical bonds were constrained at their equilibrium length using the LINCS algorithm~\cite{Hess2008b} allowing an integration time step of 2 fs.
-Temperature was controlled at 310 K using the v-rescale algorithm~\cite{Bussi2007a} and a coupling time of 0.5 ps.
-The pressure was controlled at 1 bar using the Parrinello-Rahman algorithm~\cite{Parrinello1981a} with a time constant of 10 ps.
+All systems were initially submitted to energy minimization with the steepest descent algorithm until convergence to simple precision.
+For the cytoplasm model, a short shrinking step of 500 ps was performed in order to allow the box volume to adjust to its optimal value.
+For all systems, initial velocities were sampled from a Maxwell-Boltzmann distribution at 310 K and we ran a 1 ns NVT thermalization step followed by 10 ns of NPT equilibration.
+Production runs for the dilute systems, in which each crowder is simulated alone in a water box, had the length of 200 ns.
+The cytoplasm model was run for 1 \SI{}{\micro\second}.
+All systems were simulated in three replicas.
+While the replicas for the diluted systems were generated by different initial velocities, each replica of the cytoplasm model is a completely different box in which the components of the model are in different positions and orientations.
 
-For error analysis, each simulations were repeated three times with independent starting velocities.
+\subsection{Analyses}
+For statistics, the average and standard error of the three trajectories are reported.
+For values that are specific to each chain, we used the average over the individual chains in each trajectory of the replicas.
+Additionally, since there is more than one copy of some crowders in the cytoplasm models, average values and standard errors are calculated considering each copy and each replica independently.
+Similarly to the dilute system, separate chains of each copy of the crowders were considered as independent samples when running analyses that focus on one single chain.
+The first half of all simulations was discarded as equilibration time and analyses used only the second half.
 
-All simulations were performed with Gromacs 2018.
-Single simulations were started from crystal conformations.
-The cytoplasm simulations starting conformation were taken from the equilibrated conformation of each single simulation.
-
-{\bf Single Simulation Setup: } Each component was simulated with the same parameter as the cytoplasm for $200 ns$.
-
-\subsection*{Analysis}
-Before any analysis the periodic boundary condition (pbc) artifacts have been removed.
-We used GROMACS tools to do the analysis.
-For single component simulations, first the components were made whole and jump removed and then all the atom were put inside the compact box.
-The same treatment were applied to the cytoplasm simulations.
-Additionally, each component's trajectory were extracted and fitted by rotation and translation for later rotational correlation time analysis.
-
-A Mean Square Displacement (MSD) analysis was used to calculate the translational diffusion coefficient~\cite{Allen1987a}.
-The diffusion coefficients were extracted by a linear fit to MSD analysis by averaging blocks with a length of $10 \,ns$.
-In principle diffusion coefficient needs to be corrected for finite size effects~\cite{Yeh2004} but due to relatively large simulation boxes this correction is negligible.
+Analyses were made GROMACS tools after removing artifacts generated by periodic boundary conditions from the trajectories.
+The Mean Square Displacement (MSD) analysis was performed to calculate the translational diffusion coefficient, which were extracted by a linear fit to MSD by averaging blocks with a length of $10 \,ns$~\cite{Allen1987a}.
+In principle, diffusion coefficient needs to be corrected for finite size effects~\cite{Yeh2004} but due to relatively large simulation boxes this correction is negligible.
+Root Mean Square Deviation (RMSD) and Root Mean Square Fluctuation (RMSF) calculations considered only C$\upalpha$ atoms and the reference frame for RMSF was the midpoint of the time interval used for analysis in each trajectory.
+The chain interface area of a given oligomer in a given frame was calculated as the difference between the solvent exposed surface area (SASA) of the oligomer in that frame and the sum of the SASAs of each of its chains independently in that same frame.
+Other analyses, such as box volume and Radial Distribution Function (RDF), followed standard procedures.
 
@@
 Then, this number was divided by the oligomeric state of each protein.
-This was done because the original oligomeric state for each protein was kept as reported in their crystallographic structure.
+This was done because the original oligomeric state for each protein was kept as reported in their crystallographic structure (Table~\ref{tbl:protein_fraction}).
 
@@
 \caption{Absolute and cumulative fraction of the total abundance of non-ribosomal proteins in the cytosol of {\em E.
-coli} K12 to which each protein that was selected to compose our cytoplasm model corresponds to.}
+coli} K12 and oligomeric state of each protein that was selected to compose our cytoplasm model.}
 \label{tbl:protein_fraction}
@@
 We considered the most abundant molecules of each metabolite class as representatives, i.e.
-Glutamate for amino acids, ATP for nucleotides, FBP for central carbon intermediates and Glutathione redox cofactors~\cite{Bennett2009}.
+Glutamate (GLT) for amino acids, ATP for nucleotides, Fructose-2,6-Biphosphate (FBP) for central carbon intermediates and Glutathione (GSH) redox cofactors~\cite{Bennett2009}.
 The total number of molecules was calculated considering data showing that the number of metabolite molecules in the cytoplasm of \textit{E.
@@
 The number of water molecules was calculated according to the desired biomolecular concentration, which is a parameter of the cytoplasm model.
-This number ranges from 300 to 400 $mg/ml$ in biological systems such as \textit{E.
-coli} cytoplasm~\cite{Zimmerman1991,Ellis2003a}.
+This number ranges from 300 to 400 g/L in biological systems such as \textit{E.
+coli} cytoplasm~\cite{Zimmerman1991}, which corresponds to biomolecular concentrations as high as 40\%~\cite{Ellis2003a}.
 In our case, we choose the biomolecular concentration of 30\%, that is, the number of molecules necessary to reach a ratio of total biomolecular mass to water mass of 30\% was inserted into the cytoplasm model.
@@
 \hline
-Metabolite & GLU & 1436\\
+Metabolite & GLT & 1436\\
   & ATP & 144\\
@@
 \end{tabular}
-\caption{Number of copies for each component of the cytoplasm model built at the biomolecular fraction of 30\%}
+\caption{Number of copies for each component of the cytoplasm model built at the biomolecular fraction of 30\%.}
 \label{tbl:soup_components}
@@
 \subsection{Building the simulation box}
-All components can be put in the same simulation box by inserting each of them in random orientations in a cubic box of side $L$ that is initially empty.
+All components can be put in the same simulation box by inserting each of them in random orientations in a cubic volume of side $L$ that is initially empty.
 However, that is not a trivial process.
@@
 We start with a box size $L$ that is too small to allow all components to fit in the box by random insertion.
-In our case, we started with $L = 30$ nm.
+In our case, we start with $L = 30$ nm.
 Then, we allow 100 insertion trials for each element.
@@
 In order to do this, the droplets around each component are also constructed iteratively.
-The number of water molecules that we must place in the simulation box is known from the desired biomolecular fraction, which is a parameter of the model, and the total biomolecular mass of the system.
-However, we don't know the thickness of the water layer around each component, $l$, that accounts for such amount of water molecules.
+The number of water molecules that we must place in the simulation box is known from the total biomolecular mass of the system and the desired biomolecular fraction, which is a parameter of the model.
+However, we don't know the thickness of the water layer around each component, $l$, that accounts for such amount of water molecules \textit{a priori}.
 Since $l$ depends on a series of factors such as the shape, size and abundance of each element, we define it iteratively.
@@
 If it is larger, we reduce the droplet thickness by $dl$, reduce $dl$ by 10 times and then increase $l$ with the new smaller $dl$.
-We can carry this process until an arbitrary precision cutoff, such as 5\%, is satisfied.
+We can carry this process until an arbitrary precision cutoff, such as 5\%, for the number of waters in the system is satisfied.
 
@@
 \includegraphics[scale=0.5]{shrinking.pdf} 
-\caption{A) The volume of the box with length size of $L=35 \,\, nm$ reduces during the energy minimization in the first $300 \,\, ps$ B) The density of the same box increases within the first $300\,\, ps$.
-The repeated experiments are shown in different colors.
+\caption{A) The initial volume of the simulation box, $L=35.0$ nm, and after the shrinking step, $L=22.9$ nm.
+B) The evolution of the volume of the simulation box during the shrinking simulations.
+Each replica is shown in different colors.
 }
 \end{figure}
-
-Python scripts and all topology and structure files necessary to enable any researcher to build its own cytoplasm models are publicly available at github.com/dspoel/soup.
-Such models can be used as tools that answer many research questions about the effects of crowding on specific systems of interest.
-With the files we are providing it is possible, for example, to add a probe protein to investigate the effect of crowding on it, to add new macromolecular or small crowders, verify the effects of different temperatures and change the biomolecular concentration to increase or decrease the intensity of the crowding effect.
 
@@
 
-% -------------------- TRANSLATIONAL DIFFUSION --------------------
-
 \subsubsection{Translational diffusion}
 
-Under crowded conditions, molecules are constrained to a smaller volume due to collisions with the other components of the crowded environment.
+Under crowded conditions, molecules are constrained to a smaller volume by the other components of the crowded environment.
 The extent of this effect is evident when we compare the translational diffusion constant of each component in the cytoplasm model, $D_{cyto}$, and in a dilute system, $D_{dil}$ (Fig.~\ref{fig:translational_diffusion}A).
@@
 \includegraphics[scale=0.5]{diff_cyto_over_singles.pdf}
-\caption{A) Diffusion coefficient of crowders from single molecule simulations (blue bars) and cytoplasm simulations (orange bars).
-B) The ratio between diffusion coefficient obtained from cytoplasm to single molecule simulations.
+\caption{A) Diffusion coefficient of crowders from simulations under diluted conditions (blue bars) and cytoplasm simulations (orange bars).
+B) The ratio between diffusion coefficient obtained from simulations under both conditions.
 The crowders are sorted according to their sizes on the x-axis.}
@@
 
-% \colorbox{green!50}{ Are we going to calculate and discuss rotational diffusion? No, not now at least.}
-
-% -------------------- tRNA IS AN OUTLIER DUE TO AGGREGATION --------------------
-
 \subsubsection{tRNA is aggregating}
 After inspecting the trajectories of the cytoplasm models, we found that the reason why the translational diffusion constant for tRNA was reduced to a greater extent than for the other components of the cytoplasm model is that it is forming aggregates with Mg$^{2+}$, ATP and FBP (Fig.~\ref{fig:tRNA_aggregation}).
-Thus, in the next sections we will ignore the tRNA molecules in the analyses about the structural integrity of the crowders.
-Additionally, in the last section we will further investigate such aggregation and we will show a way to prevent it.
+Thus, we will omit the tRNA molecules in the analyses about the structural integrity of the crowders.
+In the next sections we will further investigate such aggregation and we will show a way to prevent it.
 
@@
 \includegraphics[scale=0.5]{rdf_RNA_metabolites.pdf}
-\caption{Radial distribution function showing the probability of finding metabolites or Magnesium cations around tRNA molecules in the cytoplasm model.}
+\caption{Radial distribution function showing the probability of finding metabolites or Magnesium cations around tRNA molecules in the cytoplasm model.
+Note that ATP and FBP are much more concentrated around tRNA than the other metabolites, GSH and GLT.}
 \label{fig:tRNA_aggregation}
@@
 
-% -------------------- STRUCTURAL INTEGRITY OF SINGLE CHAINS --------------------
-
 \subsubsection{Structural integrity}
 The structure of the individual chains of the crowders that were used in our cytoplasm model were not affected to a great extent by the crowded environment.
-%\st{The solvent accessible surface area (SASA) value for all crowders in the cytoplasm model is within 5\% of its value under the dilute condition.}
-The root mean square deviation (RMSD) considering only their C$\upalpha$ atoms show that crowders remain intacts under crowded condition, except for TufA, MetE and IcdA, which have RMSD values 105, 25 and 33\% higher in the crowded environment than in the dilute condition.
+The value of the root mean square deviation (RMSD) considering only their C$\upalpha$ in the cytoplasm model is withing 11\% of its value under dilute conditions, except for TufA, MetE and IcdA, which have RMSD values 105, 25 and 33\% higher in the crowded environment than in the dilute condition.
 Visual inspection of their trajectories showed that there is partial unfolding of their N- or C-terminal residues.
+
+The effect of crowded environment on the structural integrity of oligomers can be evaluated by their chain interface area.
+This quantity is the solvent accessible surface area that is removed from access to the solvent by oligomerization.
+Our results show that oligomers are not significantly disturbed in the cytoplasm model when compared to the diluted condition.
+Similarly to what can be seem by the RMSD values, which reflect the structural integrity of individual chains, the oligomers that undergo the most change is TufA (Fig.~\ref{fig:structural_integrity_chain}B).
+
 \begin{figure}[H]
@@
 \includegraphics[scale=0.5]{sasa.pdf}
-
-\caption{A) Root mean squared displacement (rmsd) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
-The average rmsd for each protein in single molecule simulation is over number of chains of proteins and three replica.
-The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.
-The error bars show the standard errors.
-B) Chain interface area for oligomers from single molecule simulations (blue bars) and cytoplasm simulations (orange bars).
-The average value from single molecule simulation is over three replica.
-The average value from cytoplasm simulation is over the number of proteins and three replica.
-The error bars show the standard errors.
-}
+\caption{A) RMSD for all protein crowders from simulations under diluted conditions (blue bars) and in the cytoplasm model (orange bars).
+This value reflects the structural integrity of individual chains of each crowder.
+B) Chain interface area for oligomers under diluted conditions (blue bars) and in the cytoplasm model (orange bars).
+This value reflects the structural integrity of oligomers of the crowders.}
 \label{fig:structural_integrity_chain}
@@
 
-% -------------------- STRUCTURAL DYNAMICS --------------------
-
 \subsubsection{Structural dynamics}
 Proteins are slightly more flexible in the crowded environment.
-Root mean square fluctuation (RMSF) calculations show that the flexibility profile of the proteins is not affected by crowding, except in the cases in which there are conformational changes (TufA, MetE and IcdA) (Fig.~\ref{fig:rmsf}).
-This is in agreement with our results showing that the structural integrity of the proteins is not significantly affected (Fig.~\ref{fig:structural_integrity_chain}A).
-%In addition to that, our results show that the proteins are slightly more flexible in the crowded environment than in the dilute condition. 
-This is likely due to the nonspecific interactions that often happen under crowding conditions~\cite{Feig2011, miklos2011, Wang2012b}.
-The effec of crowded environment on inter-chain interaction of type hydrophobic, hydrogen bonding and salt bridges can be assessed by integrity of oligomeric chains.
-This is measured as chain interface area (Fig.~\ref{fig:structural_integrity_chain}B).
-Slightly lower buried area in cytoplasm simulations manifests destabilizing effect of crowded environment on chains of oligomers except except IcdA, GapA and Eno (Fig.~\ref{fig:structural_integrity_chain}B).
-These three oligomers also show the least changes in their RMSF profiles (Fig.~\ref{fig:rmsf}).
+Root mean square fluctuation (RMSF) calculations show that the flexibility profile of the proteins is not affected by crowding, which is in agreement with our results showing that the structural integrity of the proteins is not significantly affected in the cytoplasm model (Fig.~\ref{fig:rmsf}).
+On the other hand, there are some little differences between the dilute and crowding conditions.
+In most cases (TufA, MetE, AhpC, CspC and Ppa), proteins are slightly more flexible on the cytoplasm model (Fig.~\ref{fig:rmsf}).
+This is in agreement with data showing that nonspecific interactions with crowders can lead to destabilization of proteins~\cite{Feig2011, miklos2011, Wang2012b}.
 
@@
 \includegraphics[scale=0.5]{rmsf.pdf}
-\caption{Root mean squared fluctuations (rmsf) for 8 crowders from single molecule simulation (blue bars) and cytoplasm simulation (orange bars).
-x-axis shows the residues for each chain of proteins.
-The average rmsf for each protein in single molecule simulation is over number of chains of proteins and three replica.
-The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica.
-The error bars show the standard errors.}
+\caption{Root mean squared fluctuations (RMSF) from simulations under diluted conditions (blue bars) and in the cytoplasm model (orange bars).
+The x-axis shows the residues for each chain of proteins.}
 \label{fig:rmsf}
 \end{figure}
-
-% -------------------- AVOIDING AGGREGATION --------------------
 
@@
 The initial detection of aggregation in our cytoplasm model was due to the outlier behavior of the translational diffusion constant of tRNA when compared to the other crowders.
-However, tRNA is not causing such phenomenon (Fig.~\ref{fig:avoiding_aggregation}A).
-Instead, it is triggered by the exaggerated interaction between the phosphate-containing metabolites, ATP and FBP, with Mg$^{2+}$.
+However, tRNA is not causing such phenomenon.
+Instead, it is triggered by the exaggerated interaction between the phosphate-containing metabolites, ATP and FBP, with Mg$^{2+}$ (Fig.~\ref{fig:avoiding_aggregation}A).
 tRNA gets involved in the aggregates because Mg$^{2+}$ is present in the box only as counter-ion for ATP and tRNA, and so tRNA is one of the few ``sources'' of Mg$^{2+}$ in the whole system.
-We looked for ways of avoiding aggregation by simulations of small systems composed of these metabolites in high concentration, Mg$^{2+}$ and water (please check the methods section for details about the parameters we used).
+We looked for ways of avoid aggregation by using simulations of small systems composed of these metabolites in high concentration, Mg$^{2+}$ and water (please check the methods section for details).
 
-% We found that completely protonating the phosphate groups of ATP and FBP is enough to prevent their aggregation with Mg$^{2+}$ (Fig.~\ref{fig:avoiding_aggregation}B-C).
+We found that completely protonating the phosphate groups of ATP and FBP is enough to prevent their aggregation with Mg$^{2+}$.
 While RDFs show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$ (Fig.~\ref{fig:avoiding_aggregation}A), this is not observed for ATPH\textsubscript{3} and FBPH\textsubscript{4}(Fig.~\ref{fig:avoiding_aggregation}B-C).
-
-While RDFs show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$ (Fig.~\ref{fig:avoiding_aggregation}A), we found that completely protonating the phosphate groups of ATP and FBP is enough to prevent their aggregation with Mg$^{2+}$ (Fig.~\ref{fig:avoiding_aggregation}B-C).
 
@@
 \hspace*{-2cm}\includegraphics[scale=0.5]{rdf_mg.pdf}
-\caption{Radial distribution function (rdf) showing the probability of finding A) metabolites, B) ATP$^{-3}$ or an ATPH\textsubscript{3}, C) FBP$^{-4}$ or an FBPH\textsubscript{4} around MG2+.
-The averages are over three replica and the shaded regions are the standard errors.}
+\caption{RDFs centered around Mg$^{2+}$ showing the probability of finding A) metabolites in the cytoplasm model, B) ATP$^{-3}$ and ATPH\textsubscript{3} or C) FBP$^{-4}$ and FBPH\textsubscript{4} in small independent systems (please see the methods section for details).}
 \label{fig:avoiding_aggregation}
 \end{figure}
-
- % -------------------- DISCUSSION --------------------
 
@@
 It has been demonstrated that each of the force fields that are most widely used to perform molecular dynamics simulations of macromolecules have different results for highly concentrated amino acid solutions~\cite{andrews2013}.
-Moreover, it has been reported that simulations of highly concentrated environment leads to aggregation by using additive force fields~\cite{Petrov2014a,Abriata2015a, Nawrocki2017a}.
-Several solutions has been proposed to eliminate aggregation~\cite{Best2014a,Piana2015a, Bashardanesh2018b}.
-Previous all-atom simulations of high scale cytoplasm models used parameters from the CHARMM family of force fields~\cite{Yu2016a}, whose solvent-solute interactions are strengthened according to the solution proposed by Best {\em et al.}~\cite{Best2014a}.
-In this work we use the AMBER force fields~\cite{Best2014a} to study crowding effects on a cytoplasm model.
+Moreover, it has been reported that simulations of highly concentrated environment leads to aggregation when additive force fields are used~\cite{Petrov2014a,Abriata2015a, Nawrocki2017a} and several approaches have been proposed to eliminate this problem~\cite{Best2014a,Piana2015a, Bashardanesh2018b}.
+Previous all-atom simulations of high scale cytoplasm models used parameters from the CHARMM family of force fields~\cite{Yu2016a}.
+In this work we use a modified version of the AMBER FF99SB force field in which the solvent-solute interactions are 10\% stronger~\cite{Best2014a} to study crowding effects on a cytoplasm model.
+This force field was chosen because it was corrected to reproduce experimental data when there are many proteins in the same simulation box.
 All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of \textit{Mycoplasma genitalium} composed of 103 and 12 million atoms that were simulated for 20 and 60 \SI{}{\nano\second}, respectively~\cite{Yu2016a}.
@@
 
-All crowders are as compact in the cytoplasm model as they are in dilute conditions (Fig.~\ref{fig:structural_integrity_chain}).
-Despite this overall stability, three crowders (TufA, MetE and IcdA) displayed greater deviation from their crystallographic structure in the cytoplasm model when compared to the dilute condition, indicating that there was some degree of local unfolding (Fig.~\ref{fig:structural_integrity_chain}).
-Interestingly, the flexibility profiles calculated considering the C$\upalpha$ for each chain of the elements of our cytoplasm model show that, in the crowded condition, most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are slightly more flexible than in the dilute condition (Fig.~\ref{fig:rmsf}).
+Three crowders (TufA, MetE and IcdA) displayed greater deviation from their crystallographic structure in the cytoplasm model when compared to the dilute condition, indicating that there was some degree of local unfolding (Fig.~\ref{fig:structural_integrity_chain}A).
+Interestingly, oligomers were marginally less stable (Fig.~\ref{fig:structural_integrity_chain}B) and the flexibility profiles calculated considering the C$\upalpha$ for each chain of the elements of our cytoplasm model show that, in the crowded condition, most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are slightly more flexible than in the dilute condition (Fig.~\ref{fig:rmsf}).
 This is in agreement with simulation~\cite{Feig2011} and experiment~\cite{miklos2011,Wang2012b} showing that crowders can destabilize the native structure of proteins due to the numerous unspecific short-lived interactions that are formed between them in a crowded environment.
-
-Having variation of oligomeric state of the crowders in our model (Table~\ref{tbl:protein_fraction}) enabled us to study the effec of crowding on inter-chain interactions.
-We found out that the oligomers are either slightly destabilized, or no effect is seen (GapA and Eno) (Fig.~\ref{fig:structural_integrity_chain}B).
-This is in agreement with rmsf results (Fig.~\ref{fig:rmsf}).
 
@@
 To the best of our knowledge, there is only one work in the literature which studied large scale cytoplasm models that included metabolites with all-atom explicit solvent molecular dynamics~\cite{Yu2016a}.
-Although they didn't report aggregation, they mentioned that the translational diffusion of highly charged phosphate-containing metabolites was ``much slower'' than expected.
+Although they didn't report aggregation, they mention that the translational diffusion of highly charged phosphate-containing metabolites is ``much slower'' than expected in their system.
 They attributed this observation to non-specific interactions between such metabolites and proteins.
@@
 
-% -------------------- CONCLUSIONS --------------------
-
 \section*{Conclusions}\label{sec:concl}
 
-In this work we built a cytoplasm model...
+In this work we built an all-atom explicit solvent cytoplasm model composed by proteins, tRNA, metabolites, inorganic ions and water that sums 1.5 million atoms (Table \ref{tbl:soup_components}).
+The identity and amount of each component of the model was chosen according to experimental data about the composition of the cytoplasm of \textit{E.
+coli}~\cite{Dong1996,Bennett2009,Link1997,Mcguffee2010}.
 
-the model was validated by reproducing experimental data about the reduction of the translational diffusion coefficient of proteins in E.coli.
+We submitted this model to a total of \SI{3}{\micro\second} of molecular dynamics simulations using a modified AMBER FF99SB force field that was tuned to reproduce experimental data of systems composed by several macromolecules~\cite{Best2014a}.
+Our model was validated by reproducing experimental data about the reduction of the translational diffusion coefficient of proteins in the cytoplasm of \textit{E.
+coli} when compared to the dilute condition (Figure \ref{fig:translational_diffusion})~\cite{Elowitz1999,Konopka2006}.
+Additionally, our data is in agreement with recent simulation and experimental data that show that the unspecific interactions between macromolecules in a crowded environment can destabilize them~\cite{Feig2011,miklos2011,Wang2012b}.
+Specifically, we can see that, despite the overall stability of our cytoplasm model even in the \SI{}{\micro\second} timescale, the proteins are slightly less stable in  the cytoplasm model than in dilute condition (Figures \ref{fig:structural_integrity_chain} and \ref{fig:rmsf}).
 
-the model is stable.
+Phosphate-containing metabolites FBP and ATP aggregated around Mg$^{2+}$ that were used as counter-ions for ATP and tRNA (Figure \ref{fig:tRNA_aggregation}).
+We used independent simulations of small systems to evaluate different strategies to prevent such aggregation and found that this can be done by completely protonating these molecules (Figure \ref{fig:avoiding_aggregation}).
+In the specific case of ATP and FBP modeled with GAFF and simulated in boxes where they are highly concentrated in the presence of Mg$^{2+}$, we advise that they should be completely protonated even though that is not their physiological protonation state.
+In principle it is possible to develop new parameters for phosphate and Mg that doesn't lead to aggregation.
+However this is outside of the scope of this study.
 
-however, aggregation!
-
-In the specific case of ATP and FBP modeled with GAFF in boxes containing Mg$^{2+}$, we advise that they should be completely protonated even though that is not their physiological protonation state.
-The topology and structure files we are providing for ATP and FBP on github are already protonated.
-
-In principle it is possible to develop new parameters for phosphate and Mg that doesn't lead to aggregation. However this is outside of the scope of this study and
+We are providing a Python script with a set of structure and topology files that any researcher can use to build their own \textit{E.
+coli} cytoplasm models.
+Such models can be used as tools that answer many research questions about the effects of crowding on specific systems of interest or about crowding itself.
+With the files we are providing it is possible, for example, to add a probe protein to investigate the effect of crowding on it, to add new macromolecular or small crowders, verify the effects of different temperatures and change the biomolecular concentration to increase or decrease the intensity of the crowding effect.
+We highlight that, since we have already validated it, researchers can take advantage from building their own cytoplasm models even if they don't have access to the infrastructure that is necessary to perform long molecular dynamics simulations.
+With modest computatioanal power it is possible, for example, to build many independent snapshots of the cytoplasm of \textit{E.
+coli}.
+All files are publicly available at github.com/dspoel/soup.
 

### 2019-09-30 paper/virtual_cytoplasm_v2.tex
final changes

@@
 As computational power increased, modeling and simulating large scale all-atom explicit solvent models of the prokaryote cytoplasm become possible.
-In this work, we build an atomistic model of the cytoplasm of \textit{Escherichia coli} composed by 1.5 million atoms and submit it to a total of \SI{3}{\micro\second} of molecular dynamics simulations, which is unprecedented for a system of this scale.
+In this work, we build an atomistic model of the cytoplasm of \textit{Escherichia coli} composed of 1.5 million atoms and submit it to a total of \SI{3}{\micro\second} of molecular dynamics simulations, which is unprecedented for a system of this scale.
 Our model reproduces experimental data about the diffusion coefficient and stability of macromolecules under crowded conditions.
@@
 
-The intracellular microenvironment is markedly different from the \textit{in vitro} simplification from which most of our current knowledge about the biochemical and biophysical behavior of macromolecules was derived~\cite{Feig2017a}.
+The intracellular environment is markedly different from the \textit{in vitro} simplification from which most of our current knowledge about the biochemical and biophysical behavior of macromolecules was derived~\cite{Feig2017a}.
 Instead of simple and diluted, the \textit{in vivo} microenvironment is highly heterogeneous and concentrated, leading to spatial constrains that directly affects biological activities~\citep{ostrowska2019}.
@@
 More recently, all-atom explicit solvent molecular dynamics simulations have been employed on crowded systems composed of several copies of one or a few macromolecules~\cite{Wang2017c}.
-However, these models doesn't reflect the true heterogeneity of biologically relevant crowded environments such as the cytoplasm.
+However, these models do not reflect the true heterogeneity of biologically relevant crowded environments such as the cytoplasm.
 This was addressed by the development of a detailed model for the cytoplasm of the bacterium \textit{Mycoplasma genitalium} containing 103 million atoms~\cite{Feig2015}.
@@
 
-Analyses were made GROMACS tools after removing artifacts generated by periodic boundary conditions from the trajectories.
+Analyses were done by GROMACS tools after removing artifacts generated by periodic boundary conditions from the trajectories.
 The Mean Square Displacement (MSD) analysis was performed to calculate the translational diffusion coefficient, which were extracted by a linear fit to MSD by averaging blocks with a length of $10 \,ns$~\cite{Allen1987a}.
@@
 On the other hand, there are some little differences between the dilute and crowding conditions.
-In most cases (TufA, MetE, AhpC, CspC and Ppa), proteins are slightly more flexible on the cytoplasm model (Fig.~\ref{fig:rmsf}).
+In most cases (TufA, MetE, AhpC, CspC and Ppa), proteins are slightly more flexible in the cytoplasm model (Fig.~\ref{fig:rmsf}).
 This is in agreement with data showing that nonspecific interactions with crowders can lead to destabilization of proteins~\cite{Feig2011, miklos2011, Wang2012b}.
@@
 
-In this work we built an all-atom explicit solvent cytoplasm model composed by proteins, tRNA, metabolites, inorganic ions and water that sums 1.5 million atoms (Table \ref{tbl:soup_components}).
+In this work we built an all-atom explicit solvent cytoplasm model composed of proteins, tRNA, metabolites, inorganic ions and water that sums 1.5 million atoms (Table \ref{tbl:soup_components}).
 The identity and amount of each component of the model was chosen according to experimental data about the composition of the cytoplasm of \textit{E.
@@
 
-We submitted this model to a total of \SI{3}{\micro\second} of molecular dynamics simulations using a modified AMBER FF99SB force field that was tuned to reproduce experimental data of systems composed by several macromolecules~\cite{Best2014a}.
+We submitted this model to a total of \SI{3}{\micro\second} of molecular dynamics simulations using a modified AMBER FF99SB force field that was tuned to reproduce experimental data of systems composed of several macromolecules~\cite{Best2014a}.
 Our model was validated by reproducing experimental data about the reduction of the translational diffusion coefficient of proteins in the cytoplasm of \textit{E.

### 2019-10-05 paper/virtual_cytoplasm_v2.tex
Updates to paper adding many questions.

@@
-\documentclass[journal=jacsat,manuscript=article]{achemso}
+\documentclass[journal=jcisd8,manuscript=article]{achemso}
 % \linespread{1.5}
@@
 
+\newpage
 \begin{abstract}
-Biomolecular crowding affects the biophysical and biochemical behavior of macromolecules when compared to the dilute environment present in experiments made with isolated proteins.
-Computational modeling and simulation are essential tools in the investigation of how crowding affects the structural dynamics and biological activities of macromolecules.
-As computational power increased, modeling and simulating large scale all-atom explicit solvent models of the prokaryote cytoplasm become possible.
-In this work, we build an atomistic model of the cytoplasm of \textit{Escherichia coli} composed of 1.5 million atoms and submit it to a total of \SI{3}{\micro\second} of molecular dynamics simulations, which is unprecedented for a system of this scale.
-Our model reproduces experimental data about the diffusion coefficient and stability of macromolecules under crowded conditions.
-We provide a Python script and a set of files that enables any researcher to build their own \textit{E.
-coli} cytoplasm models to address questions related to crowding.
-%For Nature, the abstract is really an introductory paragraph set
-%in bold type.  This paragraph must be ``fully referenced'' and
-%less than 180 words for Letters.  This is the thing that is
-%supposed to be aimed at people from other disciplines and is
-%arguably the most important part to getting your paper past the
-%editors.  End this paragraph with a sentence like ``Here we
-%show...'' or something similar.
+Biomolecular crowding affects the biophysical and biochemical behavior
+of macromolecules when compared to the dilute environment present in
+experiments made with isolated proteins. Computational modeling and
+simulation are useful tools to study how crowding affects the
+structural dynamics and biological properties of macromolecules. As
+computational power increased, modeling and simulating large scale
+all-atom explicit solvent models of the prokaryote cytoplasm become
+possible. In this work, we build an atomistic model of the cytoplasm
+of \textit{Escherichia coli} composed of 1.5 million atoms and submit
+it to a total of \SI{3}{\micro\second} of molecular dynamics
+simulations. The simulation model is consistent with experimental data
+about the diffusion coefficient and stability of macromolecules under
+crowded conditions. In order to stimulate further work we provide a
+Python script and a set of files that enables any researcher to build
+their own \textit{E. coli} cytoplasm models to address questions
+related to crowding.
+\end{abstract}
 
-\end{abstract}
+\newpage
 \section*{Introduction}
@@
 
-The intracellular environment is markedly different from the \textit{in vitro} simplification from which most of our current knowledge about the biochemical and biophysical behavior of macromolecules was derived~\cite{Feig2017a}.
-Instead of simple and diluted, the \textit{in vivo} microenvironment is highly heterogeneous and concentrated, leading to spatial constrains that directly affects biological activities~\citep{ostrowska2019}.
-This effect, called macromolecular crowding, directly influences the stability of macromolecules through an equilibrium between stabilization of their native structure due to the entropic component given by the excluded volume effect~\cite{cheung2005} and destabilizationg due to enthalpic component given by unspecific interactions with the crowders~\cite{Feig2011,miklos2011,Wang2012b}.
-Additionally, crowding affects the structure and dynamics of water around proteins, which has an indirect but important effect in their biological activities~\cite{Harada2012a,king2013}.
+The intracellular environment is markedly different from the
+\textit{in vitro} model systems from which most of our knowledge about
+the biochemical and biophysical behavior of macromolecules has been
+derived~\cite{Feig2017a}. The \textit{in vivo} microenvironment is
+heterogeneous with high concentrations of many different compounds,
+leading to spatial constrains that directly affects biological
+activities~\citep{ostrowska2019}. This effect, called macromolecular
+crowding, influences the stability of {\bf SOME/MANY} macromolecules
+because the equilibrium is shifted towards a stable native structure
+as a result of the entropic component given by the excluded volume
+effect~\cite{cheung2005}. At the same time, an effective
+destabilization due to the enthalpic component of the Gibbs energy has
+been observed, due to replacing interactions with water by weaker
+interactions with the
+crowders~\cite{Feig2011,miklos2011,Wang2012b}. Additionally, crowding
+affects the structure and dynamics of water around proteins, which has
+a significant effect on their biological
+activities~\cite{Harada2012a,king2013}.
 
-While the structure and dynamics of biomolecules is well characterized \textit{in vitro}, the understanding of the influences of crowded environments \textit{in vivo} are still evolving.
-Techniques such as nuclear magnetic resonance~\cite{reckel2007,pielak2008} and fluorescence spectroscopy~\cite{ignatova2004,xie2008,English2011} are among the most relevant to the field due to their ability to probe cells in a non destructive way.
-Another essential class of methods come from computational modeling and simulation.
-By extrapolating the advances in computational power in the last decades to the future, it was estimated that the atomistic simulation of an entire \textit{Escherichia coli} bacterium for \SI{1}{\nano\second} will be possible in 15 years from now~\cite{vanGunsteren2006a}.
+While the structure and dynamics of biomolecules is well characterized
+\textit{in vitro}, the understanding of the influences of crowded
+environments \textit{in vivo} are still evolving.  Techniques such as
+nuclear magnetic resonance~\cite{reckel2007,pielak2008} and
+fluorescence spectroscopy~\cite{ignatova2004,xie2008,English2011} are
+among the most relevant to the field due to their ability to probe
+cells in a non destructive way. Another essential class of methods
+come from computational modeling and simulation. By extrapolating the
+advances in computational power in the last decades to the future, it
+was estimated that the atomistic simulation of an entire
+\textit{Escherichia coli} bacterium for \SI{1}{\nano\second} will be
+possible in 15 years from now~\cite{vanGunsteren2006a}. {\bf THAT WAS
+  13 YEARS AGO. WHAT IS THE STATE OF THE ART?}
 
-As the available computational power increases, models become increasingly complex and are better representations of reality.
-The most simple class of computational models apply Brownian Dynamics calculations to probe macromolecules immersed in a solution of crowders that are treated as hard spheres with varying radii to represent different types of molecules~\cite{Ando2010}.
-These models usually use an implicit representation of the solvent~\cite{Mcguffee2010}.
-More recently, all-atom explicit solvent molecular dynamics simulations have been employed on crowded systems composed of several copies of one or a few macromolecules~\cite{Wang2017c}.
-However, these models do not reflect the true heterogeneity of biologically relevant crowded environments such as the cytoplasm.
-This was addressed by the development of a detailed model for the cytoplasm of the bacterium \textit{Mycoplasma genitalium} containing 103 million atoms~\cite{Feig2015}.
-This model, as well as a smaller version with 13 million atoms, was submitted to molecular dynamics simulations~\cite{Yu2016a}.
-Despite the size of these systems, the simulation times were too short, \SI{20}{\nano\second} for the complete model and \SI{140}{\nano\second} for the smaller version, to accurately characterize dynamical effects that can take place in such a large scale.
+As the available computational power increases, models become
+increasingly complex and are better representations of reality. The
+most simple class of computational models apply Brownian Dynamics
+calculations to probe macromolecules immersed in a solution of
+crowders that are treated as hard spheres with varying radii to
+represent different types of molecules~\cite{Ando2010}. These models
+usually use an implicit representation of the
+solvent~\cite{Mcguffee2010}. More recently, all-atom explicit solvent
+molecular dynamics simulations have been employed on crowded systems
+composed of several copies of one or a few
+macromolecules~\cite{Wang2017c}. However, these models do not reflect
+the true heterogeneity of biologically relevant crowded environments
+such as the cytoplasm. This was addressed by the development of a
+detailed model for the cytoplasm of the bacterium \textit{Mycoplasma
+  genitalium} containing 103 million atoms~\cite{Feig2015}. This
+model, as well as a smaller version with 13 million atoms, was
+submitted to molecular dynamics simulations~\cite{Yu2016a}. Due to the
+size of these systems, the simulation times were very short,
+\SI{20}{\nano\second} for the complete model and
+\SI{140}{\nano\second} for a smaller version, precluding an accurate
+characterization of dynamical effects that can take place in such a
+large scale. It was noted already some years ago that even
+{microsecond} simulations may be too short to probe the dynamics of
+large complexes like viruses~\cite{Larsson2012a} since dynamics is
+slower in larger systems in general.
 
-In this work, we report an all-atom model of the cytoplasm of \textit{Escherichia coli} that was built to reflect the real biological system.
-We discuss the difficulties that arise when building such system and provide a set of Python scripts and files that anyone can use to build its own crowded systems with custom parameters and elements.
-Additionally, we performed long molecular dynamics simulations of our cytoplasm model, reaching \SI{3}{\micro\second} of total simulation time.
-This work is the first to explore the structural dynamics of an all-atom explicit solvent cytoplasm model with microsecond scale molecular dynamics simulations.
+In this work, we report an all-atom model of part of the cytoplasm of
+\textit{Escherichia coli} was built to reflect the real biological
+system, in particular as regards a realistic composition of this
+``soup''. We discuss the difficulties that arise when building such a
+system and provide a set of Python scripts and files that anyone can
+use to build its own crowded systems with custom parameters and
+compounds. Additionally, we performed long molecular dynamics
+simulations of our cytoplasm model, reaching \SI{3}{\micro\second} of
+total simulation time. This work is the first to explore the
+structural dynamics of an all-atom explicit solvent cytoplasm model
+with microsecond scale molecular dynamics simulations. {\bf IS THAT
+  REALLY TRUE? A REFEREE MAY GET ANGRY WHEN READING THIS.}
 
@@
 \subsection{Initial structures}
-The proteins and tRNA were downloaded from Protein Data Bank (PDB).
-We looked for the proteins structures that were either from {\em E.
-coli} or its closest homologues (Table \ref{tbl:protein_fraction}).
-In case of PDB IDs 1U22 (MetE protein) and 2EIP (Ppa protein), we used a loop-closure modeling tool based on the Random Coordinate Descent (RCD) method~\cite{Chys2013} to model missing residues.
-The four metabolites were parametrized using the General Amber Force Field (GAFF)~\cite{sprenger2015general} and Antechamber~\cite{Wang2005b} using RESP charges~\cite{Bayly1993a} that were derived from electron density calculated at the HF/6-31G* theory level with Gaussian16.
+The protein and tRNA structures were downloaded from the Protein Data
+Bank (PDB)~\cite{pdb}. Proteins structures that were either from {\em
+  E. coli} or its closest homologues (Table
+\ref{tbl:protein_fraction}) were selected for this work. In case of
+PDB IDs 1U22 (MetE protein) and 2EIP (Ppa protein), we used a
+loop-closure modeling tool based on the Random Coordinate Descent
+(RCD) method~\cite{Chys2013} to model missing residues. The four
+metabolites were parametrized using the General Amber Force Field
+(GAFF)~\cite{Wang2004a} and Antechamber~\cite{Wang2005b} using RESP
+charges~\cite{Bayly1993a} that were derived from electron density
+calculated at the HF/6-31G* theory level with Gaussian16.  Properties
+of compounds modeled using the GAFF have been evaluated in a number of
+papers~\cite{Caleman2012a,sprenger2015general,Fischer2015a,JZhang2015a,JZhang2015,Spoel2018a}
+and strengths and weaknesses are well understood. Here, GAFF is used
+for compatibility with the Amber99SB-ws force field~\cite{Best2014a}
+used for the biomolecules. {\bf DESCRIBE WHICH MODELS WERE USED FOR
+  THE IONS.}
 
 \subsection{General simulation setup}
-The cytoplasm model was built at 30\% biomolecular mass fraction at the physiological salt concentration of 0.15 mol/L KCl.
-All simulations were performed with the GROMACS 2018 package~\cite{Abraham2015a} with Amber99SB-ws~\cite{Best2014a}, a modified version of the Amber99SB force field, in combination with the TIP4P/2005 water model~\cite{Abascal2005b}.
-All interactions were calculated explicitly inside a 1 nm radius and long range electrostatic interactions were treated using the particle mesh Ewald algorithm~\cite{Essmann1995a}.
-Corrections due to long range effects of dispersion interactions were made to both energy and pressure.
-All chemical bonds were constrained at their equilibrium length using the LINCS algorithm~\cite{Hess2008b} allowing an integration time step of 2 fs.
-Temperature was kept at 310 K by the v-rescale thermostat~\cite{Bussi2007a} with a coupling constant of 0.5 ps.
-Pressure was kept at 1 bar by the Berendsen barostat~\cite{Berendsen1984a} with the coupling constant of 10.0 ps during thermalization steps and by the Parrinello-Rahman barostat~\cite{Parrinello1981a} with the same coupling constant during production steps.
+The cytoplasm model was built at a 30\% biomolecular mass fraction and
+a physiological salt concentration of 0.15 mol/L KCl.  {\bf DESCRIBE
+  BRIEFLY HOW COMPOUNDS WERE PLACED IN THE BOX AND THAT THE VOLUME WAS
+  WAY LARGER. DO ALSO WRITE WHETHER MOLECULES WERE ROTATED randomly
+  PRIOR TO PLACEMENT IN THE BOX. ALTERNATIVELY REFER TO RESULTS.}  All
+simulations were performed with the GROMACS 2018
+package~\cite{Abraham2015a,Pronk2013a} using the Amber99SB-ws force
+field~\cite{Best2014a}, a modified version of the Amber99SB force
+field, in combination with the TIP4P/2005 water
+model~\cite{Abascal2005b}. All interactions were calculated explicitly
+inside a 1 nm radius and long range electrostatic interactions were
+treated using the particle mesh Ewald
+algorithm~\cite{Essmann1995a}. Corrections due to long range effects
+of dispersion interactions were made to both energy and
+pressure~\cite{Allen1987a}. All chemical bonds were constrained at
+their equilibrium length using the parallel LINCS
+algorithm~\cite{Hess1997a,Hess2008b} allowing an integration time step
+of 2 fs. Temperature was kept at 310 K by the v-rescale
+thermostat~\cite{Bussi2007a} with a coupling constant of 0.5
+ps. Pressure was kept at 1 bar using the Berendsen
+barostat~\cite{Berendsen1984a} with the coupling constant of 10.0 ps
+during thermalization steps and by the Parrinello-Rahman
+barostat~\cite{Parrinello1981a} with the same coupling constant during
+production steps.
 
-All systems were initially submitted to energy minimization with the steepest descent algorithm until convergence to simple precision.
-For the cytoplasm model, a short shrinking step of 500 ps was performed in order to allow the box volume to adjust to its optimal value.
-For all systems, initial velocities were sampled from a Maxwell-Boltzmann distribution at 310 K and we ran a 1 ns NVT thermalization step followed by 10 ns of NPT equilibration.
-Production runs for the dilute systems, in which each crowder is simulated alone in a water box, had the length of 200 ns.
-The cytoplasm model was run for 1 \SI{}{\micro\second}.
-All systems were simulated in three replicas.
-While the replicas for the diluted systems were generated by different initial velocities, each replica of the cytoplasm model is a completely different box in which the components of the model are in different positions and orientations.
+All systems were initially submitted to energy minimization with the
+steepest descent algorithm until convergence to machine (single)
+precision. For the cytoplasm model, a short shrinking step of 500 ps
+was performed in order to allow the box volume to adjust to its
+optimal value. For all systems, initial velocities were sampled from a
+Maxwell-Boltzmann distribution at 310 K and we ran a 1 ns NVT
+thermalization step followed by 10 ns of NPT equilibration. Production
+runs for the dilute systems, in which each crowder is simulated alone
+in a water box, had a- length of 200 ns. The cytoplasm model was run
+for 1 \SI{}{\micro\second}. All systems were simulated in three
+replicas. While the replicas for the diluted systems were generated by
+different initial velocities, each replica of the cytoplasm model is a
+completely different box in which the components of the model are in
+different positions and orientations.
 
 \subsection{Analyses}
-For statistics, the average and standard error of the three trajectories are reported.
-For values that are specific to each chain, we used the average over the individual chains in each trajectory of the replicas.
-Additionally, since there is more than one copy of some crowders in the cytoplasm models, average values and standard errors are calculated considering each copy and each replica independently.
-Similarly to the dilute system, separate chains of each copy of the crowders were considered as independent samples when running analyses that focus on one single chain.
-The first half of all simulations was discarded as equilibration time and analyses used only the second half.
+For statistics, the average and standard error {\bf standard error or
+  deviation?} of the three trajectories are reported. For values that
+are specific to each chain, we used the average over the individual
+chains in each trajectory of the replicas. Additionally, since there
+is more than one copy of some crowders in the cytoplasm models,
+average values and standard errors are calculated {\bf DO YOU MEAN THE
+  AVERAGE OVER ALL COPIES? considering each copy and each replica}
+independently. As in the dilute system, separate chains of each copy
+of the crowders were considered as independent samples when running
+analyses that focus on one single chain. The first half of all
+simulations was discarded as equilibration time and analyses used only
+the second half.
 
-Analyses were done by GROMACS tools after removing artifacts generated by periodic boundary conditions from the trajectories.
-The Mean Square Displacement (MSD) analysis was performed to calculate the translational diffusion coefficient, which were extracted by a linear fit to MSD by averaging blocks with a length of $10 \,ns$~\cite{Allen1987a}.
-In principle, diffusion coefficient needs to be corrected for finite size effects~\cite{Yeh2004} but due to relatively large simulation boxes this correction is negligible.
-Root Mean Square Deviation (RMSD) and Root Mean Square Fluctuation (RMSF) calculations considered only C$\upalpha$ atoms and the reference frame for RMSF was the midpoint of the time interval used for analysis in each trajectory.
-The chain interface area of a given oligomer in a given frame was calculated as the difference between the solvent exposed surface area (SASA) of the oligomer in that frame and the sum of the SASAs of each of its chains independently in that same frame.
-Other analyses, such as box volume and Radial Distribution Function (RDF), followed standard procedures.
+Analyses were done using GROMACS tools after removing artifacts
+generated by periodic boundary conditions from the trajectories. A
+Mean Square Displacement (MSD) analysis was performed to calculate the
+translational diffusion coefficient, which were extracted by a linear
+fit to MSD by averaging blocks with a length of {\bf 10 ns even for
+  proteins?}  $10\,ns$~\cite{Allen1987a}. In principle, diffusion
+coefficient needs to be corrected for finite size
+effects~\cite{Yeh2004} but due to relatively large simulation boxes
+this correction is negligible {\bf NOT FOR PROTEINS THAT DIFFUSE
+  SLOWLY!}. Root Mean Square Deviation (RMSD) and Root Mean Square
+Fluctuation (RMSF) calculations considered the C$\upalpha$ atoms only
+and the reference frame for RMSF was the midpoint of the time interval
+{\bf ARE YOU SURE? THE DEFAULT IS TO TAKE THE AVERAGE STRUCTURE. OR DO
+  YOU MEAN THE STRUCTURE FOR FITTING?} used for analysis in each
+trajectory. The chain interface area of a given oligomer in a given
+frame was calculated as the difference between the solvent exposed
+surface area (SASA) of the oligomer in that frame and the sum of the
+SASAs of each of its chains independently in that same frame. Other
+analyses, such as box volume and Radial Distribution Function (RDF),
+followed standard procedures.
 
@@
 \subsection{Cytoplasm model}
-In this section we describe the rationale behind the composition of our model, which has five fractions: protein, RNA, metabolites, water and ions.
-We highlight that we didn't add lipids and DNA to our model because we are considering only elements that are free to diffuse through the cytosol.
-We gathered data from several sources in order to build a computational model that is representative of the cytoplasm of {\em Escheria coli}~\cite{Dong1996,Bennett2009,Link1997,Mcguffee2010}.
+In this section we describe the rationale behind the composition of
+our model, which consists of five fractions: protein, RNA, metabolites, water
+and ions. It should be noted that we did not add lipids and DNA to our
+model because we consider only elements that are free to diffuse
+through the cytosol. We gathered data from several sources in order to
+build a computational model that is representative of the cytoplasm of
+{\em Escheria coli}~\cite{Dong1996,Bennett2009,Link1997,Mcguffee2010}.
 
 \subsubsection{Protein fraction}
-We selected a group of eight proteins that account for 50\% of the abundance of non-ribosomal proteins in the cytoplasm of {\em Escherichia coli} \cite{Link1997}.
-The two most abundant proteins, TufA and MetE, account for about 20\% and 12\% of the total protein abundance in {\em E.
-coli}, respectively, while the other six proteins contribute with less than 5\% each (Table~\ref{tbl:protein_fraction}).
-Their crystallographic structures were used to insert them in the cytoplasm model.
-The number of copies of each protein in the model was calculated as the rounded up fold increase in their abundance when compared to the least abundant protein, which was set to have the copy number of 1.
-Then, this number was divided by the oligomeric state of each protein.
-This was done because the original oligomeric state for each protein was kept as reported in their crystallographic structure (Table~\ref{tbl:protein_fraction}).
+Eight different proteins were selected that together account for 50\%
+of the non-ribosomal proteins in the cytoplasm of {\em Escherichia
+  coli} \cite{Link1997}. The two most abundant proteins, TufA and
+MetE, account for about 20\% and 12\% of the total protein abundance
+in {\em E. coli}, respectively, while the other six proteins
+contribute with less than 5\% each (Table~\ref{tbl:protein_fraction}).
+The least abundant protein was present with 1 copy and the amount of
+each of the other proteins was taken to be an integer number of
+oligomers. This was done because the original oligomeric state for
+each protein was kept as reported in their crystallographic structure
+(Table~\ref{tbl:protein_fraction}).
 
@@
 \centering
-\begin{tabular}{ l c c c}
+\caption{Protein, PDB ID, number N of proteins in the simulation. 
+Fraction of the total abundance of non-ribosomal proteins in the cytosol of {\em
+    E. coli} K12 and oligomeric state of each protein that was
+  selected to compose our cytoplasm model. {\bf WRITE OUT WHOLE
+    PROTEIN NAMES}}
+\label{tbl:protein_fraction}
+\begin{tabular}{lcccc}
 \hline
-\multirow{2}{*}{Protein} & 	\multicolumn{2}{c}{Fraction [\%]} & \multirow{2}{*}{Oligomeric State}\\
-& Absolute & Cumulative \\ 
+Protein & PDB & N & Fraction [\%] & State\\
 \hline
-TufA & 19.7 & 19.7 & 2-mer \\
-MetE & 11.6 & 31.4 & 1-mer \\
-IcdA & 4.7  & 36.1 & 2-mer \\
-AhpC & 4.1  & 40.2 & 10-mer \\
-CspC & 4.0  & 44.2 & 1-mer \\
-Ppa  & 2.9  & 47.0 & 6-mer \\
-GapA & 2.1  & 49.1 & 4-mer \\
-Eno  & 1.9  & 51.1 & 2-mer \\
+TufA & 1DG1~\cite{Abel1996}       & 6 & 19.7 &  dimer \\
+MetE & 1U22~\cite{Ferrer2004}     & 7 & 11.6 &  monomer \\
+IcdA & 1P8F~\cite{Mesecar2000}    & 2 & 4.7  &  dimer \\
+AhpC & 1YEP~\cite{Parsonage2005}  & 1 & 4.1  &  decamer \\
+CspC & 1MJC~\cite{Schindelin1994} & 3 & 4.0  &  monomer \\
+Ppa  & 2EIP~\cite{Kankare1996}    & 1 & 2.9  &  hexamer \\
+GapA & 1S7C~\cite{ShinXXX}        & 1 & 2.1  &  tetramer \\
+Eno  & 1E9I~\cite{Kuhnel2001}     & 1 & 1.9  &  dimer \\
+\hline
+Total &                           & 48& 51.1 & \\
 \hline
 \end{tabular}
-\caption{Absolute and cumulative fraction of the total abundance of non-ribosomal proteins in the cytosol of {\em E.
-coli} K12 and oligomeric state of each protein that was selected to compose our cytoplasm model.}
-\label{tbl:protein_fraction}
 \end{table}
@@
 \subsubsection{RNA fraction}
-Transporter RNAs (tRNAs) account for 74\% of the dry weight of non-ribosomal RNAs~\cite{phillips2012}.
-Thus, we chose to model the RNA presence in the cytoplasm with tRNA molecules.
-Specifically, we considered the tRNA(Phe) molecule as a representative of tRNAs due to the availability of a recent crystallographic structure~\cite{Byrne2015}.
-The protein and RNA content of the total dry weight of \textit{E.coli} is 55\% and 2.9\%, respectively~\cite{phillips2012}.
-That is, the total RNA weight corresponds to 5\% of the total protein weight.
-This protein/RNA weight ratio was used to calculate the correct number of tRNA(Phe) molecules that were added to the cytoplasm model \tabref{tbl:soup_components}.
+Transporter RNAs (tRNAs) account for 74\% of the dry weight of
+non-ribosomal RNAs~\cite{phillips2012}. Thus, we chose to model the
+RNA presence in the cytoplasm with tRNA molecules. Specifically, 
+ the tRNA(Phe) molecule as a representative of tRNAs due to
+the availability of a recent crystal
+structure (4YCO~\cite{Byrne2015}). The protein and RNA content of the total
+dry weight of \textit{E.coli} is 55\% and 2.9\%,
+respectively~\cite{phillips2012}. That is, the total RNA weight
+corresponds to 5\% of the total protein weight. This protein/RNA
+weight ratio was used to calculate the correct number of tRNA(Phe)
+molecules that were added to the cytoplasm model
+\tabref{tbl:soup_components}.
 
 \subsubsection{Metabolites fraction}
-We considered the most abundant molecules of each metabolite class as representatives, i.e.
-Glutamate (GLT) for amino acids, ATP for nucleotides, Fructose-2,6-Biphosphate (FBP) for central carbon intermediates and Glutathione (GSH) redox cofactors~\cite{Bennett2009}.
-The total number of molecules was calculated considering data showing that the number of metabolite molecules in the cytoplasm of \textit{E.
-coli} is about 42.86 times higher than the number of proteins~\cite{Bennett2009}.
-The copy number for each molecule was calculated from the ratios between their experimentally observed concentration in {\em E.
-coli}~\cite{Bennett2009}.
+The most abundant molecules of each metabolite class was used to 
+represent the class, i.e. Glutamate (GLT) for amino acids, ATP for
+nucleotides, Fructose-2,6-Biphosphate (FBP) for central carbon
+intermediates and Glutathione (GSH) redox
+cofactors~\cite{Bennett2009}. The total number of molecules was
+calculated considering data showing that the number of metabolite
+molecules in the cytoplasm of \textit{E. coli} is $\approx$ 43 times
+higher than the number of proteins~\cite{Bennett2009}. The copy number
+for each molecule was calculated from the ratios between their
+experimentally observed concentration in {\em
+  E. coli}~\cite{Bennett2009}.
 
 \subsubsection{Water fraction}
-The number of water molecules was calculated according to the desired biomolecular concentration, which is a parameter of the cytoplasm model.
-This number ranges from 300 to 400 g/L in biological systems such as \textit{E.
-coli} cytoplasm~\cite{Zimmerman1991}, which corresponds to biomolecular concentrations as high as 40\%~\cite{Ellis2003a}.
-In our case, we choose the biomolecular concentration of 30\%, that is, the number of molecules necessary to reach a ratio of total biomolecular mass to water mass of 30\% was inserted into the cytoplasm model.
+The number of water molecules was calculated according to the desired
+biomolecular concentration, which is a parameter of the cytoplasm
+model. This number ranges from 300 to 400 g/L in biological systems
+such as \textit{E. coli} cytoplasm~\cite{Zimmerman1991}, which {\bf
+  WHAT DOES WHICH CORRESPOND TO?}  corresponds to biomolecular
+concentrations as high as 40\%~\cite{Ellis2003a}. In our case, we
+choose a biomolecular concentration of 30\%, that is, the number of
+molecules necessary to reach a ratio of total biomolecular mass to
+water mass of 30\% was inserted into the cytoplasm model.
 
 \subsubsection{Inorganic ions fraction}
-Finally, Mg\textsuperscript{2+} was used as counter-ions for tRNA and ATP.
-K$^{+}$ and Cl$^{-}$ were added to neutralize the charges of the simulation box and to reach the ionic strength of 0.150 mol/L by substitution of randomly selected water molecules.
+Finally, Mg\textsuperscript{2+} was used as counter-ions for tRNA and
+ATP. K$^{+}$ and Cl$^{-}$ were added to neutralize the charges of the
+simulation box and to reach the ionic strength of 0.150 mol/L by
+substitution of randomly selected water molecules.
 
@@
 \hline
-Class & Name (PDB ID) & Number\\
+Class & Name & Number\\
 \hline
-Protein & TufA (1DG1~\cite{Abel1996}) & 6\\
-  & MetE (1U22~\cite{Ferrer2004}) & 7\\
-  & IcdA (1P8F~\cite{Mesecar2000}) & 2\\
-  & AhpC (1YEP~\cite{Parsonage2005}) & 1\\
-  & CspC (1MJC~\cite{Schindelin1994}) & 3\\
-  & Ppa (2EIP~\cite{Kankare1996}) & 1\\
-  & GapA (1S7C~\cite{ShinXXX}) & 1\\
-  & Eno (1E9I~\cite{Kuhnel2001}) & 1\\
-\hline
-RNA & tRNA$^{\text{Phe}}$ (4YCO~\cite{Byrne2015}) & 5\\
+RNA & tRNA$^{\text{Phe}}$ & 5\\
 \hline
@@
 \end{tabular}
-\caption{Number of copies for each component of the cytoplasm model built at the biomolecular fraction of 30\%.}
+\caption{Number of copies for non-protein components of the cytoplasm model built at the biomolecular fraction of 30\%.}
 \label{tbl:soup_components}
@@
 \subsection{Building the simulation box}
-All components can be put in the same simulation box by inserting each of them in random orientations in a cubic volume of side $L$ that is initially empty.
-However, that is not a trivial process.
-We need to use a box size that is big enough to allow the random insertions to succeed without structural overlapping and without creating artificial interactions between the elements by placing them too close from each other.
-On the other hand, as the box gets bigger, it gets harder to equilibrate the system because the empty space between the components will induce the barostat to reduce the box volume quickly during the simulation.
+All components can be put in the same simulation box by inserting each
+of them in random orientations in a cubic volume of side $L$ that is
+initially empty. However, that is not a trivial process. We need to
+use a box size that is big enough to allow the random insertions to
+succeed without structural overlapping and without creating artificial
+interactions between the elements by placing them too close from each
+other. On the other hand, as the box gets bigger, it gets harder to
+equilibrate the system because the empty space between the components
+will induce the barostat to reduce the box volume quickly during the
+simulation.
 
-We devised an iterative process that solves both problems simultaneously.
-We start with a box size $L$ that is too small to allow all components to fit in the box by random insertion.
-In our case, we start with $L = 30$ nm.
-Then, we allow 100 insertion trials for each element.
-If all trials fail for any of them, we start the whole process again with an empty cubic box that is larger by a step size $dL$ of 1 nm.
-We repeat this process until all insertions succeed.
-In our case, all insertions succeeded after increasing $L$ to 35 nm.
-Additionally, instead of adding only the protein, tRNA or metabolite molecules in the empty box in each trial, we actually add a droplet of water and counter ions in which the molecule of interest is embedded.
-Such droplets are taken from molecular dynamics simulations in which each component was previously equilibrated.
-The benefits of using such droplets are threefold: \textit{i.} it acts as natural protecting layer that prevents artificial contacts between the components that could be created due to the random insertions.
-\textit{ii.} it is a natural way to place water molecules and counter-ions in the simulation box around each component.
-\textit{iii.} the components are already pre-equilibrated, which will help us to perform the equilibration of the whole cytoplasm model.
+We devised an iterative process that solves both problems
+simultaneously. We start with a box size $L$ that is too small to
+allow all components to fit in the box by random insertion. In our
+case, we start with $L = 30$ nm. Then, we allow 100 insertion trials
+for each element. If all trials fail for any of them, we start the
+whole process again with an empty cubic box that is larger by a step
+size $dL$ of 1 nm. We repeat this process until all insertions
+succeed. In our case, all insertions succeeded after increasing $L$ to
+35 nm. Additionally, instead of adding only the protein, tRNA or
+metabolite molecules in the empty box in each trial, we actually add a
+droplet of water and counter ions in which the molecule of interest is
+embedded. Such droplets are taken from molecular dynamics simulations
+in which each component was previously equilibrated. The benefits of
+using such droplets are threefold: \textit{i.} it acts as natural
+protecting layer that prevents artificial contacts between the
+components that could be created due to the random
+insertions. \textit{ii.} it is a natural way to place water molecules
+and counter-ions in the simulation box around each
+component. \textit{iii.} the components are already pre-equilibrated,
+which will help us to perform the equilibration of the whole cytoplasm
+model.
 
-In order to do this, the droplets around each component are also constructed iteratively.
-The number of water molecules that we must place in the simulation box is known from the total biomolecular mass of the system and the desired biomolecular fraction, which is a parameter of the model.
-However, we don't know the thickness of the water layer around each component, $l$, that accounts for such amount of water molecules \textit{a priori}.
-Since $l$ depends on a series of factors such as the shape, size and abundance of each element, we define it iteratively.
-We start by taking a droplet of thickness $l$ = 3 \r{A} around all elements and counting the number of water molecules that we would add to the box with such droplets.
-If it is smaller than the number we need, $l$ is increased by $dl$ = 1 \r{A}.
-If it is larger, we reduce the droplet thickness by $dl$, reduce $dl$ by 10 times and then increase $l$ with the new smaller $dl$.
-We can carry this process until an arbitrary precision cutoff, such as 5\%, for the number of waters in the system is satisfied.
+In order to do this, the droplets around each component are also
+constructed iteratively. The number of water molecules that we must
+place in the simulation box is known from the total biomolecular mass
+of the system and the desired biomolecular fraction, which is a
+parameter of the model. However, we don't know the thickness of the
+water layer around each component, $l$, that accounts for such amount
+of water molecules \textit{a priori}. Since $l$ depends on a series of
+factors such as the shape, size and abundance of each element, we
+define it iteratively. We start by taking a droplet of thickness $l$ =
+3 \r{A} around all elements and counting the number of water molecules
+that we would add to the box with such droplets. If it is smaller than
+the number we need, $l$ is increased by $dl$ = 1 \r{A}. If it is
+larger, we reduce the droplet thickness by $dl$, reduce $dl$ by 10
+times and then increase $l$ with the new smaller $dl$. We can carry
+this process until a user-defined precision cutoff, here 5\%, for the
+number of waters in the system is satisfied.
 
-After all droplets are successfully inserted in a simulation box by the iterative process described above, we proceed to add ions to neutralize the net charge of the system and to reach the desired ionic strength.
-Then, we perform energy minimization and a short simulation step of 500ps in which all molecules of the box are free to move.
-In this step the box shrinks to its optimum size.
-In our case, the simulation box shrank from the initial box size of 35 nm to 22.9 nm, which correspond to a volume change from 42875 nm\textsuperscript{3} to 12009 nm\textsuperscript{3}.
-From this point, the system is ready to be submitted to the default simulation steps such as thermalization and production run (please check the materials section for details about the parameters we used).
+After all droplets are successfully inserted in a simulation box by
+the iterative process described above, we proceed to add ions to
+neutralize the net charge of the system and to reach the desired ionic
+strength. Then, we perform energy minimization and a short simulation
+ of 500ps in which all molecules of the box are free to move. In
+this step the box shrinks to its optimum size. In our case, the
+simulation box shrank from the initial box size of 35 nm to 22.9 nm,
+which corresponds to a volume change from 42875 nm\textsuperscript{3}
+to 12009 nm\textsuperscript{3}. From this point, the system is ready
+to be submitted to the default simulation steps such as thermalization
+and production run (please check the materials section for details
+about the parameters used).
 
@@
 \includegraphics[scale=0.5]{shrinking.pdf} 
-\caption{A) The initial volume of the simulation box, $L=35.0$ nm, and after the shrinking step, $L=22.9$ nm.
-B) The evolution of the volume of the simulation box during the shrinking simulations.
-Each replica is shown in different colors.
-}
+\caption{A) The initial simulation box with $L=35.0$ nm and the box
+  after the shrinking step, $L=22.9$ nm. B) The evolution of the
+  volume of the simulation box during the shrinking simulations. Each
+  replica is shown in different colors. }
 \end{figure}
@@
 
-In order to investigate the effects of crowding in the \textit{E.
-coli} cytoplasm model on the structural integrity and dynamics of its elements, we constructed three completely independent cytoplasm models that have different orientations for each of its elements.
-We submitted these systems, each composed of more than 1.5 million atoms, to molecular dynamics simulations of \SI{1}{\micro\second}.
-We also performed \SI{200}{\nano\second} molecular dynamics simulations for each isolated element (proteins and tRNA) in order to represent the non-crowded, i.e.
-dilute environment (please check the materials section for details about the parameters).
+In order to investigate the effects of crowding in the
+\textit{E. coli} cytoplasm model on the structural integrity and
+dynamics of its elements, we constructed three completely independent
+cytoplasm models that have different orientations for each of its
+elements. We submitted these systems, each composed of more than 1.5
+million atoms, to molecular dynamics simulations of
+\SI{1}{\micro\second}. We also performed \SI{200}{\nano\second}
+molecular dynamics simulations for each isolated element (proteins and
+tRNA) in order to represent the non-crowded, i.e. dilute environment
+(please check the materials section for details about the parameters).
 
@@
 
-Under crowded conditions, molecules are constrained to a smaller volume by the other components of the crowded environment.
-The extent of this effect is evident when we compare the translational diffusion constant of each component in the cytoplasm model, $D_{cyto}$, and in a dilute system, $D_{dil}$ (Fig.~\ref{fig:translational_diffusion}A).
-The ratio of these quantities is independent of the protein size and is always close to 0.13 (Fig.~\ref{fig:translational_diffusion}B).
-However, the effect of crowding on the translational diffusion of tRNA is two times larger.
-Its $D_{cyto}/D_{dil}$ is 0.07, indicating that its movement is specially constrained in the crowded environment (Fig.~\ref{fig:translational_diffusion}B, crowder with 25 kDa).
+Under crowded conditions, molecules are confined to a smaller
+effective volume by the other components of the environment.  The
+extent of this effect is evident when we compare the translational
+diffusion constant of each component in the cytoplasm model,
+$D_{cyto}$, and in a dilute system, $D_{dil}$
+(Fig.~\ref{fig:translational_diffusion}A). The ratio of these
+quantities is independent of the protein size and is always close to
+0.13 (Fig.~\ref{fig:translational_diffusion}B). However, the effect of
+crowding on the translational diffusion of tRNA is two times
+larger. Its $D_{cyto}/D_{dil}$ is 0.07, indicating that its movement
+is especially constrained in the crowded environment
+(Fig.~\ref{fig:translational_diffusion}B, crowder with 25 kDa),
+however, see below.
 
@@
 \includegraphics[scale=0.5]{diff_cyto_over_singles.pdf}
-\caption{A) Diffusion coefficient of crowders from simulations under diluted conditions (blue bars) and cytoplasm simulations (orange bars).
-B) The ratio between diffusion coefficient obtained from simulations under both conditions.
-The crowders are sorted according to their sizes on the x-axis.}
+\caption{A) Diffusion coefficient of crowders from simulations under
+  diluted conditions (blue bars) and cytoplasm simulations (orange
+  bars). B) The ratio between diffusion coefficient obtained from
+  simulations under both conditions. The crowders are sorted according
+  to their sizes on the x-axis.}
 \label{fig:translational_diffusion}
@@
 \subsubsection{tRNA is aggregating}
-After inspecting the trajectories of the cytoplasm models, we found that the reason why the translational diffusion constant for tRNA was reduced to a greater extent than for the other components of the cytoplasm model is that it is forming aggregates with Mg$^{2+}$, ATP and FBP (Fig.~\ref{fig:tRNA_aggregation}).
-Thus, we will omit the tRNA molecules in the analyses about the structural integrity of the crowders.
-In the next sections we will further investigate such aggregation and we will show a way to prevent it.
+After inspecting the trajectories of the cytoplasm models, we found
+that the reason why the translational diffusion constant for tRNA was
+reduced to a greater extent than for the other components of the
+cytoplasm model is that it was forming aggregates with Mg$^{2+}$, ATP
+and FBP (Fig.~\ref{fig:tRNA_aggregation}). Thus, we will omit the tRNA
+molecules in the analyses about the structural integrity of the
+crowders. In the next sections we will further investigate such
+aggregation and we will show a way to prevent it.
 
@@
 \includegraphics[scale=0.5]{rdf_RNA_metabolites.pdf}
-\caption{Radial distribution function showing the probability of finding metabolites or Magnesium cations around tRNA molecules in the cytoplasm model.
-Note that ATP and FBP are much more concentrated around tRNA than the other metabolites, GSH and GLT.}
+\caption{Radial distribution function showing the probability of
+  finding metabolites or Magnesium cations around tRNA molecules in
+  the cytoplasm model. Note that ATP and FBP are much more
+  concentrated around tRNA than the other metabolites, GSH and GLT.}
 \label{fig:tRNA_aggregation}
@@
 \subsubsection{Structural integrity}
-The structure of the individual chains of the crowders that were used in our cytoplasm model were not affected to a great extent by the crowded environment.
-The value of the root mean square deviation (RMSD) considering only their C$\upalpha$ in the cytoplasm model is withing 11\% of its value under dilute conditions, except for TufA, MetE and IcdA, which have RMSD values 105, 25 and 33\% higher in the crowded environment than in the dilute condition.
-Visual inspection of their trajectories showed that there is partial unfolding of their N- or C-terminal residues.
+The structure of the individual chains of the crowders that were used
+in our cytoplasm model were not affected to a great extent by the
+crowded environment. The value of the root mean square deviation
+(RMSD, Fig.~\ref{fig:structural_integrity_chain}A) considering only
+their C$\upalpha$ in the cytoplasm model is within 11\% of its value
+under dilute conditions {\bf WHAT DOES WITHIN 11\% MEAN?}, except for
+TufA, MetE and IcdA, which have RMSD values 105, 25 and 33\% higher in
+the crowded environment than in the dilute condition. {\bf ARE THERE
+  RMSD PLOTS? REFER TO THEM.} Visual inspection of their trajectories
+showed that there is partial unfolding of their N- or C-terminal
+residues. {\bf IS THERE A KNOWN PHYSIOLOGICAL REASON FOR THAT?. IS
+  MetE A MULTIDOMAIN PROTEIN WHERE DOMAINS MOVE WITH RESPECT TO EACH
+  OTHER?}
 
-The effect of crowded environment on the structural integrity of oligomers can be evaluated by their chain interface area.
-This quantity is the solvent accessible surface area that is removed from access to the solvent by oligomerization.
-Our results show that oligomers are not significantly disturbed in the cytoplasm model when compared to the diluted condition.
-Similarly to what can be seem by the RMSD values, which reflect the structural integrity of individual chains, the oligomers that undergo the most change is TufA (Fig.~\ref{fig:structural_integrity_chain}B).
+The effect of crowded environment on the structural integrity of
+oligomers can be evaluated by their chain interface area as well. This
+quantity is the solvent accessible surface area that is removed from
+access to the solvent by oligomerization. Our results show that
+oligomers are not significantly disturbed in the cytoplasm model when
+compared to the diluted condition. Similarly to what can be seem by
+the RMSD values, which reflect the structural integrity of individual
+chains, the oligomers that undergo the most change is TufA
+(Fig.~\ref{fig:structural_integrity_chain}B).
 
@@
 \includegraphics[scale=0.5]{sasa.pdf}
-\caption{A) RMSD for all protein crowders from simulations under diluted conditions (blue bars) and in the cytoplasm model (orange bars).
-This value reflects the structural integrity of individual chains of each crowder.
-B) Chain interface area for oligomers under diluted conditions (blue bars) and in the cytoplasm model (orange bars).
-This value reflects the structural integrity of oligomers of the crowders.}
+\caption{A) RMSD for all protein crowders from simulations under
+  diluted conditions (blue bars) and in the cytoplasm model (orange
+  bars). This value reflects the structural integrity of individual
+  chains of each crowder. B) Chain interface area for oligomers under
+  diluted conditions (blue bars) and in the cytoplasm model (orange
+  bars). This value reflects the structural integrity of oligomers of
+  the crowders.}
 \label{fig:structural_integrity_chain}
@@
 \subsubsection{Structural dynamics}
-Proteins are slightly more flexible in the crowded environment.
-Root mean square fluctuation (RMSF) calculations show that the flexibility profile of the proteins is not affected by crowding, which is in agreement with our results showing that the structural integrity of the proteins is not significantly affected in the cytoplasm model (Fig.~\ref{fig:rmsf}).
-On the other hand, there are some little differences between the dilute and crowding conditions.
-In most cases (TufA, MetE, AhpC, CspC and Ppa), proteins are slightly more flexible in the cytoplasm model (Fig.~\ref{fig:rmsf}).
-This is in agreement with data showing that nonspecific interactions with crowders can lead to destabilization of proteins~\cite{Feig2011, miklos2011, Wang2012b}.
+{\bf THIS PARAGRAPH REPEATS AND CONTRADICTS ITSELF. IS FLEXIBILITY
+  AFFECTED OR NOT? FROM THE GRAPH I WOULD SAY THE THE FIVE PROTEINS
+  NOTED BELOW ARE MUCH MORE FLEXIBLE, NOT SLIGHTLY. ONE WAY OF
+  QUANTIFYING THIS IS TO COMPUTE THE AVERAGE OF THE MEAN SQUARE
+  FLUCTUATION OVER THE WHOLE PROTEIN, THAT IS BEFORE TAKING SQUARE
+  ROOTS OF COURSE AND PLOTTING THAT IN A BAR GRAPH. THEN THIS FIGURE
+  CAN GO TO SUPPORTING.}  Proteins are slightly more flexible in the
+crowded than in the dilute environment. Root mean square fluctuation
+(RMSF) calculations show that the flexibility profile of the proteins
+is not affected by crowding, which is in agreement with our results
+showing that the structural integrity of the proteins is not changed
+significantly in the cytoplasm model (Fig.~\ref{fig:rmsf}).  On the
+other hand, there are some small differences between the dilute and
+crowding conditions. In most cases (TufA, MetE, AhpC, CspC and Ppa),
+proteins are slightly more flexible in the cytoplasm model
+(Fig.~\ref{fig:rmsf}). This is in agreement with data showing that
+nonspecific interactions with crowders can lead to destabilization of
+proteins~\cite{Feig2011, miklos2011, Wang2012b}.
 
@@
 \includegraphics[scale=0.5]{rmsf.pdf}
-\caption{Root mean squared fluctuations (RMSF) from simulations under diluted conditions (blue bars) and in the cytoplasm model (orange bars).
-The x-axis shows the residues for each chain of proteins.}
+\caption{Root mean squared fluctuations (RMSF) from simulations under
+  diluted conditions (blue bars) and in the cytoplasm model (orange
+  bars). The x-axis shows the residues for each chain of proteins.}
 \label{fig:rmsf}
@@
 
-The initial detection of aggregation in our cytoplasm model was due to the outlier behavior of the translational diffusion constant of tRNA when compared to the other crowders.
-However, tRNA is not causing such phenomenon.
-Instead, it is triggered by the exaggerated interaction between the phosphate-containing metabolites, ATP and FBP, with Mg$^{2+}$ (Fig.~\ref{fig:avoiding_aggregation}A).
-tRNA gets involved in the aggregates because Mg$^{2+}$ is present in the box only as counter-ion for ATP and tRNA, and so tRNA is one of the few ``sources'' of Mg$^{2+}$ in the whole system.
-We looked for ways of avoid aggregation by using simulations of small systems composed of these metabolites in high concentration, Mg$^{2+}$ and water (please check the methods section for details).
+The initial detection of aggregation in the cytoplasm model was due to
+the outlier behavior of the translational diffusion constant of tRNA
+when compared to the other crowders. However, tRNA was not the cause of this
+phenomenon. Instead, it was triggered by the exaggerated interactions
+between the phosphate-containing metabolites, ATP and FBP, with
+Mg$^{2+}$ (Fig.~\ref{fig:avoiding_aggregation}A). tRNA gets involved
+in the aggregates because Mg$^{2+}$ is present in the box only as
+counter-ion for ATP and tRNA, and so tRNA is one of the few
+``sources'' of Mg$^{2+}$ in the whole system. We looked for ways of
+avoid aggregation by using simulations of small systems composed of
+these metabolites in high concentration, Mg$^{2+}$ and water (please
+check the methods section for details).
 
-We found that completely protonating the phosphate groups of ATP and FBP is enough to prevent their aggregation with Mg$^{2+}$.
-While RDFs show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$ (Fig.~\ref{fig:avoiding_aggregation}A), this is not observed for ATPH\textsubscript{3} and FBPH\textsubscript{4}(Fig.~\ref{fig:avoiding_aggregation}B-C).
+We found that completely protonating the phosphate groups of ATP and
+FBP is enough to prevent their aggregation with Mg$^{2+}$. While RDFs
+show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$
+(Fig.~\ref{fig:avoiding_aggregation}A), this is not observed for
+ATPH\textsubscript{3} and
+FBPH\textsubscript{4}(Fig.~\ref{fig:avoiding_aggregation}B-C).  {\bf
+  WERE ANY INTERMEDIATE PROTONATIONS TESTED? E.G. FBPH2(2-) AND ATPH(2-)?
+  COMPLETELY PROTONATING IS UNPHYSICAL AS WELL AT THE PH STUDIED. FOR E COLI THE PH SHOULDBE ROUGHLY 7.4 https://en.wikipedia.org/wiki/Cytosol}
 
@@
 \hspace*{-2cm}\includegraphics[scale=0.5]{rdf_mg.pdf}
-\caption{RDFs centered around Mg$^{2+}$ showing the probability of finding A) metabolites in the cytoplasm model, B) ATP$^{-3}$ and ATPH\textsubscript{3} or C) FBP$^{-4}$ and FBPH\textsubscript{4} in small independent systems (please see the methods section for details).}
+\caption{RDFs centered around Mg$^{2+}$ showing the probability of
+  finding A) metabolites in the cytoplasm model, B) ATP$^{-3}$ and
+  ATPH\textsubscript{3} or C) FBP$^{-4}$ and FBPH\textsubscript{4} in
+  small independent systems (please see the methods section for
+  details).}
 \label{fig:avoiding_aggregation}
@@
 \section*{Discussion}\label{sec:dissc}
+The main difference between the behavior of the macromolecules that
+compose our model in the crowded system and in the dilute condition is
+their translational diffusion coefficient, which drops by around
+85\% for the proteins and 93\% for the tRNA
+(Fig.~\ref{fig:translational_diffusion}A). The slope of the line
+generated by fitting a linear regression model to values of
+$D_{cyto}/D_{dil}$ is close to zero (slope $= 0.00028$), showing that
+the drop in diffusion coefficient is independent of the size of the
+macromolecules. We found that the aggregation of ATP$^{-3}$ and
+FBP$^{-4}$ around Mg$^{2+}$, which was used as counter-ion for tRNA,
+leads to an exaggerated confinement of the tRNA molecules.  Therefor
+tRNA is considered as an outlier in the duffusion calculations. In
+fact, the slope of the linear regression line fitted to
+$D_{cyto}/D_{dil}$ is reduced by by 50\% and gets even closer to zero
+(slope $= 0.00014$) after removing the diffusion coefficient of tRNA
+from the data. These data are in agreement with experimental results
+showing that the translational diffusion coefficient of Green
+Fluorescent Protein (GFP) in {\em E. coli}, as measured by
+fluorescence recovery after photo-bleaching (FRAP), decreased 90\%
+when compared to {\em in vitro}
+measurements~\cite{Elowitz1999,Konopka2006}.
+{\bf IT MAY BE NECESSARY TO DISCUSS RMSF HERE AS WELL}
 
-The main difference between the behavior of the macromolecules that compose our model in the crowded system and in the dilute condition is their translational diffusion coefficient, which drops by more around 85\% for the proteins and 93\% for the tRNA (Fig.~\ref{fig:translational_diffusion}A).
-The slope of the line generated by fitting a linear regression model to values of $D_{cyto}/D_{dil}$ is close to zero (slope $= 0.00028$), showing that the drop in diffusion coefficient is independent of the size of the macromolecules.
-We found that the aggregation of ATP$^{-3}$ and FBP$^{-4}$ around Mg$^{2+}$, which was used as counter-ion for tRNA, leads to its exaggerated confinement.
-Thus, tRNA is considered as an outlier in our cytoplasm model.
-In fact, the slope of the linear regression line fitted to $D_{cyto}/D_{dil}$ is reduced by by 50\% and gets even closer to zero (slope $= 0.00014$) after removing the diffusion coefficient of tRNA from the data.
-Our data is in agreement with experimental results showing that the translational diffusion coefficient of Green Fluorescent Protein (GFP) in {\em E.
-coli}, as measured by fluorescence recovery after photo-bleaching (FRAP), decreased 90\% when compared to {\em in vitro} measurements~\cite{Elowitz1999,Konopka2006}.
+It has been demonstrated that the force fields that are most widely
+used to perform molecular dynamics simulations of macromolecules yield
+different results for highly concentrated amino acid
+solutions~\cite{andrews2013}. Moreover, it has been reported that
+simulations of highly concentrated environment leads to aggregation
+when additive force fields are used~\cite{Petrov2014a,Abriata2015a,
+  Nawrocki2017a} and several approaches have been proposed to
+eliminate this
+problem~\cite{Best2014a,Piana2015a,Bashardanesh2018b}. Apart from
+protein-water interactions, the London-dispersion coefficients in the
+force fields have been investigated as well but their effect on
+protein-protein interactions seems to be
+minor~\cite{Mohebifar2017a,Walters2018a,Bashardanesh2018b}.  A
+comparison at the amino acid level shows that none of the Amber family
+of force fields combined with different water model can simultaneously
+produce high accuracy Gibbs energies of hydration and amino acid side
+chain mobility~\cite{HZhang2018a}. Of the different combinations
+tested in that work the Amber99sb-ws~\cite{Best2014a} in conjunction
+with the TIP4P/2005 water model~\cite{Abascal2005b} was deemed the
+best compromise~\cite{HZhang2018a}. Nevertheless, in a related paper
+we find signs that the increased interaction between protein and water
+may be slightly too strong as it leads to partial unfolding of a
+protein~\cite{Bashardanesh2019a}.
 
-It has been demonstrated that each of the force fields that are most widely used to perform molecular dynamics simulations of macromolecules have different results for highly concentrated amino acid solutions~\cite{andrews2013}.
-Moreover, it has been reported that simulations of highly concentrated environment leads to aggregation when additive force fields are used~\cite{Petrov2014a,Abriata2015a, Nawrocki2017a} and several approaches have been proposed to eliminate this problem~\cite{Best2014a,Piana2015a, Bashardanesh2018b}.
-Previous all-atom simulations of high scale cytoplasm models used parameters from the CHARMM family of force fields~\cite{Yu2016a}.
-In this work we use a modified version of the AMBER FF99SB force field in which the solvent-solute interactions are 10\% stronger~\cite{Best2014a} to study crowding effects on a cytoplasm model.
-This force field was chosen because it was corrected to reproduce experimental data when there are many proteins in the same simulation box.
-All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of \textit{Mycoplasma genitalium} composed of 103 and 12 million atoms that were simulated for 20 and 60 \SI{}{\nano\second}, respectively~\cite{Yu2016a}.
-The cytoplasm model described in this work has 1.5 million atoms and was simulated for the total of \SI{3}{\micro\second}, an unprecedented time scale to study crowding effects on a heterogeneous cytoplasm model in atomic detail.
+Previous
+all-atom simulations of high scale {\bf WHAT DO YOU MEAN WITH HIGH
+  SCALE?} cytoplasm models used parameters from the CHARMM family of
+force fields~\cite{Yu2016a}. In this work we use a modified version of
+the AMBER FF99SB force field in which the solvent-solute interactions
+are 10\% stronger~\cite{Best2014a} to study crowding effects on a
+cytoplasm model. This force field was chosen because it was corrected
+to reproduce experimental data when there are many proteins in the
+same simulation box. All-atom explicit solvent molecular dynamics
+simulations have been reported for large-scale cytoplasm models of
+\textit{Mycoplasma genitalium} composed of 103- and 12-million atoms
+that were simulated for 20 and 60 \SI{}{\nano\second},
+respectively~\cite{Yu2016a}. The cytoplasm model described in this
+work has 1.5 million atoms and was simulated for the total of
+\SI{3}{\micro\second}, an unprecedented time scale to study crowding
+effects on a heterogeneous cytoplasm model in atomic detail.  {\bf ADD
+  REFERENCES TO THE NEW FEIG PAPERS AND ZAHEDEHS ACS OMEGA PAPER UNDER
+  REVIEW. WE WILL SEND IN THAT ONE AS SUPPORTING MATERIAL FOR REVIEW
+  ONLY.}
 
-Three crowders (TufA, MetE and IcdA) displayed greater deviation from their crystallographic structure in the cytoplasm model when compared to the dilute condition, indicating that there was some degree of local unfolding (Fig.~\ref{fig:structural_integrity_chain}A).
-Interestingly, oligomers were marginally less stable (Fig.~\ref{fig:structural_integrity_chain}B) and the flexibility profiles calculated considering the C$\upalpha$ for each chain of the elements of our cytoplasm model show that, in the crowded condition, most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are slightly more flexible than in the dilute condition (Fig.~\ref{fig:rmsf}).
-This is in agreement with simulation~\cite{Feig2011} and experiment~\cite{miklos2011,Wang2012b} showing that crowders can destabilize the native structure of proteins due to the numerous unspecific short-lived interactions that are formed between them in a crowded environment.
+Three crowders (TufA, MetE and IcdA) displayed greater deviation from
+their crystallographic structure in the cytoplasm model when compared
+to the dilute condition, indicating that there was some degree of
+local unfolding
+(Fig.~\ref{fig:structural_integrity_chain}A). Interestingly, oligomers
+were marginally less stable {\bf THAN WHAT? Monomers?}
+(Fig.~\ref{fig:structural_integrity_chain}B) and the flexibility
+profiles calculated considering the C$\upalpha$ for each chain of the
+elements of our cytoplasm model show that, in the crowded condition,
+most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are slightly
+more flexible than in the dilute condition (Fig.~\ref{fig:rmsf}). This
+is in agreement with previous simulations~\cite{Feig2011} and
+experiment~\cite{miklos2011,Wang2012b} showing that crowders can
+destabilize the native structure of proteins due to the numerous
+unspecific short-lived interactions that are formed between them in a
+crowded environment. A study of protein properties at increasing
+concentration showed only few interactions between proteins of the
+same type, however~\cite{Bashardanesh2019a}.
+{\bf do we expect specific interactions between the proteins in our cyto model?
+In that case write that with references, otherwise write that there are no known specific interactions between these proteins.}
 
-Magnesium cations were added to our cytoplasm model as counter-ions for tRNA and ATP.
-However, as the simulation progressed, FBP and ATP aggregated around Mg$^{2+}$ and, by consequence, around tRNA.
-After identifying this problem, we used simulations of small systems containing only the metabolites, Mg$^{2+}$ and water to test different strategies to prevent aggregation in future studies using long simulations of cytoplasm models.
-We found that protonating the phosphate-containing metabolites is enough to reach this goal (Fig.~\ref{fig:avoiding_aggregation}).
-Even though ATPH$_3$ and FBPH$_4$ are not the biologically relevant forms of these molecules in physiological pH, we consider that using their protonated forms is a good compromise between parametrization complexity and practical results.
-The biggest impact of this approach is on long range interactions of these metabolites with other molecules due to the neutralization of their net charge.
-However, molecules in a crowded environment are restrained to smaller volumes and shorter inter molecular distances, which mitigates the effect of charge neutralization.
+Magnesium cations were added to our cytoplasm model as counter-ions
+for tRNA and ATP. However, as the simulation progressed, FBP and ATP
+aggregated around Mg$^{2+}$ and, by consequence, around tRNA{\bf,
+  effectively removing the counter ions from the RNA. IS THAT
+  CORRECT?}. After identifying this problem, we used simulations of
+small systems containing only the metabolites, Mg$^{2+}$ and water to
+test different strategies to prevent aggregation in future studies
+using long simulations of cytoplasm models. We found that protonating
+the phosphate-containing metabolites is enough to reach this goal
+(Fig.~\ref{fig:avoiding_aggregation}). Even though ATPH$_3$ and
+FBPH$_4$ are not the biologically relevant forms of these molecules in
+physiological pH, using their protonated forms may be an acceptable
+compromise between parametrization complexity and practical
+results. The biggest impact of this approach is on long range
+interactions of these metabolites with other molecules due to the
+neutralization of their net charge. However, molecules in a crowded
+environment are restrained to smaller volumes and shorter inter
+molecular distances, which mitigates the effect of charge
+neutralization. It has been suggested that charged species should
+employ scaled charges in order to be compatible with the empirical
+water models~\cite{Leontyev2009a,Leontyev2011a}. Due to the high local
+density of negative charges in FBP and ATP this might indeed be an
+alternative solution, barring use of polarizable
+models~\cite{Ponder2010a,Lopes2013a,Ghahremanpour2018b,Walz2018a}.
 
-To the best of our knowledge, there is only one work in the literature which studied large scale cytoplasm models that included metabolites with all-atom explicit solvent molecular dynamics~\cite{Yu2016a}.
-Although they didn't report aggregation, they mention that the translational diffusion of highly charged phosphate-containing metabolites is ``much slower'' than expected in their system.
-They attributed this observation to non-specific interactions between such metabolites and proteins.
-Maybe aggregation also contributed to their slower diffusion but it was not evident during \SI{140}{\nano\second}, which was the length of their longest simulation.
+To the best of our knowledge, {\bf AGAIN CHECK LATEST FEIG PAPERS}
+there is only published study on large scale cytoplasm models that
+included metabolites with all-atom explicit solvent molecular
+dynamics~\cite{Yu2016a}. Although these authors did not report
+aggregation, they mention that the translational diffusion of highly
+charged phosphate-containing metabolites is ``much slower'' than
+expected in their system. They attributed this observation to
+non-specific interactions between such metabolites and proteins. Maybe
+aggregation also contributed to their slower diffusion but it was not
+evident during their \SI{140}{\nano\second} simulation.
 
-\section*{Conclusions}\label{sec:concl}
+%\section*{Conclusions}\label{sec:concl}
 
-In this work we built an all-atom explicit solvent cytoplasm model composed of proteins, tRNA, metabolites, inorganic ions and water that sums 1.5 million atoms (Table \ref{tbl:soup_components}).
-The identity and amount of each component of the model was chosen according to experimental data about the composition of the cytoplasm of \textit{E.
-coli}~\cite{Dong1996,Bennett2009,Link1997,Mcguffee2010}.
+In this work we built an all-atom explicit solvent cytoplasm model
+composed of proteins, tRNA, metabolites, inorganic ions and water for a total
+of 1.5 million atoms (Table \ref{tbl:soup_components}). The identity
+and amount of each component of the model was chosen based on
+experimental data about the composition of the cytoplasm of
+\textit{E. coli}~\cite{Dong1996,Bennett2009,Link1997,Mcguffee2010}.
 
-We submitted this model to a total of \SI{3}{\micro\second} of molecular dynamics simulations using a modified AMBER FF99SB force field that was tuned to reproduce experimental data of systems composed of several macromolecules~\cite{Best2014a}.
-Our model was validated by reproducing experimental data about the reduction of the translational diffusion coefficient of proteins in the cytoplasm of \textit{E.
-coli} when compared to the dilute condition (Figure \ref{fig:translational_diffusion})~\cite{Elowitz1999,Konopka2006}.
-Additionally, our data is in agreement with recent simulation and experimental data that show that the unspecific interactions between macromolecules in a crowded environment can destabilize them~\cite{Feig2011,miklos2011,Wang2012b}.
-Specifically, we can see that, despite the overall stability of our cytoplasm model even in the \SI{}{\micro\second} timescale, the proteins are slightly less stable in  the cytoplasm model than in dilute condition (Figures \ref{fig:structural_integrity_chain} and \ref{fig:rmsf}).
+We submitted this model to a total of \SI{3}{\micro\second} of
+molecular dynamics simulations using a modified AMBER FF99SB force
+field that was tuned to reproduce experimental data of systems
+composed of several macromolecules~\cite{Best2014a}. The model was
+validated by comparing to experimental data about the reduction of the
+translational diffusion coefficient of proteins in the cytoplasm of
+\textit{E. coli} when compared to the dilute condition (Figure
+\ref{fig:translational_diffusion})~\cite{Elowitz1999,Konopka2006}. Additionally,
+our data is in agreement with recent simulation work and experimental data
+that show that the unspecific interactions between macromolecules in a
+crowded environment can destabilize
+them~\cite{Feig2011,miklos2011,Wang2012b}. Specifically, we can see
+that, despite the overall stability of our cytoplasm model even in the
+\SI{}{\micro\second} timescale, the proteins are slightly less stable
+in the cytoplasm model than in dilute condition (Figures
+\ref{fig:structural_integrity_chain} and \ref{fig:rmsf}).
 
-Phosphate-containing metabolites FBP and ATP aggregated around Mg$^{2+}$ that were used as counter-ions for ATP and tRNA (Figure \ref{fig:tRNA_aggregation}).
+%Phosphate-containing metabolites FBP and ATP aggregated around Mg$^{2+}$ that were used as counter-ions for ATP and tRNA (Figure \ref{fig:tRNA_aggregation}).
 We used independent simulations of small systems to evaluate different strategies to prevent such aggregation and found that this can be done by completely protonating these molecules (Figure \ref{fig:avoiding_aggregation}).
@@
 
-We are providing a Python script with a set of structure and topology files that any researcher can use to build their own \textit{E.
-coli} cytoplasm models.
-Such models can be used as tools that answer many research questions about the effects of crowding on specific systems of interest or about crowding itself.
-With the files we are providing it is possible, for example, to add a probe protein to investigate the effect of crowding on it, to add new macromolecular or small crowders, verify the effects of different temperatures and change the biomolecular concentration to increase or decrease the intensity of the crowding effect.
-We highlight that, since we have already validated it, researchers can take advantage from building their own cytoplasm models even if they don't have access to the infrastructure that is necessary to perform long molecular dynamics simulations.
-With modest computatioanal power it is possible, for example, to build many independent snapshots of the cytoplasm of \textit{E.
-coli}.
-All files are publicly available at github.com/dspoel/soup.
+As part of this work we are providing Python scripts with a set of
+structure and topology files that any researcher can use to build
+their own \textit{E. coli} cytoplasm models. Such models can be used
+to investigate research questions about the effects of
+crowding on specific systems of interest or about crowding
+itself. With the files we are providing it is possible, for example,
+to add a probe protein to investigate the effect of crowding on it, to
+add new macromolecular or small crowders, evaluate the effects of
+ temperatures and change the biomolecular concentration to
+increase or decrease the intensity of the crowding effect. We
+highlight that, since we have already validated it, researchers can
+take advantage from building their own cytoplasm models even if they
+do not have access to the infrastructure that is necessary to perform
+long molecular dynamics simulations. With modest computatioanal power
+it is possible, for example, to build many independent snapshots of
+the cytoplasm of \textit{E. coli}. All files are publicly available at
+http://github.com/dspoel/soup.
 

### 2019-10-05 paper/virtual_cytoplasm_v2.tex
Updated

@@
   FLUCTUATION OVER THE WHOLE PROTEIN, THAT IS BEFORE TAKING SQUARE
-  ROOTS OF COURSE AND PLOTTING THAT IN A BAR GRAPH. THEN THIS FIGURE
-  CAN GO TO SUPPORTING.}  Proteins are slightly more flexible in the
+  ROOTS OF COURSE AND PLOTTING THAT IN A BAR GRAPH.}  Proteins are slightly more flexible in the
 crowded than in the dilute environment. Root mean square fluctuation

### 2019-10-07 paper/virtual_cytoplasm_v2.tex
answered some of the comments

@@
 total simulation time. This work is the first to explore the
-structural dynamics of an all-atom explicit solvent cytoplasm model
+structural dynamics of an all-atom explicit solvent cytoplasm model {\color{blue} of {\em E. coli}}
 with microsecond scale molecular dynamics simulations. {\bf IS THAT
@@
 For statistics, the average and standard error {\bf standard error or
-  deviation?} of the three trajectories are reported. For values that
+  deviation? \color{blue} standard error} of the three trajectories are reported. For values that
 are specific to each chain, we used the average over the individual
@@
 translational diffusion coefficient, which were extracted by a linear
-fit to MSD by averaging blocks with a length of {\bf 10 ns even for
-  proteins?}  $10\,ns$~\cite{Allen1987a}. In principle, diffusion
+fit to MSD by averaging blocks with a length of $10\, ns$ for dilute systems and $100\, ns$ 
+for cytoplasm system~\cite{Allen1987a}. In principle, diffusion
 coefficient needs to be corrected for finite size
@@
 \label{tbl:protein_fraction}
-\begin{tabular}{lcccc}
+\begin{tabular}{ccccc}
 \hline
@@
 \hline
-TufA & 1DG1~\cite{Abel1996}       & 6 & 19.7 &  dimer \\
-MetE & 1U22~\cite{Ferrer2004}     & 7 & 11.6 &  monomer \\
-IcdA & 1P8F~\cite{Mesecar2000}    & 2 & 4.7  &  dimer \\
-AhpC & 1YEP~\cite{Parsonage2005}  & 1 & 4.1  &  decamer \\
-CspC & 1MJC~\cite{Schindelin1994} & 3 & 4.0  &  monomer \\
-Ppa  & 2EIP~\cite{Kankare1996}    & 1 & 2.9  &  hexamer \\
-GapA & 1S7C~\cite{ShinXXX}        & 1 & 2.1  &  tetramer \\
-Eno  & 1E9I~\cite{Kuhnel2001}     & 1 & 1.9  &  dimer \\
+{\color{blue}Elongation factor TU} (TufA) & 1DG1~\cite{Abel1996}       & 6 & 19.7 &  dimer \\
+{\color{blue}Cobalamin-independent methionine synthase} (MetE) & 1U22~\cite{Ferrer2004}     & 7 & 11.6 &  monomer \\
+{\color{blue}Isocitrate dehydrogenase} (IcdA) & 1P8F~\cite{Mesecar2000}    & 2 & 4.7  &  dimer \\
+{\color{blue}Alkyl hydroperoxide reductase subunit C} (AhpC) & 1YEP~\cite{Parsonage2005}  & 1 & 4.1  &  decamer \\
+{\color{blue}Cold-shock protein} (CspC) & 1MJC~\cite{Schindelin1994} & 3 & 4.0  &  monomer \\
+{\color{blue}Pyrophosphatase} (Ppa)  & 2EIP~\cite{Kankare1996}    & 1 & 2.9  &  hexamer \\
+{\color{blue}Glyceraldehyde 3-phosphate dehydrogenase A} (GapA) & 1S7C~\cite{ShinXXX}        & 1 & 2.1  &  tetramer \\
+{\color{blue}Enolase} (Eno)  & 1E9I~\cite{Kuhnel2001}     & 1 & 1.9  &  dimer \\
 \hline
@@
 in our cytoplasm model were not affected to a great extent by the
-crowded environment. The value of the root mean square deviation
-(RMSD, Fig.~\ref{fig:structural_integrity_chain}A) considering only
+crowded environment. The value of the root mean square deviation considering only
 their C$\upalpha$ in the cytoplasm model is within 11\% of its value
@@
 TufA, MetE and IcdA, which have RMSD values 105, 25 and 33\% higher in
-the crowded environment than in the dilute condition. {\bf ARE THERE
+the crowded environment than in the dilute condition {\color{blue}(Fig.~\ref{fig:structural_integrity_chain}A)}. {\bf ARE THERE
   RMSD PLOTS? REFER TO THEM.} Visual inspection of their trajectories
@@
 
-To the best of our knowledge, {\bf AGAIN CHECK LATEST FEIG PAPERS}
+To the best of our knowledge, {\bf AGAIN CHECK LATEST FEIG PAPERS} {\color{blue} Latest Feig paper is not a cytoplasm model~\cite{nawrocki2019}}
 there is only published study on large scale cytoplasm models that

### 2019-10-10 paper/virtual_cytoplasm_v2.tex
Review

@@
 activities~\citep{ostrowska2019}. This effect, called macromolecular
-crowding, influences the stability of {\bf SOME/MANY} macromolecules
+crowding, influences the stability of {\color{blue} many} {\bf SOME/MANY} macromolecules
 because the equilibrium is shifted towards a stable native structure
@@
 possible in 15 years from now~\cite{vanGunsteren2006a}. {\bf THAT WAS
-  13 YEARS AGO. WHAT IS THE STATE OF THE ART?}
+  13 YEARS AGO.
+WHAT IS THE STATE OF THE ART?} {\color{blue} The paper predicts that simulating a whole E.
+coli bacterium with $10^{11}$ atoms for 1 ns will be possible in 2034, thus the "15 years from now" in our text.
+That was meant as an introduction to the simulation of big systems.
+The evolution of computational models to study crowing and the state-of-the-art is discussed in the next paragraph.
+Maybe we can move this phrase to the beginning of the next paragraph }
 
@@
 simulations of our cytoplasm model, reaching \SI{3}{\micro\second} of
-total simulation time. This work is the first to explore the
-structural dynamics of an all-atom explicit solvent cytoplasm model {\color{blue} of {\em E. coli}}
+total simulation time. {\color{blue}To the best of our knowledge,} this work is the first to explore the
+structural dynamics of an all-atom explicit solvent cytoplasm model 
 with microsecond scale molecular dynamics simulations. {\bf IS THAT
-  REALLY TRUE? A REFEREE MAY GET ANGRY WHEN READING THIS.}
+  REALLY TRUE?
+A REFEREE MAY GET ANGRY WHEN READING THIS.} {\color{blue} That's true.
+We didn't find other all-atom explicit solvent cytoplasm models that were simulated for such a long time.
+Added "To the best of our knowledge" to the text.}
 
@@
 used for the biomolecules. {\bf DESCRIBE WHICH MODELS WERE USED FOR
-  THE IONS.}
+  THE IONS.} {\color{blue} it was inserted in the next section with the description of the force field and water model}
 
@@
 The cytoplasm model was built at a 30\% biomolecular mass fraction and
-a physiological salt concentration of 0.15 mol/L KCl.  {\bf DESCRIBE
+a physiological salt concentration of 0.15 mol/L KCl.
+{\color{blue} Dropplets containing each component of the model surronded by water and counter-ions were inserted in an empty box in random positions and orientations.
+Since the volume of this box was initially bigger than what was strictly necessary to accomodate all components, the system was allowed to shrink in a short simulation.
+Details about how the simulation box was built and optimized are in the results section} {\bf DESCRIBE
   BRIEFLY HOW COMPOUNDS WERE PLACED IN THE BOX AND THAT THE VOLUME WAS
@@
 field, in combination with the TIP4P/2005 water
-model~\cite{Abascal2005b}. All interactions were calculated explicitly
+model~\cite{Abascal2005b} {\color{blue} and improved K$^+$ parameters~\cite{Dang1995a} to avoid the formation of KCl crystals that is observed when using the default parameters of the Amber99SB force field~\cite{Auffinger2007a}}.
+All interactions were calculated explicitly
 inside a 1 nm radius and long range electrostatic interactions were
@@
 average values and standard errors are calculated {\bf DO YOU MEAN THE
-  AVERAGE OVER ALL COPIES? considering each copy and each replica}
-independently. As in the dilute system, separate chains of each copy
+  AVERAGE OVER ALL COPIES? considering each copy of each replica}
+independently {\color{blue} Yes, the averages are calculated over all copies of each crowder in all replicas}. As in the dilute system, separate chains of each copy
 of the crowders were considered as independent samples when running
@@
   SLOWLY!}. Root Mean Square Deviation (RMSD) and Root Mean Square
-Fluctuation (RMSF) calculations considered the C$\upalpha$ atoms only
-and the reference frame for RMSF was the midpoint of the time interval
+Fluctuation (RMSF) calculations considered the C$\upalpha$ atoms only for analysis in each trajectory
 {\bf ARE YOU SURE? THE DEFAULT IS TO TAKE THE AVERAGE STRUCTURE. OR DO
-  YOU MEAN THE STRUCTURE FOR FITTING?} used for analysis in each
-trajectory. The chain interface area of a given oligomer in a given
+  YOU MEAN THE STRUCTURE FOR FITTING?} {\color{blue} I checked and indeed we used the average structure calculated with the second half of the trajectories.
+I removed "the reference frame for RMSF was the midpoint of the time interval" from the text.
+The information about the second half of the trajectory is in the previous paragraph, which is about statistics}.
+The chain interface area of a given oligomer in a given
 frame was calculated as the difference between the solvent exposed
@@
 in our cytoplasm model were not affected to a great extent by the
-crowded environment. The value of the root mean square deviation considering only
-their C$\upalpha$ in the cytoplasm model is within 11\% of its value
-under dilute conditions {\bf WHAT DOES WITHIN 11\% MEAN?}, except for
+crowded environment. The values of their root mean square deviation considering only
+their C$\upalpha$ in the cytoplasm model {\color{blue} increased by less than 11\% when compared to the dilute condition} {\bf WHAT DOES WITHIN 11\% MEAN?}, except for
 TufA, MetE and IcdA, which have RMSD values 105, 25 and 33\% higher in
 the crowded environment than in the dilute condition {\color{blue}(Fig.~\ref{fig:structural_integrity_chain}A)}. {\bf ARE THERE
-  RMSD PLOTS? REFER TO THEM.} Visual inspection of their trajectories
-showed that there is partial unfolding of their N- or C-terminal
-residues. {\bf IS THERE A KNOWN PHYSIOLOGICAL REASON FOR THAT?. IS
+  RMSD PLOTS?
+REFER TO THEM.} {\color{blue} Visual inspection of the final frames of the simulations of these crowders in the cytoplasm model and in dilute condition revealed that, for TufA, the conformational differences are on the $\upbeta$-hairpin around residues 40-62.
+This region is called Switch I and is known to be flexible and to display conformational changes due to interactions with the ribosome or other partners~\cite{Abel1996}.
+For MetE, there are differences in the conformation of the loop around residues 450-460 and in the C-terminal $\upalpha$-helices.
+This loop is known to be flexible and was not modeled in the crystallographic structure due to such flexibility~\cite{Ferrer2004}.
+For IcdA, the differences are located at the N-terminal coil formed by residues 1-8.} {\bf IS THERE A KNOWN PHYSIOLOGICAL REASON FOR THAT?.
+IS
   MetE A MULTIDOMAIN PROTEIN WHERE DOMAINS MOVE WITH RESPECT TO EACH
-  OTHER?}
+  OTHER?} {\color{blue} The differences of TufA and MetE can be expected from some experimental data, but for IcdA it is just partial unfolding of a short N-terminal coil region.}
 
@@
   FLUCTUATION OVER THE WHOLE PROTEIN, THAT IS BEFORE TAKING SQUARE
-  ROOTS OF COURSE AND PLOTTING THAT IN A BAR GRAPH.}  Proteins are slightly more flexible in the
-crowded than in the dilute environment. Root mean square fluctuation
-(RMSF) calculations show that the flexibility profile of the proteins
-is not affected by crowding, which is in agreement with our results
-showing that the structural integrity of the proteins is not changed
-significantly in the cytoplasm model (Fig.~\ref{fig:rmsf}).  On the
-other hand, there are some small differences between the dilute and
-crowding conditions. In most cases (TufA, MetE, AhpC, CspC and Ppa),
-proteins are slightly more flexible in the cytoplasm model
-(Fig.~\ref{fig:rmsf}). This is in agreement with data showing that
-nonspecific interactions with crowders can lead to destabilization of
+  ROOTS OF COURSE AND PLOTTING THAT IN A BAR GRAPH.} {\color{blue} Overall, flexibility (meaning the values of RMSF for each residue) is affected in some cases, but the flexibility profile (meaning the overall shape of the RMSF plot) is not.
+Maybe that was the source of the apparent contradiction.
+I changed the text to clarify that.
+I also removed words such as "slightly" and "small" from the text.}
+
+Root mean square fluctuation (RMSF) calculations show that, overall, the shape of residue-wise flexibility profile of the proteins is not affected by crowding, which is in agreement with our results showing that their structural integrity does not change significantly in the cytoplasm model in most cases (Fig.~\ref{fig:rmsf}).
+Accordingly, we can see significant changes in the shape of the flexibility profile for regions that undergo conformational changes, such as the Switch I $\upbeta$-hairpin of TufA around residues 40-62 and the flexible loop of MetE around residues 450-460 (Fig.~\ref{fig:rmsf}).
+
+In most cases (TufA, MetE, AhpC, CspC and Ppa), the residue-wise RMSF values are higher in the cytoplasm model than under diluted conditions, which indicates that proteins are more flexible in the cytoplasm model (Fig.~\ref{fig:rmsf}).
+This is in agreement with computational and experimental data showing that nonspecific interactions with crowders can lead to destabilization of
 proteins~\cite{Feig2011, miklos2011, Wang2012b}.
@@
   WERE ANY INTERMEDIATE PROTONATIONS TESTED? E.G. FBPH2(2-) AND ATPH(2-)?
-  COMPLETELY PROTONATING IS UNPHYSICAL AS WELL AT THE PH STUDIED. FOR E COLI THE PH SHOULDBE ROUGHLY 7.4 https://en.wikipedia.org/wiki/Cytosol}
+  COMPLETELY PROTONATING IS UNPHYSICAL AS WELL AT THE PH STUDIED.
+FOR E COLI THE PH SHOULDBE ROUGHLY 7.4 https://en.wikipedia.org/wiki/Cytosol} {\color{blue}I didn't try intermediate protonations, but this can be done in about 1 week if necessary.
+The physiological states of FBP and ATP at pH 7.4 would be the ones we used originally in the soup.
+So, in this sense, any protonated state could be called unphysical or unrealistic.
+We discuss the tradeoffs of using this unphysical protonated state in the discussion section}
 
@@
 FBP$^{-4}$ around Mg$^{2+}$, which was used as counter-ion for tRNA,
-leads to an exaggerated confinement of the tRNA molecules.  Therefor
+leads to an exaggerated confinement of the tRNA molecules.  Therefore,
 tRNA is considered as an outlier in the duffusion calculations. In
@@
 measurements~\cite{Elowitz1999,Konopka2006}.
-{\bf IT MAY BE NECESSARY TO DISCUSS RMSF HERE AS WELL}
+{\bf IT MAY BE NECESSARY TO DISCUSS RMSF HERE AS WELL} {\color{blue} This first paragraph is focused on validation of the cytoplasm model by comparison with experimental results, which are available for the translational diffusion coefficient.
+We don't have specific experimental data that we can compare our RMSF data to.
+Should we add RMSF here anyway?
+There is a discussion about our RMSF results some paragraphs ahead in which we compare our data with general trends that were experimentally observed.}
 
@@
 Previous
-all-atom simulations of high scale {\bf WHAT DO YOU MEAN WITH HIGH
-  SCALE?} cytoplasm models used parameters from the CHARMM family of
+all-atom simulations of {\color{blue} large} scale {\bf WHAT DO YOU MEAN WITH HIGH
+  SCALE?} {\color{blue} corrected to large scale} cytoplasm models used parameters from the CHARMM family of
 force fields~\cite{Yu2016a}. In this work we use a modified version of
@@
 (Fig.~\ref{fig:structural_integrity_chain}A). Interestingly, oligomers
-were marginally less stable {\bf THAN WHAT? Monomers?}
+were marginally less stable {\color{blue} in the cytoplasm model} {\bf THAN WHAT?
+Monomers?} {\color{blue} oligomers in the cytoplasm were marginally less stable than oligomers in the dillute condition.
+Added "in the cytoplasm model" to the text.}
 (Fig.~\ref{fig:structural_integrity_chain}B) and the flexibility
@@
 same type, however~\cite{Bashardanesh2019a}.
-{\bf do we expect specific interactions between the proteins in our cyto model?
-In that case write that with references, otherwise write that there are no known specific interactions between these proteins.}
+{\color{blue} We highlight that there are no specific interactions between the proteins that were used in our cytoplasm model.} {\bf do we expect specific interactions between the proteins in our cyto model?
+In that case write that with references, otherwise write that there are no known specific interactions between these proteins.} {\color{blue} The only specific interaction we antecipate between the components of the soup is that TufA interacts with tRNA (PDB ID 1TTT is an example of such complex).
+However, since we are ignoring tRNA in our analyses, I did't mention it.
+PS: I checked and the TufA-tRNA complex is not present in the last frames of our replicas.}
 
@@
   effectively removing the counter ions from the RNA. IS THAT
-  CORRECT?}. After identifying this problem, we used simulations of
+  CORRECT?} {\color{blue} No, Figure \ref{fig:tRNA_aggregation} shows that Mg$^{2+}$ is still concentrated around tRNA in the soup }. After identifying this problem, we used simulations of
 small systems containing only the metabolites, Mg$^{2+}$ and water to

### 2019-10-10 paper/virtual_cytoplasm_v2.tex
Review again

@@
 elements of our cytoplasm model show that, in the crowded condition,
-most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are slightly
-more flexible than in the dilute condition (Fig.~\ref{fig:rmsf}). This
+most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are more flexible than in the dilute condition (Fig.~\ref{fig:rmsf}). This
 is in agreement with previous simulations~\cite{Feig2011} and
@@
 concentration showed only few interactions between proteins of the
-same type, however~\cite{Bashardanesh2019a}.
-{\color{blue} We highlight that there are no specific interactions between the proteins that were used in our cytoplasm model.} {\bf do we expect specific interactions between the proteins in our cyto model?
+same type~\cite{Bashardanesh2019a}.
+{\color{blue} We highlight that there are no specific interactions described between the proteins that were used in our cytoplasm model.} {\bf do we expect specific interactions between the proteins in our cyto model?
 In that case write that with references, otherwise write that there are no known specific interactions between these proteins.} {\color{blue} The only specific interaction we antecipate between the components of the soup is that TufA interacts with tRNA (PDB ID 1TTT is an example of such complex).

### 2019-10-13 paper/virtual_cytoplasm_v2.tex
msf data added

@@
 activities~\citep{ostrowska2019}. This effect, called macromolecular
-crowding, influences the stability of {\color{blue} many} {\bf SOME/MANY} macromolecules
+crowding, influences the stability of many macromolecules
 because the equilibrium is shifted towards a stable native structure
@@
 simulations of our cytoplasm model, reaching \SI{3}{\micro\second} of
-total simulation time. {\color{blue}To the best of our knowledge,} this work is the first to explore the
+total simulation time. To the best of our knowledge, this work is the first to explore the
 structural dynamics of an all-atom explicit solvent cytoplasm model 
-with microsecond scale molecular dynamics simulations. {\bf IS THAT
-  REALLY TRUE?
-A REFEREE MAY GET ANGRY WHEN READING THIS.} {\color{blue} That's true.
-We didn't find other all-atom explicit solvent cytoplasm models that were simulated for such a long time.
-Added "To the best of our knowledge" to the text.}
+with microsecond scale molecular dynamics simulations.
 
@@
 of compounds modeled using the GAFF have been evaluated in a number of
-papers~\cite{Caleman2012a,sprenger2015general,Fischer2015a,JZhang2015a,JZhang2015,Spoel2018a}
+papers~\cite{Caleman2012a,sprenger2015general,Fischer2015a,JZhang2015a,Spoel2018a}
 and strengths and weaknesses are well understood. Here, GAFF is used
 for compatibility with the Amber99SB-ws force field~\cite{Best2014a}
-used for the biomolecules. {\bf DESCRIBE WHICH MODELS WERE USED FOR
-  THE IONS.} {\color{blue} it was inserted in the next section with the description of the force field and water model}
+used for the biomolecules.
 
@@
 a physiological salt concentration of 0.15 mol/L KCl.
-{\color{blue} Dropplets containing each component of the model surronded by water and counter-ions were inserted in an empty box in random positions and orientations.
+Dropplets containing each component of the model surronded by water and counter-ions were inserted in an empty box in random positions and orientations.
 Since the volume of this box was initially bigger than what was strictly necessary to accomodate all components, the system was allowed to shrink in a short simulation.
-Details about how the simulation box was built and optimized are in the results section} {\bf DESCRIBE
-  BRIEFLY HOW COMPOUNDS WERE PLACED IN THE BOX AND THAT THE VOLUME WAS
-  WAY LARGER. DO ALSO WRITE WHETHER MOLECULES WERE ROTATED randomly
-  PRIOR TO PLACEMENT IN THE BOX. ALTERNATIVELY REFER TO RESULTS.}  All
+Details about how the simulation box was built and optimized are in the results section.
+All
 simulations were performed with the GROMACS 2018
@@
 field, in combination with the TIP4P/2005 water
-model~\cite{Abascal2005b} {\color{blue} and improved K$^+$ parameters~\cite{Dang1995a} to avoid the formation of KCl crystals that is observed when using the default parameters of the Amber99SB force field~\cite{Auffinger2007a}}.
+model~\cite{Abascal2005b} and improved K$^+$ parameters~\cite{Dang1995a} to avoid the formation of KCl crystals that is observed when using the default parameters of the Amber99SB force field~\cite{Auffinger2007a}.
 All interactions were calculated explicitly
@@
 \subsection{Analyses}
-For statistics, the average and standard error {\bf standard error or
-  deviation? \color{blue} standard error} of the three trajectories are reported. For values that
+For statistics, the average and standard error of the three trajectories are reported. For values that
 are specific to each chain, we used the average over the individual

### 2019-10-13 paper/barplots.py

@@ -50,6 +50,8 @@ if __name__ == '__main__':
             ylabel = "Chain Interface Area ($nm^{2}$)"
         elif args.prop == "rmsd": 
             ylabel = "RMSD ($nm$)"
+        elif args.prop == "msf": 
+            ylabel = "MSF ($nm^2$)"
         elif args.prop == "rg":
             ylabel = "Rg ($nm$)"
         elif args.prop == "msd":
@@ -80,7 +82,7 @@ if __name__ == '__main__':
     
     fig, ax = plt.subplots()
     rects1 = ax.bar(ind - width/2, singles_mean, width, yerr=singles_std,
-                    label='Sing.')
+                    label='Dilute')
     rects2 = ax.bar(ind + width/2, cyto_mean, width, yerr=cyto_std,
                     label='Cyto.')
     

### 2019-10-13 paper/virtual_cytoplasm_v2.tex
changes in the text

@@
 this correction is negligible {\bf NOT FOR PROTEINS THAT DIFFUSE
-  SLOWLY!}. Root Mean Square Deviation (RMSD) and Root Mean Square
+  SLOWLY!}.
+{\color{blue} The correcion in Yeh2004 is derived for cubic simulation boxes, in our single molecule simulations we have non-cubic boxes, one needs to derive the correction from Eq 11 in Yeh2004.
+My Mathematica skills are rusty now, could you do that?
+If you think it's absolutely necessary} Root Mean Square Deviation (RMSD) and Root Mean Square
 Fluctuation (RMSF) calculations considered the C$\upalpha$ atoms only for analysis in each trajectory
@@
 biomolecular concentration, which is a parameter of the cytoplasm
-model. This number ranges from 300 to 400 g/L in biological systems
-such as \textit{E. coli} cytoplasm~\cite{Zimmerman1991}, which {\bf
-  WHAT DOES WHICH CORRESPOND TO?}  corresponds to biomolecular
-concentrations as high as 40\%~\cite{Ellis2003a}. In our case, we
+model. Macromolecular concentration ranges from 300 to 400 g/L in biological systems
+such as \textit{E. coli} cytoplasm~\cite{Zimmerman1991}, the highest concentration corresponds to volume occupation as high as 40\%~\cite{Ellis2003a}. In our case, we
 choose a biomolecular concentration of 30\%, that is, the number of

### 2019-10-16 paper/virtual_cytoplasm_v2.tex
changes

@@
 This loop is known to be flexible and was not modeled in the crystallographic structure due to such flexibility~\cite{Ferrer2004}.
-For IcdA, the differences are located at the N-terminal coil formed by residues 1-8.} {\bf IS THERE A KNOWN PHYSIOLOGICAL REASON FOR THAT?.
-IS
-  MetE A MULTIDOMAIN PROTEIN WHERE DOMAINS MOVE WITH RESPECT TO EACH
-  OTHER?} {\color{blue} The differences of TufA and MetE can be expected from some experimental data, but for IcdA it is just partial unfolding of a short N-terminal coil region.}
+For IcdA, the differences are located at the N-terminal coil formed by residues 1-8.}
+
+  %{\bf IS THERE A KNOWN PHYSIOLOGICAL REASON FOR THAT?.
+IS  MetE A MULTIDOMAIN PROTEIN WHERE DOMAINS MOVE WITH RESPECT TO EACH  OTHER?} {\color{blue} The differences of TufA and MetE can be expected from some experimental data, but for IcdA it is just partial unfolding of a short N-terminal coil region.}
 
@@
 were marginally less stable {\color{blue} in the cytoplasm model} {\bf THAN WHAT?
-Monomers?} {\color{blue} oligomers in the cytoplasm were marginally less stable than oligomers in the dillute condition.
+Monomers?} {\color{blue} oligomers in the cytoplasm were marginally less stable than oligomers in the dilute condition.
 Added "in the cytoplasm model" to the text.}

### 2019-10-17 paper/virtual_cytoplasm_v2.tex
comments are removed from the text except one which needs to be addressed by Leandro

@@
 \textit{Escherichia coli} bacterium for \SI{1}{\nano\second} will be
-possible in 15 years from now~\cite{vanGunsteren2006a}. {\bf THAT WAS
-  13 YEARS AGO.
+possible in 15 years from now~\cite{vanGunsteren2006a}.
+% {\bf THAT WAS 13 YEARS AGO.
 WHAT IS THE STATE OF THE ART?} {\color{blue} The paper predicts that simulating a whole E.
@@
 is more than one copy of some crowders in the cytoplasm models,
-average values and standard errors are calculated {\bf DO YOU MEAN THE
-  AVERAGE OVER ALL COPIES? considering each copy of each replica}
-independently {\color{blue} Yes, the averages are calculated over all copies of each crowder in all replicas}. As in the dilute system, separate chains of each copy
+average values and standard errors are calculated considering each copy of each replica. As in the dilute system, separate chains of each copy
 of the crowders were considered as independent samples when running
@@
 coefficient needs to be corrected for finite size
-effects~\cite{Yeh2004} but due to relatively large simulation boxes
-this correction is negligible {\bf NOT FOR PROTEINS THAT DIFFUSE
-  SLOWLY!}.
+effects~\cite{Yeh2004,Abraham2018}. Due to relatively large simulation boxes
+this correction is negligible. 
+% {\bf NOT FOR PROTEINS THAT DIFFUSE
+  % SLOWLY!}.
 {\color{blue} The correcion in Yeh2004 is derived for cubic simulation boxes, in our single molecule simulations we have non-cubic boxes, one needs to derive the correction from Eq 11 in Yeh2004.
 My Mathematica skills are rusty now, could you do that?
-If you think it's absolutely necessary} Root Mean Square Deviation (RMSD) and Root Mean Square
-Fluctuation (RMSF) calculations considered the C$\upalpha$ atoms only for analysis in each trajectory
-{\bf ARE YOU SURE? THE DEFAULT IS TO TAKE THE AVERAGE STRUCTURE. OR DO
-  YOU MEAN THE STRUCTURE FOR FITTING?} {\color{blue} I checked and indeed we used the average structure calculated with the second half of the trajectories.
+If you think it's absolutely necessary}
+Root Mean Square Deviation (RMSD) and Root Mean Square
+Fluctuation (RMSF) calculations considered the C$\upalpha$ atoms only for analysis in each trajectory.
+% {\bf ARE YOU SURE?
+THE DEFAULT IS TO TAKE THE AVERAGE STRUCTURE.
+OR DO  YOU MEAN THE STRUCTURE FOR FITTING?} {\color{blue} I checked and indeed we used the average structure calculated with the second half of the trajectories.
 I removed "the reference frame for RMSF was the midpoint of the time interval" from the text.
@@
     E. coli} K12 and oligomeric state of each protein that was
-  selected to compose our cytoplasm model. {\bf WRITE OUT WHOLE
-    PROTEIN NAMES}}
+  selected to compose our cytoplasm model.}
 \label{tbl:protein_fraction}
@@
 \hline
-{\color{blue}Elongation factor TU} (TufA) & 1DG1~\cite{Abel1996}       & 6 & 19.7 &  dimer \\
-{\color{blue}Cobalamin-independent methionine synthase} (MetE) & 1U22~\cite{Ferrer2004}     & 7 & 11.6 &  monomer \\
-{\color{blue}Isocitrate dehydrogenase} (IcdA) & 1P8F~\cite{Mesecar2000}    & 2 & 4.7  &  dimer \\
-{\color{blue}Alkyl hydroperoxide reductase subunit C} (AhpC) & 1YEP~\cite{Parsonage2005}  & 1 & 4.1  &  decamer \\
-{\color{blue}Cold-shock protein} (CspC) & 1MJC~\cite{Schindelin1994} & 3 & 4.0  &  monomer \\
-{\color{blue}Pyrophosphatase} (Ppa)  & 2EIP~\cite{Kankare1996}    & 1 & 2.9  &  hexamer \\
-{\color{blue}Glyceraldehyde 3-phosphate dehydrogenase A} (GapA) & 1S7C~\cite{ShinXXX}        & 1 & 2.1  &  tetramer \\
-{\color{blue}Enolase} (Eno)  & 1E9I~\cite{Kuhnel2001}     & 1 & 1.9  &  dimer \\
+Elongation factor TU (TufA) & 1DG1~\cite{Abel1996}       & 6 & 19.7 &  dimer \\
+Cobalamin-independent methionine synthase (MetE) & 1U22~\cite{Ferrer2004}     & 7 & 11.6 &  monomer \\
+Isocitrate dehydrogenase (IcdA) & 1P8F~\cite{Mesecar2000}    & 2 & 4.7  &  dimer \\
+Alkyl hydroperoxide reductase subunit C (AhpC) & 1YEP~\cite{Parsonage2005}  & 1 & 4.1  &  decamer \\
+Cold-shock protein (CspC) & 1MJC~\cite{Schindelin1994} & 3 & 4.0  &  monomer \\
+Pyrophosphatase (Ppa)  & 2EIP~\cite{Kankare1996}    & 1 & 2.9  &  hexamer \\
+Glyceraldehyde 3-phosphate dehydrogenase A (GapA) & 1S7C~\cite{ShinXXX}        & 1 & 2.1  &  tetramer \\
+Enolase (Eno)  & 1E9I~\cite{Kuhnel2001}     & 1 & 1.9  &  dimer \\
 \hline
@@
 crowded environment. The values of their root mean square deviation considering only
-their C$\upalpha$ in the cytoplasm model {\color{blue} increased by less than 11\% when compared to the dilute condition} {\bf WHAT DOES WITHIN 11\% MEAN?}, except for
+their C$\upalpha$ in the cytoplasm model increased by less than 11\% when compared to the dilute condition, except for
 TufA, MetE and IcdA, which have RMSD values 105, 25 and 33\% higher in
-the crowded environment than in the dilute condition {\color{blue}(Fig.~\ref{fig:structural_integrity_chain}A)}. {\bf ARE THERE
-  RMSD PLOTS?
-REFER TO THEM.} {\color{blue} Visual inspection of the final frames of the simulations of these crowders in the cytoplasm model and in dilute condition revealed that, for TufA, the conformational differences are on the $\upbeta$-hairpin around residues 40-62.
+the crowded environment than in the dilute condition (Fig.~\ref{fig:structural_integrity_chain}A). 
+Visual inspection of the final frames of the simulations of these crowders in the cytoplasm model and in dilute condition revealed that, for TufA, the conformational differences are on the $\upbeta$-hairpin around residues 40-62.
 This region is called Switch I and is known to be flexible and to display conformational changes due to interactions with the ribosome or other partners~\cite{Abel1996}.
@@
 This loop is known to be flexible and was not modeled in the crystallographic structure due to such flexibility~\cite{Ferrer2004}.
-For IcdA, the differences are located at the N-terminal coil formed by residues 1-8.}
+For IcdA, the differences are located at the N-terminal coil formed by residues 1-8.
 
@@
 \subsubsection{Structural dynamics}
-{\bf THIS PARAGRAPH REPEATS AND CONTRADICTS ITSELF. IS FLEXIBILITY
-  AFFECTED OR NOT? FROM THE GRAPH I WOULD SAY THE THE FIVE PROTEINS
-  NOTED BELOW ARE MUCH MORE FLEXIBLE, NOT SLIGHTLY. ONE WAY OF
-  QUANTIFYING THIS IS TO COMPUTE THE AVERAGE OF THE MEAN SQUARE
-  FLUCTUATION OVER THE WHOLE PROTEIN, THAT IS BEFORE TAKING SQUARE
-  ROOTS OF COURSE AND PLOTTING THAT IN A BAR GRAPH.} {\color{blue} Overall, flexibility (meaning the values of RMSF for each residue) is affected in some cases, but the flexibility profile (meaning the overall shape of the RMSF plot) is not.
+% {\bf THIS PARAGRAPH REPEATS AND CONTRADICTS ITSELF. IS FLEXIBILITY
+%   AFFECTED OR NOT? FROM THE GRAPH I WOULD SAY THE THE FIVE PROTEINS
+%   NOTED BELOW ARE MUCH MORE FLEXIBLE, NOT SLIGHTLY. ONE WAY OF
+%   QUANTIFYING THIS IS TO COMPUTE THE AVERAGE OF THE MEAN SQUARE
+%   FLUCTUATION OVER THE WHOLE PROTEIN, THAT IS BEFORE TAKING SQUARE
+%   ROOTS OF COURSE AND PLOTTING THAT IN A BAR GRAPH.} {\color{blue} Overall, flexibility (meaning the values of RMSF for each residue) is affected in some cases, but the flexibility profile (meaning the overall shape of the RMSF plot) is not.
 Maybe that was the source of the apparent contradiction.
@@
 ATPH\textsubscript{3} and
-FBPH\textsubscript{4}(Fig.~\ref{fig:avoiding_aggregation}B-C).  {\bf
-  WERE ANY INTERMEDIATE PROTONATIONS TESTED? E.G. FBPH2(2-) AND ATPH(2-)?
-  COMPLETELY PROTONATING IS UNPHYSICAL AS WELL AT THE PH STUDIED.
+FBPH\textsubscript{4}(Fig.~\ref{fig:avoiding_aggregation}B-C).  
+% {\bf
+%   WERE ANY INTERMEDIATE PROTONATIONS TESTED? E.G. FBPH2(2-) AND ATPH(2-)?
+%   COMPLETELY PROTONATING IS UNPHYSICAL AS WELL AT THE PH STUDIED.
 FOR E COLI THE PH SHOULDBE ROUGHLY 7.4 https://en.wikipedia.org/wiki/Cytosol} {\color{blue}I didn't try intermediate protonations, but this can be done in about 1 week if necessary.
@@
 measurements~\cite{Elowitz1999,Konopka2006}.
-{\bf IT MAY BE NECESSARY TO DISCUSS RMSF HERE AS WELL} {\color{blue} This first paragraph is focused on validation of the cytoplasm model by comparison with experimental results, which are available for the translational diffusion coefficient.
+
+% {\bf IT MAY BE NECESSARY TO DISCUSS RMSF HERE AS WELL} {\color{blue} This first paragraph is focused on validation of the cytoplasm model by comparison with experimental results, which are available for the translational diffusion coefficient.
 We don't have specific experimental data that we can compare our RMSF data to.
@@
 Previous
-all-atom simulations of {\color{blue} large} scale {\bf WHAT DO YOU MEAN WITH HIGH
-  SCALE?} {\color{blue} corrected to large scale} cytoplasm models used parameters from the CHARMM family of
+all-atom simulations of large scale cytoplasm models used parameters from the CHARMM family of
 force fields~\cite{Yu2016a}. In this work we use a modified version of
@@
 to reproduce experimental data when there are many proteins in the
-same simulation box. All-atom explicit solvent molecular dynamics
-simulations have been reported for large-scale cytoplasm models of
+same simulation box.
+This force field or similarly adjusted one were also used for studying crowding effect on simpler systems of high concentration of proteins~\cite{Nawrocki2017a,nawrocki2019,Bashardanesh2019a}.
+All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of
 \textit{Mycoplasma genitalium} composed of 103- and 12-million atoms
@@
 work has 1.5 million atoms and was simulated for the total of
-\SI{3}{\micro\second}, an unprecedented time scale to study crowding
-effects on a heterogeneous cytoplasm model in atomic detail.  {\bf ADD
-  REFERENCES TO THE NEW FEIG PAPERS AND ZAHEDEHS ACS OMEGA PAPER UNDER
-  REVIEW. WE WILL SEND IN THAT ONE AS SUPPORTING MATERIAL FOR REVIEW
-  ONLY.}
+\SI{3}{\micro\second}, an unprecedented time scale to study crowding effects on a heterogeneous cytoplasm model in atomic detail.
 
@@
 (Fig.~\ref{fig:structural_integrity_chain}A). Interestingly, oligomers
-were marginally less stable {\color{blue} in the cytoplasm model} {\bf THAN WHAT?
-Monomers?} {\color{blue} oligomers in the cytoplasm were marginally less stable than oligomers in the dilute condition.
-Added "in the cytoplasm model" to the text.}
+were marginally less stable in the cytoplasm model than in the dilute condition.
 (Fig.~\ref{fig:structural_integrity_chain}B) and the flexibility
@@
 for tRNA and ATP. However, as the simulation progressed, FBP and ATP
-aggregated around Mg$^{2+}$ and, by consequence, around tRNA{\bf,
-  effectively removing the counter ions from the RNA. IS THAT
-  CORRECT?} {\color{blue} No, Figure \ref{fig:tRNA_aggregation} shows that Mg$^{2+}$ is still concentrated around tRNA in the soup }. After identifying this problem, we used simulations of
+aggregated around Mg$^{2+}$ and, by consequence, around tRNA
+% {\bf,
+%   effectively removing the counter ions from the RNA. IS THAT
+%   CORRECT?} {\color{blue} No, Figure \ref{fig:tRNA_aggregation} shows that Mg$^{2+}$ is still concentrated around tRNA in the soup }. 
+  After identifying this problem, we used simulations of
 small systems containing only the metabolites, Mg$^{2+}$ and water to
@@
 
-To the best of our knowledge, {\bf AGAIN CHECK LATEST FEIG PAPERS} {\color{blue} Latest Feig paper is not a cytoplasm model~\cite{nawrocki2019}}
-there is only published study on large scale cytoplasm models that
+To the best of our knowledge, there is only one published study on large scale cytoplasm models that
 included metabolites with all-atom explicit solvent molecular

### 2019-10-17 paper/virtual_cytoplasm_v2.tex
Review

@@
 same type~\cite{Bashardanesh2019a}.
-{\color{blue} We highlight that there are no specific interactions described between the proteins that were used in our cytoplasm model.} {\bf do we expect specific interactions between the proteins in our cyto model?
+The only specific interaction we antecipate between the components of our cytoplasm model is between TufA and tRNA (PDB ID 1TTT is an example of such complex)~\cite{nissen1995crystal}.
+The TufA-tRNA complex is not present in our simulation boxes even after \SI{1}{\micro\second} of simulation starting from random orientations.
+
+% {\color{blue} We highlight that there are no specific interactions described between the proteins that were used in our cytoplasm model.} {\bf do we expect specific interactions between the proteins in our cyto model?
 In that case write that with references, otherwise write that there are no known specific interactions between these proteins.} {\color{blue} The only specific interaction we antecipate between the components of the soup is that TufA interacts with tRNA (PDB ID 1TTT is an example of such complex).

### 2019-10-17 paper/virtual_cytoplasm_v2.tex
Updated for submission

@@
 it to a total of \SI{3}{\micro\second} of molecular dynamics
-simulations. The simulation model is consistent with experimental data
-about the diffusion coefficient and stability of macromolecules under
-crowded conditions. In order to stimulate further work we provide a
-Python script and a set of files that enables any researcher to build
-their own \textit{E. coli} cytoplasm models to address questions
-related to crowding.
+simulations. The properties of biomolecules under crowding conditions
+are compared to those from simulations of the individual compounds
+under dilute conditions.  The simulation model is found to be
+consistent with experimental data about the diffusion coefficient and
+stability of macromolecules under crowded conditions. In order to
+stimulate further work we provide a Python script and a set of files
+that enables other researchers to build their own \textit{E. coli}
+cytoplasm models to address questions related to crowding.
 \end{abstract}
@@
 activities~\cite{Harada2012a,king2013}.
-
 While the structure and dynamics of biomolecules is well characterized
@@
 among the most relevant to the field due to their ability to probe
-cells in a non destructive way. Another essential class of methods
-come from computational modeling and simulation. By extrapolating the
-advances in computational power in the last decades to the future, it
-was estimated that the atomistic simulation of an entire
-\textit{Escherichia coli} bacterium for \SI{1}{\nano\second} will be
-possible in 15 years from now~\cite{vanGunsteren2006a}.
-% {\bf THAT WAS 13 YEARS AGO.
-WHAT IS THE STATE OF THE ART?} {\color{blue} The paper predicts that simulating a whole E.
-coli bacterium with $10^{11}$ atoms for 1 ns will be possible in 2034, thus the "15 years from now" in our text.
-That was meant as an introduction to the simulation of big systems.
-The evolution of computational models to study crowing and the state-of-the-art is discussed in the next paragraph.
-Maybe we can move this phrase to the beginning of the next paragraph }
+cells in a non destructive way.
 
+A powerful class of methods is computational modeling and simulation.
 As the available computational power increases, models become
@@
 characterization of dynamical effects that can take place in such a
-large scale. It was noted already some years ago that even
-{microsecond} simulations may be too short to probe the dynamics of
-large complexes like viruses~\cite{Larsson2012a} since dynamics is
-slower in larger systems in general.
+large scale. It was estimated, by extrapolating the advances in
+computational power in the last decades to the future, that the
+atomistic simulation of an entire \textit{Escherichia coli} bacterium
+(about 10$^{11}$ atoms) for \SI{1}{\nano\second} will be possible
+around the year 2034~\cite{vanGunsteren2006a}. This prediction may be
+too pessimistic, seeing that ``only'' a factor of 50 is lacking
+between the simulations by Feig and co-workers~\cite{Feig2015} (from
+2015) and the target of a whole bacterium for 1 ns. It was noted
+already some years ago, however, that even {microsecond} simulations
+may be too short to probe the dynamics of large complexes like
+viruses~\cite{Larsson2012a,Larsson2012b} since dynamics is slower in
+larger systems in general. Although a meaningful simulation of a
+bacterium at the atomistic level does not seem to be within reach in
+the foreseeable future, simulations of parts of a cell may still be
+useful to adress biological questions.
 
-In this work, we report an all-atom model of part of the cytoplasm of
-\textit{Escherichia coli} was built to reflect the real biological
+In this work, we report on an all-atom model of part of the cytoplasm
+of \textit{Escherichia coli}, built to reflect the real biological
 system, in particular as regards a realistic composition of this
 ``soup''. We discuss the difficulties that arise when building such a
-system and provide a set of Python scripts and files that anyone can
-use to build its own crowded systems with custom parameters and
+system and provide a set of Python scripts and files that can be used
+to build crowded systems with custom parameters and
 compounds. Additionally, we performed long molecular dynamics
 simulations of our cytoplasm model, reaching \SI{3}{\micro\second} of
-total simulation time. To the best of our knowledge, this work is the first to explore the
-structural dynamics of an all-atom explicit solvent cytoplasm model 
-with microsecond scale molecular dynamics simulations.
+total simulation time. To the best of our knowledge, this work is the
+first to explore the structural dynamics of an all-atom explicit
+solvent cytoplasm model with microsecond scale molecular dynamics
+simulations.
 
@@
 FBP is enough to prevent their aggregation with Mg$^{2+}$. While RDFs
-show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$
+show that ATP$^{3-}$ and FBP$^{4-}$ aggregate around Mg$^{2+}$
 (Fig.~\ref{fig:avoiding_aggregation}A), this is not observed for
@@
 \caption{RDFs centered around Mg$^{2+}$ showing the probability of
-  finding A) metabolites in the cytoplasm model, B) ATP$^{-3}$ and
-  ATPH\textsubscript{3} or C) FBP$^{-4}$ and FBPH\textsubscript{4} in
+  finding A) metabolites in the cytoplasm model, B) ATP$^{3-}$ and
+  ATPH\textsubscript{3} or C) FBP$^{4-}$ and FBPH\textsubscript{4} in
   small independent systems (please see the methods section for
@@
 \section*{Discussion}\label{sec:dissc}
-The main difference between the behavior of the macromolecules that
-compose our model in the crowded system and in the dilute condition is
+The main difference in behavior of the macromolecules that
+compose our model between the crowded system and the dilute condition is
 their translational diffusion coefficient, which drops by around
@@
 the drop in diffusion coefficient is independent of the size of the
-macromolecules. We found that the aggregation of ATP$^{-3}$ and
-FBP$^{-4}$ around Mg$^{2+}$, which was used as counter-ion for tRNA,
+macromolecules. We found that the aggregation of ATP$^{3-}$ and
+FBP$^{4-}$ around Mg$^{2+}$, which was used as counter-ion for tRNA,
 leads to an exaggerated confinement of the tRNA molecules.  Therefore,
-tRNA is considered as an outlier in the duffusion calculations. In
+tRNA is considered an outlier in the duffusion calculations. In
 fact, the slope of the linear regression line fitted to
-$D_{cyto}/D_{dil}$ is reduced by by 50\% and gets even closer to zero
+$D_{cyto}/D_{dil}$ is reduced by by 50\% and  even closer to zero
 (slope $= 0.00014$) after removing the diffusion coefficient of tRNA
@@
 solutions~\cite{andrews2013}. Moreover, it has been reported that
-simulations of highly concentrated environment leads to aggregation
+simulations of a highly concentrated environment leads to aggregation
 when additive force fields are used~\cite{Petrov2014a,Abriata2015a,
@@
 
-Previous
-all-atom simulations of large scale cytoplasm models used parameters from the CHARMM family of
-force fields~\cite{Yu2016a}. In this work we use a modified version of
-the AMBER FF99SB force field in which the solvent-solute interactions
-are 10\% stronger~\cite{Best2014a} to study crowding effects on a
-cytoplasm model. This force field was chosen because it was corrected
-to reproduce experimental data when there are many proteins in the
-same simulation box.
-This force field or similarly adjusted one were also used for studying crowding effect on simpler systems of high concentration of proteins~\cite{Nawrocki2017a,nawrocki2019,Bashardanesh2019a}.
-All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of
-\textit{Mycoplasma genitalium} composed of 103- and 12-million atoms
-that were simulated for 20 and 60 \SI{}{\nano\second},
-respectively~\cite{Yu2016a}. The cytoplasm model described in this
-work has 1.5 million atoms and was simulated for the total of
-\SI{3}{\micro\second}, an unprecedented time scale to study crowding effects on a heterogeneous cytoplasm model in atomic detail.
+Previous all-atom simulations of large scale cytoplasm models used
+parameters from the CHARMM family of force fields~\cite{Yu2016a}. In
+this work we use a modified version of the AMBER FF99SB force field in
+which the solvent-solute interactions are 10\%
+stronger~\cite{Best2014a} to study crowding effects on a cytoplasm
+model. This force field was chosen because it was corrected to
+reproduce experimental data when there are many proteins in the same
+simulation box. This force field or similarly adjusted one were also
+used for studying crowding effect on simpler systems of high
+concentration of
+proteins~\cite{Nawrocki2017a,nawrocki2019,Bashardanesh2019a}. All-atom
+explicit solvent molecular dynamics simulations have been reported for
+large-scale cytoplasm models of \textit{Mycoplasma genitalium}
+composed of 103- and 12-million atoms that were simulated for 20 and
+60 \SI{}{\nano\second}, respectively~\cite{Yu2016a}. The cytoplasm
+model described in this work has 1.5 million atoms and was simulated
+for the total of \SI{3}{\micro\second}, an unprecedented time scale to
+study crowding effects on a heterogeneous cytoplasm model in atomic
+detail.
 
@@
 (Fig.~\ref{fig:structural_integrity_chain}A). Interestingly, oligomers
-were marginally less stable in the cytoplasm model than in the dilute condition.
-(Fig.~\ref{fig:structural_integrity_chain}B) and the flexibility
-profiles calculated considering the C$\upalpha$ for each chain of the
-elements of our cytoplasm model show that, in the crowded condition,
-most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are more flexible than in the dilute condition (Fig.~\ref{fig:rmsf}). This
+were marginally less stable in the cytoplasm model than in the dilute
+condition.  (Fig.~\ref{fig:structural_integrity_chain}B) and the
+flexibility profiles calculated considering the C$\upalpha$ for each
+chain of the elements of our cytoplasm model show that, in the crowded
+condition, most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are
+more flexible than in the dilute condition (Fig.~\ref{fig:rmsf}). This
 is in agreement with previous simulations~\cite{Feig2011} and
@@
 concentration showed only few interactions between proteins of the
-same type~\cite{Bashardanesh2019a}.
-The only specific interaction we antecipate between the components of our cytoplasm model is between TufA and tRNA (PDB ID 1TTT is an example of such complex)~\cite{nissen1995crystal}.
-The TufA-tRNA complex is not present in our simulation boxes even after \SI{1}{\micro\second} of simulation starting from random orientations.
+same type~\cite{Bashardanesh2019a}. The only specific interaction we
+antecipate between the components of our cytoplasm model is between
+TufA and tRNA (PDB ID 1TTT is an example of such
+complex)~\cite{nissen1995crystal}. The TufA-tRNA complex is not
+present in our simulation boxes even after \SI{1}{\micro\second} of
+simulation starting from random orientations.
 
@@
 
-To the best of our knowledge, there is only one published study on large scale cytoplasm models that
-included metabolites with all-atom explicit solvent molecular
-dynamics~\cite{Yu2016a}. Although these authors did not report
-aggregation, they mention that the translational diffusion of highly
-charged phosphate-containing metabolites is ``much slower'' than
-expected in their system. They attributed this observation to
-non-specific interactions between such metabolites and proteins. Maybe
-aggregation also contributed to their slower diffusion but it was not
-evident during their \SI{140}{\nano\second} simulation.
+To the best of our knowledge, there is only one published study on
+large scale cytoplasm models that included metabolites with all-atom
+explicit solvent molecular dynamics~\cite{Yu2016a}. Although these
+authors did not report aggregation, they mention that the
+translational diffusion of highly charged phosphate-containing
+metabolites is ``much slower'' than expected in their system. They
+attributed this observation to non-specific interactions between such
+metabolites and proteins. Maybe aggregation also contributed to their
+slower diffusion but it was not evident during their
+\SI{140}{\nano\second} simulation.
 
@@
 their own \textit{E. coli} cytoplasm models. Such models can be used
-to investigate research questions about the effects of
-crowding on specific systems of interest or about crowding
-itself. With the files we are providing it is possible, for example,
-to add a probe protein to investigate the effect of crowding on it, to
-add new macromolecular or small crowders, evaluate the effects of
- temperatures and change the biomolecular concentration to
-increase or decrease the intensity of the crowding effect. We
-highlight that, since we have already validated it, researchers can
-take advantage from building their own cytoplasm models even if they
-do not have access to the infrastructure that is necessary to perform
-long molecular dynamics simulations. With modest computatioanal power
-it is possible, for example, to build many independent snapshots of
-the cytoplasm of \textit{E. coli}. All files are publicly available at
-http://github.com/dspoel/soup.
+to investigate research questions about the effects of crowding on
+specific systems of interest or about crowding itself. With the files
+we are providing it is possible, for example, to add a probe protein
+to investigate the effect of crowding on it, to add new macromolecular
+or small crowders, evaluate the effects of temperatures and change the
+biomolecular concentration to increase or decrease the intensity of
+the crowding effect. We highlight that, since we have already
+validated it, researchers can take advantage from building their own
+cytoplasm models even if they do not have access to the infrastructure
+that is necessary to perform long molecular dynamics simulations. With
+modest computatioanal power it is possible, for example, to build many
+independent snapshots of the cytoplasm of \textit{E. coli}. All files
+are publicly available at http://github.com/dspoel/soup facilitating
+contributions to the scripts and data files from other workers in the field.
 

### 2019-10-19 scripts/soup-maker.py
Updated soup maker

@@ -4,448 +4,360 @@ import argparse, os, glob, math
 import subprocess
 import numpy as np
 
-export GMX_MAXBACKUP=-1
+os.system('export GMX_MAXBACKUP=-1')
+
 
 def parseArguments():
-    parser = argparse.ArgumentParser()
-    parser.add_argument("-gmx", "--gromacs_command",  type=str,  default="gmx")
-    # parser.add_argument("-cr_n", "--n", help="crowder index",  type=int,  default=0)
+	parser = argparse.ArgumentParser()
+	parser.add_argument("-gmx", "--gmx_path",  type=str,  default="gmx")
     
-    args = parser.parse_args()
-    return args
+	args = parser.parse_args()
+	return args
 
-box_size_initial=35		# in nm
+
+
+
+box_size_initial=35 #35		# in nm
 box_size_step=2	# nm
 
-droplet_thickness_initial=0.6
-
+droplet_thickness_initial=0.6 #0.6
 droplet_shell_precision=2	# percentage as integer
 
 crowder_total=13
-fraction=0.30
+biomolecular_fraction=0.30
 ionic_strength=0.150	# M
 
-make_droplets=0
-make_box=1
-make_top=1
-make_neutral=1
-make_is=1
-check_fraction=1
+make_droplets = False
+make_box      = False
+make_topology = True
 
-path=os.getcwd()
-path_crowders=path
 path_mdp='mdp'
+path_output='soup'
 
 mdp_min='min.mdp'
 
-path_output='soup_4'
 
-# number of gro files for each crowder
-num_crowder[1]=6
-num_crowder[2]=7
-num_crowder[3]=2
-num_crowder[4]=1
-num_crowder[5]=3
-num_crowder[6]=1
-num_crowder[7]=1
-num_crowder[8]=1
-num_crowder[9]=5
-num_crowder[10]=1436
-num_crowder[11]=144
-num_crowder[12]=225
-num_crowder[13]=255
-crowder_type_arr = []
-crowder_mass = []
+
 # number of gro files for each crowder
 
 
-print('Preliminary steps\n')
-print('\tGathering data about each crowder\n')
+crowders = {}
+for c in range(1,crowder_total+1):
+	crowders[c] = {}
 
-crowder=1
-for crowder in range(1, crowder_total+1):
+crowders[1]['copies']  = 6
+crowders[2]['copies']  = 7
+crowders[3]['copies']  = 2
[…]

### 2019-10-25 ingredients-maker.py
Some cleanups and documentation

@@ -1,286 +1,0 @@
-#!/usr/bin/env python3
-
-import argparse, os, glob, math
-import subprocess
-import numpy as np
-
-path="~/soup/"
-path_mdp= "../mdp/" 
-# path_mdp= "%s/mdp/" % (path)
-
-def parseArguments():
-	parser = argparse.ArgumentParser()
-	parser.add_argument("-gmx", "--gromacs_command",  type=str,  default="gmx")
-	parser.add_argument("-cr_n", "--n", help="crowder index",  type=int,  default=0)
-	
-	args = parser.parse_args()
-	return args
-
-
-
-# export GMX_MAXBACKUP=-1
-
-if __name__ == '__main__':
-    args  = parseArguments()
-    gromacs_command = args.gromacs_command
-    n=args.n 				#crowder index
-    crowder_total=n
-    rep_first=1
-    rep_total=3
-    flag_build=1
-    flag_eq=0
-    flag_eq_nvt=0
-    flag_eq_npt=0
-    flag_prod=1
-    mdp_min_none='min_none.mdp'
-    mdp_min='min.mdp'
-    mdp_nvt='nvt.mdp'
-    mdp_npt='npt.mdp'
-    mdp_prod='prod.mdp'
-    topology_template='%s/topology_template.top' % (path)
-    while n <= crowder_total:
-        print('Crowder No. %s:\n' % (str(n)))
-        print('\tChecking the ingredient ... ')
-        # Creates a temporary index file to check what type of molecule is the crowder
-        with open('q.txt', 'w') as outfile:
-            outfile.write("q\n")
-        os.system('%s make_ndx -f crowder_n%s/%s_*pdb -o temporary.ndx < q.txt  >/dev/null 2>/dev/null ' % (gromacs_command, str(n), str(n) ))
-        os.system('rm q.txt')
-        with open('temporary.ndx', 'r') as infile:
-            l=[]
-            for line in infile:
-                if line.find('[') > -1:
-                    l.append(line.strip())
-            if '[ Protein ]' in l:
-                crowder_type = "Protein"
-            elif '[ RNA ]' in l:
-                crowder_type = "RNA"
-            elif '[ Other ]' in l and '[ ATP ]' in l:
-                crowder_type = "Other"
-                is_atp = 1
-            elif '[ Other ]' in l and not '[ ATP ]' in l:
-                crowder_type = "Other"
-                is_atp = 0
-            else:
-                print("Unknown type for crowder_n%s" % str(n))
[…]

### 2019-10-25 soup-maker.py

@@ -1,451 +1,0 @@
-#!/usr/bin/env python3
-
-import argparse, os, glob, math
-import subprocess
-import numpy as np
-
-export GMX_MAXBACKUP=-1
-
-def parseArguments():
-    parser = argparse.ArgumentParser()
-    parser.add_argument("-gmx", "--gromacs_command",  type=str,  default="gmx")
-    # parser.add_argument("-cr_n", "--n", help="crowder index",  type=int,  default=0)
-    
-    args = parser.parse_args()
-    return args
-
-box_size_initial=35		# in nm
-box_size_step=2	# nm
-
-droplet_thickness_initial=0.6
-
-droplet_shell_precision=2	# percentage as integer
-
-crowder_total=13
-fraction=0.30
-ionic_strength=0.150	# M
-
-make_droplets=0
-make_box=1
-make_top=1
-make_neutral=1
-make_is=1
-check_fraction=1
-
-path=os.getcwd()
-path_crowders=path
-path_mdp='mdp'
-
-mdp_min='min.mdp'
-
-path_output='soup_4'
-
-# number of gro files for each crowder
-num_crowder[1]=6
-num_crowder[2]=7
-num_crowder[3]=2
-num_crowder[4]=1
-num_crowder[5]=3
-num_crowder[6]=1
-num_crowder[7]=1
-num_crowder[8]=1
-num_crowder[9]=5
-num_crowder[10]=1436
-num_crowder[11]=144
-num_crowder[12]=225
-num_crowder[13]=255
-crowder_type_arr = []
-crowder_mass = []
-# number of gro files for each crowder
-
-
-print('Preliminary steps\n')
-print('\tGathering data about each crowder\n')
-
-crowder=1
-for crowder in range(1, crowder_total+1):
-	# Creates a temporary index file to check what type of molecule is the crowder
-	print( 'q\n | gmx make_ndx -f crowder_n%s/npt.1.gro -o temporary.ndx >/dev/null 2>/dev/null ' %(crowder))
-	with open('q.txt', 'w') as outfile:
-            outfile.write("q\n")
-    os.system('%s make_ndx -f crowder_n%s/npt.1.gro -o temporary.ndx < q.txt  >/dev/null 2>/dev/null ' % (gromacs_command, str(n), str(n) ))
-    os.remove('q.txt')
-    with open('temporary.ndx', 'r') as infile:
-        l=[]
-	# Is this crowder a protein, RNA or metabolite?
-        for line in infile:
-            if line.find('[') > -1:
[…]

### 2019-10-28 scripts/ingredients-maker.py
Fixing ingredients-maker

@@ -4,13 +4,13 @@ import argparse, os, glob, math
 import subprocess
 import numpy as np
 
-path="~/soup/"
+
 path_mdp= "../mdp/" 
-# path_mdp= "%s/mdp/" % (path)
+
 
 def parseArguments():
 	parser = argparse.ArgumentParser()
-	parser.add_argument("-gmx", "--gromacs_command",  type=str,  default="gmx")
+	parser.add_argument("-gmx", "--gmx_path",  type=str,  default="gmx")
 	parser.add_argument("-cr_n", "--n", help="crowder index",  type=int,  default=0)
 	
 	args = parser.parse_args()
@@ -18,269 +18,214 @@ def parseArguments():
 
 
 
-# export GMX_MAXBACKUP=-1
+
+
+def grp_list( gro ):
+	# Creates a temporary index file to check what type of molecule is the crowder
+#	print( 'q\n | gmx make_ndx -f crowder_n%d/npt.1.gro -o temporary.ndx >/dev/null 2>/dev/null ' %(crowder))
+#	with open('q.txt', 'w') as outfile:
+ #           outfile.write("q\n")
+	os.system('echo "q""\\n" | %s make_ndx -f %s -o temporary.ndx >/dev/null 2>/dev/null' % (gmx_path, gro) )
+	with open('temporary.ndx', 'r') as infile:
+		lines = infile.readlines()
+		grps = [x.split()[1] for x in lines if x[0]=='[']
+	
+	return grps
+	os.remove('temporary.ndx')
+
+
+
+
+
+
+def crowder_type( group_list ):
+	if 'Protein' in group_list:
+		return 'Protein', 'protein'
+	elif 'RNA' in group_list:
+		return 'RNA', 'nucleic acid'
+	elif 'Other' in group_list:
+		return 'Other', 'metabolite'
+	else:
+		return 'Unknown', 'unknown'
+
+
+
+
+
 
 if __name__ == '__main__':
-    args  = parseArguments()
-    gromacs_command = args.gromacs_command
-    n=args.n 				#crowder index
-    crowder_total=n
-    rep_first=1
-    rep_total=3
-    flag_build=1
-    flag_eq=0
-    flag_eq_nvt=0
-    flag_eq_npt=0
-    flag_prod=1
-    mdp_min_none='min_none.mdp'
-    mdp_min='min.mdp'
-    mdp_nvt='nvt.mdp'
-    mdp_npt='npt.mdp'
-    mdp_prod='prod.mdp'
[…]

### 2019-11-20 paper/rebuttal.txt
First update to answer referees

Dear Dr. van der Spoel:

Thank you for submitting your manuscript to Journal of Chemical
 Information and Modeling. Please find the enclosed reviews of your
 manuscript. The reviewers have found significant problems with the
 manuscript in its current form, but we are willing to consider a
 revised manuscript in which you have made major revisions and
 responded to the reviewers' concerns. If you resubmit a revised
 manuscript, there may be a second round of external review, perhaps
 including new reviewers.

We would like to receive your revision as soon as possible, by
17-Jan-2020 at the latest.

In addition to addressing the reviewers' comments, please make each of the
technical corrections listed below:

MANUSCRIPT
- Please remove "et al." from your references and list all authors' names.
ANSWER:

- Update the publication information in Reference 56 and Reference 73 when
available.
ANSWER:

- Please include a Table of Contents graphic on the very last page of
your manuscript.  This graphic, in conjunction with the manuscript
title, should give the reader a summary of the research described in
the paper. The Table of Contents graphic will also be published in the
Abstract of the manuscript. This graphic can be no wider than 8 cm and
no taller than 4 cm and should be prepared following the
specifications under Artwork. Text should be limited to essential
technical labels.  More guidance is available at
http://pubs.acs.org/paragonplus/submission/toc_abstract_graphics_guidelines.pdf.
ANSWER:

- The title page of your manuscript file needs more complete
 affiliation information for all authors. Could you please provide
 updated affiliations including all of the following information
 (where possible)? Department and institute (or company name), city,
 state/region, postal code, country.  Please indicate the
 corresponding author's email address with an asterisk.
ANSWER:

- Funding Sources: Authors are required to report ALL funding sources
  and grant/award numbers relevant to this manuscript. Enter all
  sources of funding for ALL authors relevant to this manuscript in
  BOTH the Open Funder Registry tool in ACS Paragon Plus and in the
  manuscript to meet this requirement. See
  http://pubs.acs.org/page/4authors/funder_options.html for complete
  instructions.
ANSWER:

- ORCID: Authors submitting manuscript revisions are required to
  provide their own validated ORCID iDs before completing the
  submission, if an ORCID iD is not already associated with their ACS
  Paragon Plus user profiles. This iD may be provided during original
  manuscript submission or when submitting the manuscript
  revision. You can provide only your own ORCID iD, a unique
  researcher identifier. If your ORCID iD is not already validated and
  associated with your ACS Paragon Plus user profile, you may do so by
  following the ORCID-related links in the Email/Name section of your
  ACS Paragon Plus account. All authors are encouraged to register for
  and associate their own ORCID iDs with their ACS Paragon Plus
  profiles. The ORCID iD will be displayed in the published article
  for any author on a manuscript who has a validated ORCID iD
  associated with ACS Paragon Plus when the manuscript is
  accepted. Learn more at http://www.orcid.org.
ANSWER:

- On resubmission you must submit the following:
a) A point-by-point response to the reviewers.
b) A clean, unmarked copy of the manuscript document file.
c) A marked copy of the revised manuscript in which all the changes are clearly tracked or highlighted. Upload this file as Supporting Information for Review.
d) A clean, unmarked Supporting Information for Publication file, without any highlighting or tracked changes.

- The reviewers indicate that the quality of the written English could
  be improved. Please consider using a native English speaker or an
  English-language editing service to assist with this. The ACS
  maintains a list of editing services:
  https://authoringservices.acs.org/.

JCIM is now accepting images to appear on the front cover of the
journal. If you have a visually arresting and scientifically
interesting image, please email it and a brief (80 word) description
of the image to eic@jcim.acs.org. Additional information about the
technical specifications for the image can be found at
https://pubsapp.acs.org/paragonplus/submission/jcisd8/jcisd8_authguide.pdf.

Please contact the JCIM editorial office at eic@jcim.acs.org if you
have any specific questions regarding these guidelines.

JCIM has a Twitter feed (@JCIM_ACS; https://twitter.com/jcim_acs)
covering the latest news and highlights in the field. If you would
like us to feature your manuscript on this platform, please share a
Tweet of no more than 140 characters. Please feel free to share your
Twitter handle (if you have one) or any other relevant
department/institutional handles. Include this information in the
cover letter of your revised manuscript submission.

ACS Publications uses Crossref Similarity Check Powered by iThenticate
to detect instances of similarity in submitted manuscripts. In
publishing only original research, ACS is committed to deterring
plagiarism, including self-plagiarism. Your manuscript may be screened
for similarity to published material.

Sincerely,

Prof. Rommie Amaro
Associate Editor Journal of Chemical Information and Modeling
Urey Hall 3234, MS - 0340,
La Jolla, CA 92093-0340,
United States
Email : amaro-office@jcim.acs.org

------------------------------------ To revise your manuscript, log
into ACS Paragon Plus with your ACS ID at
http://paragonplus.acs.org/login and select "My Authoring Activity."
Your manuscript will be listed under "Revisions and Resubmissions
Requested by Editorial Office." Click on "Submit Revision" to begin
the submission process. Your manuscript number will be appended to
denote a revision.

When submitting your revised manuscript, you will be able to respond
to the comments made by the reviewer(s) in the text box provided or by
attaching a file containing your response letter.

------------------------------------

Reviewers' Comments to Author:

Reviewer: 1

Comments: Van der Spoel and co-workers built an atomistic model of the
cytoplasm of Escherichia coli composed of 1.5 million atoms and
investigated its behavior based on 3x1 us of molecular dynamics
(MD) simulations. Some interesting observations were made. The most
important aspects of this work are that i) the authors demonstrated
that it has become possible to perform MD simulations of cytoplasms
involving millions of atoms on the microsecond timescale, and ii) they
provide a Python script and a set of files that enable other
researchers to build their own E. coli cytoplasm models.

The work is of highest importance and novelty. Moreover, the
manuscript is nicely written and easy to follow. My recommendation
thus is to accept this manuscript for publication after following
minor comments have been addressed:

ANSWER: Thanks for the encouraging commments.

1) Why did the authors choose K+ and not Na+ to model the ionic
strength of 150 mmol/L? Please comment.

ANSWER: Like most organisms E. coli has a higher concentration of K+
than Na+ of 30-300 mmol/L versus 10 mmol/L Na+. The number 150 is
halfway the interval. A reference to a text book is added.

2) Figure 1A: Even though it is not of importance for the message
conveyed by this figure, this reviewer nonetheless wonders what the
difference between the red and green particles is in that figure. A
short note on it in the caption would be appreciated.

ANSWER: Red is water and green is biomolecules, this has been added to
the caption.

3) p. 13-15: It is mentioned that the RMSD of TufA, MetE and IcdA is
larger in the cytoplasm than in diluted solutions. Based on the RMSF
they also identified which part of these proteins become more
flexible. The authors later speculate that this is due to some
nonspecific interactions. Here, some more analysis is required. The
authors should demonstrate that these changes indeed arise from such
kind of interactions and not more specific interactions. And why are
only certain loops of each protein involved while all other loops seem
almost unaffected. And why does the RMSF of residues 450-460 of MetE
decrease?

ANSWER:

4) Figure 5: Why are the standard errors of the RMSF for CspC so much
larger than for the other systems. Please comment.

ANSWER:

5) In the discussion on p. 17 about the challenge of reproducing
protein-protein interactions one might also cite
DOI:10.1021/acs.jctc.8b00579 where it was shown that current force
fields have problems to distinguish between peptides known from
experiment to aggregate or not to aggregate. Moreover, increasing the
protein-water interactions as done in Amber03-ws may abolish protein
aggregation almost completely.

ANSWER: This paper was cited as well. Nevertheless, our earlier study
(Zhang et al, JCIM 58, 2018, 1037) showed that the DeltaGhydration are
slightly better using Amber99-sbws than Amber03-ws which is why we
used particular combination.

6) Following spelling/grammatical mistakes should be corrected:
p. 8, l. 29: A verb is missing in that sentence.
p. 10, l. 47: Change "don't" to "do not". =20
p. 11, l. 17: Add a space between "500" and "ps".
p. 15, l. 54: Change "of avoid" to "to avoid".
p. 16, l. 50: Change "duffusion" to "diffusion".

ANSWER: All fixed, thanks for finding these.

Reviewer: 2

Comments: Bortot et al. studied protein crowded systems using
molecular simulations with explicit solvent. MD simulations of
macromolecules in crowded conditions help us to understand protein
behaviors in a living cell. Recently, several computational scientists
performed MD simulations of multiple proteins in crowded conditions
and examined the differences of protein structure and dynamics in
crowded condition compared with that in dilute solution. In spite of
the importance of the research field, the paper has three critical
issues. At first, they used only small number of proteins and
tRNA(Phe) in the simulation system, although they call the set up as
the E. coli cytoplasm. Since the proteins can cover only 50 % of the
fraction of E. coli cytoplasm, it is difficult to call the system as
the E. coli cytoplasm. It should be called as a crowded condition that
mimic 50% of volume fraction of the cytosol of E. coli.

ANSWER:

They did not include ribosomes in the system. However, the size
differences between ribosomes and other proteins should be essential
in the cellular crowding effect.

ANSWER:

Secondly, they observed tRNA
aggregation with Mg2+, ATP, and FBP in their systems. They concluded
that completely protonating the phosphate groups of ATP and FBP is
necessary to prevent their aggregation with Mg2+. However, there is no
rationale or experimental evidences on the protonation state of
ATP. Presumably, the Mg2+ model or phosphate model may have
problems. Another possibility is that they used Amber99sb-ws, where
the 1.10 scaling is used for LJ interactions between water and
macromolecules. If they applied the scaling also to tRNA, tRNA
structure could be unfolded. However, there is no data about tRNA
structural changes in crowded condition compared to that in
solution.

ANSWER:

Thirdly, except for the protonation state, there are few novel
biophysical insights in the current work. Essentially same data on
diffusion or RMSD were already shown in the previous crowding papers,
although the length of simulations and the force fields are
different. Although the modeling strategy (Figure 1) seems to be
interesting, the rest of scientific results and discussion may have
rooms for improvements. In particular, the issue of the aggregation of
tRNA in the simulations should be overcome by the refinement of ion
models or the tRNA-water interaction scaling factors.

ANSWER:

### 2019-11-20 paper/virtual_cytoplasm_v2.tex
Some cleanups and documentation
First update to answer referees

@@
 \subsection{General simulation setup}
-The cytoplasm model was built at a 30\% biomolecular mass fraction and
-a physiological salt concentration of 0.15 mol/L KCl.
+The cytoplasm model was built at a 30\% biomolecular mass fraction.
+Like most organisms E. coli has a higher concentration of K$^+$
+(30-300 mmol/L) than Na$^+$ (10 mmol/L)~\cite{Milo2017a}. Therefore a salt
+concentration of 150 mmol/L KCl was used.
+
 Dropplets containing each component of the model surronded by water and counter-ions were inserted in an empty box in random positions and orientations.
@@
 RNA presence in the cytoplasm with tRNA molecules. Specifically, 
- the tRNA(Phe) molecule as a representative of tRNAs due to
+ the tRNA(Phe) molecule was used as a representative of tRNAs due to
 the availability of a recent crystal
@@
 of the system and the desired biomolecular fraction, which is a
-parameter of the model. However, we don't know the thickness of the
+parameter of the model. However, we do not know the thickness of the
 water layer around each component, $l$, that accounts for such amount
@@
 strength. Then, we perform energy minimization and a short simulation
- of 500ps in which all molecules of the box are free to move. In
+ of 500 ps in which all molecules of the box are free to move. In
 this step the box shrinks to its optimum size. In our case, the
@@
 \caption{A) The initial simulation box with $L=35.0$ nm and the box
-  after the shrinking step, $L=22.9$ nm. B) The evolution of the
+  after the shrinking step, $L=22.9$ nm. Biomolecules in green, water in red. B) The evolution of the
   volume of the simulation box during the shrinking simulations. Each
@@
 counter-ion for ATP and tRNA, and so tRNA is one of the few
-``sources'' of Mg$^{2+}$ in the whole system. We looked for ways of
+``sources'' of Mg$^{2+}$ in the whole system. We looked for ways to
 avoid aggregation by using simulations of small systems composed of
@@
 leads to an exaggerated confinement of the tRNA molecules.  Therefore,
-tRNA is considered an outlier in the duffusion calculations. In
+tRNA is considered an outlier in the diffusion calculations. In
 fact, the slope of the linear regression line fitted to
@@
 simulations of a highly concentrated environment leads to aggregation
-when additive force fields are used~\cite{Petrov2014a,Abriata2015a,
-  Nawrocki2017a} and several approaches have been proposed to
+when additive force fields are used~\cite{Petrov2014a,Abriata2015a,Nawrocki2017a,Carballo2018a} and several approaches have been proposed to
 eliminate this
-problem~\cite{Best2014a,Piana2015a,Bashardanesh2018b}. Apart from
+problem~\cite{Best2014a,Piana2015a,Bashardanesh2018b,Carballo2018a}. Apart from
 protein-water interactions, the London-dispersion coefficients in the
@@
 
+\begin{acknowledgement}
+The Swedish research council is acknowledged for a grant of computer time
+(SNIC2018-2-42) through the High Performance Computing
+Center North in Ume{\aa}, Sweden. Funding from eSSENCE - The e-Science
+Collaboration (Uppsala-Lund-Ume{\aa}, Sweden) is gratefully acknowledged.
+\end{acknowledgement}
+
 \bibliography{library}

### 2019-11-20 paper/rebuttal.txt
Answered referee 2

@@
 - Please remove "et al." from your references and list all authors' names.
-ANSWER:
+ANSWER: Done.
 
@@
 available.
-ANSWER:
+ANSWER: Done.
 
@@
   instructions.
-ANSWER:
+ANSWER: Funding information is and was present.
 
@@
 
-ANSWER: This paper was cited as well. Nevertheless, our earlier study
+ANSWER: This paper is now cited as well. Nevertheless, our earlier study
 (Zhang et al, JCIM 58, 2018, 1037) showed that the DeltaGhydration are
@@
 
-ANSWER:
+ANSWER: Obviously our model is not complete, to reflect this we
+changed the title of the paper to "Making Soup: Preparing and
+Validating Models of the Bacterial Cytoplasm for Molecular
+Simulation". In the abstract the following sentence is added "The
+model consists of 8 different proteins and one type of t-RNA molecule,
+representing about 50% of the cytoplasmic proteins."
 
@@
 
-ANSWER:
+ANSWER: This was a deliberate omission. A sentence was added to
+explain it: "It has been estimated that a {\em E. coli} cell contains
+about 50,000 ribosomes per $\mu^3$~\cite{Milo2017a}. The volume of our
+box is about 12,000 nm$^3$ (\figref{shrink}), therefore in principle
+one ribosome molecule should be added, however to reduce the
+complexity this was omitted."
 
@@
 
-ANSWER:
+ANSWER: There may be a number of possible solutions to the issue. The
+work by Leotyev et al. (Refs 78 and 79) offers a plausible explanation
+of the problem, namely that in fixed-charge models, that is
+non-polarizable models, the infinite frequency dielectric constant
+should be incorporated, either by introducing a relative dielectric
+constant eps(inf) of approximately 1.7, or by scaling the charges down
+by a factor of sqrt(1.7). Leontyev and co-workers advocate scaling the
+charges since eps(inf) is somewhat material dependent. Such a solution
+was described in the manuscript already, but a new sentence was added:
+"Due to the omission of the electronic polarization, the screening is
+underestimated and this corresponds to applying an effective
+dielectric constant of about 1.7, alternatively scale the charges down
+by $\sqrt{1.7}$ which would mean that FBP would get a charge of
+approximately -3 instead of -4."
 
@@
 
-ANSWER:
+ANSWER: We agree that the amount of new biophysical insights is
+limited. However, the purpose of this paper was rather to pave the way
+for further in-depth studies by collecting information and scripts
+allowing others to build on this work. For instance, as suggested by
+this referee, one could add ribosomes in larger models for the E. coli
+cytoplasm. We therefore think the present paper is particularly
+suitable for the journal of chemical information and modeling.

### 2019-11-20 paper/virtual_cytoplasm_v2.tex
Answered referee 2

@@
-\documentclass[journal=jcisd8,manuscript=article]{achemso}
+\documentclass[journal=jpcisd8,manuscript=article]{achemso}
 % \linespread{1.5}
+\setkeys{acs}{etalmode=truncate,maxauthors=0}
 \usepackage{amssymb,amsmath,graphics,epsfig}
@@
 
-\title{Making Soup: Preparing and Validating Molecular Simulations of the Bacterial Cytoplasm}
+\title{Making Soup: Preparing and Validating Models of the Bacterial Cytoplasm for Molecular Simulation}
 
@@
 \author{Zahedeh Bashardanesh}
-\affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology. Uppsala University, SE-751 05 Uppsala, Sweden}
+\affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology, Uppsala University, Box 596, SE-75124 Uppsala, Sweden}
 \altaffiliation{These authors contributed equally to this work}
@@
 \author{David van der Spoel}
-\affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology. Uppsala University, SE-751 05 Uppsala, Sweden}
+\affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology, Uppsala University, Box 596, SE-75124 Uppsala, Sweden}
 \email{david.vanderspoel@icm.uu.se}
@@
 it to a total of \SI{3}{\micro\second} of molecular dynamics
-simulations. The properties of biomolecules under crowding conditions
-are compared to those from simulations of the individual compounds
-under dilute conditions.  The simulation model is found to be
-consistent with experimental data about the diffusion coefficient and
-stability of macromolecules under crowded conditions. In order to
-stimulate further work we provide a Python script and a set of files
-that enables other researchers to build their own \textit{E. coli}
-cytoplasm models to address questions related to crowding.
+simulations. 
+The model consists of 8 different proteins and one type of t-RNA molecule,
+representing about 50\% of the cytoplasmic proteins. 
+Properties of biomolecules under crowding conditions are compared to
+those from simulations of the individual compounds under dilute
+conditions.  The simulation model is found to be consistent with
+experimental data about the diffusion coefficient and stability of
+macromolecules under crowded conditions. In order to stimulate further
+work we provide a Python script and a set of files that enables other
+researchers to build their own \textit{E. coli} cytoplasm models to
+address questions related to crowding.
 \end{abstract}
@@
 {\em Escheria coli}~\cite{Dong1996,Bennett2009,Link1997,Mcguffee2010}.
+It has been estimated that a {\em E.
+coli} cell contains about 50,000 ribosomes per $\mu^3$~\cite{Milo2017a}.
+The volume of our box is about 12,000 nm$^3$ (\figref{shrink}), therefore in principle one ribosomes molecule should be added, however to reduce the complexity this was omitted.
 
@@
   replica is shown in different colors. }
+\label{shrink}
 \end{figure}
@@
 employ scaled charges in order to be compatible with the empirical
-water models~\cite{Leontyev2009a,Leontyev2011a}. Due to the high local
+water models~\cite{Leontyev2009a,Leontyev2011a}. Due to the omission
+of the electronic polarization, the screening is underestimated and this
+corresponds to applying an effective dielectric constant of about 1.7, alternatively scale the charges down by $\sqrt{1.7}$ which would mean that FBP would get a charge of approximately -3 instead of -4.
+Due to the high local
 density of negative charges in FBP and ATP this might indeed be an
@@
 We submitted this model to a total of \SI{3}{\micro\second} of
-molecular dynamics simulations using a modified AMBER FF99SB force
+molecular dynamics simulations using a modified Amber99sb force
 field that was tuned to reproduce experimental data of systems

### 2019-11-20 paper/virtual_cytoplasm_v2.tex
Added TOC

@@
 alternative solution, barring use of polarizable
-models~\cite{Ponder2010a,Lopes2013a,Ghahremanpour2018b,Walz2018a}.
+models~\cite{Ponder2010a,Lopes2013a,Ghahremanpour2018b,Walz2018a,Walz2019b}.
 
@@
 % \end{addendum}
+\subsection*{For Table of Contents Only}
+\begin{figure}[htp] 
+ \centering
+  \includegraphics[height=4.45cm]{soup-toc.pdf}
+\end{figure}
 

### 2019-11-21 paper/virtual_cytoplasm_v2.tex
Added reference

@@
 \author{Zahedeh Bashardanesh}
-\affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology, Uppsala University, Box 596, SE-75124 Uppsala, Sweden}
+\affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology, Uppsala University, Box 596, SE-75124 Uppsala, Sweden.}
 \altaffiliation{These authors contributed equally to this work}
@@
 \author{David van der Spoel}
-\affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology, Uppsala University, Box 596, SE-75124 Uppsala, Sweden}
+\affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology, Uppsala University, Box 596, SE-75124 Uppsala, Sweden.}
 \email{david.vanderspoel@icm.uu.se}
@@
 employ scaled charges in order to be compatible with the empirical
-water models~\cite{Leontyev2009a,Leontyev2011a}. Due to the omission
+water models~\cite{Leontyev2009a,Leontyev2011a}, for a recent review, 
+see~\cite{Kirby2019a}. Due to the omission
 of the electronic polarization, the screening is underestimated and this

### 2019-11-21 paper/rmsf.py
rmsf script

@@ -1,0 +1,83 @@
+#!/usr/bin/env python3
+
+import argparse, os, glob, math
+import matplotlib.pyplot as plt
+import numpy as np
+
+
+width  = 7.84
+height = 13/2.
+crowder_name=["TufA", "MetE", "IcdA", "AhpC", "CspC", "Ppa ", "GapA", "Eno", "tRNA$^{Phe}$"]
+
+fig, (ax1, ax2, ax3, ax4, ax5, ax6, ax7, ax8) = plt.subplots(8, sharey=False, sharex=True)
+
+
+
+fig.set_figheight(height)
+fig.set_figwidth(width)
+fs=16
+c=0
+# for n in range(1,9):
+#     c=1+c
+#     for base_dir in ["../singles", "../cyto"]:
+#         L = base_dir.split("../")[1]
+#         if L == "singles":
+#             lab="Dilute"
+#         elif L == "cyto":
+#             lab = "Cyto."
+#         xvg = "%s/rmsf_ca*crowder_n%s*xvg" % (base_dir, str(n))
+#         rmsf_xvg = glob.glob(xvg)
+#         N = len(rmsf_xvg)
+#         V=[]
+#         for file in rmsf_xvg:
+#             v = []
+#             with open(file, 'r') as infile:
+#                 for line in infile:
+#                     if line.find("@") < 0 and line.find("#") < 0:
+#                         data = line.split()
+#                         v.append(float(data[1]))
+#             if n == 4:
+#                 v = v[0:163]
+#             V.append(v)
+#         fig_s = plt.subplot(4,2,c)
+
+#         plt.errorbar(list(range(1,len(v)+1)), np.mean(V, axis=0), np.std(V, axis=0)/math.sqrt(N-1), label=lab)
+#         plt.tick_params(labelsize=fs-2);
+#         plt.ylabel(crowder_name[c-1],fontsize=fs)
+#         if n == 1:
+#             fig_s.legend(fontsize=fs-3)
+#         if n == 2:
+#             plt.yticks([0, .5, 1])
+#         elif n == 7 or n == 8:
+#             plt.xlabel('Residue', fontsize=fs)
+# fig.text(-0.03,0.6,'RMSF ($nm$)',fontsize=fs+2, rotation=90)        
+# fig.tight_layout()
+# plt.savefig("rmsf.pdf", bbox_inches="tight")
+
+
+
+for base_dir in ["../singles", "../cyto"]:
[…]

### 2019-11-29 paper/rebuttal.txt
Added TOC
Updated cover
answers to referees

@@
 http://pubs.acs.org/paragonplus/submission/toc_abstract_graphics_guidelines.pdf.
-ANSWER:
+ANSWER: Done.
 
@@
   instructions.
-ANSWER: Funding information is and was present.
+ANSWER: Funding information was updated.
 
@@
 
-ANSWER:
+ANSWER: As explained in the text, the region comprised by residues 40-62 of TufA is a conformational switch that is known to be flexible, residues 450-460 of MetE are a flexible loop and residues 1-8 of IcdA is a flexible coil.
+Thus, one can expect these regions to display variable conformational dynamics under different conditions.
+We removed the speculation about the unspecific interactions and the discussion about the RMSF difference (please see the answer to the next answer for details about this).
 
@@
 
-ANSWER:
+ANSWER: Thank you for pointingt that out. There was a mistake in the statistics of our RMSF calculations.
+We corrected it and updated the RMSF plots.
+The errors are much smaller now and the overall shift towards higher RMSF values for the proteins cytoplasm models is no longer present.
+Accordingly, the following sentences were removed from the text:
+
+"In most cases (TufA, MetE, AhpC, CspC and Ppa), the residue-wise RMSF values are higher in the cytoplasm model than under diluted conditions, which indicates that proteins are more flexible in the cytoplasm model (Fig.~\ref{fig:rmsf}).
+This is in agreement with computational and experimental data showing that nonspecific interactions with crowders can lead to destabilization of proteins~\cite{Feig2011, miklos2011, Wang2012b}." from the results section.
+
+" and the flexibility profiles calculated considering the C$\upalpha$ for each
+chain of the elements of our cytoplasm model show that, in the crowded
+condition, most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are
+more flexible than in the dilute condition (Fig.~\ref{fig:rmsf})." from the conclusion section.
+
+"and \ref{fig:rmsf})" was removed from "the proteins are slightly less stable
+in the cytoplasm model than in dilute condition (Figures
+\ref{fig:structural_integrity_chain} and \ref{fig:rmsf})." in the conclusion section
 

### 2019-11-29 paper/virtual_cytoplasm_v2.tex
answers to referees

--- 2019-11-20 paper/virtual_cytoplasm_v2.tex
@@
 
-In most cases (TufA, MetE, AhpC, CspC and Ppa), the residue-wise RMSF values are higher in the cytoplasm model than under diluted conditions, which indicates that proteins are more flexible in the cytoplasm model (Fig.~\ref{fig:rmsf}).
-This is in agreement with computational and experimental data showing that nonspecific interactions with crowders can lead to destabilization of
-proteins~\cite{Feig2011, miklos2011, Wang2012b}.
-
 \begin{figure}[H]
@@
 were marginally less stable in the cytoplasm model than in the dilute
-condition.  (Fig.~\ref{fig:structural_integrity_chain}B) and the
-flexibility profiles calculated considering the C$\upalpha$ for each
-chain of the elements of our cytoplasm model show that, in the crowded
-condition, most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are
-more flexible than in the dilute condition (Fig.~\ref{fig:rmsf}). This
+condition  (Fig.~\ref{fig:structural_integrity_chain}B). This
 is in agreement with previous simulations~\cite{Feig2011} and
@@
 in the cytoplasm model than in dilute condition (Figures
-\ref{fig:structural_integrity_chain} and \ref{fig:rmsf}).
+\ref{fig:structural_integrity_chain}.
 
@@
 Center North in Ume{\aa}, Sweden. Funding from eSSENCE - The e-Science
-Collaboration (Uppsala-Lund-Ume{\aa}, Sweden) is gratefully acknowledged.
+Collaboration (Uppsala-Lund-Ume{\aa}, Sweden) and FAPESP - S{\~a}o Paulo Research Foundation (grants 2013/00927-7 and 2016/04958-2) are gratefully acknowledged.
 \end{acknowledgement}

### 2019-12-02 paper/rebuttal.txt
Final version

@@
  corresponding author's email address with an asterisk.
-ANSWER:
+ANSWER: It is complete.
 
@@
   https://authoringservices.acs.org/.
+ANSWER: TODO
 
@@
 have any specific questions regarding these guidelines.
+ANSWER: A cover image will be suggested.
 
@@
 
-ANSWER: As explained in the text, the region comprised by residues 40-62 of TufA is a conformational switch that is known to be flexible, residues 450-460 of MetE are a flexible loop and residues 1-8 of IcdA is a flexible coil.
-Thus, one can expect these regions to display variable conformational dynamics under different conditions.
-We removed the speculation about the unspecific interactions and the discussion about the RMSF difference (please see the answer to the next answer for details about this).
+ANSWER: As explained in the text, the region comprised by residues
+40-62 of TufA is a conformational switch that is known to be flexible,
+residues 450-460 of MetE are a flexible loop and residues 1-8 of IcdA
+form a flexible coil. Thus, one can expect these regions to display
+variable conformational dynamics under different conditions. We
+removed the speculation about the unspecific interactions and the
+discussion about the RMSF difference (please see the answer to the
+next answer for details about this).
 
@@
 
-ANSWER: Thank you for pointingt that out. There was a mistake in the statistics of our RMSF calculations.
-We corrected it and updated the RMSF plots.
-The errors are much smaller now and the overall shift towards higher RMSF values for the proteins cytoplasm models is no longer present.
-Accordingly, the following sentences were removed from the text:
+ANSWER: Thank you for pointingt that out. There was a mistake in the
+statistics of our RMSF calculations.  We corrected it and updated the
+RMSF plots. The errors are much smaller now and the overall shift
+towards higher RMSF values for the proteins cytoplasm models is no
+longer present. Accordingly, the following sentences were removed from
+the text:
 
-"In most cases (TufA, MetE, AhpC, CspC and Ppa), the residue-wise RMSF values are higher in the cytoplasm model than under diluted conditions, which indicates that proteins are more flexible in the cytoplasm model (Fig.~\ref{fig:rmsf}).
-This is in agreement with computational and experimental data showing that nonspecific interactions with crowders can lead to destabilization of proteins~\cite{Feig2011, miklos2011, Wang2012b}." from the results section.
+"In most cases (TufA, MetE, AhpC, CspC and Ppa), the residue-wise RMSF
+values are higher in the cytoplasm model than under diluted
+conditions, which indicates that proteins are more flexible in the
+cytoplasm model (Fig.~\ref{fig:rmsf}). This is in agreement with
+computational and experimental data showing that nonspecific
+interactions with crowders can lead to destabilization of
+proteins~\cite{Feig2011, miklos2011, Wang2012b}." from the results
+section.
 
-" and the flexibility profiles calculated considering the C$\upalpha$ for each
-chain of the elements of our cytoplasm model show that, in the crowded
-condition, most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are
-more flexible than in the dilute condition (Fig.~\ref{fig:rmsf})." from the conclusion section.
+" and the flexibility profiles calculated considering the C$\upalpha$
+for each chain of the elements of our cytoplasm model show that, in
+the crowded condition, most of the proteins (TufA, MetE, AhpC, CspC
+and Ppa) are more flexible than in the dilute condition
+(Fig.~\ref{fig:rmsf})." from the conclusion section.
 
-"and \ref{fig:rmsf})" was removed from "the proteins are slightly less stable
-in the cytoplasm model than in dilute condition (Figures
-\ref{fig:structural_integrity_chain} and \ref{fig:rmsf})." in the conclusion section
+"and \ref{fig:rmsf})" was removed from "the proteins are slightly less
+stable in the cytoplasm model than in dilute condition (Figures
+\ref{fig:structural_integrity_chain} and \ref{fig:rmsf})." in the
+conclusion section
 

### 2019-12-02 paper/virtual_cytoplasm_v2.tex
Final version

@@
 Biomolecular crowding affects the biophysical and biochemical behavior
-of macromolecules when compared to the dilute environment present in
-experiments made with isolated proteins. Computational modeling and
+of macromolecules when compared to the dilute environment  in
+experiments on isolated proteins. Computational modeling and
 simulation are useful tools to study how crowding affects the
@@
 computational power increased, modeling and simulating large scale
-all-atom explicit solvent models of the prokaryote cytoplasm become
+all-atom explicit solvent models of the prokaryote cytoplasm has now become
 possible. In this work, we build an atomistic model of the cytoplasm
@@
 simulations. 
-The model consists of 8 different proteins and one type of t-RNA molecule,
-representing about 50\% of the cytoplasmic proteins. 
+The model consists of 8 different proteins representing about 50\% of the cytoplasmic proteins and one type of t-RNA molecule. 
 Properties of biomolecules under crowding conditions are compared to
@@
 macromolecules under crowded conditions. In order to stimulate further
-work we provide a Python script and a set of files that enables other
+work we provide a Python script and a set of files to enable other
 researchers to build their own \textit{E. coli} cytoplasm models to
@@
 heterogeneous with high concentrations of many different compounds,
-leading to spatial constrains that directly affects biological
-activities~\citep{ostrowska2019}. This effect, called macromolecular
+leading to spatial constraints that directly affects biological
+activity~\citep{ostrowska2019}. This effect, called macromolecular
 crowding, influences the stability of many macromolecules
 because the equilibrium is shifted towards a stable native structure
-as a result of the entropic component given by the excluded volume
+as a result of the entropic component of the Gibbs energy, given by the excluded volume
 effect~\cite{cheung2005}. At the same time, an effective
@@
 While the structure and dynamics of biomolecules is well characterized
-\textit{in vitro}, the understanding of the influences of crowded
-environments \textit{in vivo} are still evolving.  Techniques such as
+\textit{in vitro}, the understanding of the influence of crowded
+environments \textit{in vivo} is still evolving.  Techniques such as
 nuclear magnetic resonance~\cite{reckel2007,pielak2008} and
 fluorescence spectroscopy~\cite{ignatova2004,xie2008,English2011} are
-among the most relevant to the field due to their ability to probe
+among the most important to the field due to their ability to probe
 cells in a non destructive way.
 
-A powerful class of methods is computational modeling and simulation.
+Another powerful class of methods is computational modeling and simulation.
 As the available computational power increases, models become
-increasingly complex and are better representations of reality. The
-most simple class of computational models apply Brownian Dynamics
+increasingly complex yielding better representations of reality. The
+ simplest class of computational models applies Brownian Dynamics
 calculations to probe macromolecules immersed in a solution of
@@
 \SI{140}{\nano\second} for a smaller version, precluding an accurate
-characterization of dynamical effects that can take place in such a
-large scale. It was estimated, by extrapolating the advances in
+characterization of dynamical effects that  take place in such a
+large system. It was estimated, by extrapolating the advances in
 computational power in the last decades to the future, that the
@@
 charges~\cite{Bayly1993a} that were derived from electron density
-calculated at the HF/6-31G* theory level with Gaussian16.  Properties
+calculated at the HF/6-31G* theory level with Gaussian16~\cite{g16}.  Properties
 of compounds modeled using the GAFF have been evaluated in a number of
@@
 
-Dropplets containing each component of the model surronded by water and counter-ions were inserted in an empty box in random positions and orientations.
-Since the volume of this box was initially bigger than what was strictly necessary to accomodate all components, the system was allowed to shrink in a short simulation.
+Droplets containing each component of the model surronded by water and counter-ions were inserted in an empty box in random positions and orientations.
+Since the volume of this box  initially was larger than  strictly necessary to accomodate all components, the system was allowed to shrink during a short simulation.
 Details about how the simulation box was built and optimized are in the results section.
@@
 runs for the dilute systems, in which each crowder is simulated alone
-in a water box, had a- length of 200 ns. The cytoplasm model was run
+in a water box, had a length of 200 ns. The cytoplasm models were run
 for 1 \SI{}{\micro\second}. All systems were simulated in three
-replicas. While the replicas for the diluted systems were generated by
-different initial velocities, each replica of the cytoplasm model is a
+replicas.  The replicas for the dilute systems were generated by
+different initial velocities, but each replica of the cytoplasm model was taken to be a
 completely different box in which the components of the model are in
@@
 
-Analyses were done using GROMACS tools after removing artifacts
+Analyses were done using GROMACS tools~\cite{Lindahl2001a,Spoel2005a,Hess200a} after removing artifacts
 generated by periodic boundary conditions from the trajectories. A
@@
 for cytoplasm system~\cite{Allen1987a}. In principle, diffusion
-coefficient needs to be corrected for finite size
+coefficients need to be corrected for finite size
 effects~\cite{Yeh2004,Abraham2018}. Due to relatively large simulation boxes
@@
 In this section we describe the rationale behind the composition of
-our model, which consists of five fractions: protein, RNA, metabolites, water
-and ions. It should be noted that we did not add lipids and DNA to our
-model because we consider only elements that are free to diffuse
-through the cytosol. We gathered data from several sources in order to
-build a computational model that is representative of the cytoplasm of
-{\em Escheria coli}~\cite{Dong1996,Bennett2009,Link1997,Mcguffee2010}.
-It has been estimated that a {\em E.
-coli} cell contains about 50,000 ribosomes per $\mu^3$~\cite{Milo2017a}.
-The volume of our box is about 12,000 nm$^3$ (\figref{shrink}), therefore in principle one ribosomes molecule should be added, however to reduce the complexity this was omitted.
+our model, which consists of five fractions: protein, RNA,
+metabolites, water and ions. It should be noted that we did not add
+lipids and DNA to our model because we consider only elements that are
+free to diffuse through the cytosol. We gathered data from several
+sources in order to build a computational model that is representative
+of the cytoplasm of {\em Escheria
+  coli}~\cite{Dong1996,Bennett2009,Link1997,Mcguffee2010}.  It has
+been estimated that a {\em E. coli} cell contains about 50,000
+ribosomes per $\mu^3$~\cite{Milo2017a}. The volume of our box is about
+12,000 nm$^3$ (\figref{shrink}), therefore in principle one (0.6 to be
+precise) ribosome molecule should be added, however to reduce the
+complexity this was omitted.
 
@@
 oligomers. This was done because the original oligomeric state for
-each protein was kept as reported in their crystallographic structure
+each protein was kept as reported in their respective crystallographic structure
 (Table~\ref{tbl:protein_fraction}).
@@
 weight ratio was used to calculate the correct number of tRNA(Phe)
-molecules that were added to the cytoplasm model
-\tabref{tbl:soup_components}.
+molecules that were added to the cytoplasm model (\tabref{tbl:soup_components}).
 
 \subsubsection{Metabolites fraction}
-The most abundant molecules of each metabolite class was used to 
+The most abundant molecule of each metabolite class was used to 
 represent the class, i.e. Glutamate (GLT) for amino acids, ATP for
@@
 higher than the number of proteins~\cite{Bennett2009}. The copy number
-for each molecule was calculated from the ratios between their
+for each molecule (\tabref{tbl:soup_components}) was calculated from the ratios between their
 experimentally observed concentration in {\em
@@
 model. Macromolecular concentration ranges from 300 to 400 g/L in biological systems
-such as \textit{E. coli} cytoplasm~\cite{Zimmerman1991}, the highest concentration corresponds to volume occupation as high as 40\%~\cite{Ellis2003a}. In our case, we
-choose a biomolecular concentration of 30\%, that is, the number of
+such as the \textit{E. coli} cytoplasm~\cite{Zimmerman1991}, the highest concentration corresponds to volume occupation as high as 40\%~\cite{Ellis2003a}. In our case, we
+choose a biomolecular concentration of 30\%, that is, the number of water
 molecules necessary to reach a ratio of total biomolecular mass to
@@
 ATP. K$^{+}$ and Cl$^{-}$ were added to neutralize the charges of the
-simulation box and to reach the ionic strength of 0.150 mol/L by
+simulation box and to reach the ionic strength of 150 mmol/L by
 substitution of randomly selected water molecules.
@@
 \subsection{Building the simulation box}
-All components can be put in the same simulation box by inserting each
+All components were put in the same simulation box by inserting each
 of them in random orientations in a cubic volume of side $L$ that is
-initially empty. However, that is not a trivial process. We need to
+initially empty. However, that is not a trivial process. It is necessary  to
 use a box size that is big enough to allow the random insertions to
 succeed without structural overlapping and without creating artificial
-interactions between the elements by placing them too close from each
+interactions between the elements by placing them too close to each
 other. On the other hand, as the box gets bigger, it gets harder to
@@
 embedded. Such droplets are taken from molecular dynamics simulations
-in which each component was previously equilibrated. The benefits of
+in which each component was equilibrated previously. The benefits of
 using such droplets are threefold: \textit{i.} it acts as natural
@@
 component. \textit{iii.} the components are already pre-equilibrated,
-which will help us to perform the equilibration of the whole cytoplasm
+which will speed up the equilibration of the whole cytoplasm
 model.
@@
 parameter of the model. However, we do not know  the thickness of the
-water layer around each component, $l$, that accounts for such amount
+water layer around each component, $l$, that yields such amount
 of water molecules \textit{a priori}. Since $l$ depends on a series of
@@
 this process until a user-defined precision cutoff, here 5\%, for the
-number of waters in the system is satisfied.
+number of water molecules in the system is satisfied.
 
@@
 diffusion constant of each component in the cytoplasm model,
-$D_{cyto}$, and in a dilute system, $D_{dil}$
+$D_{cyto}$ to that in a dilute system, $D_{dil}$
 (Fig.~\ref{fig:translational_diffusion}A). The ratio of these
@@
 crowding on the translational diffusion of tRNA is two times
-larger. Its $D_{cyto}/D_{dil}$ is 0.07, indicating that its movement
+larger. It's $D_{cyto}/D_{dil}$ is 0.07, indicating that it's movement
 is especially constrained in the crowded environment
@@
 \caption{A) Diffusion coefficient of crowders from simulations under
-  diluted conditions (blue bars) and cytoplasm simulations (orange
-  bars). B) The ratio between diffusion coefficient obtained from
+  dilute conditions (blue bars) and cytoplasm simulations (orange
+  bars). B) The ratio between diffusion coefficients obtained from
   simulations under both conditions. The crowders are sorted according
@@
 cytoplasm model is that it was forming aggregates with Mg$^{2+}$, ATP
-and FBP (Fig.~\ref{fig:tRNA_aggregation}). Thus, we will omit the tRNA
+and FBP (Fig.~\ref{fig:tRNA_aggregation}). Thus, we decided to omit the tRNA
 molecules in the analyses about the structural integrity of the
@@
 \subsubsection{Structural integrity}
-The structure of the individual chains of the crowders that were used
+The structures of the individual chains of the crowders that were used
 in our cytoplasm model were not affected to a great extent by the
-crowded environment. The values of their root mean square deviation considering only
-their C$\upalpha$ in the cytoplasm model increased by less than 11\% when compared to the dilute condition, except for
+crowded environment. The values of their root mean square deviation, considering only
+their C$\upalpha$ atoms, in the cytoplasm model increased by less than 11\% when compared to the dilute condition, except for
 TufA, MetE and IcdA, which have RMSD values 105, 25 and 33\% higher in
@@
 
-  %{\bf IS THERE A KNOWN PHYSIOLOGICAL REASON FOR THAT?.
-IS  MetE A MULTIDOMAIN PROTEIN WHERE DOMAINS MOVE WITH RESPECT TO EACH  OTHER?} {\color{blue} The differences of TufA and MetE can be expected from some experimental data, but for IcdA it is just partial unfolding of a short N-terminal coil region.}
-
 The effect of crowded environment on the structural integrity of
@@
 access to the solvent by oligomerization. Our results show that
-oligomers are not significantly disturbed in the cytoplasm model when
-compared to the diluted condition. Similarly to what can be seem by
+oligomers are not disturbed significantly in the cytoplasm model when
+compared to the dilute condition. Similar to what can be seem from
 the RMSD values, which reflect the structural integrity of individual
-chains, the oligomers that undergo the most change is TufA
+chains, the oligomer that undergoes the largest change is TufA
 (Fig.~\ref{fig:structural_integrity_chain}B).
@@
 \caption{A) RMSD for all protein crowders from simulations under
-  diluted conditions (blue bars) and in the cytoplasm model (orange
+  dilute conditions (blue bars) and in the cytoplasm model (orange
   bars). This value reflects the structural integrity of individual
   chains of each crowder. B) Chain interface area for oligomers under
-  diluted conditions (blue bars) and in the cytoplasm model (orange
+  dilute conditions (blue bars) and in the cytoplasm model (orange
   bars). This value reflects the structural integrity of oligomers of
@@
 \subsubsection{Structural dynamics}
-% {\bf THIS PARAGRAPH REPEATS AND CONTRADICTS ITSELF. IS FLEXIBILITY
-%   AFFECTED OR NOT? FROM THE GRAPH I WOULD SAY THE THE FIVE PROTEINS
-%   NOTED BELOW ARE MUCH MORE FLEXIBLE, NOT SLIGHTLY. ONE WAY OF
-%   QUANTIFYING THIS IS TO COMPUTE THE AVERAGE OF THE MEAN SQUARE
-%   FLUCTUATION OVER THE WHOLE PROTEIN, THAT IS BEFORE TAKING SQUARE
-%   ROOTS OF COURSE AND PLOTTING THAT IN A BAR GRAPH.} {\color{blue} Overall, flexibility (meaning the values of RMSF for each residue) is affected in some cases, but the flexibility profile (meaning the overall shape of the RMSF plot) is not.
-Maybe that was the source of the apparent contradiction.
-I changed the text to clarify that.
-I also removed words such as "slightly" and "small" from the text.}
-
-Root mean square fluctuation (RMSF) calculations show that, overall, the shape of residue-wise flexibility profile of the proteins is not affected by crowding, which is in agreement with our results showing that their structural integrity does not change significantly in the cytoplasm model in most cases (Fig.~\ref{fig:rmsf}).
-Accordingly, we can see significant changes in the shape of the flexibility profile for regions that undergo conformational changes, such as the Switch I $\upbeta$-hairpin of TufA around residues 40-62 and the flexible loop of MetE around residues 450-460 (Fig.~\ref{fig:rmsf}).
+Root mean square fluctuation (RMSF) calculations show that, overall,
+the shape of residue-wise flexibility profiles of the proteins is not
+affected by crowding (Fig.~\ref{fig:rmsf}), which is in agreement with our results showing
+that their structural integrity does not change significantly in the
+cytoplasm model in most cases (Fig.~\ref{fig:structural_integrity_chain}A). 
+Accordingly, we
+can see significant changes in the shape of the flexibility profile
+for regions that undergo conformational changes, such as the Switch I
+$\upbeta$-hairpin of TufA around residues 40-62 and the flexible loop
+of MetE around residues 450-460 (Fig.~\ref{fig:rmsf}).
 
@@
 \caption{Root mean squared fluctuations (RMSF) from simulations under
-  diluted conditions (blue bars) and in the cytoplasm model (orange
+  dilute conditions (blue bars) and in the cytoplasm model (orange
   bars). The x-axis shows the residues for each chain of proteins.}
@@
 \subsubsection{Aggregation can be avoided by protonating the metabolites}
-
 The initial detection of aggregation in the cytoplasm model was due to
-the outlier behavior of the translational diffusion constant of tRNA
+the translational diffusion constant of tRNA being much lower than expected
 when compared to the other crowders. However, tRNA was not the cause of this
@@
 ATPH\textsubscript{3} and
-FBPH\textsubscript{4}(Fig.~\ref{fig:avoiding_aggregation}B-C).  
-% {\bf
-%   WERE ANY INTERMEDIATE PROTONATIONS TESTED? E.G. FBPH2(2-) AND ATPH(2-)?
-%   COMPLETELY PROTONATING IS UNPHYSICAL AS WELL AT THE PH STUDIED.
-FOR E COLI THE PH SHOULDBE ROUGHLY 7.4 https://en.wikipedia.org/wiki/Cytosol} {\color{blue}I didn't try intermediate protonations, but this can be done in about 1 week if necessary.
-The physiological states of FBP and ATP at pH 7.4 would be the ones we used originally in the soup.
-So, in this sense, any protonated state could be called unphysical or unrealistic.
-We discuss the tradeoffs of using this unphysical protonated state in the discussion section}
+FBPH\textsubscript{4} (Fig.~\ref{fig:avoiding_aggregation}B-C).
 
@@
 
-% {\bf IT MAY BE NECESSARY TO DISCUSS RMSF HERE AS WELL} {\color{blue} This first paragraph is focused on validation of the cytoplasm model by comparison with experimental results, which are available for the translational diffusion coefficient.
-We don't have specific experimental data that we can compare our RMSF data to.
-Should we add RMSF here anyway?
-There is a discussion about our RMSF results some paragraphs ahead in which we compare our data with general trends that were experimentally observed.}
-
 It has been demonstrated that the force fields that are most widely
@@
 eliminate this
-problem~\cite{Best2014a,Piana2015a,Bashardanesh2018b,Carballo2018a}. Apart from
+problem~\cite{Best2014a,Piana2015a,Bashardanesh2018b,Carballo2018a}. 
+Apart from
 protein-water interactions, the London-dispersion coefficients in the
@@
 parameters from the CHARMM family of force fields~\cite{Yu2016a}. In
-this work we use a modified version of the AMBER FF99SB force field in
+this work we use a modified version of the Amber ff99sb force field in
 which the solvent-solute interactions are 10\%
@@
 reproduce experimental data when there are many proteins in the same
-simulation box. This force field or similarly adjusted one were also
+simulation box. This force field or similarly adjusted ones were also
 used for studying crowding effect on simpler systems of high
@@
 TufA and tRNA (PDB ID 1TTT is an example of such
-complex)~\cite{nissen1995crystal}. The TufA-tRNA complex is not
-present in our simulation boxes even after \SI{1}{\micro\second} of
-simulation starting from random orientations.
-
-% {\color{blue} We highlight that there are no specific interactions described between the proteins that were used in our cytoplasm model.} {\bf do we expect specific interactions between the proteins in our cyto model?
-In that case write that with references, otherwise write that there are no known specific interactions between these proteins.} {\color{blue} The only specific interaction we antecipate between the components of the soup is that TufA interacts with tRNA (PDB ID 1TTT is an example of such complex).
-However, since we are ignoring tRNA in our analyses, I did't mention it.
-PS: I checked and the TufA-tRNA complex is not present in the last frames of our replicas.}
+complex~\cite{nissen1995crystal}). However, no TufA-tRNA complex is found
+present in our simulation boxes within three times \SI{1}{\micro\second} of
+simulations.
 
@@
 for tRNA and ATP. However, as the simulation progressed, FBP and ATP
-aggregated around Mg$^{2+}$ and, by consequence, around tRNA
-% {\bf,
-%   effectively removing the counter ions from the RNA. IS THAT
-%   CORRECT?} {\color{blue} No, Figure \ref{fig:tRNA_aggregation} shows that Mg$^{2+}$ is still concentrated around tRNA in the soup }. 
+aggregated around Mg$^{2+}$ and, as a result, around tRNA.
   After identifying this problem, we used simulations of
@@
 test different strategies to prevent aggregation in future studies
-using long simulations of cytoplasm models. We found that protonating
+of cytoplasm models. We found that protonating
 the phosphate-containing metabolites is enough to reach this goal
 (Fig.~\ref{fig:avoiding_aggregation}). Even though ATPH$_3$ and
-FBPH$_4$ are not the biologically relevant forms of these molecules in
+FBPH$_4$ are not the biologically relevant forms of these molecules at a
 physiological pH, using their protonated forms may be an acceptable
@@
 neutralization of their net charge. However, molecules in a crowded
-environment are restrained to smaller volumes and shorter inter
-molecular distances, which mitigates the effect of charge
+environment are restrained to smaller volumes and shorter intermolecular distances, which mitigates the effect of charge
 neutralization. It has been suggested that charged species should
@@
 of the electronic polarization, the screening is underestimated and this
-corresponds to applying an effective dielectric constant of about 1.7, alternatively scale the charges down by $\sqrt{1.7}$ which would mean that FBP would get a charge of approximately -3 instead of -4.
+corresponds to applying an effective dielectric constant of about 1.7, alternatively scaling the charges down by $\sqrt{1.7}$ which would mean that FBP would get a charge of approximately -3 instead of -4.
 Due to the high local
@@
 attributed this observation to non-specific interactions between such
-metabolites and proteins. Maybe aggregation also contributed to their
-slower diffusion but it was not evident during their
-\SI{140}{\nano\second} simulation.
+metabolites and proteins. Maybe aggregation also contributed to 
+slower diffusion found in that study but it was not evident during their
+\SI{140}{\nano\second} simulation~\cite{Yu2016a}.
 
@@
 composed of proteins, tRNA, metabolites, inorganic ions and water for a total
-of 1.5 million atoms (Table \ref{tbl:soup_components}). The identity
-and amount of each component of the model was chosen based on
+of 1.5 million atoms (Table \ref{tbl:soup_components}). The identities
+and amounts of each component of the model were chosen based on
 experimental data about the composition of the cytoplasm of
@@
 
-%Phosphate-containing metabolites FBP and ATP aggregated around Mg$^{2+}$ that were used as counter-ions for ATP and tRNA (Figure \ref{fig:tRNA_aggregation}).
-We used independent simulations of small systems to evaluate different strategies to prevent such aggregation and found that this can be done by completely protonating these molecules (Figure \ref{fig:avoiding_aggregation}).
-In the specific case of ATP and FBP modeled with GAFF and simulated in boxes where they are highly concentrated in the presence of Mg$^{2+}$, we advise that they should be completely protonated even though that is not their physiological protonation state.
-In principle it is possible to develop new parameters for phosphate and Mg that doesn't lead to aggregation.
-However this is outside of the scope of this study.
-
 As part of this work we are providing Python scripts with a set of
-structure and topology files that any researcher can use to build
+structure- and topology files that any researcher can use to build
 their own \textit{E. coli} cytoplasm models. Such models can be used
@@
 we are providing it is possible, for example, to add a probe protein
-to investigate the effect of crowding on it, to add new macromolecular
-or small crowders, evaluate the effects of temperatures and change the
+to investigate the effect of crowding on it, to add new small- or macromolecular
+ crowders, evaluate the effect of temperatures and change the
 biomolecular concentration to increase or decrease the intensity of
@@
 that is necessary to perform long molecular dynamics simulations. With
-modest computatioanal power it is possible, for example, to build many
+modest computational power it is possible, for example, to build many
 independent snapshots of the cytoplasm of \textit{E. coli}. All files
-are publicly available at http://github.com/dspoel/soup facilitating
+are publicly available at http://github.com/dspoel/soup, facilitating
 contributions to the scripts and data files from other workers in the field.
@@
 \begin{acknowledgement}
-The Swedish research council is acknowledged for a grant of computer time
-(SNIC2018-2-42) through the High Performance Computing
-Center North in Ume{\aa}, Sweden. Funding from eSSENCE - The e-Science
-Collaboration (Uppsala-Lund-Ume{\aa}, Sweden) and FAPESP - S{\~a}o Paulo Research Foundation (grants 2013/00927-7 and 2016/04958-2) are gratefully acknowledged.
+The Swedish research council is acknowledged for a grant of computer
+time (SNIC2018-2-42) through the High Performance Computing Center
+North in Ume{\aa}, Sweden. Funding from eSSENCE - The e-Science
+Collaboration (Uppsala-Lund-Ume{\aa}, Sweden) and FAPESP - S{\~a}o
+Paulo Research Foundation (grants 2013/00927-7 and 2016/04958-2) are
+gratefully acknowledged.
 \end{acknowledgement}

### 2019-12-03 paper/rebuttal.txt
Updated rebuttal

@@
-Dear Dr. van der Spoel:
+Dear Prof. Amaro,
 
-Thank you for submitting your manuscript to Journal of Chemical
- Information and Modeling. Please find the enclosed reviews of your
- manuscript. The reviewers have found significant problems with the
- manuscript in its current form, but we are willing to consider a
- revised manuscript in which you have made major revisions and
- responded to the reviewers' concerns. If you resubmit a revised
- manuscript, there may be a second round of external review, perhaps
- including new reviewers.
+we were happy to receive constructive comments on our manuscript. We
+have edited the manuscript according to the comments, and answered all
+the questions. Please find a detailed list of replies below.
 
-We would like to receive your revision as soon as possible, by
-17-Jan-2020 at the latest.
+I hope the manuscript is acceptable in the present form and await your
+decision.
 
-In addition to addressing the reviewers' comments, please make each of the
-technical corrections listed below:
+Best regards,
+
+David van der Spoel,
+Uppsala University
+Sweden.
 
@@
   accepted. Learn more at http://www.orcid.org.
-ANSWER:
+ANSWER: We have added one ORCID.
 
@@
   https://authoringservices.acs.org/.
-ANSWER: TODO
-
-JCIM is now accepting images to appear on the front cover of the
-journal. If you have a visually arresting and scientifically
-interesting image, please email it and a brief (80 word) description
-of the image to eic@jcim.acs.org. Additional information about the
-technical specifications for the image can be found at
-https://pubsapp.acs.org/paragonplus/submission/jcisd8/jcisd8_authguide.pdf.
+ANSWER: We carefully went over the manuscript text and fixed errors.
 
@@
 have any specific questions regarding these guidelines.
-ANSWER: A cover image will be suggested.
-
-JCIM has a Twitter feed (@JCIM_ACS; https://twitter.com/jcim_acs)
-covering the latest news and highlights in the field. If you would
-like us to feature your manuscript on this platform, please share a
-Tweet of no more than 140 characters. Please feel free to share your
-Twitter handle (if you have one) or any other relevant
-department/institutional handles. Include this information in the
-cover letter of your revised manuscript submission.
-
-ACS Publications uses Crossref Similarity Check Powered by iThenticate
-to detect instances of similarity in submitted manuscripts. In
-publishing only original research, ACS is committed to deterring
-plagiarism, including self-plagiarism. Your manuscript may be screened
-for similarity to published material.
-
-Sincerely,
-
-Prof. Rommie Amaro
-Associate Editor Journal of Chemical Information and Modeling
-Urey Hall 3234, MS - 0340,
-La Jolla, CA 92093-0340,
-United States
-Email : amaro-office@jcim.acs.org
-
------------------------------------- To revise your manuscript, log
-into ACS Paragon Plus with your ACS ID at
-http://paragonplus.acs.org/login and select "My Authoring Activity."
-Your manuscript will be listed under "Revisions and Resubmissions
-Requested by Editorial Office." Click on "Submit Revision" to begin
-the submission process. Your manuscript number will be appended to
-denote a revision.
-
-When submitting your revised manuscript, you will be able to respond
-to the comments made by the reviewer(s) in the text box provided or by
-attaching a file containing your response letter.
-
-------------------------------------
+ANSWER: We did not manage to make a cover image unfortunately.
 
@@
 discussion about the RMSF difference (please see the answer to the
-next answer for details about this).
+next question for details about this).
 
@@
 
-ANSWER: Thank you for pointingt that out. There was a mistake in the
+ANSWER: Thank you for pointing that out. There was a mistake in the
 statistics of our RMSF calculations.  We corrected it and updated the
 RMSF plots. The errors are much smaller now and the overall shift
-towards higher RMSF values for the proteins cytoplasm models is no
+towards higher RMSF values for the protein cytoplasm models is no
 longer present. Accordingly, the following sentences were removed from
@@
 Simulation". In the abstract the following sentence is added "The
-model consists of 8 different proteins and one type of t-RNA molecule,
-representing about 50% of the cytoplasmic proteins."
+model consists of 8 different proteins representing about 50% of the
+cytoplasmic proteins and one type of t-RNA molecule."
 
@@
 ANSWER: There may be a number of possible solutions to the issue. The
-work by Leotyev et al. (Refs 78 and 79) offers a plausible explanation
+work by Leontyev et al. (Refs 78 and 79) offers a plausible explanation
 of the problem, namely that in fixed-charge models, that is
